### 2021-05-12 Latex/main.tex
Removed history

\documentclass[5p]{elsarticle}

\usepackage{lineno,hyperref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Packages and custom commands added by Authors %%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{todonotes}
\usepackage{xifthen}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{breqn}
\usepackage{siunitx}
\usepackage[ruled,vlined]{algorithm2e}

\renewcommand{\vec}[1]{\mathbf{#1}}

\def\block(#1,#2)#3{\multicolumn{#2}{c}{\multirow{#1}{*}{$ #3 $}}}

\newcommand{\norm}[2][]{%
\ifthenelse{\equal{#1}{}}{\left|\left| #2 \right|\right|}{\left|\left| #2 \right|\right|_{#1}}}
\newcommand{\pardiff}[1]{\frac{\partial}{\partial #1}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\modulolinenumbers[5]

\journal{IOJPRS}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numberedss
%\bibliographystyle{model1-num-names}

%% Numbered without titlesbb
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{Towards Spherical Robots For Mobile Mapping In Human Made Environments\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{To abide the \href{https://www.go-fair.org/fair-principles/}{FAIR-Principles} of science, all data and code used to produce the results are given in this \href{https://github.com/fallow24/SphereTDP}{github} repository.}

%% Group authors per affiliation:
\author{Fabian Arzberger, Anton Bredenbeck, Jasper Zevering, Dorit Borrmann, Andreas Nüchter}
\address{Informatics VII: Robotics and Telematics,
Julius-Maximilians-University Würzburg, Germany\\
%One contact suffices
e-mail: dorit.borrmann@uni-wuerzburg.de}

%% or include affiliations in footnotes:
%\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
%\ead[url]{www.elsevier.com}

%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
%\cortext[mycorrespondingauthor]{Corresponding author}
%\ead{support@elsevier.com}

%\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
%\address[mysecondaryaddress]{360 Park Avenue South, New York}

\begin{abstract}
Spherical robots are a format that hasn't been thoroughly explored for the application of mobile mapping. 
In contrast to other designes it provides some unique advantages. 
Among those are a protective spherical shell which protects internal sensors and actuators from possible harsh environments as well as an inherent rotation for locomotion that enables measurements in all directions.
Mobile mapping always requires a high-precise pose knowledge to obtain consistent and correct environment maps. 
This is typically be done by a combination of external reference sensors such as GNSS measurements and inertial measurements, or by post processing the data and registering the different measurements to each other. 
In indoor environments the GNSS reference is not an option, hence many mobile mapping applications turn to the second option.
An advantage of indoor environments is, that human made environments usually have a certain structure, such as parallel planes and right angles.
We propose a registration procedure that exploits this structure by minimizing the distance of measured points to a corresponding plane. 
Further we evaluate the procedure on a simulated dataset of an ideal corridor as well as in experimentally acquired datasets with different motion profiles. 
We show that we nearly reproduce the ground truth for the simulated dataset and improve the average point-to-point distance to a reference scan in the experimental dataset.
The presented algorithms are required to work completely autonomous.
\end{abstract}

\begin{keyword}
\texttt{Mobile Mapping}\sep \texttt{SLAM}\sep \texttt{Spherical Robot}
\MSC[2021] 00-01\sep  99-00
\end{keyword}

\end{frontmatter}

\linenumbers

\input{sections/introduction}
\input{sections/stateOfTheArt}
\input{sections/registration}
\input{sections/simulation}
\input{sections/realData}
\input{sections/results}
\input{sections/conclusion}

\bibliography{mybibfile,andreas_publications}

\end{document}

### 2021-05-12 Latex/sections/conclusion.tex
Removed history

%!TEX root = ../main.tex
\section{Conclusion}

In this paper, we propose a registration method for man made environments that exploits the structure of the environments. 
Given in man-made environments parallel planes and right angles are abundantly available we employ point-to-plane correspondences to improve a pre-registered 3D point-cloud. 
We evaluate the procedure on a simulated dataset as well as on a experimentally acquired dataset with different motion profiles. 
In this evaluation we show that the procedure improves both datasets and yields more human-made structure like maps. 
Currently not all steps in the procedure are autonomous, in particular the parameter tuning. 
In the future one goal is to increase the autonomy of the system.
One approach is to introduce soft-locks for the optimization dimensions.
I.e., instead of locking the some dimensions entirely from being used for optimization, adding weighting factors based on the dynamics of the system that encode which noise source is more likely.

### 2021-05-12 Latex/sections/introduction.tex
Removed history

%!TEX root = ../main.tex
\section{Introduction}

Today's robots for mobile mapping come in all shapes and sizes.
The state of the art for urban environments are laser scanners mounted to cars.
Smaller robotic systems are particularly used when cars no longer have access.
Examples for this are human operated systems such as Zebedee~\cite{FIXME}, a Hokuyo scanner on a spring, VILMA~\cite{JPRS2016}, a rolling FARO scanner, RADLER~\cite{Borrmann2020-RADLER}, a 2D laser scanner mounted to a unicycle, or a backpack mounted ``personal laser scanning system'' as in~\cite{LauterbackEtAl2015-Backpack}.
Recently more and more autonomous systems gained maturity.
A stunning example is Boston Dynamics' quadruped ``Spot'' that autonomously navigates and maps human environments~\cite{SpotRobot}.
Also the mobile mapping approaches implemented on the ANYmal platform such as~\cite{Fankhauser2018-ANYmal} were very successful.
Of all these formats, one hasn't been explored thoroughly in the scientific community: The spherical mobile mapping robot.
Yet this provides some very promising advantages over the other formats.
For one, the locomotion of a spherical robot inherently results in a rotation.
That way a sensor fixed inside the spherical structure will cover the entire environment, given the required locomotion without the need for additional actuators for the sensors.
This requires a solution for the spherical simultaneous localization and mapping (SLAM) problem given the six degrees of freedom of the robot.
Secondly a spherical shell, that encloses all sensors protects these from possible hazardous environments.
For example, any dust that deteriorates sensors or actuators when settling at sensitive locations is stopped by a shell.
This is be particularly useful for unknown and dangerous environments.
E.g., for space applications this is also proposed in the DAEDALUS study~\cite{RossiMaurelliUnnithanetal.2021}, a spherical robot that is to be lowered into a lunar cave and create a 3D map of the environment. The authors choose this format as the present moon regolith is known to damage instruments and other components.

This paper proposes to use such a spherical robot for mobile mapping man made environments.
In such environments one advantage are architectural shapes following standard conventions arising from tradition or utility, i.e., there are many flat surfaces such as walls, floors etc. that are sensed.
Exploiting this fact yields more opportunities for registration as point-to-plane correspondences can be used.
The proposed registration method minimizes the distances of each point to its corresponding plane as objective function.
The method is evaluated on a simulated dataset as well as on experimentally acquired datasets.

### 2021-05-12 Latex/sections/realData.tex
Removed history

%!TEX root = ../main.tex
\section{Experimental Setup}\label{sec:experimentalSetup}

Our algorithm was also tested on a number of real-world datasets that adhere to different motion profiles and laser scanners with different scanning patterns.

\subsection{Floating Sphere}

The first dataset is collected by a line scanner, in particular a SICK LMS141 industrial scanner, inside a sphere of a diameter of \SI{20}{\centi\meter} that lies on a floating desk, which is air-pressurized, such that the sphere is hovering.
On this floating desk the sphere can rotate freely about all axis while being fixed in position. 
Hence, in this experiment the optimization space can be reduced to a rotation but also the motion is required to obtain a 3D model from the 2D laser scanner. 
The sphere is equipped with a low-cost IMU, i.e., a PhidgetSpatial Precision 3/3/3, to estimate the initial orientation.

Figure~\ref{fig:float-sat-sphere} shows the experimental setup. TThe left column of figure~\ref{fig:cylon-corrected} shows the pre-registered resulting point cloud . 
We see, that the point cloud is rather noisy and in particular the walls are sensed multiple times, hence appearing very blurry.

\begin{figure}
	\centering
	\begin{minipage}[c]{0.5\textwidth}
		\includegraphics[width=0.33\textwidth]{./images/sphere-frame-1.eps}\hfill
		\includegraphics[width=0.33\textwidth]{./images/sphere-frame-2.eps}\hfill
		\includegraphics[width=0.33\textwidth]{./images/sphere-frame-3.eps}\hfill
	\end{minipage}\\
	\begin{minipage}[c]{0.5\textwidth}
		\includegraphics[width=0.33\textwidth]{./images/sphere-frame-4.eps}\hfill
		\includegraphics[width=0.33\textwidth]{./images/sphere-frame-5.eps}\hfill
		\includegraphics[width=0.33\textwidth]{./images/sphere-frame-6.eps}\hfill
	\end{minipage}\\
	\caption{Data acquisition using floating sphere. We show a sequence of orientations the sphere assumes.}
	\label{fig:float-sat-sphere}
\end{figure}

\subsection{RADLER}

The second dataset is recorded using the RADLER system as described in~\cite{Borrmann2020-RADLER}.
While not a spherical robot, this system shows many similarities with respect to its motion with a spherical robot.
The system was manually steered along a squared set of hallways that loop back to the initial starting point inside the old mathematics building at the University of Würzburg (denoted ``The Circle'' in~\cite{Borrmann2020-RADLER}).

Figure~\ref{fig:radler-mathe} shows the system setup as well as the resulting point cloud.
We see that with continuing measurement time the drift increases and the hallways are senses multiple times with a clear offset to each other.

\begin{figure*}
	\centering
	\begin{minipage}[c]{\textwidth}
		\includegraphics[width=0.2925\textwidth]{./images/radler_setup}\hfill
		\includegraphics[width=0.7\textwidth]{./images/mathe_top_view}
	\end{minipage}
	\caption{Data acquisition using the RADLER system. Left: System setup (courtesy of Borrmann~\cite{Borrmann2020-RADLER}). Right: Unprocessed resulting point cloud.}
	\label{fig:radler-mathe}
\end{figure*}

\subsection{Descend Sphere REMOVE THIS SECTION}

This was recorded using a LIVOX-Mid 100 laser scanner mounted inside a transparent shell. 
The robot contained three Phidget 3/3/3 Spatial IMUs and was attached to a \SI{50}{\meter} tether cable that was rolled on a spool which in turn was connected to a angular encoder. 
From these sensors we obtained orientation information as well as height information. 
The setup was then brought into a high indoor building (\href{https://www.sfs-w.de/feuerwehrschule/virtueller-rundgang.html}{Firefighter School in Würzburg}) where the system was manually descended from a balcony scanning the building interior.
The robot covered a distance of approximately \SI{22}{\meter} in about \SI{402}{\second}.
Additionally the building was scanned using a terrestrial laser scanner to provide a reference.
The measurements were then rigidly corrected before applying our plane based correction. 
Figure~\ref{fig:experimental-setup} shows the acquired 3D point clouds and the experimental setup.

\begin{figure}
	\centering
	\begin{minipage}[c]{0.25\textwidth}
		\includegraphics[width=\textwidth]{images/scan-riegl-fire}\\
		\includegraphics[width=\textwidth]{images/scan-sphere-fire}
  	\end{minipage}\hfill
  	\begin{minipage}[c]{0.2075\textwidth}
  		\includegraphics[width=\textwidth]{./images/lidarsetup}
  	\end{minipage}
	\caption{Reference 3D point cloud acquired with a RIEGL 3D terrestrial laser scanner (top left). The 3D point cloud acquired by the test sphere after the rigid registration (bottom left) and a picture of the test setup (right).}
	\label{fig:experimental-setup}
\end{figure}

### 2021-05-12 Latex/sections/registration.tex
Removed history

%!TEX root = ../main.tex
\section{Registration Algorithm}

We start with transforming each line scan into the project coodinate system , which is defined by the pose of the first acquired line scan.
With line scan, we mean the smallest chunk of range data, we obtain from the scanner device driver.
In case of a SICK LMS1xx it is a line and in case of a Livox scanner it has a flower shape.
More details are provided in the following sections.
%
For map improvement the individual scans need to be registered to another. 
We propose an algorithm that consist of multiple steps outlined in algorithm~\ref{algo:registration-algorithm}. 
Based on ideas described in~\cite{Borrmann2010} we first rigidly register the scans and then further improve the overall map by exploiting the fact that man made environments often consist of planes. 
We then find the planes in the pre-registered point-cloud and then optimize the poses associated with the scans to minimize the distance of all points to their respective planes.

\begin{algorithm}
    \SetAlgoLined
    \KwResult{A corrected map}
    1. Rigidly register the scans\;
    2. Extract planes from the registered map\;
    3. Further improve map by solving the optimization that minimizes the distance of all points to their respective planes\;
    \caption{Registration algorithm for man-made environments}
    \label{algo:registration-algorithm}
\end{algorithm}

\subsection{Plane Extraction}

After having done the rigid registration the scans are aligned well-enough to make statements about the potential planes in the environment.
To find the planes in the environment a Hough transform with an accumulator ball as described in~\cite{Borrmann2011The3H} is used. 
In the next subsection the Hesse-normal form of the plane is required.
For an ideal plane the distance parameter $\rho_{\mathcal{P}_i}$ is computed via $\vec{n}_{\mathcal{P}_i}\cdot\vec{x}_i$, where $\vec{x}_i$ is an arbitrary point on the plane.
To find this point we first define a convex hull that encloses the plane.
It is defined by the points that lie on the plane with the furthest distance to one another.
We then choose the center point of the convex hull of the plane as $\vec{x}_i$.

\subsection{Point-to-Plane Correspondences}

There are many ways of finding correspondences between points and planes.
However, even with the result not being perfect, we assume already globally registered scans, which is why a simplistic algorithm is sufficient:
For each plane, consider adding a correspondence for all points that are closer to the plane than a threshold $\epsilon$.
Depending on the threshold value, some points now have multiple correspondences if their distance to multiple planes is less then $\epsilon$. We then simply omit all points that have correspondences to multiple planes from the optimizations, to avoid introducing errors due to the plane ambiguity.

\todo[inline]{Add section about the local plane, k-means normal based region growing aka clustering for point2plane correspondences. 
Basically, if a point is correspondence to a global plane than all other points that lie on its local plane should be too. The local plane is constructed by region growing, i.e. if the difference in normal and the distance of the points are close enough they are clustered to one local plane.}

\subsection{Optimization}

Assuming we now know the Hesse-normal form of all planes and all points that are assigned to these planes, we register the points to have a minimal distance to their respective planes.
The transformation $T(\vec{p}_k)$ of each point $\vec{p}_k$ with respect to a 6 DoF motion is described in homogeneous coordinates using the roll-pitch-yaw ($\varphi-\vartheta-\psi$) Tait-Brian angles as in~\cite{diebel2006representing}. Transforming the result back from homogeneous coordinates and using $C_k$ and $S_k$ to denote the cosine and sine of the angle in the subscript yields:
\begin{align}
	T(\vec{p}_k)  =
    \resizebox{0.35\textwidth}{!}{$\begin{bmatrix}
        x C_\vartheta C_\psi - yC_\varphi S_\psi + z S_\vartheta + t_x\\
        x (C_\varphi S_\psi + C_\psi S_\varphi S_\vartheta) + y(C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi) - z C_\vartheta S_\varphi + t_y\\
        x(S_\varphi S_\psi - C_\varphi C_\psi S_\vartheta) + y(C_\psi S_\varphi + C_\varphi S_\vartheta S_\psi) + z C_\varphi C_\vartheta + t_z
    \end{bmatrix}$}
\end{align}
From this we define the function $D(\varphi,\vartheta,\psi,t_x,t_y,t_z, \vec{p}_k)$ that computes the distance of a point $\vec{p}_k$ to its corresponding plane $\mathcal{P}_i$. Omitting the arguments of the function for simplicity:
\begin{align}
\begin{aligned}
    D &= T(\vec{p}_k) \cdot \vec{n}_{\mathcal{P}_i} \\
      &= n_{\mathcal{P}_i}^x (x C_\vartheta C_\psi - yC_\varphi S_\psi + z S_\vartheta + t_x)\\
       &+ n_{\mathcal{P}_i}^y(x (C_\varphi S_\psi + C_\psi S_\varphi S_\vartheta)\\
       &\qquad+ y(C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi) - z C_\vartheta S_\varphi + t_y)\\
       &+ n_{\mathcal{P}_i}^z (x(S_\varphi S_\psi - C_\varphi C_\psi S_\vartheta)\\
       &\qquad+ y(C_\psi S_\varphi + C_\varphi S_\vartheta S_\psi) + z C_\varphi C_\vartheta + t_z)\\
       &- \rho_{\mathcal{P}_i}
\end{aligned}
\end{align}

This distance function is what we want to minimize for all points and their respective planes.
Hence the error function $E$ is chosen as follows:
\begin{align}
	E = \sum_{\forall \mathcal{P}_i}\sum_{\vec{p}_k \in \mathcal{P}_i} \norm[2]{D(\varphi,\vartheta,\psi,t_x,t_y,t_z, \vec{p}_k)}^2
\end{align}
Its gradient follows then immediately:
\begin{align}
	&\nabla E =  \sum_{\forall \mathcal{P}_i}\sum_{\vec{p}_k \in \mathcal{P}_i} \resizebox{0.28\textwidth}{!}{$\begin{bmatrix}\pardiff{\varphi} E&\pardiff{\vartheta} E&\pardiff{\psi} E&\pardiff{t_x} E&\pardiff{t_y} E&\pardiff{t_z} E\end{bmatrix}^T$}\\
    &\Rightarrow\nabla E = \sum_{\forall \mathcal{P}_i} \sum_{\vec{p}_k \in \mathcal{P}_i}\resizebox{0.28\textwidth}{!}{$2D(\vec{\Pi}, \vec{p}_k)\begin{bmatrix}\nabla E_\varphi &\nabla E_\vartheta & \nabla E_\psi & n_{\mathcal{P}_i}^x&n_{\mathcal{P}_i}^y&n_{\mathcal{P}_i}^z\end{bmatrix}^T$}
\end{align}
Where
\begin{align}
    &\begin{alignedat}{3}
        \nabla E_\varphi = & n_{\mathcal{P}_i}^x y S_\varphi S_\psi \\
        &+ n_{\mathcal{P}_i}^y&&(x[-S_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] \\&&&+y[-S_\varphi C_\psi -C_\varphi S_\vartheta S\psi] \\&&&- z C_\varphi C_\vartheta)\\
        &+ n_{\mathcal{P}_i}^z&&(x[C_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] \\&&&+y[C_\varphi C_\psi -S_\varphi S_\vartheta S\psi] \\&&&- z S_\varphi C_\vartheta)
    \end{alignedat} \\
    &\begin{alignedat}{3}
        \nabla E_\vartheta = & n_{\mathcal{P}_i}^x(-xS_\vartheta C_\psi + zC_\vartheta) \\
        &+ n_{\mathcal{P}_i}^y(xC_\psi S_\varphi C_\vartheta - yS_\varphi C_\vartheta S_\psi + z S_\vartheta S_\varphi)  \\
        &+ n_{\mathcal{P}_i}^z(-xC_\varphi C_\psi C_\vartheta + yC_\varphi C_\vartheta S_\psi - z C_\varphi S_\vartheta)
    \end{alignedat}\\
    &\begin{alignedat}{3}
       \nabla E_\psi =& n_{\mathcal{P}_i}^x\left(-xC_\vartheta S_\psi - yC_\varphi C_\psi\right) \\
       &+ n_{\mathcal{P}_i}^y\left(x[C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi] + y[-C_\varphi S_\psi - S_\varphi S_\theta C_\psi]\right)  \\
       &+ n_{\mathcal{P}_i}^z\left(x[S_\varphi C_\psi + C_\varphi S_\psi S_\theta] + y[-S_\psi S_ \varphi + C_\varphi S_\vartheta C_\psi] \right) 
    \end{alignedat}
\end{align}
and $\vec{\Pi}=\begin{bmatrix}\varphi & \vartheta & \psi & t_x & t_y & t_z\end{bmatrix}^T$.

As the gradient is well-defined we minimize the error function with any gradient based method. 
The commonly used, well-known stochastic gradient descent (SDG) algorithm computes 
\begin{align}
    \vec{\Pi}_{k+1} = \vec{\Pi}_{k} - \alpha \nabla E
\end{align}
To accelerate convergence and to improve the found solution further modifications are made.

Since we have vastly different effects on the error function by each dimension, the first consideration for improving the SDG is the following:
Typically, changes in orientation, i.e., the first three elements of the gradient vector $\pardiff{\varphi}E$, $\pardiff{\vartheta}E$, and $\pardiff{\psi}E$, have much more impact on the error function than a change in position.
This is intuitively explained since translating the scan makes the error grow linearly for all points.
However, when rotating the scan, points with a larger distance to the robot are moved drastically, leading to a higher sensibility on the error function.
For this reason, the $\alpha$ applied on orientation has to be much smaller than the $\alpha$ applied on the position.
It becomes obvious that $\alpha$ needs to be extended into vector form, $\boldsymbol\alpha$, therefore weighting each dimension differently.

Another consideration to speed up SDG is to adaptively recalculate $\boldsymbol\alpha$ for each iteration. 
We employ and modify ADADELTA as a technique to do so, which is described in detail in~\cite{zeiler2012adadelta}.
The main idea is the following:
It extends the SDG algorithm by two terms.
First, an exponentially decaying average of past gradients, which is recursively defined as
\begin{align}
    \vec{G}_{k+1} = \rho \vec{G}_{k} + (1 - \rho) {\nabla E}^2
\end{align}
and second, an exponentially decaying average of past changes, which is defined as
\begin{align}
    \vec{X}_{k+1} = \rho \vec{X}_{k} + (1 - \rho) {\boldsymbol\alpha \nabla E}^2
\end{align}
where $\rho \leq 1$ is a decay constant, typically close to $1$.
The root mean squared (RMS) of these quantities are
\begin{align}
    RMS[\vec{G}]_{k} = \sqrt{\vec{G}_{k} + \epsilon}
\end{align}
and 
\begin{align}
    RMS[\vec{X}]_{k} = \sqrt{\vec{X}_{k} + \epsilon}
\end{align}
where $\epsilon > 0$ is a very small constant, typically close to $0$.
It will prevent dividing by zero in the recalculation of $\boldsymbol\alpha$, which is as follows:
\begin{align}
    \boldsymbol\alpha_{k} = \frac{RMS[X]_{k-1}}{RMS[G]_{k}}
    \label{eq:adadeltaalphaupdate}
\end{align} 
For our particular application ADADELTA behaves a little too aggressively.
Despite giving a good measure on how to adabt $\boldsymbol\alpha$, the algorithm sometimes overshoots, and convergence doesn't happen.
Therefore, we employ another scaling factor, typically not found in ADADELTA, extending eq.~\eqref{eq:adadeltaalphaupdate} to:
\begin{align}
    \boldsymbol\alpha_{k} = \boldsymbol\alpha_0 \cdot \frac{RMS[X]_{k-1}}{RMS[G]_{k}}
    \label{eq:adafinalized}
\end{align} 
where $\boldsymbol\alpha_0 $ holds the scaling factors for each dimension.

Finally, the SDG model is improved using eq.~\eqref{eq:adafinalized} and extends to 
\begin{align}
    \vec{\Pi}_{k+1} = \vec{\Pi}_{k} - \boldsymbol\alpha_0  \frac{RMS[X]_{k-1}}{RMS[G]_{k}} \cdot \nabla E
\end{align}
Using this algorithm once after finding correspondences from points to planes leads to convergence to a local minimum, which often isn't an optimal solution.
Even if the number of iterations is dramatically increased, no better solution than the local minimum is found.
That is, unless you consider updating the correspondence model after $i$ iterations of gradient descent.
Re-assigning point-to-plane correspondences this way $k$ times, and if $k$ is chosen large enough, it leads to an optimal solution after maximum $n = k\cdot i$ iterations of gradient descent.

\subsection{Further Optimizations of the Algorithm}

The algorithm was extended by two further optmizations which are useful in different scenarios.

Firstly, we introduce a lock on some dimensions from being optimized by setting the corresponding $\alpha_i$ to zero.
Although 6D optimization generally works, reducing the optimization space is particularly useful if the source of error in the system is known and a model exists.
That way, e.g. a point cloud from a terrestrial lasescanner that experiences irregularites in its spinning mirror is registered by optimizing only over the axis of rotation as no other movement is to be expected.

Secondly we employ a continuous iteration, where the scans are processed sequentially. 
This is especially useful for mobile systems, where pose error accumulates due to increasing tracking uncertainties.
The assumption is that the the error in one scan is also present in the next scan, plus some unknown new error.
We eliminate the error from scan $m$ which is also present in scan $m+1$ by applying the pose change from scan $m$, $\vec{\Pi}_{n,m} - \vec{\Pi}_{0,m}$ after $n$ gradient descent iterations, to scan $m+1$, before restarting gradient descent.

### 2021-05-12 Latex/sections/results.tex
Removed history

%!TEX root = ../main.tex
\section{Results}

\subsection{Simulated Results}

The plane-based registration is applied to the simulated dataset with noisy pose and range measurements (cf. Figure~\ref{fig:simulatedDatasets}) without further processing.
Assuming this represents a coarsely pre-registered 3D point cloud the distances to the ground truth were evaluated before and after the plane-based registration. 
Figure~\ref{fig:simulatedEvaluation} shows the different point-to-point distances.

\begin{figure*}
 	\centering
 	\begin{minipage}[c]{0.49\textwidth}
 		\centering
		\includegraphics[width=\textwidth]{./images/uncorr_bottom_pose}\\
		\includegraphics[width=\textwidth]{./images/uncorr_side_view}\\
  		\includegraphics[width=\textwidth]{./images/uncorr_hist}
  	\end{minipage}\hfill
  	\begin{minipage}[c]{0.49\textwidth}
  		\centering
		\includegraphics[width=\textwidth]{./images/corr_bottom_pose}\\
		\includegraphics[width=\textwidth]{./images/corr_side_view}\\
  		\includegraphics[width=\textwidth]{./images/corr_hist}
  	\end{minipage} 	
 	\caption{Evaluation of point distances before (left) and after (right) the plane based registration on a simulated dataset. A maximal distance of \SI{2}{m} is set, such that all points that display a higher distance value are excluded from analysis. Further the histogram joins all values with a distance greater than \SI{1}{m} into the last bucket. The top two column shows a heat map of distances while the bottom shows the corresponding histogram. The color mapping is equivalent in both. An animation of the matching process is given at~\href{https://youtu.be/0ps5Pg4qo4E}{here}}
 	\label{fig:simulatedEvaluation}
\end{figure*}

Before the registration the corridor is only represented acceptably in the front part. 
The further into the corridor, i.e. the longer the robot accumulates error, the more imprecise the date becomes. 
Finally we see that a large portion of the data even exceeds either the \SI{1}{\meter} error, contributing to the large spike in the histogram or even the \SI{2}{\meter} error limit thus, being cut from the representation. 
After registration, we see that, qualitatively, the ideal corridor was nearly restored from the noisy data. 
In particular a very large portion (95\%) have less than \SI{17}{\centi\meter} distances.
Further the square and straight shape of the corridor is restored well and especially the large spike at errors of greater than \SI{1}{m} is remove. 
Any such errors tend to occur at the back and the front of the corridor where the measured range is the largest, hence has the largest contribution of the range error.

\subsection{Floating Sphere Results}

Figure~\ref{fig:cylon-corrected} shows the results obtained after employing the plane base registration on the dataset acquired by the floating sphere.
\begin{figure*}
	\centering
	\includegraphics[width=0.49\textwidth]{./images/cylon_uncorr_corner}\hfill
	\includegraphics[width=0.49\textwidth]{./images/cylon_corr_corner}\\
	\includegraphics[width=0.49\textwidth]{./images/cylon_uncorr_top}\hfill
	\includegraphics[width=0.49\textwidth]{./images/cylon_corr_top}
	\caption{The point cloud acquired by the floating sphere before (left) and after (right) applying the plane based registration. View from the interior that shows the used planes (top) and a birds-eye view (bottom). An animation of the registration process is given at \href{https://youtu.be/8XdIUN_9VpY}{here}.}
	\label{fig:cylon-corrected}
\end{figure*}
After the registration the walls of the room are significantly more prominent in the point-cloud. 
In particular the noise in the top left corner of the top view (cf. figure~\ref{fig:cylon-corrected}) has been visibly reduced.
Further, the noise around the walls is notably reduced as the points are moved on the plane.

However, some imprecisions remain. 
At the same location a few points cannot be matched to the plane as they are clustered with some other points that are bound to a different pose. 
Hence this error stems from the imprecise interpolation in between scans. 
\todo[inline]{More comments here}

\subsection{RADLER Results}

\subsection{Experimental Results REMOVE THIS SECTION}

For the experimentally acquired datasets a similar procedure is followed. 
The main difference being, that the scan was pre-registered using a point-to-point based method as described in section~\ref{sec:experimentalSetup}. 
Here also the point distances to the ground truth (in this case the terrestrially acquired 3D laser scan) is evaluated before and after the plane-based registration process.

\begin{figure}
	\centering
	\begin{minipage}[c]{0.25\textwidth}
		\centering	
		\includegraphics[width = \textwidth]{./images/distances_fire}\\
		\includegraphics[width = \textwidth]{./images/histogram_fire}
  	\end{minipage}\hfill
  	\begin{minipage}[c]{0.25\textwidth}
  	\end{minipage}

	\todo[inline]{Evaluate Plane matching on firefighter school data; histogram after etc}
	\caption{Evaluation of point distances before (left) and after (right) the plane based registration on a experimentally acquired dataset. A maximal distance of \SI{5}{m} is set, such that all points that display a higher distance value are excluded. The left column shows a heat map of distances while the right shows the corresponding histogram. The color mapping is equivalent in both. An animation of the matching process can be seen \href{todo}{here}}
	\label{fig:experimentalEvaluation}
\end{figure}

### 2021-05-12 Latex/sections/simulation.tex
Removed history

To quantify the quality of the proposed registration method, it is tested on simulated as well as on real world data.

%!TEX root = ../main.tex
\section{Simulation}

For the simulated data we implemented a noisy world-robot-sensor model.  
We are simulating a LIVOX-Mid100~\cite{LivoxMid40-100} laser scanner with customizable noise level on the range measurements inside a robot with different motion capabilities, subject to noise in its pose estimation. This yields scan results similar to the ones obtained in the real world.

\subsection{Range Noise Model}

Most real laser scanners are subjected to range measurement noise that is proportional to the measured range. 
Therefore the simulated sensors are as well. 
To achieve this we sample a noise percentage $n_P$ from some normal distribution $\mathcal{N}(\mu,\sigma^2)$. 
The resulting range $r$ given the true range $r_t$ is then
\begin{align}
	r = r_t(1+n_P)
\end{align}
For each measurement ray the noise percentage is sampled independently.

\subsection{Pose Estimation Noise Model}

For the pose measurement one cannot simply add white noise to the current pose estimate as this does not capture the integration error which is common among inertial measurement units (IMU). 
Therefore we assume a disturbance torque about the two axes that lie in the ground plane. 
This is equivalent to a case, where a slightly shifted ground plane or a unbalanced locomotion system of the robot is present.

Accordingly, a spherical robot that is assumed to be translating exactly along one axis by rotation exactly about one other axis at an angular velocity $\omega$  experiences random disturbance torques at each time instant that accelerate the sphere's rotation about the respective axes. 
This additional motion is determined by integrating the disturbance torques which are sampled from some normal distribution $n_\phi, n_\psi \in \mathcal{N}(\mu,\sigma^2)$ each.  
The pose update in the discrete simulation is therefore given by: 
\begin{align}
	\begin{bmatrix}\varphi\\\vartheta\\\psi\\x\\y\\z\end{bmatrix}_{k+1} = 
	\begin{bmatrix}\varphi\\\vartheta\\\psi\\x\\y\\z\end{bmatrix}_{k} 
	+ \begin{bmatrix}1\\0\\0\\R\\0\\0\end{bmatrix}\omega \Delta t  + \Delta t \underbrace{\sum_{i = 0}^k \begin{bmatrix}n_\varphi[i] \Delta t\\0\\n_\psi[i]\Delta t\\0\\0\\0\end{bmatrix}}_{\substack{\text{Difference in}\\\text{angular velocity}}}
\end{align}

Assuming intended locomotion along the $z$-axis and unintended locomotion along the $x$-axis where $R$ is the sphere radius.
No movement upwards is possible since the robot moves on a flat ground, hence the entry is zero.

\subsection{Simulated Datasets}

Figure~\ref{fig:simulatedDatasets} shows a noise free trajectory through a simulated environment as well as a trajectory with noisy sensors.
In the noise-free case, the trajectory slightly bends in the direction of disturbance torque sideways while also varying in velocity in the intended direction of travel.
In the noisy case the ideal straight trajectory is assumed, hence the planes enclosing the room are sensed multiple times. 
In particular we see that the trajectory of the robot leads through one of the sensed planes. 
Also, the further the sensed points are the noisier they appear, which is consistent with the larger influence of the pose error at higher distances.

\begin{figure*}
	\centering
	\includegraphics[width=0.51\textwidth]{images/perfect_bottom}\hfill
	\includegraphics[width=0.48\textwidth]{images/noisy_pose_and_range_bottom}\\
	\includegraphics[width=0.51\textwidth]{images/perfect_top}\hfill
	\includegraphics[width=0.48\textwidth]{images/noisy_pose_and_range_top}\\
	\includegraphics[width=0.51\textwidth]{images/perfect_top_view}\hfill
	\includegraphics[width=0.48\textwidth]{images/noisy_pose_and_range_top_view}\\\vspace{0.5cm}
	\includegraphics[width=0.25\textwidth]{images/sim_frame_00.png}\hfill
	\includegraphics[width=0.25\textwidth]{images/sim_frame_02.png}\hfill
	\includegraphics[width=0.25\textwidth]{images/sim_frame_06.png}\hfill
	\includegraphics[width=0.25\textwidth]{images/sim_frame_08.png}
	\caption{Two simulated datasets of a long hallway of size $\SI{4}{\meter}\times\SI{3}{\meter}\times\SI{100}{\meter}$. Ideal noise free (left), noisy pose ($\mathcal{N}_\varphi = \mathcal{N}_\psi(\mu = 0.0001, \sigma= 0.00001)$) and noisy range measurements ($\mathcal{N}_r(\mu = 0, \sigma = 0.001)$) (right). Below: Initial sequence of the simulation of the ideal dataset showing the simulated motion of the robot.}
	\label{fig:simulatedDatasets}
\end{figure*}

### 2021-05-12 Latex/sections/stateOfTheArt.tex
Removed history

%!TEX root = ../main.tex
\section{State Of The Art}

\subsection{Spherical SLAM}

Up to the best of our knowledge, SLAM with spherical robots and laser scanners was not done before, while there exists camera based approaches~\cite{9233654}.
However, laser-based SLAM algorithms for motions in six degrees of freedom (DoF) have been thoroughly studied.
For outdoor environments~\cite{JFR2006} provides a first baseline. Adding a heuristic for closed loop detection and a global relaxation Borrmann et al. yield highly precise maps of the scanned environment~\cite{RAS2007}.
Zhang et al. propose a real-time version of the SLAM algorithm in~\cite{Zhang2014}.
They achieve the performance at a lower computational load by dividing the SLAM algorithm into two different algorithms: one performing odometry at a high frequency but low fidelity and another running at a lower frequency performing fine matching and registration of the point clouds.
The more recently Dröschel et al. also proposes an online method using a novel combination of a hierarchical graph structure with local multi-resolution maps to overcome problems due to sparse scans~\cite{Droeschel2018}.

Since these approaches are based on point-to-point correspondences they require rather high point density to achieve precise registration.
For low-cost LiDARs this implies a slow motion and long integration time.
Further, none of the approaches exploit the structure inherent to man-made environments, hence missing out on possible advantages.

\subsection{Point Cloud Registration Using Plane Based Correspondences}

The de-facto standard for many SLAM algorithms is the Iterative-Closest-Point (ICP) algorithm~\cite{Besl1992} that employs point-to-point correspondences using closest points, as the name suggests. 
To overcome the requirements on point-density imposed by the point-to-point correspondences instead other correspondences are used. 
In man made environments planes are abundantly available, and hence provide an attractive base for correspondences.
Förster et al. use this property of man made environment successfully in~\cite{Foerstner2017}.
They register point clouds using plane-to-plane correspondences and include uncertainty measures for the detected planes as well as the estimated motion.
Thereby, they propose a costly exact algorithm and cheaper approximations that yield high quality maps.
Favre et al.~\cite{favre2021} use point-to-plane correspondences after preprocessing the point clouds using plane-to-plane correspondences to successfully register two scans with each other.

Both approaches use plane-to-plane correspondences to pre-register the scans. However for pre-registration the classical point-to-point registration is also very effective. 
One advantage that point-to-point correspondences have over plane-to-plane correspondences is, that they do not require a long stop in each pose. For plane-to-plane correspondences this is required to gather enough data to measure planes in each individual scan robustly.
The resulting scan procedure is stop-scan-and-go.
In particular for the application of a spherical robot this standstill in each pose cannot be guaranteed or even approximated making continuous-time approaches using point-to-plane correspondences the method of choice.

In the following we propose this combination of point-to-point based pre-registration followed by a point-to-plane based optimization.

### 2021-05-12 tdp/DataSetSim/Miscellaneous/pattern_test.py

@@ -1,0 +1,22 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+elev_max = np.pi / 180.0 * 38.4
+azi_max = np.pi / 180.0 * 38.4
+
+n = np.sqrt(20)
+s = 100
+t = np.arange(0, 1, 1e-5)
+
+x = np.tan(elev_max / 2) * np.sin(s *
+                                  (n * t + 0.25 * np.pi)) * np.sin(s * t)  #
+y = np.tan(azi_max / 2) * np.sin(s * (n * t + 0.25 * np.pi)) * np.cos(s * t)  #
+
+
+elev = np.arctan2(y, 1)
+azi = np.arctan2(x, 1)
+
+fig, (ax1, ax2) = plt.subplots(2, 1)
+ax1.plot(x, y)
+ax2.plot(180.0 / np.pi * elev, 180.0 / np.pi * azi)
+plt.show()

### 2021-05-12 tdp/DataSetSim/constants.hpp

@@ -1,0 +1,8 @@
+#ifndef CONSTANTS_H
+#define CONSTANTS_H 
+
+const int VERBOSITY_LEVEL = 1;
+const double MIN_SIM_RESOLUTION = 0.1;
+
+
+#endif //CONSTANTS_H

### 2021-05-12 tdp/DataSetSim/dataset_sim.cpp

@@ -1,0 +1,188 @@
+#include <iostream>
+#include <iomanip>
+#include <fstream>
+#include <sstream>
+#include <omp.h>
+
+#include "dataset_sim.hpp"
+#include "constants.hpp"
+#include "parsing.hpp"
+
+using namespace std;
+
+string Dataset_Simulator::get_ray_result(double elevation, double azimuth)
+{
+    vector<double> root{0.0,0.0,0.0},
+                   slope{0.0,0.0,0.0},
+                   intersection{0.0,0.0,0.0};
+    /* Compute robot frame representation of ray:
+     * The root is always in the origin of the robot, hence we can leave it as is,
+     * For the slope we use the azimuth and elevation to find the ray
+     * We define
+     *    azimuth := angle from z-axis to point in z-x-axis
+     *    elevation := angle from z-axis to point in z-y-axis */
+    slope[0] = - cos(elevation)*sin(azimuth);
+    slope[1] = sin(elevation);
+    slope[2] = cos(azimuth)*cos(elevation);
+    // Normalize slope
+    double norm_factor = 1/sqrt(slope[0]*slope[0] + slope[1]*slope[1] + slope[2]*slope[2]); 
+    slope[0] *= norm_factor;
+    slope[1] *= norm_factor;
+    slope[2] *= norm_factor; 
+    //Get robot pose
+    auto pose = this->get_robot()->get_mover()->get_real_pose();
+    // transfer to world coordinate frame
+    robot2world_frame(pose, false, root);
+    // Slope only needs to be rotated, not translated
+    robot2world_frame(pose, true, slope);
+    // find the intersection vector
+    this->get_world()->get_ray_intersection(root,slope, intersection);
+
+    if(VERBOSITY_LEVEL == 2)
+    {
+      cout << "Intersection World coordinate: " << intersection[0] << " "
+                                                << intersection[1] << " "
+                                                << intersection[2] << endl;
+    }
+    if(intersection.empty()){return "";}
+    else
+    {
+      /* Add range noise to intersection by adding a random portion
+       * of the slope to it. Probability distribution is defined in
+       * the sensor model. Since the slope is a unit vector the
+       * magnitude is defined my the magnitude of the sensor noise model
+       * inside the sensor class.
+       *  noisy_intersection = intersection + random_variable*slope
+       */
[…]

### 2021-05-12 tdp/DataSetSim/dataset_sim.hpp

@@ -1,0 +1,56 @@
+#ifndef DATASET_SIMULATOR_H
+#define DATASET_SIMULATOR_H
+
+#include <cassert>
+#include <string>
+
+#include "worlds/world.hpp"
+#include "utils.hpp"
+#include "robot.hpp"
+
+class Dataset_Simulator
+{
+public:
+    Dataset_Simulator(World *world, Robot *robot, double sim_time):
+        world(world), robot(robot), sim_time(sim_time)
+    {
+        // check that sensor is inside world
+        assert(world->is_inside_world(robot->get_mover()->get_real_pose()));
+    };
+    ~Dataset_Simulator()
+    {
+        delete world;
+        delete robot;
+    }
+
+    /** Returns the string to be written into a scan file
+     * from the current pose using the azimuth and elevation
+     * of the ray from the given orientation
+     * In particular this function transforms the azitmuth and
+     * elevation into a ray described in the world coordinate
+     * system and forwards this to the ray intersection method
+     * of the world
+     */
+    string get_ray_result(double azimuth, double elevation);
+
+
+    /* Writes all measured points at the current pose of the robot
+     * into the given file stream */
+    void write_all_sensor_points(fstream &scan_file, double dt);
+
+    /* Writes the current pose of the robot
+     * into the given file stream */
+    void write_pose(ostream &pose_file, bool write_true_pose = false);
+
+    // Accessors
+    double get_sim_time(){return this->sim_time;}
+    World *get_world(){return this->world;}
+    Robot *get_robot(){return this->robot;}
+
+private:
+    World *world;
+    Robot *robot;
+    double sim_time;
+};
+
+#endif //DATASET_SIMULATOR_H

### 2021-05-12 tdp/DataSetSim/movers/descender.cpp

@@ -1,0 +1,39 @@
+#include "descender.hpp"
+
+void Descender::move(double dt)
+{
+    vector<double> curr_real_pose = this->get_real_pose();
+    vector<double> curr_assumed_pose = this->get_assumed_pose();
+    double angle_increment = this->get_rot_rate()*dt;
+    double descend_increment = this->get_desc_rate()*dt;
+
+    /* We descend only in y direction and rotate about the y axis
+     * noise occures by increasing or decreasing the expection
+     * rotation and descend by a random factor */
+
+    curr_real_pose[1] -= descend_increment*
+                           (1+this->get_hpose_noise());
+    curr_real_pose[4] = to_pi_range(curr_real_pose[4] +
+                angle_increment*(1+this->get_rpose_noise()));
+
+    curr_assumed_pose[1] -= descend_increment;
+    curr_assumed_pose[4] = to_pi_range(curr_assumed_pose[4] + angle_increment);
+
+    this->set_real_pose(curr_real_pose);
+    this->set_assumed_pose(curr_assumed_pose);
+}
+
+double Descender::get_rpose_noise()
+{
+    // This randomly sampled value is a percentage of the
+    // actually desired movement increment
+    return this->rpose_distribution(this->randomness_source);
+}
+
+double Descender::get_hpose_noise()
+{
+    // This randomly sampled value is a percentage of the
+    // actually desired movement increment
+    return this->hpose_distribution(this->randomness_source);
+}
+

### 2021-05-12 tdp/DataSetSim/movers/descender.hpp

@@ -1,0 +1,41 @@
+#ifndef DESCENDER_H
+#define DESCENDER_H 
+
+#include "mover.hpp"
+
+class Descender: public Mover
+{
+public:
+    Descender(vector<double> init_pose,
+          double rot_rate,
+          double desc_rate,
+          vector<double> pose_noise_mean,
+          vector<double> pose_noise_std): Mover(init_pose)
+    {
+        this->desc_rate = desc_rate;
+        this->rot_rate = rot_rate;
+        this->rpose_distribution =
+            normal_distribution<double>(pose_noise_mean[0], pose_noise_std[0]);
+        this->hpose_distribution =
+            normal_distribution<double>(pose_noise_mean[1], pose_noise_std[1]);
+    }
+    ~Descender(){};
+
+    double get_rot_rate(){return this->rot_rate;}
+    double get_desc_rate(){return this->desc_rate;}
+
+    virtual void move(double dt);
+private:
+    // Mersenne twister PRNG, initialized with seed from previous
+    // random device instance
+    mt19937 randomness_source;
+    normal_distribution<double> rpose_distribution;
+    normal_distribution<double> hpose_distribution;
+
+    double desc_rate;
+    double rot_rate;
+    virtual double get_rpose_noise();
+    virtual double get_hpose_noise();
+};
+
+#endif //DESCENDER_H

### 2021-05-12 tdp/DataSetSim/movers/mover.hpp

@@ -1,0 +1,40 @@
+#ifndef MOVER_H
+#define MOVER_H
+
+#include <vector>
+#include <random>
+
+#include "../utils.hpp"
+
+using namespace std;
+
+class Mover
+{
+public:
+    Mover(vector<double> init_pose):
+       real_pose(init_pose), assumed_pose(init_pose){};
+
+    virtual ~Mover(){};
+
+    /* Defines the movement of the associated robot.
+     * Changes the stored pose by the movement that happens over
+     * @dt seconds */
+    virtual void move(double dt) = 0;
+
+    /* Returns the true current pose of the sensor */
+    virtual vector<double> get_real_pose(){return this->real_pose;}
+
+    /* Sets the true current pose of the sensor the given pose*/
+    virtual void set_real_pose(vector<double> new_pose){this->real_pose = new_pose;}
+
+    /* Returns the assumed current pose of the sensor */
+    virtual vector<double> get_assumed_pose(){return this->assumed_pose;}
+
+    /* Sets the true current pose of the sensor the given pose*/
+    virtual void set_assumed_pose(vector<double> new_pose){this->assumed_pose = new_pose;}
+private:
+    vector<double> real_pose;
+    vector<double> assumed_pose;
+};
+
+#endif // MOVER_H

### 2021-05-12 tdp/DataSetSim/movers/roller.cpp

@@ -1,0 +1,73 @@
+#include "roller.hpp"
+
+/* We assume the noisy spherical robot mainly rotates about one axis.
+ * At each timestep there is also a disturbance torque about the 
+ * x and z axes (i.e. intended rotation axis and other ground rotation)
+ * that is determined my a normal distribution N(mu,sigma). 
+ * from this force we can compute the change in the current velocities
+ * and then compute the pose by integrating.
+ *
+ * We assume the constant desired rotation rate is an eights rotation in 1 sec, i.e.#
+ * about the x axis thus translating into the room towars the y-axis with
+ *      pi/4 per s
+ */
+void Roller::move(double dt)
+{
+    vector<double> curr_real_pose = this->get_real_pose();
+    vector<double> curr_assumed_pose = this->get_assumed_pose();
+    double angle_increment = this->rot_rate*dt; // rad
+
+    /* Update the current disturbance velocities by integrating
+     * the sampled disturbance torque 
+     * torque about the x-axis leads to angular velocity about x 
+     * torque about the z-axis leads to angular velocity about z */ 
+    this->curr_x_dist_ang_vel += this->get_xpose_noise()*dt; // in rad/s
+    this->curr_z_dist_ang_vel += this->get_zpose_noise()*dt; // in rad/s
+    
+    /* We move in z and x direction. In z direction purposefully and randomly,
+     * in x-direction only randomly
+     * Doing a eights rotation per second
+     */
+    vector<double> real_delta_euler_angle =
+            {angle_increment + this->curr_x_dist_ang_vel*dt,
+            0,
+            // This noise is one order of magnitude higher 
+            this->curr_z_dist_ang_vel*dt};
+
+    vector<double> assumed_delta_euler_angle =
+            {angle_increment, 0, 0};
+
+    // Transform Rotaion into robot frame, only needed if we want to rotate 
+    // about world axis axis.
+    //world2robot_frame(this->get_real_pose(), true, real_delta_euler_angle);
+    //world2robot_frame(this->get_real_pose(), true, assumed_delta_euler_angle);
+    
+    // z-axis rotation corresponds to x-axis translation
+    curr_real_pose[0] -= real_delta_euler_angle[2]*this->radius;
+    curr_assumed_pose[0] -= assumed_delta_euler_angle[2]*this->radius;
[…]

### 2021-05-12 tdp/DataSetSim/movers/roller.hpp

@@ -1,0 +1,50 @@
+#ifndef ROLLER_H
+#define ROLLER_H 
+
+#include "mover.hpp"
+
+class Roller: public Mover
+{
+public:
+    Roller(vector<double> init_pose,
+          double radius,
+          vector<double> pose_noise_mean,
+          vector<double> pose_noise_std): 
+        Mover(init_pose),
+        radius(radius),
+        rot_rate(M_PI_4), // rad/s
+        curr_x_dist_ang_vel(0),
+        curr_z_dist_ang_vel(0) 
+    {
+        if(VERBOSITY_LEVEL == 1)
+        {
+            cout << "Initializing random distributions:" << endl;
+            cout << "Pose noise about X: mean " << pose_noise_mean[0] 
+                 << " std: " << pose_noise_std[0] << "\n"
+                 << "Pose noise about Z: mean " << pose_noise_mean[1]
+                 << " std: " << pose_noise_std[1] << endl;  
+        }
+        this->xpose_distribution =
+            normal_distribution<double>(pose_noise_mean[0], pose_noise_std[0]);
+        this->zpose_distribution =
+            normal_distribution<double>(pose_noise_mean[1], pose_noise_std[1]);            
+    }
+    ~Roller(){};
+
+    virtual void move(double dt);
+private:
+    // Mersenne twister PRNG, initialized with seed from previous
+    // random device instance
+    mt19937 randomness_source;
+    normal_distribution<double> xpose_distribution;
+    normal_distribution<double> zpose_distribution;
+
+    double radius;
+    double rot_rate;
+    double curr_x_dist_ang_vel;
+    double curr_z_dist_ang_vel;
+    virtual double get_xpose_noise();
+    virtual double get_zpose_noise();
+};
+
+#endif //ROLLER_H

### 2021-05-12 tdp/DataSetSim/movers/stepwise_descender.cpp

@@ -1,0 +1,60 @@
+#include "stepwise_descender.hpp"
+
+void Stepwise_Descender::move(double dt)
+{
+    vector<double> curr_real_pose = this->get_real_pose();
+    vector<double> curr_assumed_pose = this->get_assumed_pose();
+
+    // We're still in the rotating state
+    if(this->get_state())
+    {
+        curr_real_pose[4] += this->get_rot_rate()*dt*
+                                (1 + this->get_rpose_noise());
+        curr_assumed_pose[4] += this->get_rot_rate()*dt;
+
+        if(curr_assumed_pose[4] >= 2*M_PI)
+        {
+            this->toggle_state();
+            curr_real_pose[4] = to_pi_range(curr_real_pose[4]);
+            curr_assumed_pose[4] = to_pi_range(curr_assumed_pose[4]);
+            this->last_height = curr_assumed_pose[1];
+        }
+    }
+    else // We're descending
+    {
+        curr_real_pose[1] -= this->get_desc_rate()*dt*
+                                (1+this->get_hpose_noise());
+        curr_assumed_pose[1] -= this->get_desc_rate()*dt;
+
+        if(abs(curr_assumed_pose[1] - this->last_height) >= this->get_desc_step())
+        {
+            this->toggle_state();
+        }
+    }
+
+    this->set_real_pose(curr_real_pose);
+    this->set_assumed_pose(curr_assumed_pose);
+}
+
+
+double Stepwise_Descender::get_rpose_noise()
+{
+    // This randomly sampled value is a percentage of the
+    // actually desired movement increment
+    return this->rpose_distribution(this->randomness_source);
+}
+
+
+double Stepwise_Descender::get_hpose_noise()
+{
+    // This randomly sampled value is a percentage of the
+    // actually desired movement increment
+    return this->hpose_distribution(this->randomness_source);
+}
+
+
+inline void Stepwise_Descender::toggle_state()
+{
+    bool temp = !this->get_state();
+    this->rotating = temp;
+}

### 2021-05-12 tdp/DataSetSim/movers/stepwise_descender.hpp

@@ -1,0 +1,52 @@
+#ifndef STEPWISE_DESCENDER_H
+#define STEPWISE_DESCENDER_H
+
+#include "mover.hpp"
+
+class Stepwise_Descender: public Mover
+{
+public:
+    Stepwise_Descender(vector<double> init_pose,
+          double rot_rate,
+          double desc_rate,
+          double desc_step,
+          vector<double> pose_noise_mean,
+          vector<double> pose_noise_std): Mover(init_pose)
+    {
+        this->desc_rate = desc_rate;
+        this->desc_step = desc_step;
+        this->rot_rate = rot_rate;
+        this->rpose_distribution =
+            normal_distribution<double>(pose_noise_mean[0], pose_noise_std[0]);
+        this->hpose_distribution =
+            normal_distribution<double>(pose_noise_mean[1], pose_noise_std[1]);
+    }
+    ~Stepwise_Descender(){};
+
+    double get_rot_rate(){return this->rot_rate;}
+    double get_desc_rate(){return this->desc_rate;}
+    double get_desc_step(){return this->desc_step;}
+    bool get_state(){return this->rotating;}
+
+    void toggle_state();
+    virtual void move(double dt);
+
+private:
+    // Mersenne twister PRNG, initialized with seed from previous
+    // random device instance
+    mt19937 randomness_source;
+    normal_distribution<double> rpose_distribution;
+    normal_distribution<double> hpose_distribution;
+
+    double desc_rate;
+    double rot_rate;
+    double desc_step;
+    double last_height;
+    // True == we want to rotate, false == we want to descend
+    bool rotating =  true;
+
+    virtual double get_rpose_noise();
+    virtual double get_hpose_noise();
+};
+
+#endif //STEPWISE_DESCENDER_H

### 2021-05-12 tdp/DataSetSim/parsing.cpp

@@ -1,0 +1,319 @@
+#include "parsing.hpp"
+
+namespace po = boost::program_options;
+
+enum world_type {RECTANGLE_WORLD = 0, CYLINDER_WORLD};
+enum sensor_type {LINE_SCANNER = 0, NOISY_LINE_SCANNER, LIVOXMID40, LIVOXMID100, LIVOXMID70C};
+enum mover_type {ROLLER = 0, DESCENDER, STEPWISE_DESCENDER};
+
+istream& operator>>(istream &in, sensor_type &sensor)
+{
+    string token;
+    in >> token;
+    if (token == "LINE_SCANNER")
+        sensor = LINE_SCANNER;
+    else if (token == "NOISY_LINE_SCANNER")
+        sensor = NOISY_LINE_SCANNER;
+    else if (token == "LIVOXMID40")
+        sensor = LIVOXMID40;
+    else if (token == "LIVOXMID100")
+        sensor = LIVOXMID100;
+    else if (token == "LIVOXMID70C")
+        sensor = LIVOXMID70C;
+    else
+        in.setstate(std::ios_base::failbit);
+    return in;
+}
+
+istream& operator>>(istream &in, mover_type &mover)
+{
+    string token;
+    in >> token;
+    if (token == "ROLLER")
+        mover = ROLLER;
+    else if(token == "DESCENDER")
+        mover = DESCENDER;
+    else if(token == "STEPWISE_DESCENDER")
+        mover = STEPWISE_DESCENDER;
+    else
+        in.setstate(std::ios_base::failbit);
+    return in;
+}
+
+istream& operator>>(istream &in, world_type &world)
+{
+    string token;
+    in >> token;
+    if (token == "RECTANGLE_WORLD")
+        world = RECTANGLE_WORLD;
+    else if(token == "CYLINDER_WORLD")
+        world = CYLINDER_WORLD;
+    else
+        in.setstate(std::ios_base::failbit);
+    return in;
+}
+
+/* Function used to check that 'opt1' and 'opt2' are not specified
+   at the same time. */
+void sensor_option_conflict(const po::variables_map & vm,
+                     const sensor_type type,
+                     const char *option)
+{
+  if (vm.count("sensor") && vm["sensor"].as<sensor_type>() == type)
+    if (vm.count(option))
+       throw std::logic_error(std::string("This sensor type is incompatible with ")
+                         + option);
+}
+
+/* Function used to check that if 'for_what' is specified, then
+   'required_option' is specified too. */
[…]

### 2021-05-12 tdp/DataSetSim/parsing.hpp

@@ -1,0 +1,20 @@
+#ifndef PARSING_H
+#define PARSING_H
+
+#include <iostream>
+#include <iomanip>
+#include <fstream>
+#include <sstream>
+#include <boost/program_options.hpp>
+
+#include "worlds/world.hpp"
+#include "worlds/rectangle_world.hpp"
+#include "worlds/cylinder_world.hpp"
+#include "robot.hpp"
+
+bool parse_options(int argc, char const *argv[],
+                  string &out_file_path, double &sim_time, double &dt,
+                  Robot *&robot, World *&world);
+
+
+#endif // PARSING_H

### 2021-05-12 tdp/DataSetSim/robot.hpp

@@ -1,0 +1,42 @@
+#ifndef ROBOT_H
+#define ROBOT_H
+
+#include "movers/mover.hpp"
+#include "movers/roller.hpp"
+#include "movers/descender.hpp"
+#include "movers/stepwise_descender.hpp"
+
+#include "sensors/sensor.hpp"
+#include "sensors/line_scanner.hpp"
+#include "sensors/noisy_line_scanner.hpp"
+#include "sensors/livoxMid40.hpp"
+#include "sensors/livoxMid100.hpp"
+#include "sensors/livoxMid70C.hpp"
+
+class Robot
+{
+public:
+    Robot(Mover *mover, Sensor *sensor)
+    {
+        this->mover = mover;
+        this->sensor = sensor;
+    }
+    ~Robot()
+    {
+        delete this->sensor;
+        delete this->mover;
+    }
+
+    Sensor *get_sensor() {return this->sensor;}
+    Mover *get_mover() {return this->mover;}
+
+    void set_sensor(Sensor *sensor){this->sensor = sensor;}
+    void set_mover(Mover *mover){this->mover = mover;}
+
+private:
+    Sensor *sensor;
+    Mover *mover;
+};
+
+
+#endif //ROBOT_H

### 2021-05-12 tdp/DataSetSim/sensors/line_scanner.cpp

@@ -1,0 +1,33 @@
+#include <iostream>
+#include <vector>
+
+#include "line_scanner.hpp"
+
+using namespace std;
+
+/* This is an ideal sensor. Hence no noise */
+double Line_Scanner::get_range_noise()
+{
+  return 0;
+}
+
+/* Assuming an ideal line scanner that goes over the entire
+ * FOV with the resolution that's specified */
+void Line_Scanner::get_sensing_pattern(vector<pair<double, double> > &pattern,
+                                          double dt)
+{
+  double elev_max = get<0>(this->get_fov())*0.5,
+         elev_res = get<0>(this->get_resolution());
+
+  double azi_max = get<1>(this->get_fov())*0.5,
+         azi_res = get<1>(this->get_resolution());
+
+  for(double elev = - get<0>(this->get_fov())*0.5; elev <= elev_max; elev += elev_res)
+  {
+      for(double azi = - get<1>(this->get_fov())*0.5 ; azi <= azi_max; azi += azi_res)
+      {
+          pattern.push_back(make_pair(azi, elev));
+      }
+  }
+}
+

### 2021-05-12 tdp/DataSetSim/sensors/line_scanner.hpp

@@ -1,0 +1,23 @@
+#ifndef LINE_SCANNER_H
+#define LINE_SCANNER_H 
+
+#include "sensor.hpp"
+
+/* An ideal line scanner with arbitrary FOV.*/
+class Line_Scanner: public Sensor
+{
+public:
+
+    Line_Scanner(pair<double, double> range,
+        pair<double, double> fov,
+        pair<double, double> resolution):
+        Sensor(range,fov,resolution)
+    {};
+    ~Line_Scanner(){};
+
+    virtual double get_range_noise();
+    virtual void get_sensing_pattern(vector<pair<double, double> > &pattern,
+                                     double dt);
+};
+
+#endif //LINE_SCANNER_H

### 2021-05-12 tdp/DataSetSim/sensors/livoxMid100.cpp

@@ -1,0 +1,14 @@
+#include "livoxMid100.hpp"
+
+void LivoxMid100::get_sensing_pattern(vector<pair<double, double> > &pattern,
+                                      double dt)
+{
+    vector<pair<double,double> > scan1_pattern, scan2_pattern, scan3_pattern;
+    this->scan1.get_sensing_pattern(scan1_pattern, dt);
+    this->scan2.get_sensing_pattern(scan2_pattern, dt);
+    this->scan3.get_sensing_pattern(scan3_pattern, dt);
+
+    pattern.insert(pattern.end(), scan3_pattern.begin(), scan3_pattern.end());
+    pattern.insert(pattern.end(), scan2_pattern.begin(), scan2_pattern.end());
+    pattern.insert(pattern.end(), scan1_pattern.begin(), scan1_pattern.end());
+}

### 2021-05-12 tdp/DataSetSim/sensors/livoxMid100.hpp

@@ -1,0 +1,50 @@
+#ifndef LIVOX_MID_100_H
+#define LIVOX_MID_100_H 
+
+#include "livoxMid40.hpp"
+
+/* A noisy Livox-Mid-100 scanner. Consits of 3 Livox-Mid-40 scanners
+ * that are offset by azimuth to each other (but overlap slightly) to
+ * provide a wider azimuth FOV.  */
+class LivoxMid100: public Noisy_Line_Scanner
+{
+private:
+  const double AZIMUTH_OVERLAP; // [deg]
+  LivoxMid40 scan1, scan2, scan3;
+
+public:
+  LivoxMid100(pair<double, double> range,
+            pair<double, double> fov,
+            pair<double, double> resolution,
+            double range_noise_mean,
+            double range_noise_std):
+       Noisy_Line_Scanner(range, fov, resolution,
+                          range_noise_mean, range_noise_std),
+       /* We assume a constant overlap angle and
+        * calculate the offset and FOV from that
+        */
+       AZIMUTH_OVERLAP(8.4),
+       scan1(range,
+             // Calculate the mid40 fov
+             make_pair(get<0>(fov), (get<1>(fov)+2*AZIMUTH_OVERLAP)*0.333333),
+             resolution,
+             range_noise_mean, range_noise_std,
+             // Calculate the azimuth offset
+             - deg_to_rad((get<1>(fov) + 2*AZIMUTH_OVERLAP)*0.333333 - AZIMUTH_OVERLAP)),
+       scan2(range,
+             make_pair(get<0>(fov), (get<1>(fov)+2*AZIMUTH_OVERLAP)*0.333333),
+             resolution,
+             range_noise_mean, range_noise_std,
+             0),
+       scan3(range,
+             make_pair(get<0>(fov), (get<1>(fov)+2*AZIMUTH_OVERLAP)*0.333333),
+             resolution,
+             range_noise_mean, range_noise_std,
+             deg_to_rad((get<1>(fov) + 2*AZIMUTH_OVERLAP)*0.333333 -AZIMUTH_OVERLAP))
+  {}
+
+  virtual void get_sensing_pattern(vector<pair<double, double> > &pattern,
+                                   double dt);
+};
+
+#endif //LIVOX_MID_100_H

### 2021-05-12 tdp/DataSetSim/sensors/livoxMid40.cpp

@@ -1,0 +1,52 @@
+#include <iostream>
+#include "../constants.hpp"
+#include "livoxMid40.hpp"
+
+void LivoxMid40::get_sensing_pattern(vector<pair<double, double> > &pattern,
+                                     double dt)
+{
+    if(dt > MIN_SIM_RESOLUTION && VERBOSITY_LEVEL == 1)
+    {
+        cout << "WARNING: for the Livox Mid 40 scanner "
+             << "you should choose a simulation resolution "
+             << "that is smaller than " << MIN_SIM_RESOLUTION << endl;
+    }
+
+    /* Get half the fov, as only half is required to find
+     * the factor to multply the sinusoidal function with to achieve
+     * the extend of the measurement
+     */
+    double elev = deg_to_rad(get<0>(this->get_fov()))*0.5;
+    double azi = deg_to_rad(get<1>(this->get_fov()))*0.5;
+
+    /* During each simulation time step dt, we need
+     * to subdivide to find the livox pattern traced
+     * during this timestep.
+     * In 1s the livox collects 100,000 pts, therefor in
+     * dt it collects pts = dt * 100,000 / 1s.
+     * Therefore the sim_step should be 1/100,000
+     */
+    double sim_step = 0.00001;
+    /* The rotational speeds omega1 and omega2 as well as their
+     * rates of change are chosen such that we have 5 "Leaves" and
+     * within 0.5 s we have ~50% coverage, within 1 second we have ~95%
+     * coverage as specified by the livox scanner specs
+     * The r values are the maximal distance, and are computed from the FOV that
+     * is given.
+     */
+    double s = 100.0, n = sqrt(20),
+           fov_elev_factor = tan(elev), fov_azi_factor = tan(azi);
+
+    /* Adding all the elev azimuth pairs to the pattern */
+    for(double t = this->internal_time; t < this->internal_time + dt; t += sim_step)
+    {
+        double elev = fov_elev_factor*atan2(sin(s*(n*t + 0.25*M_PI))*cos(s*t), 1);
+        double azi  = fov_azi_factor*atan2(sin(s*(n*t + 0.25*M_PI))*sin(s*t), 1)
+                      +  this->azimuth_offset;
+        pattern.push_back(make_pair(rad_to_deg(elev),
+                                    rad_to_deg(azi)));
+    }
+    /* Increase the internal time such that we don't start at the same
+     * spot each timestep */
+    this->internal_time += dt;
+}

### 2021-05-12 tdp/DataSetSim/sensors/livoxMid40.hpp

@@ -1,0 +1,45 @@
+#ifndef LIVOX_MID_40_H
+#define LIVOX_MID_40_H
+
+#include "noisy_line_scanner.hpp"
+
+/* A Noisy Livox-Mid-40 scanner with the "flower"
+ * scanning pattern. */
+class LivoxMid40: public Noisy_Line_Scanner
+{
+private:
+  double azimuth_offset;
+  double internal_time = 0;
+public:
+  LivoxMid40(pair<double, double> range,
+            pair<double, double> fov,
+            pair<double, double> resolution,
+            double range_noise_mean,
+            double range_noise_std,
+            double azimuth_offset = 0):
+       Noisy_Line_Scanner(range, fov, resolution,
+                         range_noise_mean, range_noise_std)
+  {
+    if(abs(get<0>(fov) - get<1>(fov)) > 0.1)
+    {
+      cerr << "ERROR: Livox Mid 40 scanner can only provide a "
+           << "circular field of view. I.e. FOV_ELEVATION == FOV_AZIMUTH. "
+           << "Given was Elev = " << get<0>(fov) << " Azi = " << get<1>(fov)
+           <<  endl;
+      exit(1);
+    }
+    if(get<0>(fov) >= 180.0 || get<1>(fov) >= 180.0)
+    {
+      cerr << "ERROR: Cannot choose fov greater or equal to 180.0 [deg] for Livox "
+           << "Mid 40 scanner." << endl;
+      exit(1);
+    }
+
+    this->azimuth_offset = azimuth_offset;
+  }
+
+  virtual void get_sensing_pattern(vector<pair<double, double> > &pattern,
+                                   double dt);
+};
+
+#endif //LIVOX_MID_40_H

### 2021-05-12 tdp/DataSetSim/sensors/livoxMid70C.cpp

@@ -1,0 +1,12 @@
+#include "livoxMid70C.hpp"
+
+void LivoxMid70C::get_sensing_pattern(vector<pair<double, double> > &pattern,
+                                      double dt)
+{
+    vector<pair<double,double> > scan1_pattern, scan2_pattern;
+    this->scan1.get_sensing_pattern(scan1_pattern, dt);
+    this->scan2.get_sensing_pattern(scan2_pattern, dt);
+
+    pattern.insert(pattern.end(), scan2_pattern.begin(), scan2_pattern.end());
+    pattern.insert(pattern.end(), scan1_pattern.begin(), scan1_pattern.end());
+}

### 2021-05-12 tdp/DataSetSim/sensors/livoxMid70C.hpp

@@ -1,0 +1,45 @@
+#ifndef LIVOX_MID_70C_H
+#define LIVOX_MID_70C_H
+
+#include "livoxMid40.hpp"
+
+/* A noisy Livox-Mid-70C scanner. Consits of 2 Livox-Mid-70 scanners
+ * that are offset by azimuth to each other (but overlap slightly) to
+ * provide a wider azimuth FOV.  */
+class LivoxMid70C: public Noisy_Line_Scanner
+{
+private:
+  const double AZIMUTH_OVERLAP; // [deg]
+  LivoxMid40 scan1, scan2;
+
+public:
+  LivoxMid70C(pair<double, double> range,
+            pair<double, double> fov,
+            pair<double, double> resolution,
+            double range_noise_mean,
+            double range_noise_std):
+       Noisy_Line_Scanner(range, fov, resolution,
+                          range_noise_mean, range_noise_std),
+       /* We assume a constant overlap angle and
+        * calculate the offset and FOV from that
+        */
+       AZIMUTH_OVERLAP(40.4),
+       scan1(range,
+             // Calculate the mid40 fov
+             make_pair(get<0>(fov), (get<1>(fov)+AZIMUTH_OVERLAP)*0.5),
+             resolution,
+             range_noise_mean, range_noise_std,
+             // Calculate the azimuth offset
+             - deg_to_rad(AZIMUTH_OVERLAP)*0.5),
+       scan2(range,
+             make_pair(get<0>(fov), (get<1>(fov)+AZIMUTH_OVERLAP)*0.5),
+             resolution,
+             range_noise_mean, range_noise_std,
+             deg_to_rad(AZIMUTH_OVERLAP)*0.5)
+  {}
+
+  virtual void get_sensing_pattern(vector<pair<double, double> > &pattern,
+                                   double dt);
+};
+
+#endif //LIVOX_MID_70C_H

### 2021-05-12 tdp/DataSetSim/sensors/noisy_line_scanner.cpp

@@ -1,0 +1,9 @@
+#include "noisy_line_scanner.hpp"
+
+double Noisy_Line_Scanner::get_range_noise()
+{
+    // This randomly sampled value is a percentage of the measured range
+    return this->range_distribution(this->randomness_source);
+}
+
+

### 2021-05-12 tdp/DataSetSim/sensors/noisy_line_scanner.hpp

@@ -1,0 +1,33 @@
+#ifndef NOISY_LINE_SCANNER_H
+#define NOISY_LINE_SCANNER_H
+
+#include "line_scanner.hpp"
+
+/* A noisy line scanner with an arbitrary FOV */
+class Noisy_Line_Scanner: public Line_Scanner
+{
+
+public:
+    Noisy_Line_Scanner(pair<double, double> range,
+                      pair<double, double> fov,
+                      pair<double, double> resolution,
+                      double range_noise_mean,
+                      double range_noise_std):
+        Line_Scanner(range, fov, resolution)
+    {
+        this->range_distribution = normal_distribution<double>(range_noise_mean, range_noise_std);
+    }
+
+    ~Noisy_Line_Scanner(){};
+
+    /* Returns a random value sampled from a random distribution
+     * that will be added on all range measurements */
+    virtual double get_range_noise();
+private:
+    // Mersenne twister PRNG, initialized with seed from previous
+    // random device instance
+    mt19937 randomness_source;
+    normal_distribution<double> range_distribution;
+};
+
+#endif //NOISY_LINE_SCANNER_H

### 2021-05-12 tdp/DataSetSim/sensors/sensor.hpp

@@ -1,0 +1,64 @@
+#ifndef SENSOR_H
+#define SENSOR_H
+
+#include <vector>
+#include <random>
+#include <iostream>
+#include "../utils.hpp"
+using namespace std;
+
+class Sensor
+{
+public:
+
+    Sensor(pair<double,double> range,
+        pair<double, double> fov,
+        pair<double, double> resolution):
+        range(range), fov(fov), resolution(resolution){};
+
+    virtual ~Sensor(){};
+
+    /* Defines the noise model of the range measurement
+     * @returns a value in the range [-n, n] that represent
+     *  the deviation from the true measured value
+     */
+    virtual double get_range_noise() = 0;
+
+    /* Returns a vector that contains the scanning pattern coordinates
+     * as azimuth and elevation. Given the dt maybe only a partial scanning pattern
+     * can be completed. */
+    virtual void get_sensing_pattern(vector<pair<double, double> > &pattern,
+                                     double dt) = 0;
+
+    /* Returns the maximal range of the sensor */
+    virtual pair<double,double> get_range()
+    {
+        return this->range;
+    }
+    /* Returns the fov of the sensor in degree
+     * fov[0]: angle in sensor y-direction
+     * fov[1]: angle in sensor z-directin
+     */
+    virtual pair<double, double> get_fov()
+    {
+        return this->fov;
+    }
+
+    /* Returns the angular resolution between rays.
+     * can be different for horizontal and vertical
+     * rays
+     * resolution[0]: resolution in sensor y-direction
+     * resolution[1]: resolution in sensor x-direction
+     */
+    virtual pair<double, double> get_resolution()
+    {
+        return this->resolution;
+    }
+
+private:
+    pair<double, double> range;
+    pair<double, double> fov;
+    pair<double, double> resolution;
+};
+
+#endif //SENSOR_H

### 2021-05-12 tdp/DataSetSim/utils.cpp

@@ -1,0 +1,151 @@
+#include "utils.hpp"
+
+/* Coordinate transformation functions. Uses the current pose
+ * of the robot to compute the coordinates of the given point
+ * in the respective other coordinate frame */
+void robot2world_frame(vector<double> pose, bool rotate_only, vector<double> &robot_coordinates)
+{
+  // Store the current coordinate values
+  double x = robot_coordinates[0], y = robot_coordinates[1],
+         z = robot_coordinates[2];
+  // Find translation required by the transformation
+  double tx, ty, tz;
+  if(rotate_only)
+  {
+    tx = ty = tz = 0.0;
+  }
+  else
+  {
+    tx = pose[0]; ty = pose[1]; tz = pose[2];
+  }
+
+  double roll = pose[3], pitch = pose[4], yaw = pose[5];
+  /* Compute rotation matrix from euler angles. Where
+   *     |r11 r12 r13|
+   * R = |r21 r22 r23|
+   *     |r31 r32 r33|
+   */
+  // Using roll-pitch-yaw
+  double r11 = cos(pitch)*cos(yaw),
+         r12 = -cos(pitch)*sin(yaw),
+         r13 = sin(pitch),
+         r21 = cos(roll)*sin(yaw)+cos(yaw)*sin(roll)*sin(pitch),
+         r22 = cos(roll)*cos(yaw)-sin(roll)*sin(pitch)*sin(yaw),
+         r23 = -cos(pitch)*sin(roll),
+         r31 = sin(roll)*sin(yaw) - cos(roll)*cos(yaw)*sin(pitch),
+         r32 = cos(yaw)*sin(roll) + cos(roll)*sin(pitch)*sin(yaw),
+         r33 = cos(roll)*cos(pitch);
+
+  /* Execute homogeneous transformation:
+   *        x_new = | R T |  |x|
+   *         |1|    | 0 1 |  |1|
+   */
+  robot_coordinates[0] = r11*x + r12*y + r13*z + tx;
+  robot_coordinates[1] = r21*x + r22*y + r23*z + ty;
+  robot_coordinates[2] = r31*x + r32*y + r33*z + tz;
+}
+
+void world2robot_frame(vector<double> pose, bool rotate_only, vector<double> &world_coordinates)
+{
+ // Store the current coordinate values
+  double x = world_coordinates[0],
+         y = world_coordinates[1],
[…]

### 2021-05-12 tdp/DataSetSim/utils.hpp

@@ -1,0 +1,30 @@
+#ifndef UTILS_H
+#define UTILS_H
+
+#include <vector>
+#include <cmath>
+#include <iostream>
+
+#include "constants.hpp"
+
+using namespace std;
+/* Coordinate transformation functions. Uses the current pose
+ * of the robot to compute the coordinates of the given point
+ * in the respective other coordinate frame */
+void robot2world_frame(vector<double> pose, bool rotate_only, vector<double> &robot_coordinates);
+void world2robot_frame(vector<double> pose, bool rotate_only, vector<double> &world_coordinates);
+/* Helper function for computing the distance between two
+ * points in space using the euclidean distance */
+double eucl_dist(vector<double> first, vector<double> second);
+void normalize(vector<double> vec);
+
+/* Helper for Radians and degree conversion */
+double deg_to_rad(double deg);
+double rad_to_deg(double rad);
+double to_pi_range(double rad);
+
+/* Coordinate Transformations */
+void cartesian_to_cylindrical(vector<double> &cart_coord);
+void cylindrical_to_cartesian(vector<double> &cyli_coord);
+
+#endif // UTILS_H

### 2021-05-12 tdp/DataSetSim/worlds/cylinder_world.cpp

@@ -1,0 +1,93 @@
+#include <cassert>
+#include "cylinder_world.hpp"
+
+using namespace std;
+
+void Cylinder_World::get_ray_intersection(vector<double> root,
+                                          vector<double> slope,
+                                          vector<double> &intersection)
+{
+    /* Now we check with intersections with the top, bottom end the
+     * cylindrical shell. If one intersection is found we can stop
+     * since no further intersection can exist.
+     *
+     * Basically we parametrize a ray via root+lambda*slope = radius
+     * and find the parameter lambda that fullfilles the equation.
+     * Using this we can find the other coordinates and then check
+     * whether the resulting intersection is within our world.
+     */
+
+    /* Find intersection with cylindrical shell */
+    if(slope[0] != 0 || slope[2] != 0)
+    {
+        /* This is the magic factor that we have to multiply the slope by to
+         * find the intersection with the cylindrical shell.
+         * Basically we find the parameter lambda such that sqrt(x^2+z^2) = r
+         * How to find this? Do not ask me, ask your lord and savior
+         * wolfram alpha!
+         */
+        double radius = this->get_radius();
+        double lambda = 1/(slope[0]*slope[0] + slope[2]*slope[2]) *
+            (sqrt(slope[2]*slope[2]*(radius*radius - root[0]*root[0]) +
+                  2*root[0]*slope[0]*root[2]*slope[2] +
+                  slope[0]*slope[0]*(radius*radius - root[2]*root[2]))
+            - root[0]*slope[0] - root[2]*slope[2]);
+
+        intersection[0] = root[0] + lambda*slope[0];
+        intersection[1] = root[1] + lambda*slope[1];
+        intersection[2] = root[2] + lambda*slope[2];
+
+        // Check that we're actually at a cylindrical wall
+        assert(abs(sqrt(intersection[0]*intersection[0]
+            + intersection[2]*intersection[2]) - radius) < 0.1);
+
+        // If we're in bounds of the cylinder ...
+        if(intersection[1] <= this->get_height() &&
+           intersection[1] >= 0 &&
+           // ...and we're looking in sensor direction
+           lambda >= 0)
+        {
+            // we found an intersection, we can return
+            return;
+        }
+    }
+
+    /* Check top and bottom plane */
+    if(slope[1] != 0)
+    {
+        // "bottom" plane
+        intersection[0] = root[0] - root[1]/slope[1] * slope[0];
[…]

### 2021-05-12 tdp/DataSetSim/worlds/cylinder_world.hpp

@@ -1,0 +1,28 @@
+#ifndef CYLINDER_WORLD_H
+#define CYLINDER_WORLD_H 
+
+#include "world.hpp"
+
+class Cylinder_World: public World
+{
+public:
+    Cylinder_World(double radius, double height):
+        radius(radius), height(height) {};
+    ~Cylinder_World(){};
+
+    virtual void get_ray_intersection(vector<double> root,
+                                      vector<double> slope,
+                                      vector<double> &intersection);
+
+    virtual bool is_inside_world(vector<double> pose);
+
+    // Accessors
+    double get_radius(){return this->radius;}
+    double get_height(){return this->height;}
+
+private:
+    double radius;
+    double height;
+};
+
+#endif //CYLINDER_WORLD_H

### 2021-05-12 tdp/DataSetSim/worlds/rectangle_world.cpp

@@ -1,0 +1,112 @@
+#include "rectangle_world.hpp"
+
+void Rectangle_World::get_ray_intersection(vector<double> root,
+                                           vector<double> slope,
+                                           vector<double> &intersection)
+{
+    intersection[0] = intersection[1] = intersection[2] = 0;
+
+    /* Compute the intersection with any of the the planes of the rectangle
+     * given the ray. If one intersection is found we can stop as no further
+     * intersection can exist */
+
+    /* If we're withing rectangle
+     * bounds if x axis is zero, then there is an intersection.
+     * If the slope is parallel to the plane (i.e. slope in the given axis = 0),
+     * there is no need to check for an intersection
+     * Check "left" and "right" plane:
+     */
+    if(slope[0] != 0)
+    {
+        // "Left" plane
+        intersection[0] = -this->get_width()*0.5;
+        intersection[1] = root[1] + (-this->get_width()*0.5 - root[0])/slope[0] * slope[1];
+        intersection[2] = root[2] + (-this->get_width()*0.5 - root[0])/slope[0] * slope[2];
+        if(abs(intersection[1]) <= this->get_height()*0.5 &&
+           abs(intersection[2]) <= this->get_depth()*0.5 &&
+           // Check that the intersection is in viewing direction of the robot
+           0 <= (-this->get_width()*0.5 - root[0])/slope[0])
+        {
+            return;
+        }
+
+        // "Right" plane
+        intersection[0] = this->get_width()*0.5;
+        intersection[1] = root[1] + (this->get_width()*0.5 - root[0])/slope[0] * slope[1];
+        intersection[2] = root[2] + (this->get_width()*0.5 - root[0])/slope[0] * slope[2];
+        if(abs(intersection[1]) <= this->get_height()*0.5 &&
+           abs(intersection[2]) <= this->get_depth()*0.5 &&
+           // Check that the intersection is in viewing direction of the robot
+           0 <= (this->get_width()*0.5 - root[0])/slope[0])
+        {
+            return;
+        }
+    }
+
+    /* Check top and bottom plane */
+    if(slope[1] != 0)
+    {
+        // "bottom" plane
+        intersection[0] = root[0] + (-this->get_height()*0.5 - root[1])/slope[1] * slope[0];
+        intersection[1] = -this->get_height()*0.5;
[…]

### 2021-05-12 tdp/DataSetSim/worlds/rectangle_world.hpp

@@ -1,0 +1,28 @@
+#ifndef RECTANGLE_WORLD_H
+#define RECTANGLE_WORLD_H 
+
+#include "world.hpp"
+
+class Rectangle_World: public World
+{
+public:
+    Rectangle_World(double height, double width, double depth):
+        height(height), width(width), depth(depth){};
+    ~Rectangle_World(){};
+
+    virtual void get_ray_intersection(vector<double> root,
+                                      vector<double> slope,
+                                      vector<double> &intersection);
+
+    virtual bool is_inside_world(vector<double> pose);
+
+    // Accessors
+    double get_height(){return this->height;}
+    double get_width(){return this->width;}
+    double get_depth(){return this->depth;}
+
+private:
+    double height, width, depth;
+};
+
+#endif //RECTANGLE_WORLD_H

### 2021-05-12 tdp/DataSetSim/worlds/world.hpp

@@ -1,0 +1,28 @@
+#ifndef WORLD_H
+#define WORLD_H
+
+#include <iostream>
+#include <vector>
+
+#include "../utils.hpp"
+
+using namespace std;
+
+class World
+{
+public:
+    World(){};
+    virtual ~World(){};
+
+    /**
+     * Returns the coordinates of intersection of the given ray,
+     * defined by the root vector and the slope vector, with the world
+     */
+    virtual void get_ray_intersection(vector<double> root,
+                                      vector<double> slope,
+                                      vector<double> &intersection) = 0;
+
+    virtual bool is_inside_world(vector<double> pose) = 0;
+};
+
+#endif //WORLD_H

### 2021-05-12 tdp/codeFromPi/src/imuJasper.cpp

@@ -1,0 +1,823 @@
+/*----------------------------------
+    Code by Jasper Zevering
+    jasper.zevering@stud-mail.uni-wuerzburg.de
+    31.01.2021
+---------------------------------*/
+
+#include "imuJasper.h"
+
+
+//DEFINING EXTERN VARIABLES OF HEADER
+//quiet mode
+bool quiet = false;
+//gain of madgwickfilter
+float gain_ = DEFAULT_MADGWICK_GAIN;
+//gain of complemnetary
+float alpha = 0.02;
+//Mechnaism for checking if everything works
+int firstRead = 0;
+//slow-mode ( (g_new = (g_old +g_0+ g_1 + g_2)/4))
+bool slow = false;
+// shoulr roll pitch yw in degrees be printed every loop
+bool printRPY = false;
+//
+bool debugMode = false;
+//
+int data_rate = DATA_RATE_DEFAULT;
+//
+float data_intervall = 1 / DATA_RATE_DEFAULT;
+//
+int imu_data_rate = IMU_DATA_RATE_DEFAULT;
+//gravity ing
+float gravityValue = 1;
+//automaticgain
+float autogain = 0;
+//float radius in m
+float r = 0.29;
+// Signal-safe flag for whether shutdown is requested
+sig_atomic_t volatile g_request_shutdown = 0;
+
+//Handler for data from spatial0
+void CCONV onSpatial0_SpatialData(PhidgetSpatialHandle
+                            ch,
+                            void *ctx,
+                            const double acceleration[3],
+                            const double angularRate[3],
+                            const double magneticField[3],
+                            double timestamp
+) {
+    if (!quiet) { ROS_INFO("Getting Data"); }
+    int serialNr;
+    int spatialNr;
+
+    float *gxTemp;
+    float *gyTemp;
+    float *gzTemp;
+    float *axTemp;
+    float *ayTemp;
+    float *azTemp;
+
+    Phidget_getDeviceSerialNumber((PhidgetHandle)
+                                          ch, &serialNr);
+    switch (serialNr) {
+        case SERIAL_0:
+            spatialNr = 0;
+            gxTemp = &gx0;
[…]

### 2021-05-12 tdp/codeFromPi/src/imuJasper.h

@@ -1,0 +1,134 @@
+#ifndef IMU_MERGER_H
+#define IMU_MERGER_H
+
+
+
+#include <stdio.h>
+#include <stdlib.h>
+#include <phidget22.h>
+
+#include <unistd.h>
+
+
+#include <ros/ros.h>
+#include <geometry_msgs/PoseStamped.h>
+
+#include <math.h>
+#include <tf/transform_broadcaster.h>
+#include <tf/transform_datatypes.h>
+#include <signal.h>
+
+#define SERIAL_0 415233
+#define SERIAL_1 297652
+#define SERIAL_2 425377
+
+#define DATA_RATE_DEFAULT 250
+#define IMU_DATA_RATE_DEFAULT 250
+
+#define DEFAULT_MADGWICK_GAIN 0.1
+
+
+//often used calculations
+#define M_PI_BY_180 0.017453292
+#define precalc_180_BY_M_PI 57.29577951
+#define ONE_THIRD 0.333333333
+
+
+typedef struct quaternion
+{
+    double  w;
+    double  x;
+    double  y;
+    double  z;
+};
+
+
+extern bool quiet;
+//actual data rate of publishing. will be set to DATA_RATE_DEFAULT if not else stated
+extern int data_rate;
+
+geometry_msgs::PoseStamped output_msg;
+
+
+//orientation
+float q0; 
+float q1; 
+float q2; 
+float q3; 
+float px;
+float py;
+float pz;
+//gain of madgwickfilter
+extern float gain_; 
+//gain of complemnetary
+extern float alpha;
+//gyroscope
+float gx; 
+float gy; 
+float gz; 
+float gx1;
+float gy1;
+float gz1;
+float gx2;
+float gy2;
+float gz2;
+float gx0;
+float gy0;
+float gz0;
+//accelerometer
+float ax; 
+float ay; 
+float az;
+float ax1;
+float ay1;
+float az1;
+float ax2;
+float ay2;
+float az2;
+float ax0;
+float ay0;
+float az0; 
+
+float vel_x;
+float vel_y;
+float vel_z;
+
+
+float dt; 
+
+extern int firstRead;
+
+extern bool slow;
+
+ros::Time lastTime;
+
+//Functions
+//fast algorithm for inverse square root
+float invSqrt(float f);
+//fast algorithm for sqrtr
+float fast_sqrt(const float& f);
+//initalizing procedure
+int CCONV init();
+//handlign arguments giving to programm
[…]

### 2021-05-12 tdp/datasets/remove_empty_scans.sh

@@ -1,0 +1,48 @@
+#!/bin/bash
+
+# This file iterates over all .pose and .3d files in a folder
+# If a .3d file is empty (i.e. contains no scans) it and the
+# corresponding .pose file is removed. 
+# Further all files are renamed such that the numbering is still
+# consecutive 
+
+if [ $# -eq 0 ]
+  then
+    echo "No arguments supplied. Usage:"
+    printf "\n\t remove_empty_scans.sh <path> <start_idx> <end_idx> <pts_threshold>\n"
+  	exit 1;
+fi
+
+PATH=$1
+START=$2
+END=$3
+
+PTS_THRESHOLD=$4
+
+CURR_WRITE_INDEX=0
+
+echo "Starting with  $START";
+echo "Ending with with  $END";
+
+/usr/bin/mkdir "${PATH}reduced";
+/usr/bin/rm "${PATH}reduced/*";
+
+for ((i=START;i<=END;i++))
+do
+	padded_read_count=`printf "%03d" $i`;
+	echo "Reading scan${padded_read_count} ...";
+	curr_read_file="${PATH}scan${padded_read_count}";
+	num_lines=$( /usr/bin/wc -l < "${curr_read_file}.3d" );
+	# Delete all scans with less then PTS_THRESHOLD  points
+	if [ "$num_lines" -lt "$PTS_THRESHOLD" ];
+	then
+		echo "... not copying ";
+	else
+		padded_write_count=`printf "%03d" $CURR_WRITE_INDEX`;
+		curr_write_file="${PATH}reduced/scan${padded_write_count}";
+		echo "... copying to ${curr_write_file} "
+		/usr/bin/cp "${curr_read_file}.3d" "${curr_write_file}.3d";
+		/usr/bin/cp "${curr_read_file}.pose" "${curr_write_file}.pose";
+		((CURR_WRITE_INDEX++))
+	fi
+done

### 2021-05-12 tdp/PCLimpr/include/debugutil.h
preg6d added

@@ -1,0 +1,124 @@
+/*
+ * This code is from a stackoverflow thread.
+ * https://stackoverflow.com/questions/3018313/algorithm-to-convert-rgb-to-hsv-and-hsv-to-rgb-in-range-0-255-for-both
+ * Im sure many of you have also used this when working with rgb and hsv color spaces.
+ */
+
+
+typedef struct {
+    double r;       // a fraction between 0 and 1
+    double g;       // a fraction between 0 and 1
+    double b;       // a fraction between 0 and 1
+} rgb;
+
+typedef struct {
+    double h;       // angle in degrees
+    double s;       // a fraction between 0 and 1
+    double v;       // a fraction between 0 and 1
+} hsv;
+
+static hsv   rgb2hsv(rgb in);
+static rgb   hsv2rgb(hsv in);
+
+hsv rgb2hsv(rgb in)
+{
+    hsv         out;
+    double      min, max, delta;
+
+    min = in.r < in.g ? in.r : in.g;
+    min = min  < in.b ? min  : in.b;
+
+    max = in.r > in.g ? in.r : in.g;
+    max = max  > in.b ? max  : in.b;
+
+    out.v = max;                                // v
+    delta = max - min;
+    if (delta < 0.00001)
+    {
+        out.s = 0;
+        out.h = 0; // undefined, maybe nan?
+        return out;
+    }
+    if( max > 0.0 ) { // NOTE: if Max is == 0, this divide would cause a crash
+        out.s = (delta / max);                  // s
+    } else {
+        // if max is 0, then r = g = b = 0              
+        // s = 0, h is undefined
+        out.s = 0.0;
+        out.h = -1;                            // its now undefined
+        return out;
+    }
+    if( in.r >= max )                           // > is bogus, just keeps compilor happy
+        out.h = ( in.g - in.b ) / delta;        // between yellow & magenta
+    else
+    if( in.g >= max )
+        out.h = 2.0 + ( in.b - in.r ) / delta;  // between cyan & yellow
+    else
+        out.h = 4.0 + ( in.r - in.g ) / delta;  // between magenta & cyan
+
+    out.h *= 60.0;                              // degrees
+
+    if( out.h < 0.0 )
+        out.h += 360.0;
+
+    return out;
+}
+
+
+rgb hsv2rgb(hsv in)
+{
+    double      hh, p, q, t, ff;
+    long        i;
[…]

### 2021-05-12 tdp/PCLimpr/include/ioplanes.h

@@ -1,0 +1,50 @@
+/** @file
+ *  @brief Representation of a plane either in normal or convex hull form.
+ *  Wrapper class that extends the ConvexPlane class and adds 
+ *  IO funtionality. 
+ * 
+ *  @author Fabian Arzberger, JMU, Germany.
+ * 
+ *  Released under the GPL version 3.
+ */
+
+#ifndef __IO_CONV_PLANE_H_
+#define __IO_CONV_PLANE_H_
+
+#include "normalplane.h"
+#include "slam6d/normals.h"
+
+#include <sys/types.h>
+#include <sys/stat.h>
+#include <fstream>
+#include <sstream> 
+
+typedef std::vector<NormalPlane*> Planes;
+
+class PlaneIO {
+
+public:
+    static void read(std::string);
+    static Planes allPlanes;
+
+private: 
+    static int existsDir(const char* path)
+    {
+        struct stat info;
+        if (stat( path, &info ) != 0) return 0;
+        else if ( info.st_mode & S_IFDIR ) return 1;
+        else return 0;
+    }
+    // wrapper around Dot product function
+    static double dot(Point &p, double *p2)
+    {
+        double *p1 = new double[3];
+        p1[0] = p.x;
+        p1[1] = p.y;
+        p1[2] = p.z;
+        return Dot(p1, p2);
+    }
+};
+
+
+#endif //__NORM_CONV_PLANE_H_

### 2021-05-12 tdp/PCLimpr/include/normalplane.h

@@ -1,0 +1,29 @@
+/** @file
+ *  @brief Representation of a plane in normal form.
+ *  
+ *  @author Fabian Arzberger, JMU, Germany.
+ * 
+ *  Released under the GPL version 3.
+ */
+
+#ifndef __NORMALPLANE_H_
+#define __NORMALPLANE_H_
+
+#include "slam6d/globals.icc"
+
+class NormalPlane
+{
+public:
+    NormalPlane();
+    NormalPlane(double*, double*);
+    ~NormalPlane();
+
+    double n[3]; // normal
+    double x[3]; // reference point
+    double rho; // distance from ref to origin
+
+    double& operator()(int); // get normal xyz refs
+    double& operator[](int); // get reference xyz refs
+};
+
+#endif // __NORMALPLANE_H_

### 2021-05-12 tdp/PCLimpr/include/optimizer.h

@@ -1,0 +1,136 @@
+/** @file 
+ *  @brief This file implements the 6D gradient descent 
+ *  to minimize point to plane correspondant distances. 
+ *
+ *  @author Fabian Arzberger, JMU, Germany.
+ * 
+ *  Released under the GPL version 3.
+ */
+
+#ifndef __NEWTONS_H_
+#define __NEWTONS_H_
+
+#include <newmat/newmat.h>
+
+#include <utility>
+#include "planescan.h"
+
+using namespace NEWMAT;
+
+#define EPS_CONVERGENCE (-1)
+#define MAX_ITER (-1)
+#define ALPHA 0.001
+#define RPOS_ALPHA_SCALE 100
+
+// Used to lock some dimensions from being corrected.
+enum Dimensions {
+    ALL,            // performs 6D optimization
+    ROLLING,        // only optimizes rotX, as well as y and z (forward rolling motion, e.g. L.U.N.A sphere)
+    DESCENDING,     // only optimizes rotY and y (e.g. DAEDALUS descent phase) 
+    ROTATING        // only optimizes rotY (e.g. terrestial stationary RIEGL laserscanner)
+};
+
+class Optimizer
+{
+public:
+    Optimizer();
+    // Construct with Optimizer(x0, jacobian)
+    Optimizer(Matrix&, Matrix&);
+    // Construct from a PlaneScan directly
+    Optimizer(PlaneScan*);
+    // Construct with predefined dimensions to optimize
+    Optimizer(PlaneScan*, Dimensions);
+    ~Optimizer();
+
+    // Run <arg> iterations of gradient descent 
+    void iterate(int);
+    void operator()(int);
+
+    // Run until state difference between iterations is smaller then <arg>
+    void iterate(double);
+    void operator()(double);
+
+    // Combination of both top approaches
+    /*
+     * The overloaded operator() allows you to initialize 
+     * and run an Optimizer instance in the following manner:
+     * 
+     * Optimizer iter(X_0, jacobian);
+     * Optimizer::setMaxIter(10000);
+     * Optimizer::setEpsConvergence(0.0001);
+     * iter();
+     * 
+     * Use -E and -i to set these.
+     * Multiple criteria are supported.
+     * Running without criteria runs 1000 iterations.
+     */
+    void iterate(int, double);
+    void iterate();
+    void operator()(void);
+
+    // Auto-detect optimal initial alpha
+    void iterateAuto();
+
[…]

### 2021-05-12 tdp/PCLimpr/include/planescan.h

@@ -1,0 +1,132 @@
+/** @file 
+ *  @brief this file contains the PlaneScan Class that represents a Scan 
+ *  and holds the information about Point-to-Plane correspondences.
+ *
+ *  @author Fabian Arzberger, JMU, Germany.
+ * 
+ *  Released under the GPL version 3.
+ */
+
+#ifndef __PLANESCAN_H_
+#define __PLANESCAN_H_
+
+#include "slam6d/scan.h"
+#include "slam6d/globals.icc"
+#include "ioplanes.h"
+#include "ANN/ANN.h"
+
+#define EPS_POINT_2_PLANE_DIST 100 // distance in cm (?)
+
+typedef std::pair<double*, NormalPlane*> Correspondence;
+typedef std::vector< std::vector<int> > Clusters;
+
+class PlaneScan 
+{
+public:
+
+    static std::vector<PlaneScan*> allPlaneScans;
+    static Planes allPlanes;
+
+    ~PlaneScan();
+    PlaneScan() {};
+    PlaneScan(Scan*);
+    PlaneScan(Scan*, Planes&);
+
+    // for each point, store on which plane(s) it lies
+    // good for efficient iteration when pairs are needed
+    std::map<double*, NormalPlane*> correspondences;
+
+    // Scan data (indexed)
+    double **points;
+    double **normals;
+    int nrpts;
+    double timestamp;
+    double orig_transMat[16];
+    // Pose Data
+    double rPos[3];
+    double rPosTheta[3];
+    double transMat[16];
+    // Disk data
+    std::string basedir;
+    std::string identifier;
+
+    // Get point
+    const double* operator()(int);
+
+    bool isOkay();
+    bool isEmpty();
+    unsigned int index();
+
+    static void setUseCorrespondenceMin(bool);
+    static void setPlanes(Planes&);
+    static void setEpsDist(double);
+    static void setReduce(bool);
+    static double totalError();
+    static double dist2Plane(const double*, const double*, const NormalPlane*);
+    static double dist2Plane(const double*, const NormalPlane*);
+    // writes all the frame buffers to .frames files
+    static bool writeFrames();
+    // ensures that the .frames files have same length. this has to be done for animation
+    static void fillFrames(); 
+    static void setUseClustering(bool);
+    static void setKNearest(int);
+    static void setGrowthThresh(double);
+    static void setEpsSimilarity(double);
+    static void setMinClusterPoints(int);
+    static void setClusterVisualizationPath(string);
+
+    double calcErr();
[…]

### 2021-05-12 tdp/PCLimpr/src/ioplanes.cc

@@ -1,0 +1,67 @@
+/** @file 
+ *  @brief Class has only static members that handle Plane IO.
+ *  Reads "normals.list" and "planeXXX.n" files that have been previously
+ *  exported with bin/planes, utilizing the hough transform to detect
+ *  planes.
+ *  
+ *  @author Fabian Arzberger, JMU, Germany.
+ * 
+ *  Released under the GPL version 3.
+ */
+
+#include "ioplanes.h"
+
+Planes PlaneIO::allPlanes;
+
+void PlaneIO::read(std::string dir)
+{
+    // check if the given path exists
+    if ( !existsDir( dir.c_str() ) ) 
+    {
+        std::cout << "Using path: " << dir << std::endl;
+        std::cout << "Invalid plane path. Doesnt exist." << std::endl;
+        exit(0); 
+    }
+    dir += "normals.list";
+
+    // Read the normals.list file to get all planeXXX.n files 
+    std::vector<std::string> planefiles;
+    std::ifstream input;
+    std::string line; 
+    input.open( dir.c_str() );
+    while ( std::getline( input, line ) )
+    {
+        line = line.substr( 7, line.size() ); // remove the "Normal " entry
+        planefiles.push_back( line );
+    }
+    input.close();
+    input.clear();
+ 
+    // open planeXXX.n files and read points
+    std::vector<Point> points;
+    for ( const auto& path : planefiles )
+    {
+        double n[3];
+        double x[3];
+        std::cout << "Reading " << path << std::endl;
+        std::cout.flush();
+        // extract points
+        input.open( path.c_str() );
+        while ( input.good() )
+        {
+            try
+            {
+                input >> n[0] >> n[1] >> n[2];
+                input >> x[0] >> x[1] >> x[2];
+            }
+            catch (...) { break; }
+        }
+        input.close();
+        input.clear();
+
+        // Construct ConvexPlane objects
+        allPlanes.push_back( new NormalPlane(n, x) );
+
+    }
+
+}

### 2021-05-12 tdp/PCLimpr/src/normalplane.cc

@@ -1,0 +1,29 @@
+/** @file
+ *  @brief Representation of a plane in normal form.
+ *  
+ *  @author Fabian Arzberger, JMU, Germany.
+ * 
+ *  Released under the GPL version 3.
+ */
+
+#include "normalplane.h"
+
+NormalPlane::NormalPlane(double *n, double *x)
+{
+    for(int i = 0; i < 3; i++) 
+    {
+        this->n[i] = n[i];
+        this->x[i] = x[i];
+    }
+    rho = Dot(n, x);
+}
+
+double& NormalPlane::operator()(int idx)
+{
+    return n[idx];
+}
+
+double& NormalPlane::operator[](int idx)
+{
+    return x[idx];
+}

### 2021-05-12 tdp/PCLimpr/src/optimizer.cc

@@ -1,0 +1,425 @@
+/** @file 
+ *  @brief This file implements the 6D gradient descent 
+ *  to minimize point to plane correspondant distances. 
+ *
+ *  @author Fabian Arzberger, JMU, Germany.
+ * 
+ *  Released under the GPL version 3.
+ */
+
+#include "optimizer.h"
+#include <newmat/newmatio.h> //For debuging
+#include "planescan.h"
+#include "slam6d/globals.icc"
+
+// static members, program options
+int Optimizer::_max_iter = MAX_ITER;
+double Optimizer::_eps_convergence = EPS_CONVERGENCE;
+double Optimizer::_alpha = ALPHA;
+double Optimizer::_rPos_a_scale = RPOS_ALPHA_SCALE;
+bool Optimizer::_anim = false;
+bool Optimizer::_quiet = false;
+bool Optimizer::_auto = true;
+int Optimizer::_update_cor = 10;
+
+// Base constructor
+Optimizer::Optimizer()
+{
+    this->alpha = ColumnVector(6);
+    /*
+     * Typically, a change in orientation has much more impact on
+     * the total error than a change in position. While translating 
+     * the scan in x, y or z, the error grows linearily for all points.
+     * However, when rotating about x, y or z, points that are further
+     * away from the robot are moved even more dramatic, leading to 
+     * a higher sensibility on the error function. For this reason,
+     * the alpha applied on orientation must be much smaller than the 
+     * alpha applied on the position.
+     */ 
+    this->a = _alpha;
+    this->alpha << a 
+                << a  
+                << a
+                << a * _rPos_a_scale 
+                << a * _rPos_a_scale 
+                << a * _rPos_a_scale; 
+    this->J = ColumnVector(6);
+    this->Ga = ColumnVector(6);
+    this->Xa = ColumnVector(6);
+    // Do not buffer these from RAM ! C++ optimization is shit
+    Ga << 0 << 0 << 0 << 0 << 0 << 0;
+    Xa << 0 << 0 << 0 << 0 << 0 << 0;
+}
+
+Optimizer::Optimizer(PlaneScan* pScan) : Optimizer()
+{
+    // Setup current state vector
+    X0 = ColumnVector(6);
[…]

### 2021-05-12 tdp/PCLimpr/src/planereg.cc

@@ -1,0 +1,380 @@
+/** @file 
+ *  @brief Point-to-plane based post registration correction algorithm. 
+ *  Requires globaly registered scan archive and extracted planes.
+ * 
+ *  You can use bin/slam6d or bin/correction for registration.
+ *  Use bin/planes to extract planes from a scan.
+ *  To extract planes from the globaly registered scan, you have
+ *  to condense the scan archive into one globaly consistent .3d scan-file.
+ *  Use bin/condense for that.
+ *
+ *  @author Fabian Arzberger, JMU, Germany.
+ * 
+ *  Released under the GPL version 3.
+ */
+
+#include "planescan.h"
+#include "optimizer.h"
+#include "newmat/newmatio.h"
+#include "scanio/framesreader.h"
+#include <string>
+#include <omp.h>
+
+using namespace std;
+
+#include "slam6d/scan.h"
+#include "slam6d/io_types.h"
+#include "ioplanes.h"
+#include "planescan.h"
+#include "optimizer.h"
+
+#include <boost/program_options.hpp>
+namespace po = boost::program_options;
+
+// Boost needs to convert from string to IOType. Use this validation function.
+void validate(boost::any& v, const std::vector<std::string>& values,
+              IOType*, int) {
+  if (values.size() == 0)
+    throw std::runtime_error("Invalid model specification");
+  std::string arg = values.at(0);
+  try {
+    v = formatname_to_io_type(arg.c_str());
+  } catch (...) { // runtime_error
+    throw std::runtime_error("Format " + arg + " unknown.");
+  }
+}
+
+Dimensions formatname_to_dimensions(const char* str)
+{
+    if (strcasecmp(str, "all") == 0) return ALL;
+    else if (strcasecmp(str, "rolling") == 0) return ROLLING;
+    else if (strcasecmp(str, "descending") == 0) return DESCENDING;
+    else if (strcasecmp(str, "rotating") == 0) return ROTATING;
+    else throw std::runtime_error(string("Invalid Dimension type."));
+}
+
+// Same for Dimensions enum
+void validate(boost::any& v, const std::vector<std::string>& values,
+              Dimensions*, int) {
+  if (values.size() == 0)
+    throw std::runtime_error("Invalid model specification");
+  std::string arg = values.at(0);
+  try {
+    v = formatname_to_dimensions(arg.c_str());
+  } catch (...) { // runtime_error
+    throw std::runtime_error("Format " + arg + " unknown.");
+  }
+}
+
+/*
[…]

### 2021-05-12 tdp/PCLimpr/src/planescan.cc

@@ -1,0 +1,603 @@
+/** @file 
+ *  @brief this file contains the PlaneScan Class that represents a Scan 
+ *  and holds the information about Point-to-Plane correspondences.
+ *
+ *  @author Fabian Arzberger, JMU, Germany.
+ * 
+ *  Released under the GPL version 3.
+ */
+
+#include "planescan.h"
+#include "scanio/helper.h"
+
+#include "debugutil.h"
+
+// initialize the static members
+PlaneScans PlaneScan::allPlaneScans;
+unsigned int PlaneScan::idx_global = 0;
+double PlaneScan::_eps_dist = EPS_POINT_2_PLANE_DIST;
+bool PlaneScan::reduce = false; 
+Planes PlaneScan::allPlanes;
+bool PlaneScan::use_correspondence_min = false;
+bool PlaneScan::use_clustering = true;
+int PlaneScan::k_neighbours = 20;
+double PlaneScan::d_growth = 5.0;
+int PlaneScan::n_min_clusterpoints = 100;
+double PlaneScan::eps_similarity = 5;
+string PlaneScan::cluster_outpath = "";
+
+// Calculates the distance from a transformed point "xyz" to plane "p"
+double PlaneScan::dist2Plane(const double *xyz, const double *trans, const NormalPlane *p)
+{
+    double xyz_transformed[3];
+    transform3(trans, xyz, xyz_transformed);
+    return dist2Plane( xyz_transformed, p );
+}
+
+// Calculates the distance from a point "xyz" to plane "p"
+double PlaneScan::dist2Plane(const double *xyz, const NormalPlane *p)
+{
+    double v[3];
+    v[0] = xyz[0] - p->x[0];
+    v[1] = xyz[1] - p->x[1];
+    v[2] = xyz[2] - p->x[2];
+	return Dot(v, p->n); 
+}
+
+// Checks if a point "xyz" with transformation "trans" lies on the plane "p"
+bool inline PlaneScan::isInPlane(double *xyz, double *trans, const NormalPlane *p)
+{
+    // Calculate distance to plane and apply threshold
+    return fabs( dist2Plane(xyz, trans, p) ) < _eps_dist;
+}
+
+// Static setter that sets the distance threshold for plane correspondence.
+// Use before constructing the object!
+void PlaneScan::setEpsDist(double eps_dist)
+{
+	_eps_dist = eps_dist;
+}
+
[…]

### 2021-05-12 Latex/sections/acknowledgement.tex
Added acknowledgements

%!TEX root = ../main.tex
\section*{Acknowledgement}
\todo[inline]{Formulate these: ESA funding (?), elite netzwerk bayern (?), cylon floatsat group}

### 2021-05-12 Latex/sections/realData.tex
Removed (\iffalse ... \fi) sections about datasets without results

@@
 
-Our algorithm was also tested on a number of real-world datasets that adhere to different motion profiles and laser scanners with different scanning patterns.
+Our algorithm was also tested on a real-world datasets that adheres to different motion profile and laser scanners type with different scanning patterns than the simulated data.
 
-\subsection{Floating Sphere}
-
-The first dataset is collected by a line scanner, in particular a SICK LMS141 industrial scanner, inside a sphere of a diameter of \SI{20}{\centi\meter} that lies on a floating desk, which is air-pressurized, such that the sphere is hovering.
+The dataset is collected by a line scanner, in particular a SICK LMS141 industrial scanner, inside a sphere of a diameter of \SI{20}{\centi\meter} that lies on a floating desk, which is air-pressurized, such that the sphere is hovering.
 On this floating desk the sphere can rotate freely about all axis while being fixed in position. 
-Hence, in this experiment the optimization space can be reduced to a rotation but also the motion is required to obtain a 3D model from the 2D laser scanner. 
-The sphere is equipped with a low-cost IMU, i.e., a PhidgetSpatial Precision 3/3/3, to estimate the initial orientation.
+Hence, in this experiment the optimization space can be reduced to a rotation.
+Further, without the motion the 2D laser also cannot obtain a 3D model from the environment, thus requiring registration. 
+The sphere is equipped with a low-cost IMU, i.e., a PhidgetSpatial Precision 3/3/3, to estimate the orientation.
 
@@
 
+\iffalse % This is commented out because we couldn't produce results
 	\subsection{RADLER}
@@
 	\end{figure}
+\fi

### 2021-05-12 Latex/sections/registration.tex
Removed (\iffalse ... \fi) sections about datasets without results

@@
 
-\todo[inline]{Add section about the local plane, k-means normal based region growing aka clustering for point2plane correspondences. 
-Basically, if a point is correspondence to a global plane than all other points that lie on its local plane should be too.
-The local plane is constructed by region growing, i.e.
-if the difference in normal and the distance of the points are close enough they are clustered to one local plane.}
-
 \subsection{Optimization}
@@
 	&\nabla E =  \sum_{\forall \mathcal{P}_i}\sum_{\vec{p}_k \in \mathcal{P}_i} \resizebox{0.28\textwidth}{!}{$\begin{bmatrix}\pardiff{\varphi} E&\pardiff{\vartheta} E&\pardiff{\psi} E&\pardiff{t_x} E&\pardiff{t_y} E&\pardiff{t_z} E\end{bmatrix}^T$}\\
-    &\Rightarrow\nabla E = \sum_{\forall \mathcal{P}_i} \sum_{\vec{p}_k \in \mathcal{P}_i}\resizebox{0.28\textwidth}{!}{$2D(\vec{\Pi}, \vec{p}_k)\begin{bmatrix}\nabla E_\varphi &\nabla E_\vartheta & \nabla E_\psi & n_{\mathcal{P}_i}^x&n_{\mathcal{P}_i}^y&n_{\mathcal{P}_i}^z\end{bmatrix}^T$}
+    &\Rightarrow\nabla E = \resizebox{0.35\textwidth}{!}{$\sum_{\forall \mathcal{P}_i} \sum_{\vec{p}_k \in \mathcal{P}_i}2D(\vec{\Pi}, \vec{p}_k)\begin{bmatrix}\nabla E_\varphi &\nabla E_\vartheta & \nabla E_\psi & n_{\mathcal{P}_i}^x&n_{\mathcal{P}_i}^y&n_{\mathcal{P}_i}^z\end{bmatrix}^T$}
 \end{align}
@@
     &\begin{alignedat}{3}
-        \nabla E_\varphi = & n_{\mathcal{P}_i}^x y S_\varphi S_\psi \\
+        \nabla E_\varphi = & n_{\mathcal{P}_i}^x&&y S_\varphi S_\psi \\
         &+ n_{\mathcal{P}_i}^y&&(x[-S_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] \\&&&+y[-S_\varphi C_\psi -C_\varphi S_\vartheta S\psi] \\&&&- z C_\varphi C_\vartheta)\\
@@
     &\begin{alignedat}{3}
-       \nabla E_\psi =& n_{\mathcal{P}_i}^x\left(-xC_\vartheta S_\psi - yC_\varphi C_\psi\right) \\
-       &+ n_{\mathcal{P}_i}^y\left(x[C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi] + y[-C_\varphi S_\psi - S_\varphi S_\theta C_\psi]\right)  \\
-       &+ n_{\mathcal{P}_i}^z\left(x[S_\varphi C_\psi + C_\varphi S_\psi S_\theta] + y[-S_\psi S_ \varphi + C_\varphi S_\vartheta C_\psi] \right) 
+       \nabla E_\psi =& n_{\mathcal{P}_i}^x&&\left(-xC_\vartheta S_\psi - yC_\varphi C_\psi\right) \\
+       &+ n_{\mathcal{P}_i}^y&&(x[C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi] \\
+       &&&+ y[-C_\varphi S_\psi - S_\varphi S_\theta C_\psi]) \\
+       &+ n_{\mathcal{P}_i}^z&&(x[S_\varphi C_\psi + C_\varphi S_\psi S_\theta] \\
+       &&&+ y[-S_\psi S_ \varphi + C_\varphi S_\vartheta C_\psi]) 
     \end{alignedat}

### 2021-05-12 Latex/sections/results.tex
Removed (\iffalse ... \fi) sections about datasets without results

@@
 
-Figure~\ref{fig:cylon-corrected} shows the results obtained after employing the plane base registration on the dataset acquired by the floating sphere.
+Figure~\ref{fig:cylon-corrected} shows the results obtained before  and after employing the plane base registration on the dataset acquired by the floating sphere.
 \begin{figure*}
@@
 In particular the noise in the top left corner of the top view (cf. figure~\ref{fig:cylon-corrected}) has been visibly reduced.
-Further, the noise around the walls is notably reduced as the points are moved on the plane.
+Further, the deviation of points around the walls is notably smaller as the points are moved on the plane.
 
 However, some imprecisions remain. 
-At the same location a few points cannot be matched to the plane as they are clustered with some other points that are bound to a different pose. 
-Hence this error stems from the imprecise interpolation in between scans. 
-\todo[inline]{More comments here}
+In particular, at the initially very noisy top left corner of the top view, a few points cannot be matched to the plane. 
+The scans that cause these imprecisions are problematic as they measure a number of plane like structures in the interior of the room. 
+When relabeling those scans the interior points are associated to a plane and hence the corresponding distance minimized.
 
+% This section is commented out
+\iffalse
 \subsection{RADLER Results}
@@
 \end{figure}
+\fi

### 2021-05-12 Latex/sections/acknowledgement.tex
Updated conclusion and acknowledgment

@@
 \section*{Acknowledgement}
-\todo[inline]{Formulate these: ESA funding (?), elite netzwerk bayern (?), cylon floatsat group}
+\todo[inline]{Formulate these: ESA funding (?), elite netzwerk bayern (?)}
+
+We thank the Cylon team (Timothy Randolph, Ignacio Dorado, Camilo Andres Reyes) for providing the dataset and documentation of the experimental setup used in this work.

### 2021-05-12 Latex/sections/conclusion.tex
Updated conclusion and acknowledgment

@@
 
-In this paper, we propose a registration method for man made environments that exploits the structure of the environments. 
-Given in man-made environments parallel planes and right angles are abundantly available we employ point-to-plane correspondences to improve a pre-registered 3D point-cloud. 
-We evaluate the procedure on a simulated dataset as well as on a experimentally acquired dataset with different motion profiles. 
-In this evaluation we show that the procedure improves both datasets and yields more human-made structure like maps. 
+In this paper, we propose a registration method for man made environments that exploits the structure of those environments. 
+Given that in man-made environments parallel planes and right angles are abundantly available we employ point-to-plane correspondences to improve a pre-registered 3D point-cloud. 
+We evaluate the procedure on a simulated dataset as well as on an experimentally acquired dataset with different motion profiles. 
+In this evaluation we show that the procedure improves both datasets and yields more human-made, structure like maps. 
+In particular, the qualitative structure of the environments is reconstructed well, that is, in the resulting maps the parallel walls are clearly improved.
 Currently not all steps in the procedure are autonomous, in particular the parameter tuning. 
@@
 One approach is to introduce soft-locks for the optimization dimensions.
-I.e., instead of locking the some dimensions entirely from being used for optimization, adding weighting factors based on the dynamics of the system that encode which noise source is more likely.
+I.e., instead of locking some dimensions entirely from being used for optimization, they are weighted based on the dynamics of the system that encode which noise source is more likely.

### 2021-05-12 Latex/main.tex
Added acknowledgements
Changed author listing
acknowledgement

@@
 \begin{abstract}
-Spherical robots are a format that hasn't been thoroughly explored for the application of mobile mapping. 
+Spherical robots are a format that has not been thoroughly explored for the application of mobile mapping. 
 In contrast to other designes it provides some unique advantages. 
@@
 Mobile mapping always requires a high-precise pose knowledge to obtain consistent and correct environment maps. 
-This is typically be done by a combination of external reference sensors such as GNSS measurements and inertial measurements, or by post processing the data and registering the different measurements to each other.
+This is typically done by a combination of external reference sensors such as Global Navigation Satellite System (GNSS) measurements and inertial measurements, or by post processing the data and registering the different measurements to each other.
 In indoor environments the GNSS reference is not an option, hence many mobile mapping applications turn to the second option.
-An advantage of indoor environments is, that human made environments usually have a certain structure, such as parallel planes and right angles.
+An advantage of indoor environments is, that human made environments usually have a certain structure, such as parallel and perpendicular planes.
 We propose a registration procedure that exploits this structure by minimizing the distance of measured points to a corresponding plane. 
@@
 \input{sections/conclusion}
+\input{sections/acknowledgement}
 

### 2021-05-12 Latex/sections/acknowledgement.tex
acknowledgement

--- 2021-05-12 Latex/sections/acknowledgement.tex
@@
 \section*{Acknowledgement}
-\todo[inline]{Formulate these: ESA funding (?), elite netzwerk bayern (?), cylon floatsat group}
+\todo[inline]{check this}
+The authors thank the Cylon team (Ignacio Dorado Llamas, Timothy Randolph, Camilo Andres Reyes Mantilla, Tobias Wahl) for designing and building the prototype and Timo Burger for acquiring the floating table data set.
+Special thanks to Dieter Ziegler and Sergio Montenegro for supporting our work.
+We acknowledge funding from the ESA Contract No.
+4000130925/20/NL/GLC for the ``DAEDALUS -- Descent And Exploration in Deep Autonomy of Lava Underground Structures'' Open Space Innovation Platform (OSIP) lunar caves-system study and the Elite Network Bavaria (ENB) for providing funds for the academic program ``Satellite Technology''.

### 2021-05-12 Latex/sections/introduction.tex
acknowledgement

@@
 Smaller robotic systems are particularly used when cars no longer have access.
-Examples for this are human operated systems such as Zebedee~\cite{FIXME}, a Hokuyo scanner on a spring, VILMA~\cite{JPRS2016}, a rolling FARO scanner, RADLER~\cite{Borrmann2020-RADLER}, a 2D laser scanner mounted to a unicycle, or a backpack mounted ``personal laser scanning system'' as in~\cite{LauterbackEtAl2015-Backpack}.
+Examples for this are human operated systems such as Zebedee~\cite{bosse2012zebedee}, a Hokuyo scanner on a spring, VILMA~\cite{JPRS2016}, a rolling FARO scanner, RADLER~\cite{Borrmann2020-RADLER}, a 2D laser scanner mounted to a unicycle, or a backpack mounted ``personal laser scanning system'' as in~\cite{LauterbackEtAl2015-Backpack}.
 Recently more and more autonomous systems gained maturity.

### 2021-05-12 Latex/sections/introduction.tex

--- 2021-05-12 Latex/sections/introduction.tex
@@
 Smaller robotic systems are particularly used when cars no longer have access.
-Examples for this are human operated systems such as Zebedee~\cite{FIXME}, a Hokuyo scanner on a spring, VILMA~\cite{JPRS2016}, a rolling FARO scanner, RADLER~\cite{Borrmann2020-RADLER}, a 2D laser scanner mounted to a unicycle, or a backpack mounted ``personal laser scanning system'' as in~\cite{LauterbackEtAl2015-Backpack}.
+Examples for this are human operated systems such as Zebedee~\cite{Bosse2012-zebedee}, a Hokuyo scanner on a spring, VILMA~\cite{JPRS2016}, a rolling FARO scanner, RADLER~\cite{Borrmann2020-RADLER}, a 2D laser scanner mounted to a unicycle, or a backpack mounted ``personal laser scanning system'' as in~\cite{LauterbackEtAl2015-Backpack}.
 Recently more and more autonomous systems gained maturity.

### 2021-05-12 Latex/sections/registration.tex

@@
 Depending on the threshold value, some points now have multiple correspondences if their distance to multiple planes is less then $\epsilon$.
-We then simply omit all points that have correspondences to multiple planes from the optimizations, to avoid introducing errors due to the plane ambiguity.
+We then simply omit all points that have correspondences to multiple planes from the optimization, to avoid introducing errors due to the plane ambiguity.
 

### 2021-05-13 Latex/sections/introduction.tex
small corrections introduction

@@
 Today's robots for mobile mapping come in all shapes and sizes.
-The state of the art for urban environments are laser scanners mounted to cars.
+State of the art for urban environments are laser scanners mounted to cars.
 Smaller robotic systems are particularly used when cars no longer have access.
@@
 A stunning example is Boston Dynamics' quadruped ``Spot'' that autonomously navigates and maps human environments~\cite{SpotRobot}.
-Also the mobile mapping approaches implemented on the ANYmal platform such as~\cite{Fankhauser2018-ANYmal} were very successful.
+Also, the mobile mapping approaches implemented on the ANYmal platform such as~\cite{Fankhauser2018-ANYmal} were very successful.
 Of all these formats, one hasn't been explored thoroughly in the scientific community: The spherical mobile mapping robot.
 Yet this provides some very promising advantages over the other formats.
-For one, the locomotion of a spherical robot inherently results in a rotation.
-That way a sensor fixed inside the spherical structure will cover the entire environment, given the required locomotion without the need for additional actuators for the sensors.
+For one, the locomotion of a spherical robot inherently results in rotation.
+That way, a sensor fixed inside the spherical structure will cover the entire environment, given the required locomotion without the need for additional actuators for the sensors.
 This requires a solution for the spherical simultaneous localization and mapping (SLAM) problem given the six degrees of freedom of the robot.
-Secondly a spherical shell, that encloses all sensors protects these from possible hazardous environments.
-For example, any dust that deteriorates sensors or actuators when settling at sensitive locations is stopped by a shell.
-This is be particularly useful for unknown and dangerous environments.
-E.g., for space applications this is also proposed in the DAEDALUS study~\cite{RossiMaurelliUnnithanetal.2021}, a spherical robot that is to be lowered into a lunar cave and create a 3D map of the environment.
+Secondly, a spherical shell that encloses all sensors protects these from possible hazardous environments.
+For example, the shell stops any dust that deteriorates sensors or actuators when settling at sensitive locations.
+This is particularly useful for unknown and dangerous environments.
+E.g., for space applications, this is also proposed in the DAEDALUS study~\cite{RossiMaurelliUnnithanetal.2021}, a spherical robot that is to be lowered into a lunar cave and create a 3D map of the environment.
 The authors choose this format as the present moon regolith is known to damage instruments and other components.
 
-This paper proposes to use such a spherical robot for mobile mapping man made environments.
-In such environments one advantage are architectural shapes following standard conventions arising from tradition or utility, i.e., there are many flat surfaces such as walls, floors etc.
+This paper proposes to use such a spherical robot for mobile mapping man-made environments.
+In such environments, one advantage are architectural shapes following standard conventions arising from tradition or utility, i.e., there are many flat surfaces such as walls, floors, etc.
 that are sensed.
 Exploiting this fact yields more opportunities for registration as point-to-plane correspondences can be used.
-The proposed registration method minimizes the distances of each point to its corresponding plane as objective function.
+The proposed registration method minimizes the distances of each point to its corresponding plane as an objective function.
 The method is evaluated on a simulated dataset as well as on experimentally acquired datasets.

### 2021-05-13 Latex/main.tex

@@
 %% Group authors per affiliation:
-\author{Fabian Arzberger, Anton Bredenbeck, Jasper Zevering, Dorit Borrmann, Andreas Nüchter}
+\author{Fabian Arzberger and Anton Bredenbeck, Jasper Zevering, Dorit Borrmann, Andreas Nüchter}
 \address{Informatics VII: Robotics and Telematics,
@@
 Spherical robots are a format that has not been thoroughly explored for the application of mobile mapping. 
-In contrast to other designes it provides some unique advantages. 
-Among those are a protective spherical shell which protects internal sensors and actuators from possible harsh environments as well as an inherent rotation for locomotion that enables measurements in all directions.
+In contrast to other designes, it provides some unique advantages. 
+Among those is a protective spherical shell that protects internal sensors and actuators from possible harsh environments, as well as an inherent rotation for locomotion that enables measurements in all directions.
 Mobile mapping always requires a high-precise pose knowledge to obtain consistent and correct environment maps. 
-This is typically done by a combination of external reference sensors such as Global Navigation Satellite System (GNSS) measurements and inertial measurements, or by post processing the data and registering the different measurements to each other.
-In indoor environments the GNSS reference is not an option, hence many mobile mapping applications turn to the second option.
-An advantage of indoor environments is, that human made environments usually have a certain structure, such as parallel and perpendicular planes.
+This is typically done by a combination of external reference sensors such as Global Navigation Satellite System (GNSS) measurements and inertial measurements or by post processing the data and registering the different measurements to each other.
+In indoor environments, the GNSS reference is not an option.
+Hence many mobile mapping applications turn to the second option.
+An advantage of indoor environments is that human-made environments usually have a certain structure, such as parallel and perpendicular planes.
 We propose a registration procedure that exploits this structure by minimizing the distance of measured points to a corresponding plane. 
-Further we evaluate the procedure on a simulated dataset of an ideal corridor as well as in experimentally acquired datasets with different motion profiles. 
+Further, we evaluate the procedure on a simulated dataset of an ideal corridor and on an experimentally acquired dataset with different motion profiles. 
 We show that we nearly reproduce the ground truth for the simulated dataset and improve the average point-to-point distance to a reference scan in the experimental dataset.

### 2021-05-13 Latex/sections/stateOfTheArt.tex

@@
 
-Up to the best of our knowledge, SLAM with spherical robots and laser scanners was not done before, while there exists camera based approaches~\cite{9233654}.
+Up to the best of our knowledge, SLAM with spherical robots and laser scanners was not done before, while there exists camera-based approaches~\cite{9233654}.
 However, laser-based SLAM algorithms for motions in six degrees of freedom (DoF) have been thoroughly studied.
 For outdoor environments~\cite{JFR2006} provides a first baseline.
-Adding a heuristic for closed loop detection and a global relaxation Borrmann et al.
-yield highly precise maps of the scanned environment~\cite{RAS2007}.
+Adding a heuristic for closed-loop detection and a global relaxation Borrmann et al. yield highly precise maps of the scanned environment~\cite{RAS2007}.
 Zhang et al. propose a real-time version of the SLAM algorithm in~\cite{Zhang2014}.
 They achieve the performance at a lower computational load by dividing the SLAM algorithm into two different algorithms: one performing odometry at a high frequency but low fidelity and another running at a lower frequency performing fine matching and registration of the point clouds.
-The more recently Dröschel et al.
-also proposes an online method using a novel combination of a hierarchical graph structure with local multi-resolution maps to overcome problems due to sparse scans~\cite{Droeschel2018}.
+More recently Dröschel et al.
+also propose an online method using a novel combination of a hierarchical graph structure with local multi-resolution maps to overcome problems due to sparse scans~\cite{Droeschel2018}.
 
-Since these approaches are based on point-to-point correspondences they require rather high point density to achieve precise registration.
-For low-cost LiDARs this implies a slow motion and long integration time.
-Further, none of the approaches exploit the structure inherent to man-made environments, hence missing out on possible advantages.
+Since these approaches are based on point-to-point correspondences, they require a rather high point density to achieve precise registration.
+For low-cost LiDARs, this implies slow motion and long integration time.
+Further, none of the approaches exploit the structure inherent to human-made environments, hence missing out on possible advantages.
 
@@
 To overcome the requirements on point-density imposed by the point-to-point correspondences instead other correspondences are used. 
-In man made environments planes are abundantly available, and hence provide an attractive base for correspondences.
-Förster et al. use this property of man made environment successfully in~\cite{Foerstner2017}.
-They register point clouds using plane-to-plane correspondences and include uncertainty measures for the detected planes as well as the estimated motion.
-Thereby, they propose a costly exact algorithm and cheaper approximations that yield high quality maps.
-Favre et al.~\cite{favre2021} use point-to-plane correspondences after preprocessing the point clouds using plane-to-plane correspondences to successfully register two scans with each other.
+In human-made environments, planes are abundantly available and hence provide an attractive base for correspondences.
+Förster et al. use this property of human-made environment successfully in~\cite{Foerstner2017}.
+They register point clouds using plane-to-plane correspondences and include uncertainty measures for the detected planes and the estimated motion.
+Thereby, they propose a costly exact algorithm and cheaper approximations that yield high-quality maps.
+Favre et al.~\cite{favre2021} use point-to-plane correspondences after preprocessing the point clouds using plane-to-plane correspondences to register two scans with each other successfully.
 
-Both approaches use plane-to-plane correspondences to pre-register the scans. However for pre-registration the classical point-to-point registration is also very effective. 
-One advantage that point-to-point correspondences have over plane-to-plane correspondences is, that they do not require a long stop in each pose.
-For plane-to-plane correspondences this is required to gather enough data to measure planes in each individual scan robustly.
+Both approaches use plane-to-plane correspondences to pre-register the scans.
+However, for pre-registration the classical point-to-point registration is also very effective. 
+One advantage that point-to-point correspondences have over plane-to-plane correspondences is that they do not require a long stop in each pose.
+For plane-to-plane correspondences, this is necessary to gather enough data to measure planes in each scan robustly.
 The resulting scan procedure is stop-scan-and-go.
-In particular for the application of a spherical robot this standstill in each pose cannot be guaranteed or even approximated making continuous-time approaches using point-to-plane correspondences the method of choice.
+In particular, for the application of a spherical robot this standstill in each pose cannot be guaranteed or even approximated, making continuous-time approaches using point-to-plane correspondences the method of choice.
 
-In the following we propose this combination of point-to-point based pre-registration followed by a point-to-plane based optimization.
+In the following, we propose this combination of point-to-point based pre-registration followed by a point-to-plane based optimization.

### 2021-05-13 Latex/sections/registration.tex
small corrections registrtation

@@
 
-We start with transforming each line scan into the project coodinate system , which is defined by the pose of the first acquired line scan.
-With line scan, we mean the smallest chunk of range data, we obtain from the scanner device driver.
-In case of a SICK LMS1xx it is a line and in case of a Livox scanner it has a flower shape.
+We start with transforming each line scan into the project coordinate system, which is defined by the pose of the first acquired line scan.
+With line scan, we mean the smallest chunk of range data we obtain from the scanner device driver.
+In the case of a SICK LMS1xx it is a line, and in case of a Livox scanner it has a flower shape.
 More details are provided in the following sections.
 %
-For map improvement the individual scans need to be registered to another. 
-We propose an algorithm that consist of multiple steps outlined in algorithm~\ref{algo:registration-algorithm}. 
-Based on ideas described in~\cite{Borrmann2010} we first rigidly register the scans and then further improve the overall map by exploiting the fact that man made environments often consist of planes. 
-We then find the planes in the pre-registered point-cloud and then optimize the poses associated with the scans to minimize the distance of all points to their respective planes.
+For map improvement, the individual scans need to be registered to another. 
+We propose an algorithm that consists of multiple steps outlined in algorithm~\ref{algo:registration-algorithm}. 
+Based on ideas described in~\cite{Borrmann2010} we first rigidly register the scans and then further improve the overall map by exploiting the fact that human-made environments often consist of planes.
+We then find the planes in the pre-registered point cloud and then optimize the poses associated with the scans to minimize the distance of all points to their respective planes.
 
@@
 
-After having done the rigid registration the scans are aligned well-enough to make statements about the potential planes in the environment.
-To find the planes in the environment a Hough transform with an accumulator ball as described in~\cite{Borrmann2011The3H} is used. 
-In the next subsection the Hesse-normal form of the plane is required.
+After having done the rigid registration, the scans are aligned well enough to make statements about the potential planes in the environment.
+To find the planes in the environment, a Hough transform with an accumulator ball as described in~\cite{Borrmann2011The3H} is used. 
+In the next subsection, the Hesse-normal form of the plane is required.
 For an ideal plane the distance parameter $\rho_{\mathcal{P}_i}$ is computed via $\vec{n}_{\mathcal{P}_i}\cdot\vec{x}_i$, where $\vec{x}_i$ is an arbitrary point on the plane.
-To find this point we first define a convex hull that encloses the plane.
+To find this point, we first define a convex hull that encloses the plane.
 It is defined by the points that lie on the plane with the furthest distance to one another.
@@
 There are many ways of finding correspondences between points and planes.
-However, even with the result not being perfect, we assume already globally registered scans, which is why a simplistic algorithm is sufficient:
-For each plane, consider adding a correspondence for all points that are closer to the plane than a threshold $\epsilon$.
-Depending on the threshold value, some points now have multiple correspondences if their distance to multiple planes is less then $\epsilon$.
-We then simply omit all points that have correspondences to multiple planes from the optimization, to avoid introducing errors due to the plane ambiguity.
+However, even with the result not being perfect, we assume already globally registered scans, which is why a simple algorithm is sufficient:
+For each plane, consider adding a correspondence for all points closer to the plane than a threshold $\epsilon$.
+Depending on the threshold value, some points now have multiple correspondences if their distance to multiple planes is less than $\epsilon$.
+We then omit all points with correspondences to multiple planes from the optimization to avoid introducing errors due to the plane ambiguity.
 
@@
 
-Assuming we now know the Hesse-normal form of all planes and all points that are assigned to these planes, we register the points to have a minimal distance to their respective planes.
+Assuming we now know the Hesse-normal form of all planes and all assigned points to these planes, we register the points to have a minimal distance to their respective planes.
 The transformation $T(\vec{p}_k)$ of each point $\vec{p}_k$ with respect to a 6 DoF motion is described in homogeneous coordinates using the roll-pitch-yaw ($\varphi-\vartheta-\psi$) Tait-Brian angles as in~\cite{diebel2006representing}.
@@
 \end{align} 
-For our particular application ADADELTA behaves a little too aggressively.
-Despite giving a good measure on how to adabt $\boldsymbol\alpha$, the algorithm sometimes overshoots, and convergence doesn't happen.
+For our particular application, ADADELTA behaves a little too aggressively.
+Despite giving a good measure on how to adapt $\boldsymbol\alpha$, the algorithm sometimes overshoots, and convergence doesn't happen.
 Therefore, we employ another scaling factor, typically not found in ADADELTA, extending eq.~\eqref{eq:adadeltaalphaupdate} to:
@@
 Using this algorithm once after finding correspondences from points to planes leads to convergence to a local minimum, which often isn't an optimal solution.
-Even if the number of iterations is dramatically increased, no better solution than the local minimum is found.
-That is, unless you consider updating the correspondence model after $i$ iterations of gradient descent.
+Even if  we increase the number of iterations dramatically, no better solution than the local minimum is found.
+That is unless you consider updating the correspondence model after $i$ iterations of gradient descent.
 Re-assigning point-to-plane correspondences this way $k$ times, and if $k$ is chosen large enough, it leads to an optimal solution after maximum $n = k\cdot i$ iterations of gradient descent.
@@
 
-The algorithm was extended by two further optmizations which are useful in different scenarios.
+The algorithm was extended by two further optimizations, which are helpful in different scenarios.
 
@@
 Although 6D optimization generally works, reducing the optimization space is particularly useful if the source of error in the system is known and a model exists.
-That way, e.g.
-a point cloud from a terrestrial lasescanner that experiences irregularites in its spinning mirror is registered by optimizing only over the axis of rotation as no other movement is to be expected.
+That way, e.g., a point cloud from a terrestrial laserscanner that experiences irregularities in its spinning mirror is registered by optimizing only over the axis of rotation as no other movement is to be expected.
 
-Secondly we employ a continuous iteration, where the scans are processed sequentially. 
+Secondly, we employ a continuous iteration, where the scans are processed sequentially. 
 This is especially useful for mobile systems, where pose error accumulates due to increasing tracking uncertainties.
-The assumption is that the the error in one scan is also present in the next scan, plus some unknown new error.
+The assumption is that the error in one scan is also present in the following scan, plus some unknown new error.
 We eliminate the error from scan $m$ which is also present in scan $m+1$ by applying the pose change from scan $m$, $\vec{\Pi}_{n,m} - \vec{\Pi}_{0,m}$ after $n$ gradient descent iterations, to scan $m+1$, before restarting gradient descent.

### 2021-05-13 Latex/sections/conclusion.tex

@@
 In particular, the qualitative structure of the environments is reconstructed well, that is, in the resulting maps the parallel walls are clearly improved.
-Currently not all steps in the procedure are autonomous, in particular the parameter tuning. 
+
+However, as \ref{sec:floatres} shows, some outliers remain due to the simplistic point-to-plane correspondence model, which is to be improved in future studies.
+One idea to do so is to employ a local clustering technique onto the individual scans, before registering them globaly. 
+Points from the scan with a similar normal get clustered together if their distance is small enough.
+This way, a cluster represents a local plane, while at the same time storing all the points that belong to that plane.
+Point-to-plane correspondences then arise from the intersection of local clusters with global planes.
+
+Furthermore, currently not all steps in the procedure are autonomous, in particular the parameter tuning. 
 In the future one goal is to increase the autonomy of the system.

### 2021-05-13 Latex/sections/simulation.tex
small corrections simulation

@@
-To quantify the quality of the proposed registration method, it is tested on simulated as well as on real world data.
+To quantify the quality of the proposed registration method, it is tested on simulated and real-world data.
 
@@
 
-For the simulated data we implemented a noisy world-robot-sensor model.  
+For the simulated data, we implemented a noisy world-robot-sensor model.  
 We are simulating a LIVOX-Mid100~\cite{LivoxMid40-100} laser scanner with customizable noise level on the range measurements inside a robot with different motion capabilities, subject to noise in its pose estimation.
@@
 Therefore the simulated sensors are as well. 
-To achieve this we sample a noise percentage $n_P$ from some normal distribution $\mathcal{N}(\mu,\sigma^2)$. 
+To achieve this, we sample a noise percentage $n_P$ from some normal distribution $\mathcal{N}(\mu,\sigma^2)$. 
 The resulting range $r$ given the true range $r_t$ is then
@@
 
-For the pose measurement one cannot simply add white noise to the current pose estimate as this does not capture the integration error which is common among inertial measurement units (IMU). 
+For the pose measurement, one cannot simply add white noise to the current pose estimate as this does not capture the integration error, which is common among inertial measurement units (IMU). 
 Therefore we assume a disturbance torque about the two axes that lie in the ground plane. 
-This is equivalent to a case, where a slightly shifted ground plane or a unbalanced locomotion system of the robot is present.
+This is equivalent to a case where a slightly shifted ground plane or an unbalanced locomotion system of the robot is present.
 
 Accordingly, a spherical robot that is assumed to be translating exactly along one axis by rotation exactly about one other axis at an angular velocity $\omega$  experiences random disturbance torques at each time instant that accelerate the sphere's rotation about the respective axes.
-This additional motion is determined by integrating the disturbance torques which are sampled from some normal distribution $n_\phi, n_\psi \in \mathcal{N}(\mu,\sigma^2)$ each.  
+This additional motion is determined by integrating the disturbance torques, which are sampled from some normal distribution $n_\phi, n_\psi \in \mathcal{N}(\mu,\sigma^2)$ each.  
 The pose update in the discrete simulation is therefore given by: 
@@
 
-Figure~\ref{fig:simulatedDatasets} shows a noise free trajectory through a simulated environment as well as a trajectory with noisy sensors.
+Figure~\ref{fig:simulatedDatasets} shows a noise-free trajectory through a simulated environment and a trajectory with noisy sensors.
 In the noise-free case, the trajectory slightly bends in the direction of disturbance torque sideways while also varying in velocity in the intended direction of travel.
-In the noisy case the ideal straight trajectory is assumed, hence the planes enclosing the room are sensed multiple times. 
-In particular we see that the trajectory of the robot leads through one of the sensed planes. 
-Also, the further the sensed points are the noisier they appear, which is consistent with the larger influence of the pose error at higher distances.
+In the noisy case, the ideal straight trajectory is assumed.
+Hence the planes enclosing the room are sensed multiple times. 
+In particular, we see that the trajectory of the robot leads through one of the sensed planes. 
+Also, the further the sensed points are, the noisier they appear, which is consistent with the larger influence of the pose error at higher distances.
 

### 2021-05-13 Latex/sections/realData.tex
Added todo for fabi; added clustering info to cylon; slightly resized cylon image
small corrections real data

@@
 
-Our algorithm was also tested on a real-world datasets that adheres to different motion profile and laser scanners type with different scanning patterns than the simulated data.
+We also tested out algorithm on a real-world dataset that adheres to different motion profiles and laser scanners type with different scanning patterns than the simulated data.
 
 The dataset is collected by a line scanner, in particular a SICK LMS141 industrial scanner, inside a sphere of a diameter of \SI{20}{\centi\meter} that lies on a floating desk, which is air-pressurized, such that the sphere is hovering.
-On this floating desk the sphere can rotate freely about all axis while being fixed in position. 
+On this floating desk, the sphere can rotate freely about all axis while being fixed in position. 
 Hence, in this experiment the optimization space can be reduced to a rotation.
-Further, without the motion the 2D laser also cannot obtain a 3D model from the environment, thus requiring registration. 
+Further, without the motion, the 2D laser cannot obtain a 3D model from the environment, thus requiring registration. 
 The sphere is equipped with a low-cost IMU, i.e., a PhidgetSpatial Precision 3/3/3, to estimate the orientation.
@@
 Figure~\ref{fig:float-sat-sphere} shows the experimental setup. TThe left column of figure~\ref{fig:cylon-corrected} shows the pre-registered resulting point cloud . 
-We see, that the point cloud is rather noisy and in particular the walls are sensed multiple times, hence appearing very blurry.
+We see that the point cloud is rather noisy, and in particular, the walls are sensed multiple times, hence appearing very blurry.
 
@@
 	\centering
-	\begin{minipage}[c]{0.5\textwidth}
+	\begin{minipage}[c]{0.45\textwidth}
 		\includegraphics[width=0.33\textwidth]{./images/sphere-frame-1.eps}\hfill
@@
 	\end{minipage}\\
-	\begin{minipage}[c]{0.5\textwidth}
+	\begin{minipage}[c]{0.45\textwidth}
 		\includegraphics[width=0.33\textwidth]{./images/sphere-frame-4.eps}\hfill
@@
 	\end{minipage}\\
-	\caption{Data acquisition using floating sphere. We show a sequence of orientations the sphere assumes.}
+	\caption{Data acquisition using floating sphere. We show a sequence of orientations the sphere assumes during data acquisition.}
 	\label{fig:float-sat-sphere}
@@
 	While not a spherical robot, this system shows many similarities with respect to its motion with a spherical robot.
-	The system was manually steered along a squared set of hallways that loop back to the initial starting point inside the old mathematics building at the University of Würzburg (denoted ``The Circle'' in~\cite{Borrmann2020-RADLER}).
+	The system was manually steered along with a squared set of hallways that loop back to the initial starting point inside the old mathematics building at the University of Würzburg (denoted ``The Circle'' in~\cite{Borrmann2020-RADLER}).
 
 	Figure~\ref{fig:radler-mathe} shows the system setup as well as the resulting point cloud.
-	We see that with continuing measurement time the drift increases and the hallways are senses multiple times with a clear offset to each other.
+	We see that with continuing measurement time, the drift increases, and the hallways are senses multiple times with a clear offset to each other.
 
@@
 	This was recorded using a LIVOX-Mid 100 laser scanner mounted inside a transparent shell. 
-	The robot contained three Phidget 3/3/3 Spatial IMUs and was attached to a \SI{50}{\meter} tether cable that was rolled on a spool which in turn was connected to a angular encoder. 
-	From these sensors we obtained orientation information as well as height information. 
+	The robot contained three Phidget 3/3/3 Spatial IMUs and was attached to a \SI{50}{\meter} tether cable that was rolled on a spool which in turn was connected to an angular encoder. 
+	From these sensors, we obtained orientation information as well as height information. 
 	The setup was then brought into a high indoor building (\href{https://www.sfs-w.de/feuerwehrschule/virtueller-rundgang.html}{Firefighter School in Würzburg}) where the system was manually descended from a balcony scanning the building interior.
 	The robot covered a distance of approximately \SI{22}{\meter} in about \SI{402}{\second}.
-	Additionally the building was scanned using a terrestrial laser scanner to provide a reference.
-	The measurements were then rigidly corrected before applying our plane based correction. 
+	Additionally, the building was scanned using a terrestrial laser scanner to provide a reference.
+	The measurements were then rigidly corrected before applying our plane-based correction. 
 	Figure~\ref{fig:experimental-setup} shows the acquired 3D point clouds and the experimental setup.

### 2021-05-13 Latex/sections/conclusion.tex
small corrections conclusion

@@
 In this paper, we propose a registration method for man made environments that exploits the structure of those environments. 
-Given that in man-made environments parallel planes and right angles are abundantly available we employ point-to-plane correspondences to improve a pre-registered 3D point-cloud. 
-We evaluate the procedure on a simulated dataset as well as on an experimentally acquired dataset with different motion profiles. 
-In this evaluation we show that the procedure improves both datasets and yields more human-made, structure like maps. 
-In particular, the qualitative structure of the environments is reconstructed well, that is, in the resulting maps the parallel walls are clearly improved.
+Given that in human-made environments parallel planes and right angles are abundantly available, we employ point-to-plane correspondences to improve a pre-registered 3D point cloud. 
+We evaluate the procedure on a simulated dataset and on an experimentally acquired dataset with different motion profiles. 
+In this evaluation, we show that the procedure improves both datasets and yields more human-made, structure-slike maps. 
+In particular, the qualitative structure of the environments is reconstructed well. 
+In the resulting maps, the parallel walls are clearly improved.
 
 However, as \ref{sec:floatres} shows, some outliers remain due to the simplistic point-to-plane correspondence model, which is to be improved in future studies.
-One idea to do so is to employ a local clustering technique onto the individual scans, before registering them globaly. 
+One idea to do so is to employ a local clustering technique onto the individual scans before registering them globally. 
 Points from the scan with a similar normal get clustered together if their distance is small enough.
-This way, a cluster represents a local plane, while at the same time storing all the points that belong to that plane.
+This way, a cluster represents a local plane while at the same time storing all the points that belong to that plane.
 Point-to-plane correspondences then arise from the intersection of local clusters with global planes.
 
-Furthermore, currently not all steps in the procedure are autonomous, in particular the parameter tuning. 
-In the future one goal is to increase the autonomy of the system.
+Currently, not all steps in the procedure are autonomous, in particular the parameter tuning. 
+In the future, one goal is to increase the autonomy of the system.
 One approach is to introduce soft-locks for the optimization dimensions.

### 2021-05-13 Latex/sections/results.tex
Added todo for fabi; added clustering info to cylon; slightly resized cylon image

@@
 Figure~\ref{fig:cylon-corrected} shows the results obtained before and after employing the plane base registration on the dataset acquired by the floating sphere.
+We combine ten scans into one meta-scan, which is then globally registered.
+This does not only speed up convergence due to the proportional effect on the error function, but also decreases the risk of transforming the scan in a sensless way.
+Transformations like these happen when points that have been labled incorrectly have a huge effect on the error gradient, which is especially the case when considering only a few points.
+
 \begin{figure*}

### 2021-05-13 Latex/sections/results.tex
small corrections conclusion

--- 2021-05-12 Latex/sections/results.tex
@@
 The plane-based registration is applied to the simulated dataset with noisy pose and range measurements (cf. Figure~\ref{fig:simulatedDatasets}) without further processing.
-Assuming this represents a coarsely pre-registered 3D point cloud the distances to the ground truth were evaluated before and after the plane-based registration. 
+Assuming this represents a coarsely pre-registered 3D point cloud, the distances to the ground truth were evaluated before and after the plane-based registration. 
 Figure~\ref{fig:simulatedEvaluation} shows the different point-to-point distances.
@@
   	\end{minipage} 	
- 	\caption{Evaluation of point distances before (left) and after (right) the plane based registration on a simulated dataset.
-A maximal distance of \SI{2}{m} is set, such that all points that display a higher distance value are excluded from analysis.
-Further the histogram joins all values with a distance greater than \SI{1}{m} into the last bucket.
-The top two column shows a heat map of distances while the bottom shows the corresponding histogram.
+ 	\caption{Evaluation of point distances before (left) and after (right) the plane-based registration on a simulated dataset.
+A maximal distance of \SI{2}{m} is set, such that all points that display a higher distance value are excluded from the analysis.
+Further, the histogram joins all values with a distance greater than \SI{1}{m} into the last bucket.
+The top two columns show a heat map of distances, while the bottom shows the corresponding histogram.
 The color mapping is equivalent in both.
@@
 
-Before the registration the corridor is only represented acceptably in the front part. 
-The further into the corridor, i.e. the longer the robot accumulates error, the more imprecise the date becomes. 
-Finally we see that a large portion of the data even exceeds either the \SI{1}{\meter} error, contributing to the large spike in the histogram or even the \SI{2}{\meter} error limit thus, being cut from the representation.
+Before the registration, the corridor is only represented acceptably in the front part. 
+The further into the corridor, i.e., the longer the robot accumulates error, the more imprecise the date becomes. 
+Finally, we see that a large portion of the data even exceeds either the \SI{1}{\meter} error, contributing to the large spike in the histogram or even the \SI{2}{\meter} error limit; thus, being cut from the representation.
 After registration, we see that, qualitatively, the ideal corridor was nearly restored from the noisy data. 
 In particular a very large portion (95\%) have less than \SI{17}{\centi\meter} distances.
-Further the square and straight shape of the corridor is restored well and especially the large spike at errors of greater than \SI{1}{m} is remove. 
-Any such errors tend to occur at the back and the front of the corridor where the measured range is the largest, hence has the largest contribution of the range error.
+Further, the square and straight shape of the corridor is restored well, and especially the large spike at errors of greater than \SI{1}{m} is removed. 
+Any such errors tend to occur at the back and the front of the corridor where the measured range is the largest hence has the largest contribution of the range error.
 
@@
 Figure~\ref{fig:cylon-corrected} shows the results obtained before and after employing the plane base registration on the dataset acquired by the floating sphere.
+We combine ten scans into one meta-scan, which is then globally registered.
+This does not only speed up convergence due to the proportional effect on the error function but also decreases the risk of transforming the scan in a senseless way.
+Transformations like these happen when incorrectly labeled points have a huge effect on the error gradient, which is especially the case when considering only a few points.
+
 \begin{figure*}
@@
 \end{figure*}
-After the registration the walls of the room are significantly more prominent in the point-cloud. 
-In particular the noise in the top left corner of the top view (cf. figure~\ref{fig:cylon-corrected}) has been visibly reduced.
+After the registration, the walls of the room are significantly more prominent in the point cloud. 
+In particular, the noise in the top left corner of the top view (cf. figure~\ref{fig:cylon-corrected}) has been visibly reduced.
 Further, the deviation of points around the walls is notably smaller as the points are moved on the plane.
@@
 In particular, at the initially very noisy top left corner of the top view, a few points cannot be matched to the plane. 
-The scans that cause these imprecisions are problematic as they measure a number of plane like structures in the interior of the room. 
-When relabeling those scans the interior points are associated to a plane and hence the corresponding distance minimized.
+The scans that cause these imprecisions are problematic as they measure a number of plane-like structures in the interior of the room. 
+When relabeling those scans, the interior points are associated with a plane, and hence the corresponding distance is minimized.
 

### 2021-05-13 Latex/sections/conclusion.tex
Fixed wrong section reference

@@
 
-However, as \ref{sec:floatres} shows, some outliers remain due to the simplistic point-to-plane correspondence model, which is to be improved in future studies.
+However, as section \ref{sec:experimentalSetup} shows, some outliers remain due to the simplistic point-to-plane correspondence model, which is to be improved in future studies.
 One idea to do so is to employ a local clustering technique onto the individual scans before registering them globally. 

### 2021-05-13 Latex/sections/realData.tex
Added Hardware setup of the cylon sphere

@@
 	\begin{minipage}[c]{0.45\textwidth}
-		\includegraphics[width=0.33\textwidth]{./images/sphere-frame-1.eps}\hfill
-		\includegraphics[width=0.33\textwidth]{./images/sphere-frame-2.eps}\hfill
-		\includegraphics[width=0.33\textwidth]{./images/sphere-frame-3.eps}\hfill
+		\centering
+		\includegraphics[width=0.33\textwidth, height=0.33\textwidth]{./images/sphere-frame-1.eps}\hfill
+		\includegraphics[width=0.33\textwidth, height=0.33\textwidth]{./images/sphere-frame-2.eps}\hfill
+		\includegraphics[width=0.33\textwidth, height=0.33\textwidth]{./images/sphere-frame-3.eps}\\
+		\includegraphics[width=0.33\textwidth, height=0.33\textwidth]{./images/sphere-frame-4.eps}\hfill
+		\includegraphics[width=0.33\textwidth, height=0.33\textwidth]{./images/sphere-frame-5.eps}\hfill
+		\includegraphics[width=0.33\textwidth, height=0.33\textwidth]{./images/sphere-frame-6.eps}\hfill
 	\end{minipage}\\
+	\vspace{0.5cm}
 	\begin{minipage}[c]{0.45\textwidth}
-		\includegraphics[width=0.33\textwidth]{./images/sphere-frame-4.eps}\hfill
-		\includegraphics[width=0.33\textwidth]{./images/sphere-frame-5.eps}\hfill
-		\includegraphics[width=0.33\textwidth]{./images/sphere-frame-6.eps}\hfill
-	\end{minipage}\\
-	\caption{Data acquisition using floating sphere. We show a sequence of orientations the sphere assumes during data acquisition.}
+		\centering
+		\includegraphics[width=0.49\textwidth, trim={20cm 0 10cm 0}, clip]{./images/cylon-exploded}\hfill
+		\includegraphics[width=0.49\textwidth, trim={20cm 0 10cm 0}, clip]{./images/cylon-setup}\hfill
+	\end{minipage}
+	\caption{Data acquisition using floating sphere. Top: A sequence of orientations the sphere assumes during data acquisition. Bottom: Hardware setup of the sphere.} 
 	\label{fig:float-sat-sphere}

### 2021-05-13 Latex/sections/simulation.tex
Added Info about the Livox Scanners

@@
 For the simulated data, we implemented a noisy world-robot-sensor model.  
-We are simulating a LIVOX-Mid100~\cite{LivoxMid40-100} laser scanner with customizable noise level on the range measurements inside a robot with different motion capabilities, subject to noise in its pose estimation.
+We are simulating a Livox-Mid100~\cite{LivoxMid40-100} laser scanner with customizable noise level on the range measurements inside a robot with different motion capabilities, subject to noise in its pose estimation.
 This yields scan results similar to the ones obtained in the real world.
+
+\subsection{Simulated Sensor}
+
+The Livox-Mid100 laser scanner consists of three smaller-sized Mid40 scanners arranged to provide an overall field-of-view of \SI{98.4}{\degree} (Horizontal) by \SI{38.4}{\degree} (Vertical).
+It provides a range precision of \SI{2}{\centi\meter} up to \SI{20}{\meter} range and yields \SI[per-mode=symbol]{300,000}{\pts\per\second}.
+A particular advantage of the Livox laser scanners is their unique scanning pattern.
+The scanners use two motorized, so-called Risley prisms to steer the beam deflectance.
+For the specific ratio of rotation speeds used in the Livox scanners, the beam traces out a non-repetitive, flower-shaped scanning pattern~\cite{thorlabs}.
+Since the pattern is non-repetitive, the point density increases with integration time, enabling denser measurements.
 

### 2021-05-13 Latex/sections/conclusion.tex
Added spherical robot renders

--- 2021-05-13 Latex/sections/conclusion.tex
@@
 
-However, as \ref{sec:floatres} shows, some outliers remain due to the simplistic point-to-plane correspondence model, which is to be improved in future studies.
+However, as section~\ref{sec:experimentalSetup} shows, some outliers remain due to the simplistic point-to-plane correspondence model, which is to be improved in future studies.
 One idea to do so is to employ a local clustering technique onto the individual scans before registering them globally. 

### 2021-05-13 Latex/sections/simulation.tex
Added spherical robot renders

@@
 
-Figure~\ref{fig:simulatedDatasets} shows a noise-free trajectory through a simulated environment and a trajectory with noisy sensors.
+Figure~\ref{fig:simulatedDatasets} shows a noise-free trajectory through a simulated environment and a trajectory with noisy sensors while figure~\ref{fig:robotRender} shows a rendering of the simulated robot as well as the simulation sequence.
 In the noise-free case, the trajectory slightly bends in the direction of disturbance torque sideways while also varying in velocity in the intended direction of travel.
@@
 	\includegraphics[width=0.51\textwidth]{images/perfect_top_view}\hfill
-	\includegraphics[width=0.48\textwidth]{images/noisy_pose_and_range_top_view}\\\vspace{0.5cm}
-	\includegraphics[width=0.25\textwidth]{images/sim_frame_00.png}\hfill
-	\includegraphics[width=0.25\textwidth]{images/sim_frame_02.png}\hfill
-	\includegraphics[width=0.25\textwidth]{images/sim_frame_06.png}\hfill
-	\includegraphics[width=0.25\textwidth]{images/sim_frame_08.png}
+	\includegraphics[width=0.48\textwidth]{images/noisy_pose_and_range_top_view}
 	\caption{Two simulated datasets of a long hallway of size $\SI{4}{\meter}\times\SI{3}{\meter}\times\SI{100}{\meter}$.
-Ideal noise free (left), noisy pose ($\mathcal{N}_\varphi = \mathcal{N}_\psi(\mu = 0.0001, \sigma= 0.00001)$) and noisy range measurements ($\mathcal{N}_r(\mu = 0, \sigma = 0.001)$) (right).
-Below: Initial sequence of the simulation of the ideal dataset showing the simulated motion of the robot.}
+Ideal noise free (left), noisy pose ($\mathcal{N}_\varphi = \mathcal{N}_\psi(\mu = 0.0001, \sigma= 0.00001)$) and noisy range measurements ($\mathcal{N}_r(\mu = 0, \sigma = 0.001)$) (right).}
 	\label{fig:simulatedDatasets}
 \end{figure*}
+
+\begin{figure*}
+	\centering
+	\includegraphics[width=0.5\textwidth]{images/sim_frame_00.png}\hfill
+	\includegraphics[width=0.5\textwidth]{images/sim_frame_02.png}\\
+	\includegraphics[width=0.5\textwidth]{images/sim_frame_06.png}\hfill
+	\includegraphics[width=0.5\textwidth]{images/sim_frame_08.png}\\\vspace{0.5cm}
+	\includegraphics[width=0.5\textwidth]{images/ProtoRoll1.png}\hfill
+	\includegraphics[width=0.5\textwidth]{images/ProtoRoll2.png}
+	\caption{Top: Initial sequence of the simulation of the ideal dataset showing the simulated motion of the robot.
+	Below: Rendering of the simulated robot, a Livox laser scanner insider a spherical shell.}
+	\label{fig:robotRender}
+\end{figure*}

### 2021-05-13 Latex/sections/acknowledgement.tex
removed todo from acknowledgemennts

@@
 \section*{Acknowledgement}
-\todo[inline]{check this}
+
 The authors thank the Cylon team (Ignacio Dorado Llamas, Timothy Randolph, Camilo Andres Reyes Mantilla, Tobias Wahl) for designing and building the prototype and Timo Burger for acquiring the floating table data set.

### 2021-05-13 Latex/main.tex
Added Info about the Livox Scanners
Implemented most comments from andreas

@@
 \newcommand{\pardiff}[1]{\frac{\partial}{\partial #1}}
+
+\DeclareSIUnit\pts{points}
+
+\renewcommand\floatpagefraction{1}
+\renewcommand\topfraction{1}
+\renewcommand\bottomfraction{1}
+\renewcommand\textfraction{0}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \title{Towards Spherical Robots For Mobile Mapping In Human Made Environments\tnoteref{mytitlenote}}
-\tnotetext[mytitlenote]{To abide the \href{https://www.go-fair.org/fair-principles/}{FAIR-Principles} of science, all data and code used to produce the results are given in this \href{https://github.com/fallow24/SphereTDP}{github} repository.}
+\tnotetext[mytitlenote]{To abide the \href{https://www.go-fair.org/fair-principles/}{FAIR-Principles} of science, all data and code used to produce the results are given in this our repository (\url{https://github.com/fallow24/SphereTDP}).}
 

### 2021-05-13 Latex/sections/results.tex

@@
  	\centering
- 	\begin{minipage}[c]{0.49\textwidth}
+ 	\begin{minipage}[c]{0.495\textwidth}
  		\centering
@@
   	\end{minipage}\hfill
-  	\begin{minipage}[c]{0.49\textwidth}
+  	\begin{minipage}[c]{0.495\textwidth}
   		\centering
@@
 The color mapping is equivalent in both.
-An animation of the matching process is given at~\href{https://youtu.be/0ps5Pg4qo4E}{here}}
+An animation of the matching process is given at \url{https://youtu.be/0ps5Pg4qo4E} .}
  	\label{fig:simulatedEvaluation}
@@
 	\centering
-	\includegraphics[width=0.49\textwidth]{./images/cylon_uncorr_corner}\hfill
-	\includegraphics[width=0.49\textwidth]{./images/cylon_corr_corner}\\
-	\includegraphics[width=0.49\textwidth]{./images/cylon_uncorr_top}\hfill
-	\includegraphics[width=0.49\textwidth]{./images/cylon_corr_top}
+	\includegraphics[width=0.495\textwidth]{./images/cylon_uncorr_corner}\hfill
+	\includegraphics[width=0.495\textwidth]{./images/cylon_corr_corner}\\
+	\includegraphics[width=0.495\textwidth]{./images/cylon_uncorr_top}\hfill
+	\includegraphics[width=0.495\textwidth]{./images/cylon_corr_top}
 	\caption{The point cloud acquired by the floating sphere before (left) and after (right) applying the plane based registration.
 View from the interior that shows the used planes (top) and a birds-eye view (bottom).
-An animation of the registration process is given at \href{https://youtu.be/8XdIUN_9VpY}{here}.}
+An animation of the registration process is given at \url{https://youtu.be/8XdIUN_9VpY} .}
 	\label{fig:cylon-corrected}

### 2021-05-13 Latex/sections/simulation.tex

@@
 	\includegraphics[width=0.51\textwidth]{images/perfect_bottom}\hfill
-	\includegraphics[width=0.48\textwidth]{images/noisy_pose_and_range_bottom}\\
+	\includegraphics[width=0.485\textwidth]{images/noisy_pose_and_range_bottom}\\
 	\includegraphics[width=0.51\textwidth]{images/perfect_top}\hfill
-	\includegraphics[width=0.48\textwidth]{images/noisy_pose_and_range_top}\\
+	\includegraphics[width=0.485\textwidth]{images/noisy_pose_and_range_top}\\
 	\includegraphics[width=0.51\textwidth]{images/perfect_top_view}\hfill
-	\includegraphics[width=0.48\textwidth]{images/noisy_pose_and_range_top_view}
+	\includegraphics[width=0.485\textwidth]{images/noisy_pose_and_range_top_view}
 	\caption{Two simulated datasets of a long hallway of size $\SI{4}{\meter}\times\SI{3}{\meter}\times\SI{100}{\meter}$.
@@
 	\centering
-	\includegraphics[width=0.5\textwidth]{images/sim_frame_00.png}\hfill
-	\includegraphics[width=0.5\textwidth]{images/sim_frame_02.png}\\
-	\includegraphics[width=0.5\textwidth]{images/sim_frame_06.png}\hfill
-	\includegraphics[width=0.5\textwidth]{images/sim_frame_08.png}\\\vspace{0.5cm}
-	\includegraphics[width=0.5\textwidth]{images/ProtoRoll1.png}\hfill
-	\includegraphics[width=0.5\textwidth]{images/ProtoRoll2.png}
+	\includegraphics[width=0.495\textwidth]{images/sim_frame_00.png}\hfill
+	\includegraphics[width=0.495\textwidth]{images/sim_frame_02.png}\\
+	\includegraphics[width=0.495\textwidth]{images/sim_frame_06.png}\hfill
+	\includegraphics[width=0.495\textwidth]{images/sim_frame_08.png}\\
+	\includegraphics[width=0.495\textwidth]{images/ProtoRoll1.png}\hfill
+	\includegraphics[width=0.495\textwidth]{images/ProtoRoll2.png}
 	\caption{Top: Initial sequence of the simulation of the ideal dataset showing the simulated motion of the robot.

### 2021-05-13 Latex/sections/introduction.tex
Introduction and state of the art

@@
 Smaller robotic systems are particularly used when cars no longer have access.
-Examples for this are human operated systems such as Zebedee~\cite{Bosse2012-zebedee}, a Hokuyo scanner on a spring, VILMA~\cite{JPRS2016}, a rolling FARO scanner, RADLER~\cite{Borrmann2020-RADLER}, a 2D laser scanner mounted to a unicycle, or a backpack mounted ``personal laser scanning system'' as in~\cite{LauterbackEtAl2015-Backpack}.
+Examples for this are human operated systems such as Zebedee~\cite{Bosse2012-zebedee}, a small Hokuyo 2D scanner on a spring, VILMA~\cite{JPRS2016}, a rolling FARO scanner operating in profiler mode, RADLER~\cite{Borrmann2020-RADLER}, a SICK 2D laser scanner mounted to a unicycle, or a backpack mounted ``personal laser scanning system'' as in~\cite{LauterbackEtAl2015-Backpack} or~\cite{WWWLeicaBackpack}.
 Recently more and more autonomous systems gained maturity.
@@
 Also, the mobile mapping approaches implemented on the ANYmal platform such as~\cite{Fankhauser2018-ANYmal} were very successful.
-Of all these formats, one hasn't been explored thoroughly in the scientific community: The spherical mobile mapping robot.
+Of all these formats, one has not been explored thoroughly in the scientific community: The spherical mobile mapping robot.
 Yet this provides some very promising advantages over the other formats.
@@
 This is particularly useful for unknown and dangerous environments.
-E.g., for space applications, this is also proposed in the DAEDALUS study~\cite{RossiMaurelliUnnithanetal.2021}, a spherical robot that is to be lowered into a lunar cave and create a 3D map of the environment.
+E.g., for space applications, in the DAEDALUS study~\cite{RossiMaurelliUnnithanetal.2021}, a spherical robot is proposed that is to be lowered into a lunar cave and create a 3D map of the environment.
 The authors choose this format as the present moon regolith is known to damage instruments and other components.

### 2021-05-13 Latex/sections/realData.tex
Redid images

@@
 	\end{minipage}\\
-	\vspace{0.5cm}
+	\vspace{0.1cm}
 	\begin{minipage}[c]{0.45\textwidth}
 		\centering
-		\includegraphics[width=0.49\textwidth, trim={20cm 0 10cm 0}, clip]{./images/cylon-exploded}\hfill
-		\includegraphics[width=0.49\textwidth, trim={20cm 0 10cm 0}, clip]{./images/cylon-setup}\hfill
+		\includegraphics[width=0.4615\textwidth, trim={20cm 0 10cm 0}, clip]{./images/cylon-exploded}\hfill
+		\includegraphics[width=0.525\textwidth, trim={25cm 7.5cm 15cm 5cm}, clip]{./images/cylon-setup}\hfill
 	\end{minipage}

### 2021-05-13 Latex/sections/simulation.tex
Redid images

@@
 	\centering
-	\includegraphics[width=0.51\textwidth]{images/perfect_bottom}\hfill
-	\includegraphics[width=0.485\textwidth]{images/noisy_pose_and_range_bottom}\\
-	\includegraphics[width=0.51\textwidth]{images/perfect_top}\hfill
-	\includegraphics[width=0.485\textwidth]{images/noisy_pose_and_range_top}\\
-	\includegraphics[width=0.51\textwidth]{images/perfect_top_view}\hfill
-	\includegraphics[width=0.485\textwidth]{images/noisy_pose_and_range_top_view}
+	\includegraphics[width=0.495\textwidth]{images/perfect_bottom}\hfill
+	\includegraphics[width=0.495\textwidth]{images/noisy_pose_and_range_bottom}\\
+	\includegraphics[width=0.495\textwidth]{images/perfect_top}\hfill
+	\includegraphics[width=0.495\textwidth]{images/noisy_pose_and_range_top}\\
+	\includegraphics[width=0.495\textwidth]{images/perfect_top_view}\hfill
+	\includegraphics[width=0.495\textwidth, height=0.1825\textwidth]{images/noisy_pose_and_range_top_view}
 	\caption{Two simulated datasets of a long hallway of size $\SI{4}{\meter}\times\SI{3}{\meter}\times\SI{100}{\meter}$.

### 2021-05-13 Latex/sections/acknowledgement.tex

@@
 
-The authors thank the Cylon team (Ignacio Dorado Llamas, Timothy Randolph, Camilo Andres Reyes Mantilla, Tobias Wahl) for designing and building the prototype and Timo Burger for acquiring the floating table data set.
+The authors thank the Cylon team (Ignacio Dorado Llamas, Timothy Randolph, Camilo Andres Reyes Mantilla) for designing and building the prototype and Timo Burger for acquiring the floating table data set.
 Special thanks to Dieter Ziegler and Sergio Montenegro for supporting our work.

### 2021-05-13 Latex/sections/conclusion.tex

@@
 
-In this paper, we propose a registration method for man made environments that exploits the structure of those environments. 
+In this paper, we have proposed a registration method for man made environments that exploits the structure of those environments. 
 Given that in human-made environments parallel planes and right angles are abundantly available, we employ point-to-plane correspondences to improve a pre-registered 3D point cloud. 
-We evaluate the procedure on a simulated dataset and on an experimentally acquired dataset with different motion profiles. 
-In this evaluation, we show that the procedure improves both datasets and yields more human-made, structure-slike maps. 
+We have evaluated the procedure on a simulated dataset and on an experimentally acquired dataset with different motion profiles. 
+In this evaluation, we have shown that the procedure improves both datasets and yields more human-made, structure-slike maps. 
 In particular, the qualitative structure of the environments is reconstructed well. 
@@
 
-Currently, not all steps in the procedure are autonomous, in particular the parameter tuning. 
+Needless to say, a lot of work remains to be done. 
+%
+Right now, not all steps in the procedure are autonomous, in particular the parameter tuning. 
 In the future, one goal is to increase the autonomy of the system.
@@
 I.e., instead of locking some dimensions entirely from being used for optimization, they are weighted based on the dynamics of the system that encode which noise source is more likely.  
+%
+Curently, we are assembling the sphere of Figure~\ref{FIXME} to test it in real environments.
+Unfortunately, this sphere will be too large to fit our floating setup, thus it will be tested in an office like environment.
+Furthermore, we will test our rolling and scanning spheres on a rail system to achieve repeatable trajectories.

### 2021-05-13 Latex/sections/stateOfTheArt.tex

@@
 
-Up to the best of our knowledge, SLAM with spherical robots and laser scanners was not done before, while there exists camera-based approaches~\cite{9233654}.
+While there exist camera-based approaches~\cite{9233654}, to the best of our knowledge, SLAM with spherical robots and laser scanners was not done before.
 However, laser-based SLAM algorithms for motions in six degrees of freedom (DoF) have been thoroughly studied.
@@
 Adding a heuristic for closed-loop detection and a global relaxation Borrmann et al. yield highly precise maps of the scanned environment~\cite{RAS2007}.
-Zhang et al. propose a real-time version of the SLAM algorithm in~\cite{Zhang2014}.
+Zhang et al. propose a real-time solution to the SLAM problem in~\cite{Zhang2014}.
 They achieve the performance at a lower computational load by dividing the SLAM algorithm into two different algorithms: one performing odometry at a high frequency but low fidelity and another running at a lower frequency performing fine matching and registration of the point clouds.
@@
 In human-made environments, planes are abundantly available and hence provide an attractive base for correspondences.
-Förster et al. use this property of human-made environment successfully in~\cite{Foerstner2017}.
+
+Pathak et al.~\cite{Pathak:2010} propose to reduce the complexity of the registration by using correspondences between planar patches instead of points.
+They demonstrate the effectiveness of their approach even with noisy data in cluttered environments.
+However, their approach is designed for data acquired in stop-scan-go fashion and not for mobile mapping applications.
+As they use a region growing procedure with randomized initialization for detecting planar patches the distortions in the data introduced by pose uncertainties are likely to effect the shape of the planar patches and lead to faulty correspondences.
+
+Förster et al. use the planarity of human-made environment successfully in~\cite{Foerstner2017}.
 They register point clouds using plane-to-plane correspondences and include uncertainty measures for the detected planes and the estimated motion.

### 2021-05-13 Latex/sections/realData.tex
Changed Cylon Setup image
Added minor infos to cylon hardware

@@
 Further, without the motion, the 2D laser cannot obtain a 3D model from the environment, thus requiring registration. 
-The sphere is equipped with a low-cost IMU, i.e., a PhidgetSpatial Precision 3/3/3, to estimate the orientation.
+The sphere is equipped with a low-cost IMU, i.e., a PhidgetSpatial Precision 3/3/3, to estimate the orientation, a battery pack and a raspberry pie for processing.
 
-Figure~\ref{fig:float-sat-sphere} shows the experimental setup. TThe left column of figure~\ref{fig:cylon-corrected} shows the pre-registered resulting point cloud . 
+Figure~\ref{fig:float-sat-sphere} shows the experimental setup. The left column of figure~\ref{fig:cylon-corrected} shows the pre-registered resulting point cloud . 
 We see that the point cloud is rather noisy, and in particular, the walls are sensed multiple times, hence appearing very blurry.
@@
 		\includegraphics[width=0.4615\textwidth, trim={20cm 0 10cm 0}, clip]{./images/cylon-exploded}\hfill
-		\includegraphics[width=0.525\textwidth, trim={25cm 7.5cm 15cm 5cm}, clip]{./images/cylon-setup}\hfill
+		\includegraphics[width=0.525\textwidth]{./images/cylon_setup}\hfill
 	\end{minipage}

### 2021-05-13 Latex/sections/registration.tex
changed to doi
TODO

@@
 
-We start with transforming each line scan into the project coordinate system, which is defined by the pose of the first acquired line scan.
-With line scan, we mean the smallest chunk of range data we obtain from the scanner device driver.
+Consider a line scan as the smallest chunk of range data we obtain from the scanner device driver.
 In the case of a SICK LMS1xx it is a line, and in case of a Livox scanner it has a flower shape.
 More details are provided in the following sections.
+We start with transforming each line scan into the project coordinate system, which is defined by the pose of the first acquired line scan.
 %
@@
 
+\subsection{Rigid Registration}
+
+\todo[inline]{Die Rigid Registration sollte kurz erwähnt werden. Welcher Algorithmus? Werden dafür Linescans zusammengefasst?}
+
 \subsection{Plane Extraction}
@@
 After having done the rigid registration, the scans are aligned well enough to make statements about the potential planes in the environment.
-To find the planes in the environment, a Hough transform with an accumulator ball as described in~\cite{Borrmann2011The3H} is used. 
+To find the planes in the environment, a Randomized Hough transform with an accumulator ball as described in~\cite{3DRESEARCH2011} is used. 
 In the next subsection, the Hesse-normal form of the plane is required.
@@
 
-Assuming we now know the Hesse-normal form of all planes and all assigned points to these planes, we register the points to have a minimal distance to their respective planes.
+Assuming we know now the Hesse-normal form of all planes and all assigned points to these planes, we register the points to have a minimal distance to their respective planes.
 The transformation $T(\vec{p}_k)$ of each point $\vec{p}_k$ with respect to a 6 DoF motion is described in homogeneous coordinates using the roll-pitch-yaw ($\varphi-\vartheta-\psi$) Tait-Brian angles as in~\cite{diebel2006representing}.

### 2021-05-13 mfi21/ieeeconf/root.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper

%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{enumitem}
\newcommand{\M}[1]{\mathbf{#1}} % Bold for matrix
\newcommand{\V}[1]{\mathbf{#1}} % Bold for vector
%In case you encounter the following error:
%Error 1010 The PDF file may be corrupt (unable to open PDF file) OR
%Error 1000 An error occurred while parsing a contents stream. Unable to analyze the PDF file.
%This is a known problem with pdfLaTeX conversion filter. The file cannot be opened with acrobat reader
%Please use one of the alternatives below to circumvent this error by uncommenting one or the other
%\pdfobjcompresslevel=0
%\pdfminorversion=4

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

\title{\LARGE \bf
IMU-based pose-estimation for spherical robots with limited resources
}

\author{Jasper Zevering*, Anton Bredenbeck, Fabian Arzberger, Dorit Borrmann, Andreas Nuechter$^{1}$% <-this % stops a space
\thanks{*
        Corresponding author: }%
        \thanks{ \quad  \tt\small jasper.zevering@stud-mail.uni-wuerzburg.de}%
\thanks{$^{1}$ All authors are with Informatics VII : Robotics and Telematics - University of Wuerzburg}%
\thanks{We acknowledge funding from the ESA Contract No. 4000130925/20/NL/GLC for the “DAEDALUS – Descent And Exploration in Deep Autonomy of Lava Underground Structures” Open Space Innovation Platform (OSIP) lunar caves-system study and the Elite Network Bavaria (ENB) for providing funds for the academic program “Satellite Technology”.}
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

Spherical robots are a robot format that is not yet thoroughly studied for the application of mobile mapping.
However, in contrast to other forms they provide some unique advantages.
For one, the spherical shell provides a protection against harsh environments, e.g. guarding the sensors and actuators from dust and solid rock.
This is particularly useful in space applications. 
Furthermore, the inherent rotation the robot uses for locomotion can be exploited to measure in all directions without having the sensor itself actuated.
To use this rotation in combination with sensor data to create a consistent environment map a reasonable estimation of the robot pose is required. 
In such cases the pose estimation can be done by inertial measurements from sensors such as accelerometers and gyroscopes, as interpolating instance for calculation-intensive slow algorithms as optical localization algorithms or as rough estimation.
We propose a pose estimation procedure based on inertial measurements that exploits the known dynamics of a spherical robot. 
It emphasizes a low jitter to maintain constant world measurements during standstill and avoids exponentially growing error in position estimates. 
The evaluation of  the position and orientation estimates with given ground truth frames shows that we can reduce the jitter in orientation and can handle slip and partly slide behavior as other commonly used filters such as the Madgwick filter.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}

Spherical robots are a relatively narrow field of robotics. Still, they could be useful in situations where the measurement equipment needs to be protected against the harsh environments.
As an example, the 2021 CDF study LunarCaves by ESA about the feasibility of a spherical robot, called DAEDALUS, for exploration of lunar-lava tubes \cite{c2} shows the need for orientation and position estimation of spherical robots with the limitations on resources common for space-applications.
Although a purely Inertial Measurement Unit (IMU)-based estimation is not recommended for various reasons, the lack of an absolute reference among others, IMU-based pose estimation can be used as part of an overall multi-sensor-fusion-based estimation.
IMUs are sensors with higher refresh rates than other Simultaneous Localization and Mapping (SLAM) based algorithms, such as optical or Light Detection and Ranging (LIDAR) SLAM, making them suitable for interpolating the more precise and absolute position reference points of these slower algorithms.
The limited computational power of space-qualified CPUs further imposes the need for good orientation estimation without complex algorithms like Kalman filters.
Lastly, the geometry of a spherical robot opens the possibility for the not often used position-estimation by IMUs. \\
This paper introduces an IMU-based orientation and position algorithm for spherical robots with limited computational power and the need for real-time estimations.
Also the algorithm  does not rely on one specific locomotion approach, but it assumes a rotation of the IMU together with at least the shell of the robot if not the whole robot.

\section{Assumptions and Prelimitations}
\label{AssumptionsAndPrelimitations}
The presented approach requires multiple assumptions, to be beneficial-Therefore, we classify the calculated pose concerning its usage and define the usage as follows:

\paragraph*{Real Time Operations}The operation of the robot does not allow post-calculation of the pose. This implies the usage of the onboard computational power.
\paragraph*{Limitation of Computational Power}
The need for efficient code without massive bottleneck operations, as the inversion of a  matrix with the extended Kalman Filter. The code must minimize the usage of non-basic operations.
\paragraph*{Low Costs of Hardware}
We assume biased and noisy data in a non-negligible amount from the sensors.
\paragraph*{Sphere Shape}
The code relies heavily on the assumption of a sphere as a shape with a constant robot radius. We also assume that rotation of the sphere on the ground leads to translation.
\paragraph*{Slip and Sliding}
We assume the sphere has both slip (the sphere rotates, but there is no translation) and sliding (the sphere has translation without rotation) in a limited, non-permanent way. The limitation is the assumption that sliding and slipping reduce the amount of translation for a given rotation and vice versa, however, not for complete absence.
\paragraph*{Stability of Pose}
For simultaneous localization and mapping (SLAM) purposes using LIDARs, a stable but maybe minimal false pose is preferred over a more exact but noisy and jumping position.
\paragraph*{Further Processing of Pose}
The approach should avoid jumping values or abrupt changes of data if not clearly indicated by the sensors. Suppose there is an internal algorithmic change of behavior due to the change of state from standing to rolling. This change shall not be represented by an unnatural acceleration in the data.
\paragraph*{Uncertainty of Iterative Position Integration}
IMUs deliever no absolute reference for a position as they do for orientation in the form of the noisy but measurable gravity vector. So without external or secondary sensors relying on absolute waypoints, the position estimation will always underlie the integration of errors.
\paragraph*{Space Suitability}
As the code is considered to be applied to space missions, the magnetometer data is not considered in the algorithm, despite its positive impact on orientation estimation on earth.
\paragraph*{Short Term Mobility}
As the code still relies on integration parts, the sphere is not considered to have a straight motion for a long period but rather makes multiple short paths. This will also be necessary, as in further extension, the fusion of a pose estimation by the laser scanner will be merged. 
\paragraph*{Non-reliablity of locomotion}
The sphere from the DAEDALUS example has two completely different locomotion methods; therefore, we will not take locomotion itself into account of pose estimation. Also, the same locomotion-approach can lead to heavily different behavior with the same input to the controller. \cite{c2}
\paragraph*{Classification of Data}
The generated data is a compromise between simple filtration of data and heavy calculation post-record. It is easy to calculate and adapt but due to the number of pre-assumptions, as it makes assumptions about the dynamics of the system, it still delivers reliable data which takes physical behavior into account. It is real-time capable and does not rely on delay steps.
\paragraph*{Field of Use}
The use is restricted to spherical or cylindrical robots. It is optimized for gathering laser data where noise position while standing means a lot of calculation in the map creation, and therefore, even if slightly false, a stable position is preferred.

\section{State of the Art}
\label{sec:SOA}
For pose estimation on embedded hardware, there are three main algorithms.
\subsection{Kalman Filter}
The Kalman Filter or the more suitable Extended Kalman Filter (EKF) uses state prediction and merges with sensor data for orientation estimation.
For the defined limitations, the Kalman Filter is not suitable.
First, the EKF requires a considerable amount of computational power, mainly because of its matrix inversion. For this, the needed calculations grow exponentially for input data.
Further, the Kalman Filter relies on input to a system to calculate an estimation of the pose for the next iterative step.
But, as we described in the preliminary section, we can not take the input to locomotion control as a reference to pose estimation as the behavior is not linear or even predictable without exact knowledge of the surface.
Therefore it is not possible to implement a physical behavior of the system to the filter, as easily as it is for other robotic applications.
\subsection{Madgwick Filter}
The Madgwick filter is a widely used filter for orientation estimation with IMUs \cite{c1}.
It is efficient and easy to calculate but does not consider the physics of the motion.
It does nearly provide the sought-after classification for  the orientation.
The problem is the optimization for movement. The Madgwick filter has a jittering behavior at the standstill of the IMU.
This does not meet the desired behavior as described in the prelimitations section (\ref{AssumptionsAndPrelimitations}).
The behavior is shown in comparison to the complementary filter in Figure \ref{madgVsComp1}.
This behavior is most prominent during standstill as the Madgwick filter was designed for movement.
\subsection{Complementary Filter}
A complimentary filter is a fundamental approach of combining precise but not exact with nonprecise but exact data.
Therefore it is widely used for combining noisy accelerometer data with gravity as a reference with the precise but drifting measurement of gyroscopes.
The basic concept is to have the main part of a value dominated by a gyroscope but over time a slow drift towards the measurement of the accelerometer.
E.g. for pitch this would result in 
\begin{equation}
\theta_n = (1-\alpha) \cdot (\theta_{n-1}+ \Delta\theta_{gyro})+ \alpha \cdot \theta_{acc}
\end{equation}
where $\alpha$ is the complementary gain, it determines how much weight the accelerometer data gets and how fast the value drifts towards the measurement of the accelerometer.
A commonly used value for $\alpha$ is 0.02, meaning every iteration, 2 percent of the new values come from the accelerometer.
An $\alpha$ of  1 means the resulting data is completely the one of the accelerometers, and the gyroscope has no influence. \\
The complementary filter is best for orientation calculation without position change as the accelerometer value relies on calculating the gravity value.
With an acceleration in one direction, this leads to a miscalculation of the gravitational vector and, therefore a false orientation estimation.
Simple orientation estimations of quadrocopters often just change the $\alpha$ gain to 0 while moving to avoid the miscalculation of the gravity vector. \cite{complementary}

\begin{figure}[thpb]
\includegraphics[width=\linewidth]{./graphics/MadgwickVsComplementaryNoMovement.eps}
\caption{Quaternion values of Madgwick filter and complementary filter with default gains when IMU is at a stillstand. The Madgwick filter has clearly visible jitter, the complementary filter not }
\label{madgVsComp1}
\end{figure}

\subsection{Position Estimation}
Position estimation purely by IMU is not very recommended as the double integration of the accelerometer measurements leads to exponential error summation.
Therefore there are very few approaches targeting this way of position calculating \cite{c3}.
The very impressive \cite{Valencia2017ASM} also uses IMUs for 6D pose estimation.
They follow a pre-specified trajectory with a manipulator and manage to estimate its pose within 10\% of the ground truth.
Given their application is rather broad they do not exploit known dynamics of the system which could be useful for spherical robots.

\section{Filter-concept}
The pose orientation for the spherical robot is a concept with interconnected components.
Both orientation and position estimates are computed based on the gyroscope and accelerometer measurements taking into account limitations of the system described in Section \ref{AssumptionsAndPrelimitations}.
The generalized overview of orientation and position estimation is shown in Figure \ref{generalized}.

\subsection{Orientation}
The orientation estimation is the combination of the Madgwick and a complementary filter.
The weaknesses and strengths of both described in Section \ref{sec:SOA} can be combined. 
Therefore the general idea is to distinguish between motion and standstill.
In motion, the Madgwick filter is used, and without motion, the complementary filter.
The transition needs to be in a differentiable way, as this is one of our limitations.  
As both filters are applied with weights, this is done by an allocation of a given gain towards one or the other filter depending on the likelihood the sphere is moving or standing still. 
Therefore this requires some adapting mechanism and an estimation of how likely the sphere is to rotate and hence translate.
Both filters share the same meaning of the gain, describing the influence of the accelerometer on the orientation.
The Madgwick filter weight determines how much influence the accelerometer measurement has on the overall orientation, just in a different way than the complementary filter.
Thus the implementation of this approach does not endanger the stability of the estimation due to the change between two completely independent approaches.
The algorithm can shift the gain from one filter to the other in a smooth way without jumping values or abruptly changing behavior.

\subsection{Position}
As there is no absolute reference of position neither with calculation by gyroscope nor by the accelerometer, the filter can only try to improve the calculation of the velocity leading to the change of position. 
The straightforward approach is to double integrate the accelerometer, which leads to the earlier described unwanted behavior of exponential error integration and makes the position estimation completely unusable. 
The second straightforward approach multiplies the rotation speed with the known radius of the sphere and takes this as velocity, integrating once.
The more precise and less noisy data of the gyroscope leads to a usable estimation of position. 
There is no such extensive error integration, and more importantly: no exponentially growing error with time. However, this approach neglects slip and sliding, as it is one of our limitations and vertical movement. 
If the sphere slips, this approach will predict the full translation.
Also the other way around, if there is translation but just slight rotation because of sliding, it will predict too little translation. 
Furthermore, integrating the rotation does not let us estimate vertical movement.
Every rotation will be predicted as translation in a given plane, which is most likely parallel to the ground. 
If for instance the sphere rolls up an obstacle, the length of the path is estimated to the ground-plane.\\
Our algorithm uses both approaches of position estimation and combines them, and limits all components according to the other components. 
This implies a connection between the translation and rotation. 
Overall the extreme situation (full translation, no rotation, or the other way round) will automatically lead to wrong predictions, as there is no way to tell which of the two approaches has at that moment more reliable data and therefore the combination of both rather then only one is chosen.

\begin{figure}[thpb]
\includegraphics[width=\linewidth]{./graphics/imuJasperChartSimliefiedBIG.png}
\caption{Generalized diagram of the estimation}
\label{generalized}
\end{figure}

\section{Filter-Implementation}

\subsection{Orientation}
The orientation is a symbiosis of the Madgwick filter and complementary filter. For no- or slow-moving, the complementary filter is chosen, for fast-moving the Madgwick filter. The Madgwick filter itself is untouched.
The standard complimentary filter is adapted to work in quaternion representation.
 As the integration only works in the Euler angles domain, the jumps of values at the border of the domain like 0 and 360 degrees do not match the mostly continuous change of quaternions.
As the representation is still valid (0 degree and 360 degree roll being the same position), this is not a problem when using at one point any of the filters exclusively.
 But combining both, this leads to contradicting updates from each filter with respect to the other.
 Therefore the normal way of the complementary filter of the new orientation is 
%\begin{equation}
%\begin{bmatrix}
%\Phi_{new} \\
%\theta_{new}\\
%\Psi_{new}
%\end{bmatrix}
%= (1-\alpha)\begin{bmatrix}
%\Phi_{old} \\
%\theta_{old}\\
%\Psi_{old}
%\end{bmatrix}
% + \alpha \begin{bmatrix}
%\Phi_{acc} \\
%\theta_{acc}\\
%\Psi_{acc}
%\end{bmatrix}
%\end{equation}
not used, as it implies a transition of each value independently.
Hence, the RPY-orientation is transformed to a quaternion, and then a quaternion slerp is performed \cite{c4}.
 The quaternion slerp works over the representation of two quaternions on a sphere. The transisiton from one quaternion to the other is now not done directly but always over the surface of the sphere.
 This leads to the same handling of quaternions as the Madgwick-filter without endangering robustness due to using two different quaternions with very different values representing nearly the same position.\\
The transistion between the two filters is done by shifting a fixed factor among the $\alpha$-gain of the complementary filter and the gain factor of the Madgwick filter.
 As the common values widely used for both values differ by one order of magnitude, the gain shifted to the complementary filter is divided by ten and the other way around.
 The values of the gyroscope axes determine in which manner to shift the overall gain to both Filters. 
 The defining values are the threshold until which value the algorithm should use a purely complementary filter and at which threshold a full Madgwick filter should determine the orientation. 
 These values have been determined empirically.
 Future work will investigate on improving those values by developing a suitable heuristic for the thresholds.
 The transition between the full complementary and the full Madgwick estimation, as is required by the prelimitations,  can be done by a function suitable for the specific scenario.
 This avoids large accelerations of the orientation due to the rapid change from one algorithm to another.
 In our experiments, when starting from stillstand, the incipient Madgwick filter, has a more rapid impact on the orientation values then the incipient complementary filter, when stopping rotation. 
Therefore the transition between $\alpha$-gain of the complementary filter and the Madgwick-gain was chosen to have quadratic behavior. 
The value referred to as autogain $\Theta$ indicates how much impact both employed filters have on the orientation estimation. 
Therefore we calculate factors from the gyroscope measurements and scale the autogain with the maximum $\beta$ (cf. equation \ref{eq:factors} and \ref{eq:weight_autogain}).
The $g_{k}$ represent the gyro measurement in the corresponding an axis $k$ in rad/s. Later on, these factors will again be used for position estimation, namely to determine if the sphere rotates or not.
These factors are heuristically defined as:

\begin{equation}
f_k =
\left\{
\begin{align}
0& & g_k \leq 0.1  \\ 
0.25\cdot (g_x&-0.1)^2& \quad 0.1< g_k <2.1 \\
 1& & g_k \geq 2.1
\end{align}
\label{eq:factors}
\right\}
\end{equation}

\begin{equation}
\beta = \max(f_x,f_y,f_z)\label{eq:weight_autogain}
\end{equation}
Then the Madgwick gain $\gamma$ for a given autogain $\Theta$ is calculated by 
\begin{equation}
\gamma = \Theta \cdot \beta
\end{equation}
The complementary filter gain $\alpha$ is calculated by
\begin{equation}
\alpha= \Theta \cdot \theta \cdot(1-\beta)
\end{equation}
So the Madgwick gain has a maximum of $\Theta$ and $\alpha$ a maximum of $\theta\cdot \Theta$. This ratio $\theta$ can be adapted to the specific needs.
We used $\theta = 0.1$ because of the the two often used standard values $\gamma = 0.2$ and $\alpha= 0.02$, which have the ration $0.1$.
Hence for adapting the position estimation to a given robot, the function for calculating the factors needs to be adapted, in such a way that it is a good indication for movement or standstill.
The ratio $\theta$ needs to be chosen in a way, that the desired ratio between Madgwick and complementary gain is reached.
And lastly the autogain needs to be chosen to fit the specific needs in terms of how strong both filters should be able to influence the orientation.
We used $\Theta = 0.2$ to get the described standard values.

\subsection{Position}
The calculated orientation is directly used for position calculation.
With the orientation, a gravity vector is calculated, which, once normalized, represents the  measurement of the accelerometer in abstinence of movement. 
Let $\V g=\begin{bmatrix}0 &0 & -1 \end{bmatrix}^T$ represent the gravity vector in the world frame and the matrix $\M R$ represent the current orientation of the IMU, then

\begin{equation}
\V g\V '=\begin{bmatrix}
g'_x\\
g'_y\\
g'_z
\end{bmatrix}
= R \cdot \begin{bmatrix}
0 \\0  \\ - 1
\end{bmatrix}
\end{equation}
describes the effect of gravity in the coordinate system of the IMU.
This gravity vector is then subtracted from the accelerometer measurement. If the there is no translation, the subtraction should lead to $\begin{bmatrix}
0 & 0&0
\end{bmatrix}^T$ meaning there is no acceleration other than gravity. If there are non-zero components, there is an acceleration in that direction. Given the accelerometer measurement $\V a$ this leads to the 
\begin{equation}
\begin{bmatrix}
v_{x\text{Acc}} \\ v_{y\text{Acc}} \\v_{z\text{Acc}} 
\end{bmatrix}
= 
\int_0^T \left(\begin{bmatrix}
a_x \\ a_y \\ a_z
\end{bmatrix} 
-  \begin{bmatrix}
g'_x\\
g'_y\\
g'_z
\end{bmatrix} dt\right)
\end{equation}
These Values are now coupled to the rotation of both axes other than their own. Meaning there can be a translation in $x$ if there is rotation around $y$ or $z$.
Therefore, the factors from the orientation step, representing the likelihood of rotation are used as limiting factors.
If one of the three factors is 1, this means we let translation in this direction happen.
When there is no rotation, it will hold the velocity to 0.
During fast movement, this would allow exponential error integration. 
This exponential error is taken care of later on.

\begin{equation}
\begin{bmatrix}
v_{x\text{Acc}} \\ v_{y\text{Acc}} \\v_{z\text{Acc}} 
\end{bmatrix}
 = \begin{bmatrix}
v_{x\text{Acc}} \\ v_{y\text{Acc}} \\v_{z\text{Acc}} 
\end{bmatrix}  \cdot 
\begin{bmatrix}
\max(f_y, f_z) \\ \max(f_x, f_z) \\ \max(f_x, f_y) 
\end{bmatrix}
\end{equation}
The rotation into the world frame is computed as:
\begin{equation}
\begin{bmatrix}
v_{x\text{\text{AccWorld}}} \\ v_{y\text{AccWorld}} \\v_{z\text{AccWorld}} 
\end{bmatrix}
= \M R^T \cdot \begin{bmatrix}
v_{x\text{Acc}} \\ v_{y\text{Acc}} \\v_{z\text{Acc}} 
\end{bmatrix}
\end{equation}.
The next step requires the velocity calculated by the rotation.
Therefore the rotations (direct values from the gyro) are rotated into the world frame.
\begin{equation}
\begin{bmatrix}
\omega_{x\text{World}} \\ \omega_{y\text{World}} \\(\omega_{z\text{World}}) 
\end{bmatrix} = 
 \M R^T \cdot
\begin{bmatrix}
g_x \\ g_y\\ g_z
\end{bmatrix} 
\end{equation}
The rotation around the world $z$-axis is not used to calculate the velocity.
Its effect on the position is considered when determining the orientation in the world coordinate system.
$\omega_{z\text{world}}$ needs not to be computed as it has no influence on the position.

The resulting velocities are calculated by multiplying the circumference of the sphere:
\begin{equation}
\begin{bmatrix}
v_{xz\text{GyroWorld}} \\ v_{yz\text{GyroWorld}}
\end{bmatrix} = 
\begin{bmatrix}
\omega_{y\text{World}}/(2 \pi ) \\ \omega_{x\text{World}} / (2 \pi)
\end{bmatrix}
\cdot 2\pi r = \begin{bmatrix}
\omega_{y\text{World}}\\ \omega_{x\text{World}} 
\end{bmatrix}
\cdot r 
\end{equation}.
As the gyro measures rad/s the value is divided by $2\pi$.
Note that the two-dimensional velocities $v_{xz\text{GyroWorld}}$ and $v_{zz\text{GyroWorld}}$ do not correspond to the velocity on the ground plane but may consist of motion in $z$ direction.

 Next the $v_{x\text{AccWorld}} $and $v_{y\text{AccWorld}} $ are limited depending on $v_{xz\text{GyroWorld}} $ and $ v_{yz\text{GyroWorld}}$, in our example this was 120 percent of the gyro calculated velocity.
 This ensures no exponential error integration when in motion.
 This is coupled to the physical act of sliding.
 The higher the likelihood of sliding, the higher the velocity should be compared to the velocity based on rotation.
 This does not count for slipping, as this would imply the adaption of limitation of the velocity by rotation by the velocity measured by the accelerometer.
 But there is no need for a limitation as there is no double integration with the rotation.\\
 The last step for calculating the velocity is taking the velocity in $z$-direction into account for the velocity by rotation.
 This velocity in $z$ is from the noisy double integration, as this is the only indication for change of the $z$-axis.
 To avoid large errors and unlimited double error integration, $v_{z\text{AccWorld}} $ is limited to the mean of $v_{xz\text{GyroWorld}} $ and $ v_{yz\text{GyroWorld}}$.
With the pythagorean theorem the velocity in x and y is calculated accepting the velocity in z direction, which can not be derived by the velocity from gyro:
 \begin{equation}
\begin{bmatrix}
v_{x\text{GyroWorld}} \\ v_{y\text{GyroWorld}}
\end{bmatrix} = 
\begin{bmatrix}
\sqrt{v_{xz\text{GyroWorld}}^2-v_{z\text{AccWorld}}^2}  \\ \sqrt{v_{yz\text{GyroWorld}}^2-v_{z\text{AccWorld}}^2 } 
\end{bmatrix}
\end{equation}.
The last step to get the position is the integration of the velocity:
\begin{equation}
\begin{bmatrix}
p_x \\ p_y \\p_z
\end{bmatrix}
 = \int_0^T \begin{bmatrix}
 (1-\beta)v_{x\text{AccWorld}} + \beta \cdot v_{x\text{GyroWorld}} 
 \\
  (1-\beta)v_{y\text{AccWorld}} + \beta \cdot v_{y\text{GyroWorld}} 
 \\
 v_{z\text{AccWorld}} 
 \end{bmatrix} dt
\end{equation}.
Depending on the practical circumstances and quality of measurement, the factor $\beta$ can be chosen, that the position relies more on the accelerometer calculated velocity or the gyro.
There should always be a mixture of both if there is slipping and sliding.
$\beta =0.5$ would result in just the mean of both approaches.

\subsection{Time Complexity}

Overall an pose estimation step consist only of operations that are constant in time. 
Given that we use the fast inverse square root algorithm \cite{FastInvSqrt} for normalization the trigonometric functions are considered the remaining bottleneck.
Overall there are $4\times\arctan$, $4\times\sin$, $4\times\cos$, $1\times\arcsin$, $1\times\arccos$ and $5$ divisions. 
The IMU measurements are collected at 125 Hz and are processed without problems on a raspberry pi.
Hence no overwhelming computational load is to be expected.

\section{Evaluation}
The proposed algorithm is evaluated by two sets of experiments performed with three Phidgets 3/3/3 IMUs mounted on an round acrylic glass plate.
For position experiments the plate is put inside a acrylic glass sphere where the plate exactly fits in the middle, as the radii are the same.
First, the fusion of Madgwick and complementary filter, here denoted as Autogain, is compared to a Madgwick only estimate. Second, the trajectory computation using both gyroscope and accelerometer is evaluated.

\subsection{Orientation}
A first experiment is performed to show the overall performance of the Autogain computation for the orientation. The IMU was manually rotated in arbitrary directions and the orientation plotted using the Madgwick filter and the Autogain filter.
As expected, the overall behavior of the presented algorithm and the Madgwick filter is very similar in the dynamic case, as shown in Figure \ref{quaternion1}.
\begin{figure}[thpb]
\includegraphics[width=\linewidth]{./graphics/AutoVsMad.eps}
\caption{Quaternion-values of the presented fusion of complementary and Madgwick filter and only the Madgwick filter}
\label{quaternion1}
\end{figure}
Therefore the more interesting part is the change between motion and standstill, as in motion, both filters use the same algorithm but differ during standstill and slow motion.
As described in Section \ref{sec:SOA}, the Madgwick filter has a rather string jitter during slow movements and standstill. 
To evaluate specifically the change between slow motion and standstill, a further experiment was performed. The results in Figure \ref{quaternionq0} show a short excerpt of the experiment with four phases of slow motion.
For external reference, we used a precise optical tracking system. 
Manually transforming the measured IMU values to the optical reference frame then
yields the above plot.
The optical system has the same shape as the presented algorithm and lacks the perturbations of the pure Madgwick filter.
To stress the same behavior without relying on absolute orientation, which will be part of further research, Figure \ref{derivative} shoes the derivative of all three approaches.
Here the change of direction due to small oscillations and the smoother fusion algorithm in comparison to the tracking is even more apparent.

\begin{figure}
\includegraphics[width=\linewidth]{./graphics/opticalOrientation.eps}
\caption{q0-quaternion-value of the presented fusion of complementary and Madgwick filter, only the Madgwick filter and an evaluation via optical tracking }
\label{quaternionq0}
\end{figure}

\begin{figure}
\includegraphics[width=\linewidth]{./graphics/derivative.eps}
\caption{Derivative of q0-quaternion-value of the presented fusion of complementary and Madgwick filter, only the Madgwick filter and an evaluation via optical tracking }
\label{derivative}
\end{figure}

\subsection{Position}
The primary evaluation for position focuses on the slip- and slide-behavior.
Here the plate with the IMU(s) is fixed inside a sphere with a diameter of 29 cm and rolled along an L-shaped track of one meter by one meter.
Slip behavior was manually provoked.
Figure \ref{LTest} shows the resulting trajectories.
Pure rotation integration leads to an enlarged L by about 50 percent, as the manually provoked rotation is just integrated and therefore interpreted as linear motion.
The fusion between the accelerometer and gyroscope interprets an L with dimensions much closer to the original.
The estimation leaves room for improvement, as in one direction, it misses the actual length about -10 percent and in the other direction overestimates it by 10 percent. This should be improvable by parameter tuning.
But it shows the overall capability of the algorithm to compensate for slip.

\begin{figure}
\includegraphics[width=\linewidth]{./graphics/LTest.eps}
\caption{Ground position by integration of rotation and the fusion algorithm. An L-shape with 1 meter by 1 meter was performed with provoked slip(rotation without translation)}
\label{LTest}
\end{figure}

The same experiment was performed with added slide, i.e. more translation than just the rotation would produce.
Here the results shown in Figure \ref{LTestSlide} were not as reliably good as with the slip experiment.
This is due to the timing of the acceleration. In order to recognize the acceleration as translation, there needs to be a little bit of rotation, as the algorithm does not allow any translation without rotation at all.
As the translation computed using the accelerometer is by double integration, the acceleration of the slide movement needs to be at the same time as a minimal amount of rotation.
If the rotation starts after the sliding, the double integration of the acceleration is suppressed.
When the rotation starts and the translation could be double integrated, it is already zero as the velocity is constant.
This can be seen in Figure \ref{LTestSlide} where the one side of the L was interpreted exactly as long as the pure integration, i.e., approx. 50 cm, and therefore the slide was not recognized, and the other part was correctly interpreted longer than just the integration and with 1.02 m as long as the actual trajectory.
This is to be blamed on the execution of the experiment, as it was not perfectly ensured only to start sliding with ongoing rotation, which is hard to achieve when manually guiding a sphere.
But the sliding behavior in a real environment could be in both ways, too, so sliding while rolling and starting sliding before rolling.
With slip, this is not a problem, as, by definition, the translation cannot begin before the rotation when slipping.
If so, it would be sliding.
\begin{figure}
\includegraphics[width=\linewidth]{./graphics/LTestSlide.eps}
\caption{X-Y-Position by integration of rotation and the fusion algorithm. An L-shape with 1 meter by 1 meter was performed with provoked slide(translateion without rotation)}
\label{LTestSlide}
\end{figure}
The experiments for sliding did not reliably detect the slide event and, on some occasions, interpreted the distance shorter than just the integration, as the acceleration of the linear motion was missed.
However, the deceleration of the sphere was still recognized while rotation existed and therefore, the sphere miscalculated a couple of centimeters.

\subsection{Conclusion}
This paper introduced an IMU-based pose estimation filter based on the fusion of multiple already existing filters and approaches.
It is optimized for specific needs and circumstances of spherical robots with laser scanning tasks in a space-suitable way.
Needless to say, a lot of work remains to be done.
The fusion of complementary and Madgwick orientation filter solves the shortcomings of the Madgwick when at a standstill, particular the jitter.
A first qualitative evaluation shows the functionality of the filter.
In further experiments, a quantitative Orientation comparison with a synchronized optical system needs to be performed.
The position estimation as well as the combination of the integration by rotation of the gyroscope and the linear acceleration of the accelerometer  shows the capability to overcome the problem of the missing absolute reference of IMUs partly, which mainly  improves the measurements in the ground plane with slip behavior present.
Still, there is a need for parameter tuning. 
Sliding was not detected reliably all the time.
Here a mechanism for a delayed rotation needs to be investigated, that acceleration shortly before a rotation is integrated and taken into account once rotation starts.
The approach for, at least roughly, indicating the change in the $z$-direction, which is now done only by the accelerometer values, shows no promising results nor usable results in first qualitative experiments with rolling over obstacles and therefore needs to be re-investigated.

\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
                                  % on the last page of the document manually. It shortens
                                  % the textheight of the last page by a suitable amount.
                                  % This command does not take effect until the next page
                                  % so it should come on the page before the last. Make
                                  % sure that you do not shorten the textheight too much.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{thebibliography}{99}
\bibitem{c2} Rossi A.P. et al.. "DAEDALUS - Descent And Exploration in Deep Autonomy of Lava Underground Structures". Wuerzburger Forschungsberichte in Robotik und Telematik (21). 978-3-945459-33-1. (2021).
\bibitem{c1} Madgwick, Sebastian O. H.. "An efficient orientation filter for inertial and inertial / magnetic sensor arrays". (2010).
\bibitem{complementary}Mahony, R. ; Hamel, T.; Pflimlin, J.-M. ."Complementary filter design on the special orthogonal group SO(3)", Proceedings of the 44th IEEE Conference on Decision and Control. (2005).
\bibitem{c3} Kok, Manon; Hol, Jeroen D.;  Schoen,Thomas B. .  "Using Inertial Sensors for Position and Orientation Estimation", Foundations and Trends in Signal Processing: Vol. 11.(2017).
\bibitem{Valencia2017ASM} Botero Valencia, J.-S. ; Garc{\'i}a, M. R. ; Ceballos, J. V. . "A simple method to estimate the trajectory of a low cost mobile robotic platform using an IMU". International Journal on Interactive Design and Manufacturing (ijidem). (2017).
\bibitem{c4}Dam, Erik B.; Koch, Martin;  Lillholm, Martin. "Quaternions, Interpolation and Animation". (1998).
\bibitem{FastInvSqrt}"Origin of Quake3's Fast InvSqrt()". beyond3d.com/content/articles/8/. Accessed: 2021-05-01.

\end{thebibliography}

\end{document}

### 2021-05-13 Latex/sections/registration.tex
Replace rigid registration with coarse prere-registration; added info about pre-registration

@@
 We propose an algorithm that consists of multiple steps outlined in algorithm~\ref{algo:registration-algorithm}. 
-Based on ideas described in~\cite{Borrmann2010} we first rigidly register the scans and then further improve the overall map by exploiting the fact that human-made environments often consist of planes.
+Based on ideas described in~\cite{Borrmann2010} we first pre-register the scans and then further improve the overall map by exploiting the fact that human-made environments often consist of planes. 
 We then find the planes in the pre-registered point cloud and then optimize the poses associated with the scans to minimize the distance of all points to their respective planes.
@@
     \KwResult{A corrected map}
-    1. Rigidly register the scans\;
+    1. Pre-register the scans\;
     2. Extract planes from the registered map\;
@@
 
-\subsection{Rigid Registration}
+\subsection{Pre-Registration}
 
-\todo[inline]{Die Rigid Registration sollte kurz erwähnt werden. Welcher Algorithmus? Werden dafür Linescans zusammengefasst?}
+We perform a coarse registration with the continuous-time SLAM approach from~\cite{REMSEN2013}.
+This yields maps that can be used for plane registration and hence a point-to-plane correspondences base registration.
 
@@
 
-After having done the rigid registration, the scans are aligned well enough to make statements about the potential planes in the environment.
+After having done the pre-registration, the scans are aligned well enough to make statements about the potential planes in the environment.
 To find the planes in the environment, a Randomized Hough transform with an accumulator ball as described in~\cite{3DRESEARCH2011} is used. 

### 2021-05-13 Latex/sections/registration.tex
small typos

--- 2021-05-13 Latex/sections/registration.tex
@@
 For our particular application, ADADELTA behaves a little too aggressively.
-Despite giving a good measure on how to adapt $\boldsymbol\alpha$, the algorithm sometimes overshoots, and convergence doesn't happen.
+Despite giving a good measure on how to adapt $\boldsymbol\alpha$, the algorithm sometimes overshoots, and does not converge.
 Therefore, we employ another scaling factor, typically not found in ADADELTA, extending eq.~\eqref{eq:adadeltaalphaupdate} to:
@@
 \end{align}
-Using this algorithm once after finding correspondences from points to planes leads to convergence to a local minimum, which often isn't an optimal solution.
+Using this algorithm once after finding correspondences from points to planes leads to convergence to a local minimum, which is often not an optimal solution.
 Even if we increase the number of iterations dramatically, no better solution than the local minimum is found.
 That is unless you consider updating the correspondence model after $i$ iterations of gradient descent.
-Re-assigning point-to-plane correspondences this way $k$ times, and if $k$ is chosen large enough, it leads to an optimal solution after maximum $n = k\cdot i$ iterations of gradient descent.
+Re-assigning point-to-plane correspondences this way $k$ times, with a large enough $k$, leads to an optimal solution after maximum $n = k\cdot i$ iterations of gradient descent.
 
@@
 Secondly, we employ a continuous iteration, where the scans are processed sequentially. 
-This is especially useful for mobile systems, where pose error accumulates due to increasing tracking uncertainties.
+This is especially useful for mobile systems, where pose errors accumulate due to increasing tracking uncertainties.
 The assumption is that the error in one scan is also present in the following scan, plus some unknown new error.

### 2021-05-13 Latex/sections/results.tex
small typos

@@
 Before the registration, the corridor is only represented acceptably in the front part. 
-The further into the corridor, i.e., the longer the robot accumulates error, the more imprecise the date becomes. 
+The further into the corridor, i.e., the longer the robot accumulates errors, the more imprecise the data becomes. 
 Finally, we see that a large portion of the data even exceeds either the \SI{1}{\meter} error, contributing to the large spike in the histogram or even the \SI{2}{\meter} error limit; thus, being cut from the representation.
@@
 
-Figure~\ref{fig:cylon-corrected} shows the results obtained before and after employing the plane base registration on the dataset acquired by the floating sphere.
+Figure~\ref{fig:cylon-corrected} shows the results obtained before and after employing the plane based registration on the dataset acquired by the floating sphere.
 We combine ten scans into one meta-scan, which is then globally registered.

### 2021-05-13 Latex/sections/realData.tex
Replaced exploded view with better render from jasbaby

@@
 
-Figure~\ref{fig:float-sat-sphere} shows the experimental setup. The left column of figure~\ref{fig:cylon-corrected} shows the pre-registered resulting point cloud . 
+Figure~\ref{fig:float-sat-sphere} shows the experimental setup. The left column of figure~\ref{fig:cylon-corrected} shows the pre-registered resulting point cloud. 
 We see that the point cloud is rather noisy, and in particular, the walls are sensed multiple times, hence appearing very blurry.
@@
 		\centering
-		\includegraphics[width=0.4615\textwidth, trim={20cm 0 10cm 0}, clip]{./images/cylon-exploded}\hfill
-		\includegraphics[width=0.525\textwidth]{./images/cylon_setup}\hfill
+		\includegraphics[width=0.495\textwidth]{./images/cylon_exploded}\hfill
+		\includegraphics[width=0.495\textwidth]{./images/cylon_setup}\hfill
 	\end{minipage}
@@
 	Additionally, the building was scanned using a terrestrial laser scanner to provide a reference.
-	The measurements were then rigidly corrected before applying our plane-based correction. 
+	The measurements were then pre-registered before applying our plane-based correction. 
 	Figure~\ref{fig:experimental-setup} shows the acquired 3D point clouds and the experimental setup.
@@
 		\caption{Reference 3D point cloud acquired with a RIEGL 3D terrestrial laser scanner (top left).
-The 3D point cloud acquired by the test sphere after the rigid registration (bottom left) and a picture of the test setup (right).}
+The 3D point cloud acquired by the test sphere after the pre-registration (bottom left) and a picture of the test setup (right).}
 		\label{fig:experimental-setup}

### 2021-05-14 Latex/sections/conclusion.tex
removed weird english sentence; added 95%-üercentile error of uncorrected data

@@
 We have evaluated the procedure on a simulated dataset and on an experimentally acquired dataset with different motion profiles. 
-In this evaluation, we have shown that the procedure improves both datasets and yields more human-made, structure-slike maps. 
+In this evaluation, we have shown that the procedure improves both datasets and yields maps that better resemble human-made, structures. 
 In particular, the qualitative structure of the environments is reconstructed well. 
@@
 %
-Curently, we are assembling the sphere of Figure~\ref{FIXME} to test it in real environments.
+Curently, we are assembling the sphere of Figure~\ref{fig:robotRender} to test it in real environments.
 Unfortunately, this sphere will be too large to fit our floating setup, thus it will be tested in an office like environment.

### 2021-05-14 Latex/sections/simulation.tex

@@
 
-Accordingly, a spherical robot that is assumed to be translating exactly along one axis by rotation exactly about one other axis at an angular velocity $\omega$  experiences random disturbance torques at each time instant that accelerate the sphere's rotation about the respective axes.
+Accordingly, a spherical robot that is assumed to be translating exactly along one axis by rotation exactly about one other axis at an angular velocity $\omega$  experiences random disturbance torques at each time instant that accelerate the rotation of the sphere about the respective axes.
 This additional motion is determined by integrating the disturbance torques, which are sampled from some normal distribution $n_\phi, n_\psi \in \mathcal{N}(\mu,\sigma^2)$ each.  
@@
 Assuming intended locomotion along the $z$-axis and unintended locomotion along the $x$-axis where $R$ is the sphere radius.
-No movement upwards is possible since the robot moves on a flat ground, hence the entry is zero.
+No upwards movement is possible since the robot moves on a flat ground, hence the entry is zero.
 

### 2021-05-14 mfi21/ieeeconf/root.tex
small changes

@@
 Spherical robots are a robot format that is not yet thoroughly studied for the application of mobile mapping.
-However, in contrast to other forms they provide some unique advantages.
-For one, the spherical shell provides a protection against harsh environments, e.g. guarding the sensors and actuators from dust and solid rock.
+However, in contrast to other forms, they provide some unique advantages.
+For one, the spherical shell provides protection against harsh environments, e.g., guarding the sensors and actuators against dust and solid rock.
 This is particularly useful in space applications. 
 Furthermore, the inherent rotation the robot uses for locomotion can be exploited to measure in all directions without having the sensor itself actuated.
-To use this rotation in combination with sensor data to create a consistent environment map a reasonable estimation of the robot pose is required. 
-In such cases the pose estimation can be done by inertial measurements from sensors such as accelerometers and gyroscopes, as interpolating instance for calculation-intensive slow algorithms as optical localization algorithms or as rough estimation.
+To use this rotation in combination with sensor data to create a consistent environment map, a reasonable estimation of the robot pose is required. 
+In such cases the pose estimation can be done by inertial measurements from sensors such as accelerometers and gyroscopes, as interpolating instances for calculation-intensive slow algorithms as optical localization algorithms or as rough estimation.
 We propose a pose estimation procedure based on inertial measurements that exploits the known dynamics of a spherical robot. 
 It emphasizes a low jitter to maintain constant world measurements during standstill and avoids exponentially growing error in position estimates. 
-The evaluation of  the position and orientation estimates with given ground truth frames shows that we can reduce the jitter in orientation and can handle slip and partly slide behavior as other commonly used filters such as the Madgwick filter.
+Evaluating the position and orientation estimates with given ground truth frames shows that we can reduce the jitter in orientation and handle slip and partly slide behavior as other commonly used filters such as the Madgwick filter.
 
@@
 
-Spherical robots are a relatively narrow field of robotics. Still, they could be useful in situations where the measurement equipment needs to be protected against the harsh environments.
+Spherical robots are a relatively narrow field of robotics. Still, they could be useful in situations where the measurement equipment needs to be protected against harsh environments.
 As an example, the 2021 CDF study LunarCaves by ESA about the feasibility of a spherical robot, called DAEDALUS, for exploration of lunar-lava tubes \cite{c2} shows the need for orientation and position estimation of spherical robots with the limitations on resources common for space-applications.
 Although a purely Inertial Measurement Unit (IMU)-based estimation is not recommended for various reasons, the lack of an absolute reference among others, IMU-based pose estimation can be used as part of an overall multi-sensor-fusion-based estimation.
-IMUs are sensors with higher refresh rates than other Simultaneous Localization and Mapping (SLAM) based algorithms, such as optical or Light Detection and Ranging (LIDAR) SLAM, making them suitable for interpolating the more precise and absolute position reference points of these slower algorithms.
+IMUs are sensors with higher refresh rates than other Simultaneous Localization and Mapping (SLAM) based algorithms, such as optical or Light Detection and Ranging (LIDAR) SLAM.
+This makes them suitable for interpolating the more precise and absolute position reference points of these slower algorithms.
 The limited computational power of space-qualified CPUs further imposes the need for good orientation estimation without complex algorithms like Kalman filters.
-Lastly, the geometry of a spherical robot opens the possibility for the not often used position-estimation by IMUs. \\
+Lastly, the geometry of a spherical robot opens the possibility for the not often used position estimation by IMUs. \\
 This paper introduces an IMU-based orientation and position algorithm for spherical robots with limited computational power and the need for real-time estimations.
-Also the algorithm  does not rely on one specific locomotion approach, but it assumes a rotation of the IMU together with at least the shell of the robot if not the whole robot.
+Also, the algorithm does not rely on one specific locomotion approach, but it assumes a rotation of the IMU together with at least the shell of the robot, if not the whole robot.
 
@@
 \label{AssumptionsAndPrelimitations}
-The presented approach requires multiple assumptions, to be beneficial-Therefore, we classify the calculated pose concerning its usage and define the usage as follows:
+The presented approach requires multiple assumptions to be beneficial. Therefore, we classify the calculated pose concerning its usage and define the usage as follows:
 
@@
 \paragraph*{Uncertainty of Iterative Position Integration}
-IMUs deliever no absolute reference for a position as they do for orientation in the form of the noisy but measurable gravity vector.
+IMUs deliver no absolute reference for a position as they do for orientation in the form of the noisy but measurable gravity vector.
 So without external or secondary sensors relying on absolute waypoints, the position estimation will always underlie the integration of errors.
@@
 The sphere from the DAEDALUS example has two completely different locomotion methods; therefore, we will not take locomotion itself into account of pose estimation.
-Also, the same locomotion-approach can lead to heavily different behavior with the same input to the controller.
+Also, the same locomotion approach can lead to heavily different behavior with the same input to the controller.
 \cite{c2}
@@
 The generated data is a compromise between simple filtration of data and heavy calculation post-record.
-It is easy to calculate and adapt but due to the number of pre-assumptions, as it makes assumptions about the dynamics of the system, it still delivers reliable data which takes physical behavior into account.
+It is easy to calculate and adapt, but due to the number of pre-assumptions, as it makes assumptions about the dynamics of the system, it still delivers reliable data which takes physical behavior into account.
 It is real-time capable and does not rely on delay steps.
@@
 \subsection{Complementary Filter}
-A complimentary filter is a fundamental approach of combining precise but not exact with nonprecise but exact data.
+A complementary filter is a fundamental approach of combining precise but not exact with nonprecise but exact data.
 Therefore it is widely used for combining noisy accelerometer data with gravity as a reference with the precise but drifting measurement of gyroscopes.
-The basic concept is to have the main part of a value dominated by a gyroscope but over time a slow drift towards the measurement of the accelerometer.
-E.g. for pitch this would result in 
+The basic concept is to have the main part of a value dominated by a gyroscope but, over time, a slow drift towards the measurement of the accelerometer.
+E.g. for pitch, this would result in 
 \begin{equation}
@@
 The complementary filter is best for orientation calculation without position change as the accelerometer value relies on calculating the gravity value.
-With an acceleration in one direction, this leads to a miscalculation of the gravitational vector and, therefore a false orientation estimation.
+With an acceleration in one direction, this leads to a miscalculation of the gravitational vector and, therefore, a false orientation estimation.
 Simple orientation estimations of quadrocopters often just change the $\alpha$ gain to 0 while moving to avoid the miscalculation of the gravity vector. \cite{complementary}
@@
 \includegraphics[width=\linewidth]{./graphics/MadgwickVsComplementaryNoMovement.eps}
-\caption{Quaternion values of Madgwick filter and complementary filter with default gains when IMU is at a stillstand. The Madgwick filter has clearly visible jitter, the complementary filter not }
+\caption{Quaternion values of Madgwick filter and complementary filter with default gains when IMU is at a standstill. The Madgwick filter has clearly visible jitter, the complementary filter not}
 \label{madgVsComp1}
@@
 \subsection{Position Estimation}
-Position estimation purely by IMU is not very recommended as the double integration of the accelerometer measurements leads to exponential error summation.
+Position estimation purely by IMU is not recommended as the double integration of the accelerometer measurements leads to exponential error summation.
 Therefore there are very few approaches targeting this way of position calculating \cite{c3}.
@@
 They follow a pre-specified trajectory with a manipulator and manage to estimate its pose within 10\% of the ground truth.
-Given their application is rather broad they do not exploit known dynamics of the system which could be useful for spherical robots.
+Given their application is rather broad, they do not exploit known dynamics of the system, which could be helpful for spherical robots.
 
@@
 The pose orientation for the spherical robot is a concept with interconnected components.
-Both orientation and position estimates are computed based on the gyroscope and accelerometer measurements taking into account limitations of the system described in Section \ref{AssumptionsAndPrelimitations}.
+Both orientation and position estimates are computed based on the gyroscope and accelerometer measurements, taking into account the limitations of the system described in Section \ref{AssumptionsAndPrelimitations}.
 The generalized overview of orientation and position estimation is shown in Figure \ref{generalized}.
@@
 If the sphere slips, this approach will predict the full translation.
-Also the other way around, if there is translation but just slight rotation because of sliding, it will predict too little translation. 
+Also, the other way around, if there is translation but just slight rotation because of sliding, it will predict too little translation. 
 Furthermore, integrating the rotation does not let us estimate vertical movement.
 Every rotation will be predicted as translation in a given plane, which is most likely parallel to the ground. 
-If for instance the sphere rolls up an obstacle, the length of the path is estimated to the ground-plane.\\
+For instance, if the sphere rolls up an obstacle, the length of the path is estimated to the ground-plane.\\
 Our algorithm uses both approaches of position estimation and combines them, and limits all components according to the other components. 
-This implies a connection between the translation and rotation. 
-Overall the extreme situation (full translation, no rotation, or the other way round) will automatically lead to wrong predictions, as there is no way to tell which of the two approaches has at that moment more reliable data and therefore the combination of both rather then only one is chosen.
+This implies a connection between translation and rotation. 
+Overall the extreme situation (full translation, no rotation, or the other way round) will automatically lead to wrong predictions, as there is no way to tell which of the two approaches has at that moment more reliable data and therefore, the combination of both rather then only one is chosen.
 
@@
 The orientation is a symbiosis of the Madgwick filter and complementary filter.
-For no- or slow-moving, the complementary filter is chosen, for fast-moving the Madgwick filter.
+For no- or slow-moving, the complementary filter is chosen; for fast-moving, the Madgwick filter.
 The Madgwick filter itself is untouched.
@@
  As the integration only works in the Euler angles domain, the jumps of values at the border of the domain like 0 and 360 degrees do not match the mostly continuous change of quaternions.
-As the representation is still valid (0 degree and 360 degree roll being the same position), this is not a problem when using at one point any of the filters exclusively.
- But combining both, this leads to contradicting updates from each filter with respect to the other.
+As the representation is still valid (0 degree and 360-degree roll being the same position), this is not a problem when using at one point any of the filters exclusively.
+ But combining both leads to contradicting updates from each filter with respect to the other.
  Therefore the normal way of the complementary filter of the new orientation is 
@@
 Hence, the RPY-orientation is transformed to a quaternion, and then a quaternion slerp is performed \cite{c4}.
- The quaternion slerp works over the representation of two quaternions on a sphere. The transisiton from one quaternion to the other is now not done directly but always over the surface of the sphere.
+ The quaternion slerp works over the representation of two quaternions on a sphere. The transition from one quaternion to the other is now not done directly but always over the surface of the sphere.
  This leads to the same handling of quaternions as the Madgwick-filter without endangering robustness due to using two different quaternions with very different values representing nearly the same position.\\
-The transistion between the two filters is done by shifting a fixed factor among the $\alpha$-gain of the complementary filter and the gain factor of the Madgwick filter.
+The transition between the two filters is done by shifting a fixed factor among the $\alpha$-gain of the complementary filter and the gain factor of the Madgwick filter.
  As the common values widely used for both values differ by one order of magnitude, the gain shifted to the complementary filter is divided by ten and the other way around.
@@
  This avoids large accelerations of the orientation due to the rapid change from one algorithm to another.
- In our experiments, when starting from stillstand, the incipient Madgwick filter, has a more rapid impact on the orientation values then the incipient complementary filter, when stopping rotation. 
+ In our experiments, when starting from stillstand, the incipient Madgwick filter has a more rapid impact on the orientation values than the incipient complementary filter when stopping rotation. 
 Therefore the transition between $\alpha$-gain of the complementary filter and the Madgwick-gain was chosen to have quadratic behavior. 
@@
 Therefore we calculate factors from the gyroscope measurements and scale the autogain with the maximum $\beta$ (cf. equation \ref{eq:factors} and \ref{eq:weight_autogain}).
-The $g_{k}$ represent the gyro measurement in the corresponding an axis $k$ in rad/s.
+The $g_{k}$ represents the gyro measurement in the corresponding an axis $k$ in rad/s.
 Later on, these factors will again be used for position estimation, namely to determine if the sphere rotates or not.
@@
 We used $\theta = 0.1$ because of the the two often used standard values $\gamma = 0.2$ and $\alpha= 0.02$, which have the ration $0.1$.
-Hence for adapting the position estimation to a given robot, the function for calculating the factors needs to be adapted, in such a way that it is a good indication for movement or standstill.
-The ratio $\theta$ needs to be chosen in a way, that the desired ratio between Madgwick and complementary gain is reached.
-And lastly the autogain needs to be chosen to fit the specific needs in terms of how strong both filters should be able to influence the orientation.
-We used $\Theta = 0.2$ to get the described standard values.
+Hence for adapting the position estimation to a given robot, the function for calculating the factors needs to be adapted in such a way that it is a good indication for movement or standstill.
+The ratio $\theta$ needs to be chosen in a way that the desired ratio between Madgwick and complementary gain is reached.
+And lastly, the autogain needs to be chosen to fit the specific needs in terms of how strong both filters should be able to influence the orientation.
+We used $\Theta = 0.2$ to get the described common values.
 
@@
 describes the effect of gravity in the coordinate system of the IMU.
-This gravity vector is then subtracted from the accelerometer measurement. If the there is no translation, the subtraction should lead to $\begin{bmatrix}
+This gravity vector is then subtracted from the accelerometer measurement. If there is no translation, the subtraction should lead to $\begin{bmatrix}
 0 & 0&0
@@
 These Values are now coupled to the rotation of both axes other than their own. Meaning there can be a translation in $x$ if there is rotation around $y$ or $z$.
-Therefore, the factors from the orientation step, representing the likelihood of rotation are used as limiting factors.
+Therefore, the factors from the orientation step, representing the likelihood of rotation, are used as limiting factors.
 If one of the three factors is 1, this means we let translation in this direction happen.
@@
  To avoid large errors and unlimited double error integration, $v_{z\text{AccWorld}} $ is limited to the mean of $v_{xz\text{GyroWorld}} $ and $ v_{yz\text{GyroWorld}}$.
-With the pythagorean theorem the velocity in x and y is calculated accepting the velocity in z direction, which can not be derived by the velocity from gyro:
+With the Pythagorean theorem the velocity in x and y is calculated accepting the velocity in z-direction, which can not be derived by the velocity from gyro:
  \begin{equation}
@@
 
-Overall an pose estimation step consist only of operations that are constant in time. 
-Given that we use the fast inverse square root algorithm \cite{FastInvSqrt} for normalization the trigonometric functions are considered the remaining bottleneck.
+Overall a pose estimation step consists only of operations that are constant in time. 
+Given that we use the fast inverse square root algorithm \cite{FastInvSqrt} for normalization, the trigonometric functions are considered the remaining bottleneck.
 Overall there are $4\times\arctan$, $4\times\sin$, $4\times\cos$, $1\times\arcsin$, $1\times\arccos$ and $5$ divisions. 
@@
 \section{Evaluation}
-The proposed algorithm is evaluated by two sets of experiments performed with three Phidgets 3/3/3 IMUs mounted on an round acrylic glass plate.
-For position experiments the plate is put inside a acrylic glass sphere where the plate exactly fits in the middle, as the radii are the same.
+The proposed algorithm is evaluated by two sets of experiments performed with three Phidgets 3/3/3 IMUs mounted on a round acrylic glass plate.
+For position experiments, the plate is put inside an acrylic glass sphere where the plate exactly fits in the middle, as the radii are the same.
 First, the fusion of Madgwick and complementary filter, here denoted as Autogain, is compared to a Madgwick only estimate.
@@
 A first experiment is performed to show the overall performance of the Autogain computation for the orientation.
-The IMU was manually rotated in arbitrary directions and the orientation plotted using the Madgwick filter and the Autogain filter.
+The IMU was manually rotated in arbitrary directions, and the orientation plotted using the Madgwick filter and the Autogain filter.
 As expected, the overall behavior of the presented algorithm and the Madgwick filter is very similar in the dynamic case, as shown in Figure \ref{quaternion1}.
@@
 \end{figure}
-Therefore the more interesting part is the change between motion and standstill, as in motion, both filters use the same algorithm but differ during standstill and slow motion.
+Therefore the more exciting part is the change between motion and standstill, as in motion, both filters use the same algorithm but differ during standstill and slow motion.
 As described in Section \ref{sec:SOA}, the Madgwick filter has a rather string jitter during slow movements and standstill. 
@@
 \subsection{Position}
-The primary evaluation for position focuses on the slip- and slide-behavior.
+The primary evaluation for the position focuses on slip- and slide-behavior.
 Here the plate with the IMU(s) is fixed inside a sphere with a diameter of 29 cm and rolled along an L-shaped track of one meter by one meter.
@@
 The fusion between the accelerometer and gyroscope interprets an L with dimensions much closer to the original.
-The estimation leaves room for improvement, as in one direction, it misses the actual length about -10 percent and in the other direction overestimates it by 10 percent.
+The estimation leaves room for improvement, as in one direction, it misses the actual length about -10 percent, and in the other direction overestimates it by 10 percent.
 This should be improvable by parameter tuning.
@@
 
-The same experiment was performed with added slide, i.e. more translation than just the rotation would produce.
+The same experiment was performed with added slide, i.e., more translation than just the rotation would produce.
 Here the results shown in Figure \ref{LTestSlide} were not as reliably good as with the slip experiment.
@@
 \end{figure}
-The experiments for sliding did not reliably detect the slide event and, on some occasions, interpreted the distance shorter than just the integration, as the acceleration of the linear motion was missed.
-However, the deceleration of the sphere was still recognized while rotation existed and therefore, the sphere miscalculated a couple of centimeters.
+The experiments for sliding did not reliably detect the slide event and, on some occasions, interpreted the distance as shorter than just the integration, as the acceleration of the linear motion was missed.
+However, the deceleration of the sphere was still recognized while rotation existed, and therefore, the sphere miscalculated a couple of centimeters.
 
@@
 Needless to say, a lot of work remains to be done.
-The fusion of complementary and Madgwick orientation filter solves the shortcomings of the Madgwick when at a standstill, particular the jitter.
+The fusion of complementary and Madgwick orientation filter solves the shortcomings of the Madgwick when at a standstill, particularly the jitter.
 A first qualitative evaluation shows the functionality of the filter.
 In further experiments, a quantitative Orientation comparison with a synchronized optical system needs to be performed.
-The position estimation as well as the combination of the integration by rotation of the gyroscope and the linear acceleration of the accelerometer  shows the capability to overcome the problem of the missing absolute reference of IMUs partly, which mainly  improves the measurements in the ground plane with slip behavior present.
+The position estimation, as well as the combination of the integration by rotation of the gyroscope and the linear acceleration of the accelerometer, shows the capability to overcome the problem of the missing absolute reference of IMUs partly, which mainly improves the measurements in the ground plane with slip behavior present.
 Still, there is a need for parameter tuning. 
-Sliding was not detected reliably all the time.
+Sliding was not detected all the time reliably.
 Here a mechanism for a delayed rotation needs to be investigated, that acceleration shortly before a rotation is integrated and taken into account once rotation starts.

### 2021-05-14 Latex/main.tex
Introduction and state of the art

@@
 \newcommand{\pardiff}[1]{\frac{\partial}{\partial #1}}
+\renewcommand{\epsilon}{\varepsilon}
 
@@
 \title{Towards Spherical Robots For Mobile Mapping In Human Made Environments\tnoteref{mytitlenote}}
-\tnotetext[mytitlenote]{To abide the \href{https://www.go-fair.org/fair-principles/}{FAIR-Principles} of science, all data and code used to produce the results are given in this our repository (\url{https://github.com/fallow24/SphereTDP}).}
+\tnotetext[mytitlenote]{To abide the \href{https://www.go-fair.org/fair-principles/}{FAIR-Principles} of science, all data and code used to produce the results are given in our repository (\url{https://github.com/fallow24/SphereTDP}).}
 
@@
 We show that we nearly reproduce the ground truth for the simulated dataset and improve the average point-to-point distance to a reference scan in the experimental dataset.
-The presented algorithms are required to work completely autonomous.
+The presented algorithms are required to work completely autonomously.
 \end{abstract}

### 2021-05-14 Latex/sections/results.tex
removed weird english sentence; added 95%-üercentile error of uncorrected data
Added percentile table

@@
 After registration, we see that, qualitatively, the ideal corridor was nearly restored from the noisy data. 
-In particular a very large portion (95\%) have less than \SI{17}{\centi\meter} distances.
+In particular a very large portion of points (95\%) have less than \SI{16.53}{\centi\meter} distances.
+Table~\ref{tab:percentiles} shows the comparison of further percentiles of both datasets.
+
+\begin{table}
+	\centering
+	\begin{tabular}{@{}cccc@{}}\hline
+		& P90 & P95 & P98\\ \hline\hline
+		Uncorrected & \SI{24.12}{\centi\meter} &  \SI{38.19}{\centi\meter} &  \SI{61.46}{\centi\meter}\\
+		Corrected & \SI{12.78}{\centi\meter} &  \SI{16.53}{\centi\meter} &  \SI{21.55}{\centi\meter}\\ \hline
+	\end{tabular}
+	\caption{Comparison of percentiles of point-distances in the uncorrected and corrected simulated dataset.}
+	\label{tab:percentiles}
+\end{table}
+
 Further, the square and straight shape of the corridor is restored well, and especially the large spike at errors of greater than \SI{1}{m} is removed. 

### 2021-05-14 Latex/sections/results.tex
Added mean to stats table

--- 2021-05-13 Latex/sections/results.tex
@@
 After registration, we see that, qualitatively, the ideal corridor was nearly restored from the noisy data. 
-In particular a very large portion (95\%) have less than \SI{17}{\centi\meter} distances.
+In particular a very large portion of points (95\%) have less than \SI{16.53}{\centi\meter} distances.
+Table~\ref{tab:percentiles} shows the comparison of further percentiles of both datasets.
+
+\begin{table}
+	\centering
+	\begin{tabular}{@{}lcccc@{}}\hline
+		& P90 & P95 & P98 & Mean\\ \hline\hline
+		Uncorrected & \SI{24.12}{\centi\meter} &  \SI{38.19}{\centi\meter} &  \SI{61.46}{\centi\meter} & \SI{9.97}{\centi\meter}\\
+		Corrected & \SI{12.78}{\centi\meter} &  \SI{16.53}{\centi\meter} &  \SI{21.55}{\centi\meter} & \SI{6.36}{\centi\meter}\\ \hline
+	\end{tabular}
+	\caption{Comparison of point-distances in the uncorrected and corrected simulated dataset.}
+	\label{tab:percentiles}
+\end{table}
+
 Further, the square and straight shape of the corridor is restored well, and especially the large spike at errors of greater than \SI{1}{m} is removed. 

### 2021-05-14 Latex/sections/results.tex
Updated histograms and statistics of results

--- 2021-05-13 Latex/sections/results.tex
@@
 After registration, we see that, qualitatively, the ideal corridor was nearly restored from the noisy data. 
-In particular a very large portion (95\%) have less than \SI{17}{\centi\meter} distances.
+In particular a very large portion of points (95\%) have distances of less than \SI{16.6}{\centi\meter}.
+Table~\ref{tab:percentiles} shows the comparison of further percentiles of both datasets.
+
+\begin{table}
+	\centering
+	\begin{tabular}{@{}lcccc@{}}\hline
+		& P90 & P95 & P98 & Mean\\ \hline\hline
+		Uncorrected & \SI{24.2}{\centi\meter} &  \SI{38.2}{\centi\meter} &  \SI{61.5}{\centi\meter} & \SI{9.7}{\centi\meter}\\
+		Corrected & \SI{12.8}{\centi\meter} &  \SI{16.6}{\centi\meter} &  \SI{21.6}{\centi\meter} & \SI{6.3}{\centi\meter}\\ \hline
+	\end{tabular}
+	\caption{Comparison of point-distances in the uncorrected and corrected simulated dataset.}
+	\label{tab:percentiles}
+\end{table}
+
 Further, the square and straight shape of the corridor is restored well, and especially the large spike at errors of greater than \SI{1}{m} is removed. 

### 2021-05-14 Latex/sections/simulation.tex
Cleaned up line breaks for lines then went pst the boundary

@@
 For the simulated data, we implemented a noisy world-robot-sensor model.  
-We are simulating a Livox-Mid100~\cite{LivoxMid40-100} laser scanner with customizable noise level on the range measurements inside a robot with different motion capabilities, subject to noise in its pose estimation.
+The simulated sensor is modeled after a Livox-Mid100~\cite{LivoxMid40-100} laser scanner with customizable noise level on the range measurements inside a robot with different motion capabilities, subject to noise in its pose estimation.
 This yields scan results similar to the ones obtained in the real world.
@@
 
-For the pose measurement, one cannot simply add white noise to the current pose estimate as this does not capture the integration error, which is common among inertial measurement units (IMU). 
+Regarding the pose measurement, one cannot simply add white noise to the current pose estimate as this does not capture the integration error, which is common among inertial measurement units (IMU). 
 Therefore we assume a disturbance torque about the two axes that lie in the ground plane. 
@@
 
-Accordingly, a spherical robot that is assumed to be translating exactly along one axis by rotation exactly about one other axis at an angular velocity $\omega$  experiences random disturbance torques at each time instant that accelerate the rotation of the sphere about the respective axes.
+Accordingly, a spherical robot that is assumed to be translating exactly along one axis by rotating about one other axis at an angular velocity $\omega$  experiences random disturbance torques at each time instant that accelerate the rotation of the sphere about the respective axes.
 This additional motion is determined by integrating the disturbance torques, which are sampled from some normal distribution $n_\phi, n_\psi \in \mathcal{N}(\mu,\sigma^2)$ each.  

### 2021-05-14 Latex/sections/conclusion.tex

@@
 
-However, as section~\ref{sec:experimentalSetup} shows, some outliers remain due to the simplistic point-to-plane correspondence model, which is to be improved in future studies.
+However, as section~\ref{sec:experimentalSetup} shows, some outliers remain in the registered point cloud due to the simplistic point-to-plane correspondence model.
+This to be improved in future studies.
 One idea to do so is to employ a local clustering technique onto the individual scans before registering them globally. 
@@
 %
-Curently, we are assembling the sphere of Figure~\ref{fig:robotRender} to test it in real environments.
+Currently, we are assembling the sphere of Figure~\ref{fig:robotRender} to test it in real environments.
 Unfortunately, this sphere will be too large to fit our floating setup, thus it will be tested in an office like environment.

### 2021-05-14 Latex/sections/introduction.tex

@@
 Smaller robotic systems are particularly used when cars no longer have access.
-Examples for this are human operated systems such as Zebedee~\cite{Bosse2012-zebedee}, a small Hokuyo 2D scanner on a spring, VILMA~\cite{JPRS2016}, a rolling FARO scanner operating in profiler mode, RADLER~\cite{Borrmann2020-RADLER}, a SICK 2D laser scanner mounted to a unicycle, or a backpack mounted ``personal laser scanning system'' as in~\cite{LauterbackEtAl2015-Backpack} or~\cite{WWWLeicaBackpack}.
+Examples for this are human operated systems such as Zebedee~\cite{Bosse2012-zebedee}, a small Hokuyo 2D scanner on a spring, that is carried through the environment, VILMA~\cite{JPRS2016}, a rolling FARO scanner operating in profiler mode, RADLER~\cite{Borrmann2020-RADLER}, a SICK 2D laser scanner mounted to a unicycle, or a backpack mounted ``personal laser scanning system'' as in~\cite{LauterbackEtAl2015-Backpack} or~\cite{WWWLeicaBackpack}.
 Recently more and more autonomous systems gained maturity.

### 2021-05-14 Latex/sections/registration.tex

@@
 
-We perform a coarse registration with the continuous-time SLAM approach from~\cite{REMSEN2013}.
-This yields maps that can be used for plane registration and hence a point-to-plane correspondences base registration.
+For the plane extraction the linescans need to be transformed into the project coordinate system.
+Various pre-registration methods are suitable for this.
+The most simple method is to use the data from an intertial measurement unit for a coarse alignment.
+Alternative solutions combine multiple linescans into a metascan and perform registration methods known from terrestrial laser scanning or use a small number of iterations of continuous-time SLAM approaches such as the one presented in~\cite{REMSEN2013}.
+This yields maps that can be used for plane registration and hence a point-to-plane correspondences based registration.
 
@@
 After having done the pre-registration, the scans are aligned well enough to make statements about the potential planes in the environment.
-To find the planes in the environment, a Randomized Hough transform with an accumulator ball as described in~\cite{3DRESEARCH2011} is used. 
+To find the planes in the environment, a Randomized Hough transform with an accumulator ball as described in~\cite{3DRESEARCH2011} is used as this method prefers dominant planes such as the building structure over smaller planar surfaces.
 In the next subsection, the Hesse-normal form of the plane is required.

### 2021-05-23 mfi21/ieeeconf/root.tex
Andreas
new evaluation

@@
 To evaluate specifically the change between slow motion and standstill, a further experiment was performed.
-The results in Figure \ref{quaternionq0} show a short excerpt of the experiment with four phases of slow motion.
+The results in Figure \ref{distance} show a short excerpt of the experiment with three phases of slow motion and rotation at second 7, 9 and 11. 
 For external reference, we used a precise optical tracking system. 
-Manually transforming the measured IMU values to the optical reference frame then
-yields the above plot.
-The optical system has the same shape as the presented algorithm and lacks the perturbations of the pure Madgwick filter.
-To stress the same behavior without relying on absolute orientation, which will be part of further research, Figure \ref{derivative} shoes the derivative of all three approaches.
+The definition and proposed calculation of the distance of 2 quaternions of \cite{kuffner} was used.
+The optical system matches the overall orientation of both IMU-based algorithms quite well and lacks the perturbations of the pure Madgwick filter.
+The transformation of the orientation by the optical tracking system and the IMU-based algorithms yield two unknown transformations, the relative orientation of the coordinate system of IMU to optical coordinate system and how the axes are interpreted on the tracked structure itself.
+Therefore, the algorithm proposed in \cite{dornaika} was used to solve these two unknown transformations on basis of the resulting data, to match the orientations of optical and imu-based data.
+As this matching of the optical system happens on basis of one of the two IMU orientation data-sets, the comparison of absolute differences may be misleading, as the transformations were matched on basis of the autogain filter.
+To stress the jitter behavior of the Madgwick without relying on absolute orientation, which will be part of further research, Figure \ref{diffdistance} shows the derivative of the distance.
 Here the change of direction due to small oscillations and the smoother fusion algorithm in comparison to the tracking is even more apparent.
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{./graphics/opticalOrientation.eps}
-\caption{q0-quaternion-value of the presented fusion of complementary and Madgwick filter, only the Madgwick filter and an evaluation via optical tracking }
-\label{quaternionq0}
+\includegraphics[width=\linewidth]{./graphics/distance.eps}
+\caption{Distance of the proposed and the Madgwick algorithm to the optical tracking system, using the distance definiton of \cite{kuffner}}
+\label{distance}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{./graphics/derivative.eps}
-\caption{Derivative of q0-quaternion-value of the presented fusion of complementary and Madgwick filter, only the Madgwick filter and an evaluation via optical tracking }
-\label{derivative}
+\includegraphics[width=\linewidth]{./graphics/diffdistance.eps}
+\caption{Derivative of the distance of the proposed and the Madgwick algorithm to the optical tracking system, using the distance definiton of \cite{kuffner}}
+\label{diffdistance}
 \end{figure}
@@
 \begin{thebibliography}{99}
-\bibitem{c2} Rossi A.P.
-et al..
-"DAEDALUS - Descent And Exploration in Deep Autonomy of Lava Underground Structures".
-Wuerzburger Forschungsberichte in Robotik und Telematik (21).
-978-3-945459-33-1.
-(2021).
-\bibitem{c1} Madgwick, Sebastian O. H.. "An efficient orientation filter for inertial and inertial / magnetic sensor arrays". (2010).
+\bibitem{c2} Rossi, A.P.
+et al., "DAEDALUS - Descent And Exploration in Deep Autonomy of Lava Underground Structures", Wuerzburger Forschungsberichte in Robotik und Telematik (21), 978-3-945459-33-1, (2021).
+\bibitem{c1} Madgwick, Sebastian O. H., "An efficient orientation filter for inertial and inertial / magnetic sensor arrays", (2010).
 \bibitem{complementary}Mahony, R.
-; Hamel, T.; Pflimlin, J.-M.
-."Complementary filter design on the special orthogonal group SO(3)", Proceedings of the 44th IEEE Conference on Decision and Control.
-(2005).
-\bibitem{c3} Kok, Manon; Hol, Jeroen D.;  Schoen,Thomas B. .  "Using Inertial Sensors for Position and Orientation Estimation", Foundations and Trends in Signal Processing: Vol. 11.(2017).
+; Hamel, T.; Pflimlin, J.-M.,"Complementary filter design on the special orthogonal group SO(3)", Proceedings of the 44th IEEE Conference on Decision and Control, (2005).
+\bibitem{c3} Kok, Manon; Hol, Jeroen D.;  Schoen,Thomas B.,  "Using Inertial Sensors for Position and Orientation Estimation", Foundations and Trends in Signal Processing: Vol. 11.(2017).
 \bibitem{Valencia2017ASM} Botero Valencia, J.-S.
@@
 ; Ceballos, J.
-V.
-.
-"A simple method to estimate the trajectory of a low cost mobile robotic platform using an IMU".
-International Journal on Interactive Design and Manufacturing (ijidem).
-(2017).
-\bibitem{c4}Dam, Erik B.; Koch, Martin;  Lillholm, Martin. "Quaternions, Interpolation and Animation". (1998).
-\bibitem{FastInvSqrt}"Origin of Quake3's Fast InvSqrt()". beyond3d.com/content/articles/8/. Accessed: 2021-05-01.
+V., "A simple method to estimate the trajectory of a low cost mobile robotic platform using an IMU", International Journal on Interactive Design and Manufacturing (ijidem), (2017).
+\bibitem{c4}Dam, Erik B.; Koch, Martin;  Lillholm, Martin, "Quaternions, Interpolation and Animation", (1998).
+\bibitem{FastInvSqrt}"Origin of Quake3's Fast InvSqrt()", beyond3d.com/content/articles/8/, Accessed: 2021-05-01.
+\bibitem {kuffner}Kuffner, James, "Effective sampling and distance metrics for 3D rigid body path planning", Proceedings - IEEE International Conference on Robotics and Automation, pp.
+3993 - 3998 Vol.4, (2004).
+\bibitem{dornaika}Dornaika, F.; Horaud, R., Simultaneous robot-world and hand-eye calibration, IEEE Transactions on Robotics and Automation 14(4), pp. 617-622, (1998).
 

### 2021-05-23 Latex/sections/acknowledgement.tex

@@
 We acknowledge funding from the ESA Contract No.
-4000130925/20/NL/GLC for the ``DAEDALUS -- Descent And Exploration in Deep Autonomy of Lava Underground Structures'' Open Space Innovation Platform (OSIP) lunar caves-system study and the Elite Network Bavaria (ENB) for providing funds for the academic program ``Satellite Technology''.
+4000130925/20/NL/GLC for the study ``DAEDALUS -- Descent And Exploration in Deep Autonomy of Lava Underground Structures'' within the Open Space Innovation Platform (OSIP) lunar caves-system and the Elite Network Bavaria (ENB) for providing funds for the academic program ``Satellite Technology''.

### 2021-05-23 Latex/sections/simulation.tex

@@
 
-The Livox-Mid100 laser scanner consists of three smaller-sized Mid40 scanners arranged to provide an overall field-of-view of \SI{98.4}{\degree} (Horizontal) by \SI{38.4}{\degree} (Vertical).
-It provides a range precision of \SI{2}{\centi\meter} up to \SI{20}{\meter} range and yields \SI[per-mode=symbol]{300,000}{\pts\per\second}.
+The Livox-Mid100 laser scanner consists of three Mid40 scanners arranged to provide an overall field-of-view of \SI{98.4}{\degree} (horizontal) by \SI{38.4}{\degree} (vertical).
+With a range precision of \SI{2}{\centi\meter} up to \SI{20}{\meter} range it provides \SI[per-mode=symbol]{300,000}{\pts\per\second} with a range precision of \SI{2}{\centi\meter} up to \SI{20}{\meter} range.
 A particular advantage of the Livox laser scanners is their unique scanning pattern.

### 2021-05-24 mfi21/ieeeconf/root.tex
changed small things, updatet commited version

@@
 
-\begin{figure}[thpb]
+\begin{figure}
 \includegraphics[width=\linewidth]{./graphics/MadgwickVsComplementaryNoMovement.eps}
@@
 The proposed algorithm is evaluated by two sets of experiments performed with three Phidgets 3/3/3 IMUs mounted on a round acrylic glass plate.
-For position experiments, the plate is put inside an acrylic glass sphere where the plate exactly fits in the middle, as the radii are the same.
+For position experiments, the plate is put inside an acrylic glass sphere with the same radius as the glass plate.
 First, the fusion of Madgwick and complementary filter, here denoted as Autogain, is compared to a Madgwick only estimate.
@@
 As expected, the overall behavior of the presented algorithm and the Madgwick filter is very similar in the dynamic case, as shown in Figure \ref{quaternion1}.
-\begin{figure}[thpb]
+\begin{figure}
 \includegraphics[width=\linewidth]{./graphics/AutoVsMad.eps}
@@
 \end{figure}
-Therefore the more exciting part is the change between motion and standstill, as in motion, both filters use the same algorithm but differ during standstill and slow motion.
-As described in Section \ref{sec:SOA}, the Madgwick filter has a rather string jitter during slow movements and standstill. 
-To evaluate specifically the change between slow motion and standstill, a further experiment was performed.
-The results in Figure \ref{distance} show a short excerpt of the experiment with three phases of slow motion and rotation at second 7, 9 and 11. 
+Therefore the more expressive part is the change between motion and standstill, as in motion, both filters use the same algorithm but differ during standstill and slow motion.
+As described in Section \ref{sec:SOA}, the Madgwick filter has a rather strong jitter during slow movements and standstill. 
+To evaluate specifically the change between slow motion and standstill, a further experiment is performed.
+Figure \ref{distance} show the distance to a precise external reference over a short excerpt of the experiment with three phases of slow motion and rotation at second 7, 9 and 11. 
 For external reference, we used a precise optical tracking system. 
 The definition and proposed calculation of the distance of 2 quaternions of \cite{kuffner} was used.
+It describes the direct angle between both quaternions,
 The optical system matches the overall orientation of both IMU-based algorithms quite well and lacks the perturbations of the pure Madgwick filter.
 The transformation of the orientation by the optical tracking system and the IMU-based algorithms yield two unknown transformations, the relative orientation of the coordinate system of IMU to optical coordinate system and how the axes are interpreted on the tracked structure itself.
-Therefore, the algorithm proposed in \cite{dornaika} was used to solve these two unknown transformations on basis of the resulting data, to match the orientations of optical and imu-based data.
+Therefore, the algorithm proposed in \cite{dornaika} was used to solve these two unknown transformations on basis of the resulting data, to match the orientations of optical and IMU-based data.
 As this matching of the optical system happens on basis of one of the two IMU orientation data-sets, the comparison of absolute differences may be misleading, as the transformations were matched on basis of the autogain filter.
@@
 The estimation leaves room for improvement, as in one direction, it misses the actual length about -10 percent, and in the other direction overestimates it by 10 percent.
-This should be improvable by parameter tuning.
 But it shows the overall capability of the algorithm to compensate for slip.
@@
 This is due to the timing of the acceleration.
-In order to recognize the acceleration as translation, there needs to be a little bit of rotation, as the algorithm does not allow any translation without rotation at all.
-As the translation computed using the accelerometer is by double integration, the acceleration of the slide movement needs to be at the same time as a minimal amount of rotation.
+In order to recognize the acceleration as translation, the rotation needs to be above a certain threshold, as the algorithm does not allow any translation without rotation at all.
+The translation is computed using the accelerometer via double integration.
+Therefore the acceleration of the slide movement needs to be at the same time as a minimal amount of rotation. 
 If the rotation starts after the sliding, the double integration of the acceleration is suppressed.
@@
 This is to be blamed on the execution of the experiment, as it was not perfectly ensured only to start sliding with ongoing rotation, which is hard to achieve when manually guiding a sphere.
-But the sliding behavior in a real environment could be in both ways, too, so sliding while rolling and starting sliding before rolling.
+However, the sliding behavior in a real environment could be in both ways, too, so sliding while rolling and starting sliding before rolling.
 With slip, this is not a problem, as, by definition, the translation cannot begin before the rotation when slipping.

### 2021-05-24 mfi21/ieeeconf/root.tex
changed one comma

@@
 The definition and proposed calculation of the distance of 2 quaternions of \cite{kuffner} was used.
-It describes the direct angle between both quaternions,
+It describes the direct angle between two quaternions.
 The optical system matches the overall orientation of both IMU-based algorithms quite well and lacks the perturbations of the pure Madgwick filter.

### 2021-05-28 mfi21/ieeeconf/root.tex
Andreas' changes

@@
 \usepackage{amsmath}
-\usepackage{enumitem}
+%\usepackage{enumitem}
 \newcommand{\M}[1]{\mathbf{#1}} % Bold for matrix
@@
         Corresponding author: }%
-        \thanks{ \quad  \tt\small jasper.zevering@stud-mail.uni-wuerzburg.de}%
+        \thanks{ \quad  \tt\small info@jasperzevering.de}%
 \thanks{$^{1}$ All authors are with Informatics VII : Robotics and Telematics - University of Wuerzburg}%
 \thanks{We acknowledge funding from the ESA Contract No.
-4000130925/20/NL/GLC for the “DAEDALUS – Descent And Exploration in Deep Autonomy of Lava Underground Structures” Open Space Innovation Platform (OSIP) lunar caves-system study and the Elite Network Bavaria (ENB) for providing funds for the academic program “Satellite Technology”.}
+4000130925/20/NL/GLC for the ``DAEDALUS -- Descent And Exploration in Deep Autonomy of Lava Underground Structures” Open Space Innovation Platform (OSIP) lunar caves-system study and the Elite Network Bavaria (ENB) for providing funds for the academic program “Satellite Technology''.}
 }
@@
 Furthermore, the inherent rotation the robot uses for locomotion can be exploited to measure in all directions without having the sensor itself actuated.
-To use this rotation in combination with sensor data to create a consistent environment map, a reasonable estimation of the robot pose is required. 
-In such cases the pose estimation can be done by inertial measurements from sensors such as accelerometers and gyroscopes, as interpolating instances for calculation-intensive slow algorithms as optical localization algorithms or as rough estimation.
-We propose a pose estimation procedure based on inertial measurements that exploits the known dynamics of a spherical robot. 
+To exploit this rotation in combination with sensor data to create a consistent environment map, a reasonable estimation of the robot pose is required. 
+In such cases the pose estimation are done by inertial measurements from sensors such as accelerometers and gyroscopes, as interpolating instances for calculation-intensive slow algorithms such as optical localization algorithms or as initial estimate for subsequent SLAM.
+The paper presents a pose estimation procedure based on inertial measurements that exploits the known dynamics of a spherical robot. 
 It emphasizes a low jitter to maintain constant world measurements during standstill and avoids exponentially growing error in position estimates. 
-Evaluating the position and orientation estimates with given ground truth frames shows that we can reduce the jitter in orientation and handle slip and partly slide behavior as other commonly used filters such as the Madgwick filter.
+Evaluating the position and orientation estimates with given ground truth frames shows that we reduce the jitter in orientation and handle slip and partly slide behavior as other commonly used filters such as the Madgwick filter.
 
@@
 As an example, the 2021 CDF study LunarCaves by ESA about the feasibility of a spherical robot, called DAEDALUS, for exploration of lunar-lava tubes \cite{c2} shows the need for orientation and position estimation of spherical robots with the limitations on resources common for space-applications.
-Although a purely Inertial Measurement Unit (IMU)-based estimation is not recommended for various reasons, the lack of an absolute reference among others, IMU-based pose estimation can be used as part of an overall multi-sensor-fusion-based estimation.
-IMUs are sensors with higher refresh rates than other Simultaneous Localization and Mapping (SLAM) based algorithms, such as optical or Light Detection and Ranging (LIDAR) SLAM.
+Although a purely Inertial Measurement Unit (IMU)-based estimation is not recommended for various reasons, the lack of an absolute reference among others, IMU-based pose estimation are used as part of an overall multi-sensor-fusion-based estimation.
+IMUs are sensors with higher refresh rates than Simultaneous Localization and Mapping (SLAM) based algorithms, such as optical or Light Detection and Ranging (LIDAR) SLAM.
 This makes them suitable for interpolating the more precise and absolute position reference points of these slower algorithms.
 The limited computational power of space-qualified CPUs further imposes the need for good orientation estimation without complex algorithms like Kalman filters.
-Lastly, the geometry of a spherical robot opens the possibility for the not often used position estimation by IMUs. \\
+Lastly, the geometry of a spherical robot opens the possibility for the not often used position estimation by IMUs.
+
 This paper introduces an IMU-based orientation and position algorithm for spherical robots with limited computational power and the need for real-time estimations.
@@
 \paragraph*{Slip and Sliding}
-We assume the sphere has both slip (the sphere rotates, but there is no translation) and sliding (the sphere has translation without rotation) in a limited, non-permanent way.
+We assume the sphere has both slip, i.e., the sphere rotates, but there is no translation, and sliding, i.e., the sphere has translation without rotation in a limited, non-permanent way.
 The limitation is the assumption that sliding and slipping reduce the amount of translation for a given rotation and vice versa, however, not for complete absence.
@@
 The sphere from the DAEDALUS example has two completely different locomotion methods; therefore, we will not take locomotion itself into account of pose estimation.
-Also, the same locomotion approach can lead to heavily different behavior with the same input to the controller.
-\cite{c2}
+Also, the same locomotion approach leads to heavily different behavior with the same input to the controller~\cite{c2}.
 \paragraph*{Classification of Data}
@@
 \subsection{Madgwick Filter}
-The Madgwick filter is a widely used filter for orientation estimation with IMUs \cite{c1}.
+The Madgwick filter is a widely used filter for orientation estimation with IMUs~\cite{c1}.
 It is efficient and easy to calculate but does not consider the physics of the motion.
@@
 This does not meet the desired behavior as described in the prelimitations section (\ref{AssumptionsAndPrelimitations}).
-The behavior is shown in comparison to the complementary filter in Figure \ref{madgVsComp1}.
+The behavior is shown in comparison to the complementary filter in Figure~\ref{madgVsComp1}.
 This behavior is most prominent during standstill as the Madgwick filter was designed for movement.
@@
 The basic concept is to have the main part of a value dominated by a gyroscope but, over time, a slow drift towards the measurement of the accelerometer.
-E.g. for pitch, this would result in 
+E.g., for pitch, this would result in 
 \begin{equation}
@@
 With an acceleration in one direction, this leads to a miscalculation of the gravitational vector and, therefore, a false orientation estimation.
-Simple orientation estimations of quadrocopters often just change the $\alpha$ gain to 0 while moving to avoid the miscalculation of the gravity vector. \cite{complementary}
+Simple orientation estimations of quadrocopters often just change the $\alpha$ gain to 0 while moving to avoid the miscalculation of the gravity vector \cite{complementary}.
 
@@
 Therefore there are very few approaches targeting this way of position calculating \cite{c3}.
-The very impressive \cite{Valencia2017ASM} also uses IMUs for 6D pose estimation.
+The very impressive work in \cite{Valencia2017ASM} also uses IMUs for 6D pose estimation.
 They follow a pre-specified trajectory with a manipulator and manage to estimate its pose within 10\% of the ground truth.
@@
 The orientation estimation is the combination of the Madgwick and a complementary filter.
-The weaknesses and strengths of both described in Section \ref{sec:SOA} can be combined. 
-Therefore the general idea is to distinguish between motion and standstill.
+The weaknesses and strengths of both described in Section \ref{sec:SOA} are combined. 
+The general idea is to distinguish between motion and standstill.
 In motion, the Madgwick filter is used, and without motion, the complementary filter.
@@
 As both filters are applied with weights, this is done by an allocation of a given gain towards one or the other filter depending on the likelihood the sphere is moving or standing still. 
-Therefore this requires some adapting mechanism and an estimation of how likely the sphere is to rotate and hence translate.
+Therefore, this requires some adapting mechanism and an estimation of how likely the sphere is to rotate and hence translate.
 Both filters share the same meaning of the gain, describing the influence of the accelerometer on the orientation.
@@
 
-\begin{figure}[thpb]
+\begin{figure}[hb]
 \includegraphics[width=\linewidth]{./graphics/imuJasperChartSimliefiedBIG.png}
@@
  The quaternion slerp works over the representation of two quaternions on a sphere. The transition from one quaternion to the other is now not done directly but always over the surface of the sphere.
- This leads to the same handling of quaternions as the Madgwick-filter without endangering robustness due to using two different quaternions with very different values representing nearly the same position.\\
+ This leads to the same handling of quaternions as the Madgwick-filter without endangering robustness due to using two different quaternions with very different values representing nearly the same position.
+
 The transition between the two filters is done by shifting a fixed factor among the $\alpha$-gain of the complementary filter and the gain factor of the Madgwick filter.
@@
 \left\{
-\begin{align}
+\begin{aligned}
 0& & g_k \leq 0.1  \\ 
@@
  1& & g_k \geq 2.1
-\end{align}
+\end{aligned}
 \label{eq:factors}
-\right\}
+\right.
 \end{equation}
 
-\begin{equation}
-\beta = \max(f_x,f_y,f_z)\label{eq:weight_autogain}
-\end{equation}
-Then the Madgwick gain $\gamma$ for a given autogain $\Theta$ is calculated by 
-\begin{equation}
-\gamma = \Theta \cdot \beta
-\end{equation}
-The complementary filter gain $\alpha$ is calculated by
-\begin{equation}
-\alpha= \Theta \cdot \theta \cdot(1-\beta)
-\end{equation}
-So the Madgwick gain has a maximum of $\Theta$ and $\alpha$ a maximum of $\theta\cdot \Theta$. This ratio $\theta$ can be adapted to the specific needs.
+\begin{align}
+\beta &= \max(f_x,f_y,f_z)\label{eq:weight_autogain}\\
+\intertext{Then the Madgwick gain $\gamma$ for a given autogain $\Theta$ is calculated by} 
+\gamma &= \Theta \cdot \beta \\
+\intertext{The complementary filter gain $\alpha$ is calculated by}
+\alpha &= \Theta \cdot \theta \cdot(1-\beta).
+\end{align}
+Thus, the Madgwick gain has a maximum of $\Theta$ and $\alpha$ a maximum of $\theta\cdot \Theta$. This ratio $\theta$ can be adapted to the specific needs.
 We used $\theta = 0.1$ because of the the two often used standard values $\gamma = 0.2$ and $\alpha= 0.02$, which have the ration $0.1$.
@@
 g'_z
-\end{bmatrix} dt\right)
+\end{bmatrix} \right)\, dt
 \end{equation}
@@
 Therefore, the factors from the orientation step, representing the likelihood of rotation, are used as limiting factors.
-If one of the three factors is 1, this means we let translation in this direction happen.
-When there is no rotation, it will hold the velocity to 0.
+If one of the three factors is 1, this means there is translation in this direction.
+When there is no rotation, it will set the velocity to 0.
 During fast movement, this would allow exponential error integration. 
@@
 The rotation into the world frame is computed as:
-\begin{equation}
+\begin{align}
 \begin{bmatrix}
@@
 \end{bmatrix}
-= \M R^T \cdot \begin{bmatrix}
+&= \M R^T \cdot \begin{bmatrix}
 v_{x\text{Acc}} \\ v_{y\text{Acc}} \\v_{z\text{Acc}} 
-\end{bmatrix}
-\end{equation}.
-The next step requires the velocity calculated by the rotation.
-Therefore the rotations (direct values from the gyro) are rotated into the world frame.
-\begin{equation}
+\end{bmatrix}.
+\intertext{The next step requires the velocity calculated by the rotation.
+Therefore the rotations (direct values from the gyro) are rotated into the world frame, i.e.,}
 \begin{bmatrix}
 \omega_{x\text{World}} \\ \omega_{y\text{World}} \\(\omega_{z\text{World}}) 
-\end{bmatrix} = 
+\end{bmatrix} &= 
  \M R^T \cdot
@@
 g_x \\ g_y\\ g_z
-\end{bmatrix} 
-\end{equation}
+\end{bmatrix}. 
+\end{align}
 The rotation around the world $z$-axis is not used to calculate the velocity.

### 2021-05-28 mfi21/ieeeconf/root.tex
Andreas' changes

@@
  To avoid large errors and unlimited double error integration, $v_{z\text{AccWorld}} $ is limited to the mean of $v_{xz\text{GyroWorld}} $ and $ v_{yz\text{GyroWorld}}$.
-With the Pythagorean theorem the velocity in x and y is calculated accepting the velocity in z-direction, which can not be derived by the velocity from gyro:
+With the Pythagorean theorem the velocity in $x$ and $y$ is calculated accepting the velocity in $z$-direction, which can not be derived by the velocity from gyro:
  \begin{equation}
@@
 \end{equation}.
-Depending on the practical circumstances and quality of measurement, the factor $\beta$ can be chosen, that the position relies more on the accelerometer calculated velocity or the gyro.
+Depending on the practical circumstances and quality of measurement, the factor $\beta$ is chosen, that the position relies more on the accelerometer calculated velocity or the gyro.
 There should always be a mixture of both if there is slipping and sliding.
@@
 The IMU was manually rotated in arbitrary directions, and the orientation plotted using the Madgwick filter and the Autogain filter.
-As expected, the overall behavior of the presented algorithm and the Madgwick filter is very similar in the dynamic case, as shown in Figure \ref{quaternion1}.
+As expected, the overall behavior of the presented algorithm and the Madgwick filter is very similar in the dynamic case, as shown in Figure~\ref{quaternion1}.
 \begin{figure}
@@
 The primary evaluation for the position focuses on slip- and slide-behavior.
-Here the plate with the IMU(s) is fixed inside a sphere with a diameter of 29 cm and rolled along an L-shaped track of one meter by one meter.
+Here the plate with the IMU(s) is fixed inside a sphere with a diameter of 29 cm and rolled along an ``L''-shaped track of one meter by one meter.
 Slip behavior was manually provoked.
 Figure \ref{LTest} shows the resulting trajectories.
-Pure rotation integration leads to an enlarged L by about 50 percent, as the manually provoked rotation is just integrated and therefore interpreted as linear motion.
-The fusion between the accelerometer and gyroscope interprets an L with dimensions much closer to the original.
+Pure rotation integration leads to an enlarged ``L'' by about 50 percent, as the manually provoked rotation is just integrated and therefore interpreted as linear motion.
+The fusion between the accelerometer and gyroscope interprets an ``L'' with dimensions much closer to the original.
 The estimation leaves room for improvement, as in one direction, it misses the actual length about -10 percent, and in the other direction overestimates it by 10 percent.

### 2021-05-29 mfi21/ieeeconf/root.tex
small changes

@@
 \usepackage{amsmath}
+\usepackage{hyperref}
 %\usepackage{enumitem}
@@
 Spherical robots are a relatively narrow field of robotics. Still, they could be useful in situations where the measurement equipment needs to be protected against harsh environments.
-As an example, the 2021 CDF study LunarCaves by ESA about the feasibility of a spherical robot, called DAEDALUS, for exploration of lunar-lava tubes \cite{c2} shows the need for orientation and position estimation of spherical robots with the limitations on resources common for space-applications.
+As an example, the 2021 CDF study LunarCaves by ESA about the feasibility of a spherical robot, called DAEDALUS, for exploration of lunar-lava tubes \cite{rossi2021daedalus} shows the need for orientation and position estimation of spherical robots with the limitations on resources common for space-applications.
 Although a purely Inertial Measurement Unit (IMU)-based estimation is not recommended for various reasons, the lack of an absolute reference among others, IMU-based pose estimation are used as part of an overall multi-sensor-fusion-based estimation.
@@
 The sphere from the DAEDALUS example has two completely different locomotion methods; therefore, we will not take locomotion itself into account of pose estimation.
-Also, the same locomotion approach leads to heavily different behavior with the same input to the controller~\cite{c2}.
+Also, the same locomotion approach leads to heavily different behavior with the same input to the controller \cite{rossi2021daedalus}.
 \paragraph*{Classification of Data}
@@
 \subsection{Kalman Filter}
-The Kalman Filter or the more suitable Extended Kalman Filter (EKF) uses state prediction and merges with sensor data for orientation estimation.
+The Kalman Filter or the more suitable Extended Kalman Filter (EKF) uses state prediction and merges with sensor data for orientation estimation \cite{Kalman1960}.
 For the defined limitations, the Kalman Filter is not suitable.
@@
 \subsection{Madgwick Filter}
-The Madgwick filter is a widely used filter for orientation estimation with IMUs~\cite{c1}.
+The Madgwick filter is a widely used filter for orientation estimation with IMUs \cite{madgwick2010}.
 It is efficient and easy to calculate but does not consider the physics of the motion.
@@
 With an acceleration in one direction, this leads to a miscalculation of the gravitational vector and, therefore, a false orientation estimation.
-Simple orientation estimations of quadrocopters often just change the $\alpha$ gain to 0 while moving to avoid the miscalculation of the gravity vector \cite{complementary}.
+Simple orientation estimations of quadrocopters often just change the $\alpha$ gain to 0 while moving to avoid the miscalculation of the gravity vector \cite{mahony2005complementary}.
 
@@
 Position estimation purely by IMU is not recommended as the double integration of the accelerometer measurements leads to exponential error summation.
-Therefore there are very few approaches targeting this way of position calculating \cite{c3}.
-The very impressive work in \cite{Valencia2017ASM} also uses IMUs for 6D pose estimation.
+Therefore there are very few approaches targeting this way of position calculating \cite{kok2017IMU}.
+The very impressive work in \cite{valencia2017simpleIMU} also uses IMUs for 6D pose estimation.
 They follow a pre-specified trajectory with a manipulator and manage to estimate its pose within 10\% of the ground truth.
@@
 not used, as it implies a transition of each value independently.
-Hence, the RPY-orientation is transformed to a quaternion, and then a quaternion slerp is performed \cite{c4}.
+Hence, the RPY-orientation is transformed to a quaternion, and then a quaternion slerp is performed \cite{dam1998quaternions}.
  The quaternion slerp works over the representation of two quaternions on a sphere. The transition from one quaternion to the other is now not done directly but always over the surface of the sphere.
@@
 The IMU was manually rotated in arbitrary directions, and the orientation plotted using the Madgwick filter and the Autogain filter.
-As expected, the overall behavior of the presented algorithm and the Madgwick filter is very similar in the dynamic case, as shown in Figure~\ref{quaternion1}.
+As expected, the overall behavior of the presented algorithm and the Madgwick filter is very similar in the dynamic case, as shown in Figure \ref{quaternion1}.
 \begin{figure}
@@
 For external reference, we used a precise optical tracking system. 
-The definition and proposed calculation of the distance of 2 quaternions of \cite{kuffner} was used.
+The definition and proposed calculation of the distance of 2 quaternions of \cite{kuffner2004effectiveDistance} was used.
 It describes the direct angle between two quaternions.
@@
 The transformation of the orientation by the optical tracking system and the IMU-based algorithms yield two unknown transformations, the relative orientation of the coordinate system of IMU to optical coordinate system and how the axes are interpreted on the tracked structure itself.
-Therefore, the algorithm proposed in \cite{dornaika} was used to solve these two unknown transformations on basis of the resulting data, to match the orientations of optical and IMU-based data.
+Therefore, the algorithm proposed in \cite{dornaika1998calibration} was used to solve these two unknown transformations on basis of the resulting data, to match the orientations of optical and IMU-based data.
 As this matching of the optical system happens on basis of one of the two IMU orientation data-sets, the comparison of absolute differences may be misleading, as the transformations were matched on basis of the autogain filter.
@@
 \includegraphics[width=\linewidth]{./graphics/distance.eps}
-\caption{Distance of the proposed and the Madgwick algorithm to the optical tracking system, using the distance definiton of \cite{kuffner}}
+\caption{Distance of the proposed and the Madgwick algorithm to the optical tracking system, using the distance definiton of \cite{kuffner2004effectiveDistance}}
 \label{distance}
@@
 \includegraphics[width=\linewidth]{./graphics/diffdistance.eps}
-\caption{Derivative of the distance of the proposed and the Madgwick algorithm to the optical tracking system, using the distance definiton of \cite{kuffner}}
+\caption{Derivative of the distance of the proposed and the Madgwick algorithm to the optical tracking system, using the distance definiton of \cite{kuffner2004effectiveDistance}}
 \label{diffdistance}
@@
 In further experiments, a quantitative Orientation comparison with a synchronized optical system needs to be performed.
+Also as further work, all puzzle pieces, this filter, a locomotion system, a scanning system,  need to be combined to an overall working prototypes.
 The position estimation, as well as the combination of the integration by rotation of the gyroscope and the linear acceleration of the accelerometer, shows the capability to overcome the problem of the missing absolute reference of IMUs partly, which mainly improves the measurements in the ground plane with slip behavior present.
@@
 
-\begin{thebibliography}{99}
-\bibitem{c2} Rossi, A.P.
-et al., "DAEDALUS - Descent And Exploration in Deep Autonomy of Lava Underground Structures", Wuerzburger Forschungsberichte in Robotik und Telematik (21), 978-3-945459-33-1, (2021).
-\bibitem{c1} Madgwick, Sebastian O. H., "An efficient orientation filter for inertial and inertial / magnetic sensor arrays", (2010).
-\bibitem{complementary}Mahony, R.
-; Hamel, T.; Pflimlin, J.-M.,"Complementary filter design on the special orthogonal group SO(3)", Proceedings of the 44th IEEE Conference on Decision and Control, (2005).
-\bibitem{c3} Kok, Manon; Hol, Jeroen D.;  Schoen,Thomas B.,  "Using Inertial Sensors for Position and Orientation Estimation", Foundations and Trends in Signal Processing: Vol. 11.(2017).
-\bibitem{Valencia2017ASM} Botero Valencia, J.-S.
-; Garc{\'i}a, M.
-R.
-; Ceballos, J.
-V., "A simple method to estimate the trajectory of a low cost mobile robotic platform using an IMU", International Journal on Interactive Design and Manufacturing (ijidem), (2017).
-\bibitem{c4}Dam, Erik B.; Koch, Martin;  Lillholm, Martin, "Quaternions, Interpolation and Animation", (1998).
-\bibitem{FastInvSqrt}"Origin of Quake3's Fast InvSqrt()", beyond3d.com/content/articles/8/, Accessed: 2021-05-01.
-\bibitem {kuffner}Kuffner, James, "Effective sampling and distance metrics for 3D rigid body path planning", Proceedings - IEEE International Conference on Robotics and Automation, pp.
-3993 - 3998 Vol.4, (2004).
-\bibitem{dornaika}Dornaika, F.; Horaud, R., Simultaneous robot-world and hand-eye calibration, IEEE Transactions on Robotics and Automation 14(4), pp. 617-622, (1998).
-
-\end{thebibliography}
+\bibliographystyle{unsrt}
+\bibliography{bibliographyJasper}
 

### 2021-05-29 mfi21/ieeeconf/root.tex
added citation suggested by andreas
small typos

@@
 In motion, the Madgwick filter is used, and without motion, the complementary filter.
+This approach of distinguishing the evaluation algorithm depending on the state of motion of a robot has been done before, like in \cite{Hertzberg2012}.
 The transition needs to be in a differentiable way, as this is one of our limitations.  
@@
 In further experiments, a quantitative Orientation comparison with a synchronized optical system needs to be performed.
-Also as further work, all puzzle pieces, this filter, a locomotion system, a scanning system,  need to be combined to an overall working prototypes.
+Also, as further work, all puzzle pieces, this filter, a locomotion system, a scanning system need to be combined to an overall working prototype.
+
 The position estimation, as well as the combination of the integration by rotation of the gyroscope and the linear acceleration of the accelerometer, shows the capability to overcome the problem of the missing absolute reference of IMUs partly, which mainly improves the measurements in the ground plane with slip behavior present.

### 2021-05-30 mfi21/ieeeconf/root.tex
added prototype pictures

@@
 However, the deceleration of the sphere was still recognized while rotation existed, and therefore, the sphere miscalculated a couple of centimeters.
+To test the algorithm in a real spherical robot with working laserscanners, the prototype shown in Figure \ref{fig:prototype}
+was build.
+The Plate in the middle is equal to the one used to perform the position evaluation.
+The IMUs and the micrococontroller are at the exact same position inside the sphere.
+The Livox Lidar-Mid-100 laser-scanner was mounted under the plate and the battery above it.
+Due to the lack of a locomotion-system, the sphere needs to be rolled externally.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{./graphics/photoReal1.jpeg}
+\includegraphics[width=\linewidth]{./graphics/photoReal2.jpeg}
+\caption{Prototype to test algorithm in envisaged technical environment.}
+\label{fig:prototype}
+\end{figure}
+
+%TODO Fabian, put input of laserscanns here
 

### 2021-05-31 mfi21/ieeeconf/root.tex
changed position of picture and therefore text. Added caption

--- 2021-05-29 mfi21/ieeeconf/root.tex
@@
 Also, the algorithm does not rely on one specific locomotion approach, but it assumes a rotation of the IMU together with at least the shell of the robot, if not the whole robot. 
+Such as the prototype developed at the computer science chair VII of the University of Wuerzburg,  depicted in Figure \ref{fig:prototype}, which is inspired by previous work such as Daedalus \cite{rossi2021daedalus}.
+The evaluations made in this paper were done with the setup of the IMUs of this prototype.
+The Livox Mid-100 was used as laser scanner.
+Due to the lack of a locomotion system, the sphere needs to be rolled externally. 
+The evaluation will also show and discuss the resulting point clouds of this prototype with the presented pose estimation.
+\begin{figure}[b]
+\includegraphics[width=0.484\linewidth]{./graphics/photoReal1.jpeg}
+\includegraphics[width=0.50\linewidth]{./graphics/photoReal2.jpeg}
+\caption{Prototype to test the algorithm in an envisaged technical environment.
+The main payload is the Livox Mid-100 laser-scanner.
+For pose-estimation, three IMUs of the manufacturer Phidget are placed on the middle plate and a Raspberry Pi 4 for the calculations.
+On the top are two batteries, and on the bottom one voltage stabilizer and breakout box of the laser-scanner.
+}
+\label{fig:prototype}
+\end{figure}
 
@@
 
+%TODO Fabian, put input of laserscanns here
+
 \subsection{Conclusion}

### 2021-05-31 mfi21/ieeeconf/root.tex
changed photo
now here comes the actual input for 3d mapping

@@
 \begin{figure}[b]
-\includegraphics[width=0.484\linewidth]{./graphics/photoReal1.jpeg}
-\includegraphics[width=0.50\linewidth]{./graphics/photoReal2.jpeg}
+\includegraphics[width=0.49\linewidth]{./graphics/photoReal1.jpeg}
+\includegraphics[width=0.50\linewidth]{./graphics/photoReal3.jpeg}
 \caption{Prototype to test the algorithm in an envisaged technical environment.
@@
 
-%TODO Fabian, put input of laserscanns here
+\subsection{Impact on 3D mapping}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\linewidth]{./graphics/jasperhome1.png} \\
+\vspace{0.2cm}
+\includegraphics[width=\linewidth]{./graphics/jasperhome1madw.png}
+\caption{(Top) Resulting 3D map using the presented algorithm for pose filtering.
+(Bottom) Resulting 3D map using Madgwick + double integrator for pose filtering.
+Both images were shot from the exact same point of view.
+}
+\label{fig:mapping}
+\end{center}
+\end{figure}
+
+One major design driver for the presented algorithm is the suitability for spherical SLAM purposes.
+Therefore, a pose estimate is prefered to be slightly wrong, if it is more continuous, i.e. less noisy and with less jumping values, than the exact pose would be.
+
+To test the impact of the filter on 3D mapping, we used the prototype in Figure \ref{fig:prototype}.
+After 2 s of initial standstill, the sphere rolled aprox. 4 m within 5 s in an indoor environment from one corner of the room to the opposing one.
+We extracted the laserscan data, as well as the filtered pose using the presented algorithm.
+Without further pre- or post-processing, the pose data is combined with the laserscan data to create a 3D map.
+Figure \ref{fig:mapping} compares the result with another 3D map, which is generated the exact same way, but using a different pose filter.
+Namely a combination of the Madgwick filter for orientation and a double integration of the accelerometer mean for position.
+The result of this filter clearly got improved by the presented algorithm as it provides a better pose estimation.
+The more detailed part of the map, as seen in the bottom left corner of the top image of Figure \ref{fig:mapping}, corresponds to the initial standstill.
+It is more detailed than the rest of the map because of the non-repeating flower-shaped scanning pattern, where point density increases with time.
+Note that the same detailed part is also present in the bottom image.
+However, it is out of view (at the bottom right corner) as the Madgwick filter misses the initial pose estimation by a lot. 
+The presented algorithm also performs better when the sphere is moving, as the walls and the ceiling are more recognizable, which is favourable for many SLAM algorithms.
 

### 2021-05-31 mfi21/ieeeconf/root.tex
top/bottom -> above/below

@@
 \includegraphics[width=\linewidth]{./graphics/jasperhome1madw.png}
-\caption{(Top) Resulting 3D map using the presented algorithm for pose filtering.
-(Bottom) Resulting 3D map using Madgwick + double integrator for pose filtering.
+\caption{Above: resulting 3D map using the presented algorithm for pose filtering.
+Below: resulting 3D map using Madgwick + double integrator for pose filtering.
 Both images were shot from the exact same point of view.
@@
 The result of this filter clearly got improved by the presented algorithm as it provides a better pose estimation.
-The more detailed part of the map, as seen in the bottom left corner of the top image of Figure \ref{fig:mapping}, corresponds to the initial standstill.
+The more detailed part of the map, as seen in the bottom left corner of the above image of Figure \ref{fig:mapping}, corresponds to the initial standstill.
 It is more detailed than the rest of the map because of the non-repeating flower-shaped scanning pattern, where point density increases with time.
-Note that the same detailed part is also present in the bottom image.
+Note that the same detailed part is also present in the below image.
 However, it is out of view (at the bottom right corner) as the Madgwick filter misses the initial pose estimation by a lot. 

### 2021-05-31 mfi21/ieeeconf/root.tex
minor improvements

@@
 Without further pre- or post-processing, the pose data is combined with the laserscan data to create a 3D map.
-Figure \ref{fig:mapping} compares the result with another 3D map, which is generated the exact same way, but using a different pose filter.
-Namely a combination of the Madgwick filter for orientation and a double integration of the accelerometer mean for position.
-The result of this filter clearly got improved by the presented algorithm as it provides a better pose estimation.
+Figure \ref{fig:mapping} compares the result with another 3D map, which is generated in the same fashion.
+However, a combination of the Madgwick filter for orientation and a double integration of the accelerometer mean for position, is used.
+The presented filter clearly improves the resulting map compared to the Madgwick/integrator approach.
+In particular, the walls and the ceiling are more recognizable, which is favourable for many SLAM algorithms.
+Another improvement is that points were no longer sensed below the ground, i.e. the plane on which the sphere was moving on.
 The more detailed part of the map, as seen in the bottom left corner of the above image of Figure \ref{fig:mapping}, corresponds to the initial standstill.
 It is more detailed than the rest of the map because of the non-repeating flower-shaped scanning pattern, where point density increases with time.
-Note that the same detailed part is also present in the below image.
+Note that the same detailed part is also present in the image below.
 However, it is out of view (at the bottom right corner) as the Madgwick filter misses the initial pose estimation by a lot.
-The presented algorithm also performs better when the sphere is moving, as the walls and the ceiling are more recognizable, which is favourable for many SLAM algorithms.
 

### 2021-05-31 mfi21/ieeeconf/root.tex
small changes

@@
 Without further pre- or post-processing, the pose data is combined with the laserscan data to create a 3D map.
-Figure \ref{fig:mapping} compares the result with another 3D map, which is generated in the same fashion.
-However, a combination of the Madgwick filter for orientation and a double integration of the accelerometer mean for position, is used.
-The presented filter clearly improves the resulting map compared to the Madgwick/integrator approach.
+Figure \ref{fig:mapping} compares the result with another 3D map, using the in the same run recorded pose-estimation by the Madgwick filter and simple rotation integration.
+Therefore the same laserscan-data is mapped with different pose estimations.
+The presented filter improves the resulting map compared to the Madgwick/integrator approach.
 In particular, the walls and the ceiling are more recognizable, which is favourable for many SLAM algorithms.

### 2021-06-22 Latex/rebuttal.txt
rebuttal

 
> Manuscript Number: OPHOTO-D-21-00020  
> Towards Spherical Robots For Mobile Mapping In Human Made
> Environments
>
> Dear Mr. Bredenbeck,     
> 
> Thank you for submitting your manuscript to ISPRS Open Journal of
> Photogrammetry and Remote Sensing.
>

> I have completed my evaluation of your manuscript. The reviewers
> recommend reconsideration of your manuscript following revision. I
> invite you to resubmit your manuscript after addressing the comments
> below. Please resubmit your revised manuscript by Jul 13, 2021.  
> When revising your manuscript, please consider all issues mentioned
> in the reviewers' comments carefully: please outline in a cover
> letter every change made in response to their comments and provide
> suitable rebuttals for any comments not addressed. Please note that
> your revised submission may need to be re-reviewed.     
> 
> To submit your revised manuscript, please log in as an author at
> https://www.editorialmanager.com/ophoto/, and navigate to the
> "Submissions Needing Revision" folder under the Author Main Menu.
> 
> ISPRS Open Journal of Photogrammetry and Remote Sensing values your
> contribution and I look forward to receiving your revised
> manuscript.
>   
> Kind regards,     
> 
> George Vosselman   
> Editor-in-Chief   
> ISPRS Open Journal of Photogrammetry and Remote Sensing    
>  
> Editor and Reviewer Comments:     
> 
> Editor:
> As you will see below both reviewers appreciate your work and don't
> have concerns regarding the methodology or results. Reviewers, in
> particular Reviewer 2, however, do have a large number of
> recommendations to improve the readability of your manuscript. Given
> the number of recommendations, I ask for a major revision. The
> recommendations should, however, be easy to implement.
>  
>   
> Reviewer 1: The manuscript presents the concept of a spherical
> mobile mapping system, and a data registration method utilizing the
> geometric constraints available in human-made environments,
> e.g. planar surfaces.
> 
> The manuscript has novel content and is well written, and can be
> accepted after minor revision. My comments and suggestions to be
> considered by the authors are as follows:
> 
> Lines 27-31: Spherical shell can protect the sensors inside, but
> this does not prevent the external disturbances such as dust or
> other impurity staining the sphere itself, affecting the performance
> of the system. So in principle this does not change the situation
> from using protective covers on "traditional" mobile mapping
> systems.
> 
> Lines 111-116: Modern kinematic scanning systems collect point cloud
> data with high speed, so also kinematic point clouds are often dense
> enough for robust plane determination. And on the other hand, sparse
> point clouds are not optimal for point-to-point matching due to the
> long distances between adjacent points.
> 
> Lines 244-246: The Livox ranging precision is given twice in this
> sentence.
> 
> Figure 4: The x-scale is given in figure caption, but I would
> suggest adding it to the figure also.
> 
> Conclusion: Whereas the title, Abstract and Introduction concentrate
> mostly on the spherical mobile mapping, Conclusion deals almost
> entirely with the topic of registration, thus feeling somewhat
> separated from the rest of the manuscript. Reading the Conclusion,
> one could not expect that the title of the manuscript is "Towards
> Spherical Robots For Mobile Mapping In Human Made Environments".
> 
> 
> 
> 
> 
> Reviewer 2: The authors have written a very interesting paper on the
> topic of mobile mapping with a spherical robot in human-made
> environments. In a rather slim paper and pleasantly concise form,
> they present their approach and show first results based on
> simulated and real measurement data. The reading flow is mostly
> pleasant and no irrelevant information is given. Conversely, the
> authors leave out important details in several places, which would
> be very helpful for the reader's understanding. In the following,
> there are several points for improvement and critical questions that
> should be taken into account by the authors. Due to the large number
> and the importance of central points, I recommend a major revision.
> 
> - A short outline at the end of the introduction is missing and
> would give the reader a better overview of the further content and
> should therefore be added.
> 
> - In the entire article, no reference is made to the targeted or
> required accuracy potential with regard to the quality of the pose
> determination and the final point cloud, which in my view requires
> at least a rough classification.
> 
> - For the short state of the art of existing mobile systems, which
> is mainly based on the authors' own systems, in my opinion, at least
> the NavVis VLX system could be added, which in comparison is newer
> and as a human-operated system for mapping is one of the benchmarks.
> 
> - The authors mention unknown and dangerous environments as a
> possible field of application, in which the advantages of the
> spherical system should come into play. However, only one space
> application is mentioned as a specific example, which may be valid
> in itself, but does not fit the topic of the paper very
> well. Application scenarios should therefore be supplemented by
> terrestrial areas and in particular by human-made environments,
> which better motivate the future use and justification of such
> systems for the reader.
> 
> - The advantages of a spherical mobile platforms are discussed, but
> nothing is said about possible disadvantages or
> limitations. Moreover, I wonder whether the "unique advantages"
> mentioned in the abstract are really so remarkable. For example,
> what about the recording constellation, which in my view is
> relatively unfavourable on the ground and leads to restricted fields
> of view and weak angles of incidence when mapping the
> ground. Another advantage stated by the authors is that the shell
> protects the sensors, which is reasonable. However, this protection
> is only necessary because you are moving directly on the ground. In
> addition, can dust and dirt accumulate on the outer shell and
> influence the acquisition?
> 
> - Section 2.1 discusses existing laser scanner-based SLAM
> approaches, concluding with the statement (page 2, column 1, line
> 30-33) that all of these approaches do not include the structural
> properties of human-made environments in the calculations. This
> statement is not true. There are a variety of SLAM approaches
> (e.g. Jung et al. 2015; Nguyen et al.  2006), but also other
> applications with recursive state estimation, which make use of this
> additional information in the context of georeferencing. This should
> be investigated and reworked in more detail.
> Jung, J., Yoon, S., Ju, S., and Heo, J. (2015). Development of Kinematic
> 3D Laser Scanning System for Indoor Mapping and As-Built BIM Using
> Constrained SLAM. In: Sensors (Basel, Switzerland), 15(10), DOI:
> https://doi.org/10.3390/s151026430
> Nguyen, V., Harati, A., Martinelli, A., Siegwart, R., and Tomatis, N.
> (2006). Orthogonal SLAM: A Step Toward Lightweight Indoor Autonomous
> Navigation. In: IEEE/RSJ International Conference on Intelligent Robots
> and Systems, DOI: https://doi.org/10.1109/IROS.2006.282527
> 
> - The introduction in Section 3 should be revised as it is not very
> intuitive for the reader and something about the objectives and
> motivation for this section is missing. However, this is also
> directly related to the general lack of outline beforehand.
> 
> - In section 3, the authors introduce their own algorithm for
> registration, but they do not give any further information about the
> temporal context in which this takes place. This is probably done in
> the context of each epoch, although it is also not clear how an
> epoch is defined. Is it a single full rotation of the laser scanner
> or that of the sphere? And does the registration always take place
> only locally in relation from the current epoch to the last epoch or
> with a window approach or even globally?
> 
> - Section 3.2 deals with the extraction of planes, but the following
> questions remain unclear: Are the planes always searched for only in
> the context of the observational data of the current epoch or with
> regard to the already registered global point clouds? How many
> planes are needed at all and what demands are made on their spatial
> distribution? The abstract mentioned parallel and perpendicular
> planes, but this is not discussed further here. Are only these used
> and if so, why? Furthermore, what about planes that are not
> necessarily parallel or perpendicular (e.g. doors, cupboards)? How
> is the flatness of a plane checked? How are outliers dealt with,
> especially with regard to existing measurement and motion noise?
> 
> - Throughout Section 3, numerous variables are not introduced and
> are not explained in more detail, so that a more precise
> interpretation of the individual equations is unfortunately very
> difficult.
> 
> - What orders of magnitude were finally used for the range noise
> model in Section 4.2 and why was this not also done for the angular
> resolution?
> 
> - In the context of the real experiment in Section 5, a major
> simplification is made. Even if this simplification is certainly
> suitable for the beginning, this special case comes quite surprising
> here...especially in comparison to the simulation and the previously
> described topic. Have other measurements not yet been carried out or
> what is the background? At least this simplification should be
> mentioned beforehand, otherwise the reader will have false
> expectations.
> 
> - Wouldn't it also be interesting to compare the estimated pose
> parameters with the nominal values from the simulation within the
> framework of Section 6.1? So far, the focus has only been on the
> point cloud.
> 
> - Why is there no qualitative validation for the results of the real
> data in Section 6.2? A reference scan based on a terrestrial laser
> scanner would be a possibility or otherwise based on the flatness of
> known planes in the registered point cloud?
> 
> - In the Conclusion, the parallelism and perpendicularity within the
> environment is mentioned again, although this did not play a role in
> the main part of the paper. Therefore, the question arises why do
> the planes have to be parallel or perpendicular at all and why is
> this mentioned at the beginning? This is not clear, since no
> additional constraints on this additional information have been
> introduced so far. There are also already other methods available
> that take this idea of geometric constraints for georeferencing into
> account.
> 
> Further open questions/comments are:
> Page 1, Line 15: Double use of "protect", which should be replaced
> by another word.
> Page 1, Line 21: The link for the "second option" to the use of an
> IMU is not necessarily obvious.
> Page 1, Column 2, Line 52-56: Complicated sentence which should be
> reformulated.
> Page 2, Column 2, Line 15: The term "long stop" is not to be
> interpreted clearly in this context. How is "long" defined in this
> sense in terms of time or space?
> Page 3, Column 1, Line 6-8: A specific reference should be given for
> the first half of the sentence, as reference 19 probably only refers
> to the second part.
> Page 3, Column 1, Line 23-26: Not all variables were introduced
> accurately
> Page 3, Column 1, Line 33-34: Selection of references for this?
> Page 3, Column 1, Line 34-35: But what demands are made on the
> assignment? This is unclear to the reader, also in connection with
> the general quality requirements of the whole approach.
> Page 3, Column 1, Line 38: How can the threshold value be specified,
> since it is probably not identical for all applications and
> environments? Later, there is also no indication of the choice of
> the value in the own simulated and real data sets.
> Page 3, Column 1, Line 50: What does the index k describe and what
> is the difference to the index i from Section 3.2?
> Page 3, Equation (1): Again, some variables are not specified. For
> example, the source of tx ty tz is unclear.
> Page 3, Equation (3): Equation unclear regarding the use of the 2 in
> the exponent as well as the index.
> Page 3, Equations (4)-(5): Here I don't understand how to get to the
> elements of the normal vector through the transformation. Moreover,
> these individual elements were not introduced in advance.
> Page 3, Equation (9): The variable alpha is not described and
> explained
> Page 4, Equation (10): Description of the individual variables is
> missing and, moreover, rho has already been used as the distance
> parameter of the Hesse-normal form
> Page 4, Equation (12): Is epsilon here the same variable as the
> threshold in Section 3.3? Probably not.
> Page 4, Column 2, Line 15: "some dimensions": Which ones and why
> exactly are they chosen? Are they the same ones all the time?
> Page 4, Column 2, Line 20-25: Reference is missing
> Page 4, Column 2, Line 55: Doubled information regarding the
> distance measurement precision
> Page 5, Column 1, Equation (18): Delta t is not introduced and there
> is no indication of how large the time steps are in the simulation.
> Page 5, Column 2, Line 14: Should it be "out" or "our"?
> Page 5, Column 2, Line 27-28: What are the specifications of the
> used IMU with regard to its uncertainty budget?
> Page 6, Figure 1: How are these values to be interpreted and from
> where are they derived? This is unclear.
> Page 8, Column 1, Table 1: What is missing is a classification of
> the values in relation to the expectations / demands...which are,
> however, still missing in advance
> Page 8, Column 1, Line 37-43: At this point, an interpretation of
> the significant second peak in the right-hand diagram in Figure 4
> should be made. This has not been discussed so far.
> Page 8, Column 1, Line 51: "ten scans" means ten individual scan
> lines or 10 epochs or is this identical? What kind of temporal
> offset is there in between and wouldn't a motion model have to be
> taken into account for this or is an identical pose assumed for the
> ten scans?
> Page 8, Column 1, Line 55-56: What could be the cause of this?
> Page 8, Column 2, Line 44: This sentence always applies and should
> not remain here.
> Page 9, Figure 4: The two point clouds in the middle of figure 4 are
> difficult to interpret in terms of their orientation. Could it be an
> option to add a coordinate system?

### 2021-06-28 Latex/rebuttal.txt
Done reviewer 2 additional points corrections

@@
 > Page 1, Line 15: Double use of "protect", which should be replaced
-> by another word.
+> by another word. -> Removed first instance
 > Page 1, Line 21: The link for the "second option" to the use of an
-> IMU is not necessarily obvious.
+> IMU is not necessarily obvious. -> added IMU as sensor for coarse estimation
 > Page 1, Column 2, Line 52-56: Complicated sentence which should be
-> reformulated.
+> reformulated. -> Split into two sentences
 > Page 2, Column 2, Line 15: The term "long stop" is not to be
 > interpreted clearly in this context. How is "long" defined in this
-> sense in terms of time or space?
+> sense in terms of time or space? -> Specified stop is required to find enough points for plane detection
 > Page 3, Column 1, Line 6-8: A specific reference should be given for
 > the first half of the sentence, as reference 19 probably only refers
-> to the second part.
+> to the second part. -> Added todo for Dorit/Andreas as they wrote this
 > Page 3, Column 1, Line 23-26: Not all variables were introduced
-> accurately
-> Page 3, Column 1, Line 33-34: Selection of references for this?
+> accurately -> Introduced missing variables and and clarified rho
+> Page 3, Column 1, Line 33-34: Selection of references for this? -> Added plane segmentation sources
 > Page 3, Column 1, Line 34-35: But what demands are made on the
 > assignment? This is unclear to the reader, also in connection with
-> the general quality requirements of the whole approach.
+> the general quality requirements of the whole approach. -> Added specification "good enough for plane detection"
 > Page 3, Column 1, Line 38: How can the threshold value be specified,
@@
 > environments? Later, there is also no indication of the choice of
-> the value in the own simulated and real data sets.
+> the value in the own simulated and real data sets. -> Added todo for Fabi
 > Page 3, Column 1, Line 50: What does the index k describe and what
-> is the difference to the index i from Section 3.2?
+> is the difference to the index i from Section 3.2? -> clarified indeces k=planes i=points
 > Page 3, Equation (1): Again, some variables are not specified. For
-> example, the source of tx ty tz is unclear.
+> example, the source of tx ty tz is unclear. -> introduced tx ty tz properly
 > Page 3, Equation (3): Equation unclear regarding the use of the 2 in
-> the exponent as well as the index.
+> the exponent as well as the index. -> Specified that we're using the square of the L2 norm
 > Page 3, Equations (4)-(5): Here I don't understand how to get to the
 > elements of the normal vector through the transformation. Moreover,
-> these individual elements were not introduced in advance.
+> these individual elements were not introduced in advance. -> Corrected indeces and introduced elements
 > Page 3, Equation (9): The variable alpha is not described and
-> explained
+> explained -> explained alpha
 > Page 4, Equation (10): Description of the individual variables is
 > missing and, moreover, rho has already been used as the distance
-> parameter of the Hesse-normal form
+> parameter of the Hesse-normal form -> replaced rho with zeta, introduced variables
 > Page 4, Equation (12): Is epsilon here the same variable as the
-> threshold in Section 3.3? Probably not.
+> threshold in Section 3.3? Probably not.-> renamed epsilon, clarified difference to sec 3.3
 > Page 4, Column 2, Line 15: "some dimensions": Which ones and why
-> exactly are they chosen? Are they the same ones all the time?
-> Page 4, Column 2, Line 20-25: Reference is missing
+> exactly are they chosen? Are they the same ones all the time? -> chose better example, added todo for Fabi
+> Page 4, Column 2, Line 20-25: Reference is missing -> added source
 > Page 4, Column 2, Line 55: Doubled information regarding the
-> distance measurement precision
+> distance measurement precision -> removed duplicates
 > Page 5, Column 1, Equation (18): Delta t is not introduced and there
-> is no indication of how large the time steps are in the simulation.
-> Page 5, Column 2, Line 14: Should it be "out" or "our"?
+> is no indication of how large the time steps are in the simulation. -> introduced delta t and gave default value
+> Page 5, Column 2, Line 14: Should it be "out" or "our"? -> corrected
 > Page 5, Column 2, Line 27-28: What are the specifications of the
-> used IMU with regard to its uncertainty budget?
+> used IMU with regard to its uncertainty budget? -> added precision of IMU + source
 > Page 6, Figure 1: How are these values to be interpreted and from
-> where are they derived? This is unclear.
+> where are they derived? This is unclear.-> specified (chosen to still allow plane detection)
 > Page 8, Column 1, Table 1: What is missing is a classification of
 > the values in relation to the expectations / demands...which are,
-> however, still missing in advance
+> however, still missing in advance -> I don't understand this comment? TODO 
 > Page 8, Column 1, Line 37-43: At this point, an interpretation of
 > the significant second peak in the right-hand diagram in Figure 4
-> should be made. This has not been discussed so far.
+> should be made. This has not been discussed so far. -> added todo REDO experiment
 > Page 8, Column 1, Line 51: "ten scans" means ten individual scan
@@
 > taken into account for this or is an identical pose assumed for the
-> ten scans?
-> Page 8, Column 1, Line 55-56: What could be the cause of this?
+> ten scans? -> Specified which scans and that we assuem constant pose
+> Page 8, Column 1, Line 55-56: What could be the cause of this?-> specified outlier dependence
 > Page 8, Column 2, Line 44: This sentence always applies and should
-> not remain here.
+> not remain here. -> removed :(
 > Page 9, Figure 4: The two point clouds in the middle of figure 4 are
 > difficult to interpret in terms of their orientation. Could it be an
-> option to add a coordinate system?
+> option to add a coordinate system? -> specified in the figure label that lateral images have the same orientation

### 2021-06-28 Latex/main.tex
Changed Author List
Done reviewer 2 additional points corrections

@@
 %% Group authors per affiliation:
-\author{Fabian Arzberger and Anton Bredenbeck, Jasper Zevering, Dorit Borrmann, Andreas Nüchter}
+\author{Fabian Arzberger, Anton Bredenbeck, Jasper Zevering, Dorit Borrmann, Andreas Nüchter}
 \address{Informatics VII: Robotics and Telematics,
@@
 Spherical robots are a format that has not been thoroughly explored for the application of mobile mapping. 
-In contrast to other designes, it provides some unique advantages. 
-Among those is a protective spherical shell that protects internal sensors and actuators from possible harsh environments, as well as an inherent rotation for locomotion that enables measurements in all directions.
+In contrast to other designs, it provides some unique advantages. 
+Among those is a spherical shell that protects internal sensors and actuators from possible harsh environments, as well as an inherent rotation for locomotion that enables measurements in all directions.
 Mobile mapping always requires a high-precise pose knowledge to obtain consistent and correct environment maps. 
-This is typically done by a combination of external reference sensors such as Global Navigation Satellite System (GNSS) measurements and inertial measurements or by post processing the data and registering the different measurements to each other.
+This is typically done by a combination of external reference sensors such as Global Navigation Satellite System (GNSS) measurements and inertial measurements or by coarsely estimating the pose using inertial measurement units (IMUs) and post processing the data by registering the different measurements to each other.
 In indoor environments, the GNSS reference is not an option.

### 2021-06-28 Latex/sections/introduction.tex
Done reviewer 2 additional points corrections

@@
 This paper proposes to use such a spherical robot for mobile mapping man-made environments.
-In such environments, one advantage are architectural shapes following standard conventions arising from tradition or utility, i.e., there are many flat surfaces such as walls, floors, etc.
+In such environments, one advantage are architectural shapes following standard conventions arising from tradition or utility.
+In particular, there are many flat surfaces such as walls, floors, etc.
 that are sensed.

### 2021-06-28 Latex/sections/realData.tex
Done reviewer 2 additional points corrections

@@
 
-We also tested out algorithm on a real-world dataset that adheres to different motion profiles and laser scanners type with different scanning patterns than the simulated data.
+We also tested our algorithm on a real-world dataset that adheres to different motion profiles and laser scanners type with different scanning patterns than the simulated data.
 
@@
 The sphere is equipped with a low-cost IMU, i.e., a PhidgetSpatial Precision 3/3/3, to estimate the orientation, a battery pack and a raspberry pie for processing.
+This IMU provides a precision of \SI{76.3}{\micro g} in linear acceleration, \SI{0.02}{\degree\per\second} in angular velocity about the x and y axis, and \SI{0.013}{\degree\per\second} in angular velocity about the z axis \cite{PhidgetSpatial}.
 

### 2021-06-28 Latex/sections/registration.tex
Done reviewer 2 additional points corrections

@@
 The most simple method is to use the data from an intertial measurement unit for a coarse alignment.
-Alternative solutions combine multiple linescans into a metascan and perform registration methods known from terrestrial laser scanning or use a small number of iterations of continuous-time SLAM approaches such as the one presented in~\cite{REMSEN2013}.
+Alternative solutions combine multiple linescans into a metascan and perform registration methods known from terrestrial laser scanning \todo{add source here} or use a small number of iterations of continuous-time SLAM approaches such as the one presented in~\cite{REMSEN2013}.
 This yields maps that can be used for plane registration and hence a point-to-plane correspondences based registration.
@@
 In the next subsection, the Hesse-normal form of the plane is required.
-For an ideal plane the distance parameter $\rho_{\mathcal{P}_i}$ is computed via $\vec{n}_{\mathcal{P}_i}\cdot\vec{x}_i$, where $\vec{x}_i$ is an arbitrary point on the plane.
+For an ideal plane the orthogonal distance from the origin $\rho_{\mathcal{P}_k}$ is computed via $\vec{n}_{\mathcal{P}_k}\cdot\vec{p}_i$, where $\vec{p}_i = \begin{bmatrix}x_i&y_i&z_i\end{bmatrix}^T$ is an arbitrary point on the plane and $\vec{n}_{\mathcal{P}_k} = \begin{bmatrix}n_{\mathcal{P}_k}^x&n_{\mathcal{P}_k}^y&n_{\mathcal{P}_k}^z\end{bmatrix}^T$ is the normal vector of that plane.
 To find this point, we first define a convex hull that encloses the plane.
 It is defined by the points that lie on the plane with the furthest distance to one another.
-We then choose the center point of the convex hull of the plane as $\vec{x}_i$.
+We then choose the center point of the convex hull of the plane as $\vec{p}_i$.
 
-\subsection{Point-to-Plane Correspondences}
+\subsection{Point-to-Plane Correspondences}\label{sec:point-to-plane-correspondences}
 
-There are many ways of finding correspondences between points and planes.
-However, even with the result not being perfect, we assume already globally registered scans, which is why a simple algorithm is sufficient:
+There are many ways of solving the segmentation problem of assigning each point to a plane often based on the RANSAC method \cite{Honti2018}, \cite{Gaspers2011} but also deep learning approaches \cite{Engelmann2018}.
+The quality of the preregistration directly influences the quality of the plane detection and therefore the quality of the final registration result. 
+We assume small enough errors in pre-registration to allow for plane detection.
+Which is why a simple algorithm is sufficient:
 For each plane, consider adding a correspondence for all points closer to the plane than a threshold $\epsilon$.
-Depending on the threshold value, some points now have multiple correspondences if their distance to multiple planes is less than $\epsilon$.
+Depending on the threshold value, some points now have multiple correspondences if their distance to multiple planes is less than $\epsilon$ \todo{@Fabi add actual $\epsilon$ values to the experiments}.
 We then omit all points with correspondences to multiple planes from the optimization to avoid introducing errors due to the plane ambiguity.
@@
 Assuming we know now the Hesse-normal form of all planes and all assigned points to these planes, we register the points to have a minimal distance to their respective planes.
-The transformation $T(\vec{p}_k)$ of each point $\vec{p}_k$ with respect to a 6 DoF motion is described in homogeneous coordinates using the roll-pitch-yaw ($\varphi-\vartheta-\psi$) Tait-Brian angles as in~\cite{diebel2006representing}.
-Transforming the result back from homogeneous coordinates and using $C_k$ and $S_k$ to denote the cosine and sine of the angle in the subscript yields:
+The transformation $T(\vec{p}_i)$ of each point $\vec{p}_i$ with respect to a 6 DoF motion is described in homogeneous coordinates using the roll-pitch-yaw ($\varphi-\vartheta-\psi$) Tait-Brian angles as in~\cite{diebel2006representing}.
+Transforming the result back from homogeneous coordinates and using $C_a$ and $S_a$ to denote the cosine and sine of the angle in the subscript, and $t_a$ denoting the translation along the axis in the subscript yields:
 \begin{align}
-	T(\vec{p}_k)  =
+	T(\vec{p}_i)  =
     \resizebox{0.35\textwidth}{!}{$\begin{bmatrix}
-        x C_\vartheta C_\psi - yC_\varphi S_\psi + z S_\vartheta + t_x\\
-        x (C_\varphi S_\psi + C_\psi S_\varphi S_\vartheta) + y(C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi) - z C_\vartheta S_\varphi + t_y\\
-        x(S_\varphi S_\psi - C_\varphi C_\psi S_\vartheta) + y(C_\psi S_\varphi + C_\varphi S_\vartheta S_\psi) + z C_\varphi C_\vartheta + t_z
+        x_i C_\vartheta C_\psi - y_iC_\vartheta S_\psi + z_i S_\vartheta + t_x\\
+        x_i (C_\varphi S_\psi + C_\psi S_\varphi S_\vartheta) + y_i(C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi) - z_i C_\vartheta S_\varphi + t_y\\
+        x_i(S_\varphi S_\psi - C_\varphi C_\psi S_\vartheta) + y_i(C_\psi S_\varphi + C_\varphi S_\vartheta S_\psi) + z_i C_\varphi C_\vartheta + t_z
     \end{bmatrix}$}
 \end{align}
-From this we define the function $D(\varphi,\vartheta,\psi,t_x,t_y,t_z, \vec{p}_k)$ that computes the distance of a point $\vec{p}_k$ to its corresponding plane $\mathcal{P}_i$.
+From this we define the function $D(\varphi,\vartheta,\psi,t_x,t_y,t_z, \vec{p}_i)$ that computes the distance of a point $\vec{p}_i$ to its corresponding plane $\mathcal{P}_k$.
 Omitting the arguments of the function for simplicity:
@@
 \begin{aligned}
-    D &= T(\vec{p}_k) \cdot \vec{n}_{\mathcal{P}_i} \\
-      &= n_{\mathcal{P}_i}^x (x C_\vartheta C_\psi - yC_\varphi S_\psi + z S_\vartheta + t_x)\\
-       &+ n_{\mathcal{P}_i}^y(x (C_\varphi S_\psi + C_\psi S_\varphi S_\vartheta)\\
-       &\qquad+ y(C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi) - z C_\vartheta S_\varphi + t_y)\\
-       &+ n_{\mathcal{P}_i}^z (x(S_\varphi S_\psi - C_\varphi C_\psi S_\vartheta)\\
-       &\qquad+ y(C_\psi S_\varphi + C_\varphi S_\vartheta S_\psi) + z C_\varphi C_\vartheta + t_z)\\
-       &- \rho_{\mathcal{P}_i}
+    D &= T(\vec{p}_i) \cdot \vec{n}_{\mathcal{P}_k} \\
+      &= n_{\mathcal{P}_k}^x (x_i C_\vartheta C_\psi - y_iC_\vartheta S_\psi + z_i S_\vartheta + t_x)\\
+       &+ n_{\mathcal{P}_k}^y(x_i (C_\varphi S_\psi + C_\psi S_\varphi S_\vartheta)\\
+       &\qquad+ y_i(C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi) - z_i C_\vartheta S_\varphi + t_y)\\
+       &+ n_{\mathcal{P}_k}^z (x_i(S_\varphi S_\psi - C_\varphi C_\psi S_\vartheta)\\
+       &\qquad+ y_i(C_\psi S_\varphi + C_\varphi S_\vartheta S_\psi) + z_i C_\varphi C_\vartheta + t_z)\\
+       &- \rho_{\mathcal{P}_k}
 \end{aligned}
@@
 This distance function is what we want to minimize for all points and their respective planes.
-Hence the error function $E$ is chosen as follows:
+Hence the error function $E$ is chosen as the square of the L2-norm of the distance:
 \begin{align}
-	E = \sum_{\forall \mathcal{P}_i}\sum_{\vec{p}_k \in \mathcal{P}_i} \norm[2]{D(\varphi,\vartheta,\psi,t_x,t_y,t_z, \vec{p}_k)}^2
+	E = \sum_{\forall \mathcal{P}_k}\sum_{\vec{p}_i \in \mathcal{P}_k} \norm[2]{D(\varphi,\vartheta,\psi,t_x,t_y,t_z, \vec{p}_i)}^2
 \end{align}
@@
 \begin{align}
-	&\nabla E =  \sum_{\forall \mathcal{P}_i}\sum_{\vec{p}_k \in \mathcal{P}_i} \resizebox{0.28\textwidth}{!}{$\begin{bmatrix}\pardiff{\varphi} E&\pardiff{\vartheta} E&\pardiff{\psi} E&\pardiff{t_x} E&\pardiff{t_y} E&\pardiff{t_z} E\end{bmatrix}^T$}\\
-    &\Rightarrow\nabla E = \resizebox{0.35\textwidth}{!}{$\sum_{\forall \mathcal{P}_i} \sum_{\vec{p}_k \in \mathcal{P}_i}2D(\vec{\Pi}, \vec{p}_k)\begin{bmatrix}\nabla E_\varphi &\nabla E_\vartheta & \nabla E_\psi & n_{\mathcal{P}_i}^x&n_{\mathcal{P}_i}^y&n_{\mathcal{P}_i}^z\end{bmatrix}^T$}
+	&\nabla E =  \sum_{\forall \mathcal{P}_k}\sum_{\vec{p}_i \in \mathcal{P}_k} \resizebox{0.28\textwidth}{!}{$\begin{bmatrix}\pardiff{\varphi} E&\pardiff{\vartheta} E&\pardiff{\psi} E&\pardiff{t_x} E&\pardiff{t_y} E&\pardiff{t_z} E\end{bmatrix}^T$}\\
+    &\Rightarrow\nabla E = \resizebox{0.35\textwidth}{!}{$\sum_{\forall \mathcal{P}_k} \sum_{\vec{p}_i\in \mathcal{P}_k}2D(\vec{\Pi}, \vec{p}_i)\begin{bmatrix}\nabla E_\varphi &\nabla E_\vartheta & \nabla E_\psi & n_{\mathcal{P}_k}^x&n_{\mathcal{P}_k}^y&n_{\mathcal{P}_k}^z\end{bmatrix}^T$}
 \end{align}
@@
     &\begin{alignedat}{3}
-        \nabla E_\varphi = & n_{\mathcal{P}_i}^x&&y S_\varphi S_\psi \\
-        &+ n_{\mathcal{P}_i}^y&&(x[-S_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] \\&&&+y[-S_\varphi C_\psi -C_\varphi S_\vartheta S\psi] \\&&&- z C_\varphi C_\vartheta)\\
-        &+ n_{\mathcal{P}_i}^z&&(x[C_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] \\&&&+y[C_\varphi C_\psi -S_\varphi S_\vartheta S\psi] \\&&&- z S_\varphi C_\vartheta)
+        \nabla E_\varphi = &n_{\mathcal{P}_k}^y&&(x_i[-S_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] \\&&&+y_i[-S_\varphi C_\psi -C_\varphi S_\vartheta S\psi] \\&&&- z_i C_\varphi C_\vartheta)\\
+        &+ n_{\mathcal{P}_k}^z&&(x_i[C_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] \\&&&+y_i[C_\varphi C_\psi -S_\varphi S_\vartheta S\psi] \\&&&- z_i S_\varphi C_\vartheta)
     \end{alignedat} \\
     &\begin{alignedat}{3}
-        \nabla E_\vartheta = & n_{\mathcal{P}_i}^x(-xS_\vartheta C_\psi + zC_\vartheta) \\
-        &+ n_{\mathcal{P}_i}^y(xC_\psi S_\varphi C_\vartheta - yS_\varphi C_\vartheta S_\psi + z S_\vartheta S_\varphi)  \\
-        &+ n_{\mathcal{P}_i}^z(-xC_\varphi C_\psi C_\vartheta + yC_\varphi C_\vartheta S_\psi - z C_\varphi S_\vartheta)
+        \nabla E_\vartheta = & n_{\mathcal{P}_k}^x(-x_iS_\vartheta C_\psi + y_i S_\vartheta S_\psi + z_iC_\vartheta) \\
+        &+ n_{\mathcal{P}_k}^y(x_iC_\psi S_\varphi C_\vartheta - y_iS_\varphi C_\vartheta S_\psi + z_i S_\vartheta S_\varphi)  \\
+        &+ n_{\mathcal{P}_i}^z(-x_iC_\varphi C_\psi C_\vartheta + y_iC_\varphi C_\vartheta S_\psi - z_i C_\varphi S_\vartheta)
     \end{alignedat}\\
     &\begin{alignedat}{3}
-       \nabla E_\psi =& n_{\mathcal{P}_i}^x&&\left(-xC_\vartheta S_\psi - yC_\varphi C_\psi\right) \\
-       &+ n_{\mathcal{P}_i}^y&&(x[C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi] \\
-       &&&+ y[-C_\varphi S_\psi - S_\varphi S_\theta C_\psi]) \\
-       &+ n_{\mathcal{P}_i}^z&&(x[S_\varphi C_\psi + C_\varphi S_\psi S_\theta] \\
-       &&&+ y[-S_\psi S_ \varphi + C_\varphi S_\vartheta C_\psi]) 
+       \nabla E_\psi =& n_{\mathcal{P}_k}^x&&\left(-x_iC_\vartheta S_\psi - y_iC_\vartheta C_\psi\right) \\
+       &+ n_{\mathcal{P}_k}^y&&(x_i[C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi] \\
+       &&&+ y_i[-C_\varphi S_\psi - S_\varphi S_\theta C_\psi]) \\
+       &+ n_{\mathcal{P}_k}^z&&(x_i[S_\varphi C_\psi + C_\varphi S_\psi S_\theta] \\
+       &&&+ y_i[-S_\psi S_ \varphi + C_\varphi S_\vartheta C_\psi]) 
     \end{alignedat}
@@
 \end{align}
+where $\alpha$ is the learning rate.
 To accelerate convergence and to improve the found solution further modifications are made.
@@
 It extends the SDG algorithm by two terms.
-First, an exponentially decaying average of past gradients, which is recursively defined as
+First, an exponentially decaying average of past gradients $\vec{G}_k$, which is recursively defined as
 \begin{align}
-    \vec{G}_{k+1} = \rho \vec{G}_{k} + (1 - \rho) {\nabla E}^2
+    \vec{G}_{k+1} = \zeta \vec{G}_{k} + (1 - \zeta) {\nabla E}^2
 \end{align}
-and second, an exponentially decaying average of past changes, which is defined as
+and second, an exponentially decaying average of past changes $\vec{X}_k$, which is defined as
 \begin{align}
-    \vec{X}_{k+1} = \rho \vec{X}_{k} + (1 - \rho) {\boldsymbol\alpha \nabla E}^2
+    \vec{X}_{k+1} = \zeta \vec{X}_{k} + (1 - \zeta) {\boldsymbol\alpha \nabla E}^2
 \end{align}
-where $\rho \leq 1$ is a decay constant, typically close to $1$.
+where $\zeta \leq 1$ is a decay constant, typically close to $1$.
 The root mean squared (RMS) of these quantities are
 \begin{align}
-    RMS[\vec{G}]_{k} = \sqrt{\vec{G}_{k} + \epsilon}
+    RMS[\vec{G}]_{k} = \sqrt{\vec{G}_{k} + \epsilon_{num}}
 \end{align}
@@
 \begin{align}
-    RMS[\vec{X}]_{k} = \sqrt{\vec{X}_{k} + \epsilon}
+    RMS[\vec{X}]_{k} = \sqrt{\vec{X}_{k} + \epsilon_{num}}
 \end{align}
-where $\epsilon > 0$ is a very small constant, typically close to $0$.
+where $\epsilon_{num} > 0$ is a very small constant, typically close to $0$ (Note: this is a different threshold as the one used in section~\ref{sec:point-to-plane-correspondences}).
 It will prevent dividing by zero in the recalculation of $\boldsymbol\alpha$, which is as follows:
@@
 
-Firstly, we introduce a lock on some dimensions from being optimized by setting the corresponding $\alpha_i$ to zero.
+Firstly, we introduce the possibility to choose a lock on some dimensions from being optimized by setting the corresponding $\alpha_i$ to zero.
 Although 6D optimization generally works, reducing the optimization space is particularly useful if the source of error in the system is known and a model exists.
-That way, e.g., a point cloud from a terrestrial laserscanner that experiences irregularities in its spinning mirror is registered by optimizing only over the axis of rotation as no other movement is to be expected.
+That way, e.g., as we expect a spherical robot to move on a plane the position along the axis perpendicular to the plane is constant and should not be used for optimization.\todo{@Fabi clarify in experiments which locks were active.}
 

### 2021-06-28 Latex/sections/results.tex
Removed mean as its not representative: Includes points with distance > 100 only with distance 100
Done reviewer 2 additional points corrections

@@
 Assuming this represents a coarsely pre-registered 3D point cloud, the distances to the ground truth were evaluated before and after the plane-based registration. 
-Figure~\ref{fig:simulatedEvaluation} shows the different point-to-point distances.
+Figure~\ref{fig:simulatedEvaluation} \todo{REDO experiments!!!} shows the different point-to-point distances.
 
@@
  	\caption{Evaluation of point distances before (left) and after (right) the plane-based registration on a simulated dataset.
+Lateral images always have the same orientation.
 A maximal distance of \SI{2}{m} is set, such that all points that display a higher distance value are excluded from the analysis.
@@
 The color mapping is equivalent in both.
-An animation of the matching process is given at \url{https://youtu.be/0ps5Pg4qo4E} .}
+An animation of the matching process is given at \url{https://youtu.be/0ps5Pg4qo4E}.}
  	\label{fig:simulatedEvaluation}
@@
 	\centering
-	\begin{tabular}{@{}lcccc@{}}\hline
-		& P90 & P95 & P98 & Mean\\ \hline\hline
-		Uncorrected & \SI{24.2}{\centi\meter} &  \SI{38.2}{\centi\meter} &  \SI{61.5}{\centi\meter} & \SI{9.7}{\centi\meter}\\
-		Corrected & \SI{12.8}{\centi\meter} &  \SI{16.6}{\centi\meter} &  \SI{21.6}{\centi\meter} & \SI{6.3}{\centi\meter}\\ \hline
+	\begin{tabular}{@{}lccc@{}}\hline
+		& P90 & P95 & P98 \\ \hline\hline
+		Uncorrected & \SI{24.2}{\centi\meter} &  \SI{38.2}{\centi\meter} &  \SI{61.5}{\centi\meter} \\%& \SI{9.7}{\centi\meter}\\
+		Corrected & \SI{12.8}{\centi\meter} &  \SI{16.6}{\centi\meter} &  \SI{21.6}{\centi\meter} \\\hline%& \SI{6.3}{\centi\meter}\\ \hline
 	\end{tabular}
@@
 \end{table}
+
+The second significant peak in the right hand histogram in figure~\ref{fig:simulatedEvaluation} \todo{Explain this peak. Not sure right now.}
 
@@
 Figure~\ref{fig:cylon-corrected} shows the results obtained before and after employing the plane based registration on the dataset acquired by the floating sphere.
-We combine ten scans into one meta-scan, which is then globally registered.
+The pre-registration is obtained by determining the orientation of the sphere via Madgwick filtered~\cite{madgwick2010efficient} IMU measurements.
+To increase the number of possible point-to-plane correspondences we combine ten temporally successive scans into one meta-scan, which is then globally registered. 
+For this we must assume constant pose over this time duration. 
+However, given the high measurement frequency of the Lidar this introduces only a small error. 
 This does not only speed up convergence due to the proportional effect on the error function but also decreases the risk of transforming the scan in a senseless way.
-Transformations like these happen when incorrectly labeled points have a huge effect on the error gradient, which is especially the case when considering only a few points.
+Transformations like these happen in particular for a small collection of points as outliers have more influence.
 

### 2021-06-28 Latex/sections/simulation.tex
Done reviewer 2 additional points corrections

@@
 The Livox-Mid100 laser scanner consists of three Mid40 scanners arranged to provide an overall field-of-view of \SI{98.4}{\degree} (horizontal) by \SI{38.4}{\degree} (vertical).
-With a range precision of \SI{2}{\centi\meter} up to \SI{20}{\meter} range it provides \SI[per-mode=symbol]{300,000}{\pts\per\second} with a range precision of \SI{2}{\centi\meter} up to \SI{20}{\meter} range.
+It provides \SI[per-mode=symbol]{300,000}{\pts\per\second} with a range precision of \SI{2}{\centi\meter} up to \SI{20}{\meter} range~\cite{LivoxMid40-100}.
 A particular advantage of the Livox laser scanners is their unique scanning pattern.
@@
 This additional motion is determined by integrating the disturbance torques, which are sampled from some normal distribution $n_\phi, n_\psi \in \mathcal{N}(\mu,\sigma^2)$ each.  
-The pose update in the discrete simulation is therefore given by: 
+The pose update after a discrete time step $\Delta t$ in the simulation is therefore given by: 
 \begin{align}
@@
 \end{align}
+where $\Delta t$ is generally chosen as some small value (e.g. \SI{1}{\milli\second})
 
@@
 	\caption{Two simulated datasets of a long hallway of size $\SI{4}{\meter}\times\SI{3}{\meter}\times\SI{100}{\meter}$.
-Ideal noise free (left), noisy pose ($\mathcal{N}_\varphi = \mathcal{N}_\psi(\mu = 0.0001, \sigma= 0.00001)$) and noisy range measurements ($\mathcal{N}_r(\mu = 0, \sigma = 0.001)$) (right).}
+Ideal noise free (left), noisy pose ($\mathcal{N}_\varphi = \mathcal{N}_\psi(\mu = 0.0001, \sigma= 0.00001)$) and noisy range measurements ($\mathcal{N}_r(\mu = 0, \sigma = 0.001)$) (right).
+	The normal distributions are define in such a way, that in the resulting dataset plane detection is still possible.}
 	\label{fig:simulatedDatasets}

### 2021-06-28 Latex/sections/stateOfTheArt.tex
Done reviewer 2 additional points corrections

@@
 However, for pre-registration the classical point-to-point registration is also very effective. 
-One advantage that point-to-point correspondences have over plane-to-plane correspondences is that they do not require a long stop in each pose.
+One advantage that point-to-point correspondences have over plane-to-plane correspondences is that they do not require a long stop in order to obtain enough points to detect planes in each pose.
 For plane-to-plane correspondences, this is necessary to gather enough data to measure planes in each scan robustly.

### 2021-07-01 Latex/rebuttal.txt
Rebuttal modified

--- 2021-06-22 Latex/rebuttal.txt
@@
- 
 > Manuscript Number: OPHOTO-D-21-00020  
@@
 > 
-
 > I have completed my evaluation of your manuscript. The reviewers
@@
 > by another word.
+
+We removed first instance.
+
 > Page 1, Line 21: The link for the "second option" to the use of an
 > IMU is not necessarily obvious.
+
+We added IMU as sensor for coarse estimation.
+
 > Page 1, Column 2, Line 52-56: Complicated sentence which should be
 > reformulated.
+
+We have split into two sentences
+
 > Page 2, Column 2, Line 15: The term "long stop" is not to be
@@
 > sense in terms of time or space?
+
+The specified stop is required to find enough points for plane
+detection.
+
 > Page 3, Column 1, Line 6-8: A specific reference should be given for
@@
 > to the second part.
+
+FIXME! Added todo for Dorit/Andreas as they wrote this
+
 > Page 3, Column 1, Line 23-26: Not all variables were introduced
 > accurately
+
+We have introduced missing variables and clarified rho.
+
 > Page 3, Column 1, Line 33-34: Selection of references for this?
+
+We have added plane segmentation sources.
+
 > Page 3, Column 1, Line 34-35: But what demands are made on the
@@
 > the general quality requirements of the whole approach.
+
+We have added a specification for "good enough for plane detection".
+
 > Page 3, Column 1, Line 38: How can the threshold value be specified,
@@
 > the value in the own simulated and real data sets.
+
+FIXME: Added todo for Fabi
+
 > Page 3, Column 1, Line 50: What does the index k describe and what
 > is the difference to the index i from Section 3.2?
+
+We have clarified indeces k=planes i=points.
+
 > Page 3, Equation (1): Again, some variables are not specified. For
 > example, the source of tx ty tz is unclear.
+
+We have introduced tx ty tz properly.
+
 > Page 3, Equation (3): Equation unclear regarding the use of the 2 in
 > the exponent as well as the index.
+
+We have specified that we're using the square of the L2 norm.
+
 > Page 3, Equations (4)-(5): Here I don't understand how to get to the
@@
 > these individual elements were not introduced in advance.
+
+We have corrected indeces and introduced the elements.
+
 > Page 3, Equation (9): The variable alpha is not described and
 > explained
+
+An explaination of alpha was added.
+
 > Page 4, Equation (10): Description of the individual variables is
@@
 > parameter of the Hesse-normal form
+
+we have replaced rho with zeta and introduced the variables.
+
 > Page 4, Equation (12): Is epsilon here the same variable as the
 > threshold in Section 3.3? Probably not.
+
+We have renamed epsilon and clarified the difference to sec 3.3.
+
 > Page 4, Column 2, Line 15: "some dimensions": Which ones and why
 > exactly are they chosen? Are they the same ones all the time?
+
+FIXME: chose better example, added todo for Fabi
+
 > Page 4, Column 2, Line 20-25: Reference is missing
+
+We haded a citation.
+
 > Page 4, Column 2, Line 55: Doubled information regarding the
 > distance measurement precision
+
+We removed duplicated information.
+
 > Page 5, Column 1, Equation (18): Delta t is not introduced and there
 > is no indication of how large the time steps are in the simulation.
+
+We have introduced delta t and give a default value.
+
 > Page 5, Column 2, Line 14: Should it be "out" or "our"?
+
+We have corrected this.
+
 > Page 5, Column 2, Line 27-28: What are the specifications of the
 > used IMU with regard to its uncertainty budget?
+
+We have added precision of IMU and added the source.
+
 > Page 6, Figure 1: How are these values to be interpreted and from
 > where are they derived? This is unclear.
+
+We have specified this. It is chosen to still allow plane detection.
+
 > Page 8, Column 1, Table 1: What is missing is a classification of
@@
 > however, still missing in advance
+
+FIXME: -> I don't understand this comment? TODO
+
 > Page 8, Column 1, Line 37-43: At this point, an interpretation of
@@
 > should be made. This has not been discussed so far.
+
+FIXME: -> added todo REDO experiment
+
 > Page 8, Column 1, Line 51: "ten scans" means ten individual scan
@@
 > ten scans?
+
+We have specified which scans and that we assuem constant pose.
+
 > Page 8, Column 1, Line 55-56: What could be the cause of this?
+
+We have specified the outlier dependence.
+
 > Page 8, Column 2, Line 44: This sentence always applies and should
 > not remain here.
+
+We have removed this sentence.
+
 > Page 9, Figure 4: The two point clouds in the middle of figure 4 are
@@
 > option to add a coordinate system?
+
+It was specified in the figure label that lateral images have the same
+orientation.

### 2021-07-03 Latex/rebuttal.txt
Done reviewer one comments

@@
+ 
 > Manuscript Number: OPHOTO-D-21-00020  
@@
 > systems.
-> 
+
+We clarified that the shell can enclose all sensor without requiring a 
+number of points of connection as individual enlosings for all sensors
+would. Further we added the requirement, that the shell is very 
+durable.
+
 > Lines 111-116: Modern kinematic scanning systems collect point cloud
@@
 > long distances between adjacent points.
-> 
+
+We clarified that also the lilmited FOV plays a role in not finding 
+planes for robust correspondences. 
+FIXME @Andreas and @Dorit: I (Anton) disagree can you double check?
+
 > Lines 244-246: The Livox ranging precision is given twice in this
@@
 > suggest adding it to the figure also.
-> 
+
+FIXME Redoing the experiment anyways
+
 > Conclusion: Whereas the title, Abstract and Introduction concentrate
@@
 > Spherical Robots For Mobile Mapping In Human Made Environments".
-> 
-> 
-> 
-> 
+
+We added more context about spherical robots and put the registration 
+method into context.
+
 > 

### 2021-07-03 Latex/sections/conclusion.tex
Done reviewer 2 additional points corrections
Done reviewer one comments

@@
 
-In this paper, we have proposed a registration method for man made environments that exploits the structure of those environments. 
-Given that in human-made environments parallel planes and right angles are abundantly available, we employ point-to-plane correspondences to improve a pre-registered 3D point cloud. 
+In this paper, we proposed an approach to mobile mapping using a spherical robot. 
+Using the inherent rotation of the sphere used for locomotion to also rotate a laser scanner also for measurements of the entire environment.
+We propose to post-process the acquire data using a novel registration method for man made environments that exploits the structure of those environments. 
+Given that in human-made environments straight planes are abundantly available, we employ point-to-plane correspondences to improve a pre-registered 3D point cloud. 
 We have evaluated the procedure on a simulated dataset and on an experimentally acquired dataset with different motion profiles. 
@@
 However, as section~\ref{sec:experimentalSetup} shows, some outliers remain in the registered point cloud due to the simplistic point-to-plane correspondence model.
-This to be improved in future studies.
+This is to be improved in future studies.
 One idea to do so is to employ a local clustering technique onto the individual scans before registering them globally. 
@@
 
-Needless to say, a lot of work remains to be done. 
 %

### 2021-07-04 tdp/datasets/processing/histogram.cc
Redid Histogram and actually pushed the processing files this time

@@ -1,0 +1,106 @@
+#include <iostream>
+#include <fstream>
+#include <string> 
+#include <math.h>
+#include <sstream>
+
+using namespace std;
+
+void init(int arr[], int n)
+{
+    for (int i = 0; i < n; i++)
+        arr[i] = 0;
+}
+
+int main(int argc, char * argv[])
+{
+    string line, token;
+    string infile = "scan001.3d";
+    string outfile = "hist.txt";
+    double MAX_DIST=100.0;
+    double STEP=0.1;
+
+    if(argc > 1) infile = argv[1];
+    else cout << "Usage: " << argv[0] << " {infile outfile maxdist step} \n Continuing with defaul parameters." << endl;
+    if(argc > 2) outfile = argv[2];
+    if(argc > 3) MAX_DIST = atof(argv[3]);
+    if(argc > 4) STEP = atof(argv[4]);
+    ifstream file(infile);
+    ofstream hisfile(outfile);
+
+    int bins = (MAX_DIST/STEP) + 1;
+    int count[bins];
+    init(count, bins);
+   
+    int pointcount = 0;
+    if (file.is_open())
+    {
+        // sum up all data
+        cout << "Creating data ..." << infile << endl;
+        cout.flush();
+        double x, y, z, distance;
+
+        while(file.good()) {
+          file >> x;
+          file >> y;
+          file >> z;
+          file >> distance;
+          int index = (int)(distance / STEP);
+          count[min(index, bins-1)]++;
+          pointcount++;
+        }
+
+        //for(int i = 0; i < bins; i++) count[i] /=pointcount;
+        cout << "Writing data ..." << outfile << endl;
+        cout.flush();
+
+        int pointsum = 0;
+        double mean = 0;
+      	bool p90 = false;
+        bool p95 = false;
+        bool p98 = false;
+        for(int i = 0; i < bins; i++) {
+          hisfile << ((i * STEP) + STEP) << " " << (double)count[i] << endl;
+          pointsum += count[i];
+          if(!p90) {
+            if(pointsum > 0.9*pointcount) {
+              cout << "P90: " << ((i * STEP) + STEP) << endl;
+              p90 = true;
+            }
+          } if(!p95) {
+            if(pointsum > 0.95*pointcount) {
+              cout << "P95: " << ((i * STEP) + STEP) << endl;
+              p95 = true;
+            }
+          } if(!p98) {
[…]

### 2021-07-04 Latex/sections/results.tex
Redid experiment with simulated data

@@
 Assuming this represents a coarsely pre-registered 3D point cloud, the distances to the ground truth were evaluated before and after the plane-based registration. 
-Figure~\ref{fig:simulatedEvaluation} \todo{REDO experiments!!!} shows the different point-to-point distances.
+Figure~\ref{fig:simulatedEvaluation} \todo[inline]{Reupload matching process} shows the different point-to-point distances.
 
@@
 Lateral images always have the same orientation.
-A maximal distance of \SI{2}{m} is set, such that all points that display a higher distance value are excluded from the analysis.
-Further, the histogram joins all values with a distance greater than \SI{1}{m} into the last bucket.
+A maximal distance of \SI{30}{m} is set, such that all points that display a higher distance value are excluded from the analysis.
+Both point-clouds were reduced before evaluating point-to-point distances to the ground-truth.
+Further, the color-space maps all values with a distance greater than \SI{1}{m} to the same color.
 The top two columns show a heat map of distances, while the bottom shows the corresponding histogram.
@@
 The further into the corridor, i.e., the longer the robot accumulates errors, the more imprecise the data becomes. 
-Finally, we see that a large portion of the data even exceeds either the \SI{1}{\meter} error, contributing to the large spike in the histogram or even the \SI{2}{\meter} error limit; thus, being cut from the representation.
+Finally, we see that many points exceed the threshold of \SI{1}{\meter} and thus being mapped to the same color value.
 After registration, we see that, qualitatively, the ideal corridor was nearly restored from the noisy data. 
-In particular a very large portion of points (95\%) have distances of less than \SI{16.6}{\centi\meter}.
+In particular a very large portion of points (90\%) have distances of less than \SI{35.9}{\centi\meter}.
 Table~\ref{tab:percentiles} shows the comparison of further percentiles of both datasets.
@@
 		& P90 & P95 & P98 \\ \hline\hline
-		Uncorrected & \SI{24.2}{\centi\meter} &  \SI{38.2}{\centi\meter} &  \SI{61.5}{\centi\meter} \\%& \SI{9.7}{\centi\meter}\\
-		Corrected & \SI{12.8}{\centi\meter} &  \SI{16.6}{\centi\meter} &  \SI{21.6}{\centi\meter} \\\hline%& \SI{6.3}{\centi\meter}\\ \hline
+		Uncorrected & \SI{372.1}{\centi\meter} &  \SI{553.4}{\centi\meter} &  \SI{827.9}{\centi\meter} \\%& \SI{9.7}{\centi\meter}\\
+		Corrected & \SI{35.9}{\centi\meter} &  \SI{64.1}{\centi\meter} &  \SI{122.8}{\centi\meter} \\\hline%& \SI{6.3}{\centi\meter}\\ \hline
 	\end{tabular}
@@
 
-The second significant peak in the right hand histogram in figure~\ref{fig:simulatedEvaluation} \todo{Explain this peak. Not sure right now.}
-
-Further, the square and straight shape of the corridor is restored well, and especially the large spike at errors of greater than \SI{1}{m} is removed. 
+Further, the square and straight shape of the corridor is restored well, and especially the large amount of points with an errors of greater than \SI{1}{m} is removed. 
 Any such errors tend to occur at the back and the front of the corridor where the measured range is the largest hence has the largest contribution of the range error.

### 2021-07-04 Latex/sections/results.tex
Updated youtube link of matching simulated data

@@
 Assuming this represents a coarsely pre-registered 3D point cloud, the distances to the ground truth were evaluated before and after the plane-based registration. 
-Figure~\ref{fig:simulatedEvaluation} \todo[inline]{Reupload matching process} shows the different point-to-point distances.
+Figure~\ref{fig:simulatedEvaluation} shows the different point-to-point distances.
 
@@
 The color mapping is equivalent in both.
-An animation of the matching process is given at \url{https://youtu.be/0ps5Pg4qo4E}.}
+An animation of the matching process is given at \url{https://youtu.be/7igQdEeCsYk}.}
  	\label{fig:simulatedEvaluation}

### 2021-07-05 Latex/rebuttal.txt
Redid experiment with simulated data
second reviewers comments addressed

@@
 
-FIXME Redoing the experiment anyways
+We added a label to the x-scale
 
@@
 > should therefore be added.
-> 
+
+We added a short outline, introducing you to the next sections.
+
 > - In the entire article, no reference is made to the targeted or
@@
 > at least a rough classification.
-> 
+
+FIXME: I dont understand. How should we make any statement about the accuracy potential?
+It depends on the motion profile and hardware being used, thus is not easy to generalize.
+Or am I getting this wrong?
+
 > - For the short state of the art of existing mobile systems, which
@@
 > and as a human-operated system for mapping is one of the benchmarks.
-> 
+
+We added the NavVis VLX as a reference in the state of the art section.
+
 > - The authors mention unknown and dangerous environments as a
@@
 > systems for the reader.
-> 
+
+We added four more terrestial, human-made environments as dangerous examples where a compact, spherical robot would really unfold its potential.
+
 > - The advantages of a spherical mobile platforms are discussed, but
@@
 > influence the acquisition?
-> 
+
+We added disadvantages (including your example) in the introduction. They show the need for a reliable, robust registration procedure.
+There is an approach to prevent dust and dirt to accumulate on the outer shell. We added a note / reference to that.
+
 > - Section 2.1 discusses existing laser scanner-based SLAM
@@
 > and Systems, DOI: https://doi.org/10.1109/IROS.2006.282527
-> 
+
+Removed the false statement.
+BIG FIXME: Get more references on plane based slam.
+
 > - The introduction in Section 3 should be revised as it is not very
@@
 > directly related to the general lack of outline beforehand.
-> 
+
+We added an outline for section 3.
+
 > - In section 3, the authors introduce their own algorithm for
@@
 > with a window approach or even globally?
-> 
+
+Globally. I hope we made that clear in the introduction of section 3 now.
+
 > - Section 3.2 deals with the extraction of planes, but the following
@@
 > especially with regard to existing measurement and motion noise?
-> 
+
+We added more detail in the plane extraction subsection.
+
 > - Throughout Section 3, numerous variables are not introduced and
@@
 > difficult.
+
+TODO: Anton? Have you done it already?
+
 > 
@@
 
-FIXME: -> added todo REDO experiment
+We redid the experiment with a larger amount of points (all that have
+a point-to-point distance of less than 3000 cm). The second significant
+peak was a artifact of the previous thresholding and hence is now not 
+significant anymore.
 

### 2021-07-05 Latex/sections/introduction.tex
Done reviewer one comments
second reviewers comments addressed

@@
 For example, the shell stops any dust that deteriorates sensors or actuators when settling at sensitive locations.
-This is particularly useful for unknown and dangerous environments.
-E.g., for space applications, in the DAEDALUS study~\cite{RossiMaurelliUnnithanetal.2021}, a spherical robot is proposed that is to be lowered into a lunar cave and create a 3D map of the environment.
-The authors choose this format as the present moon regolith is known to damage instruments and other components.
+In contrast to an usual enclosing the shell can separate the sensors entirely from the environment without the need for a number of points-of-connection. 
+A strict requirement then is that the shell is very durable.
+This is particularly useful for unknown or dangerous environments.
+E.g., old buildings that are in danger of collapsing, narrow underground funnels, construction sites, or mining shafts. 
+The spherical format is, in fact, also suited for space applications.
+In the DAEDALUS study~\cite{RossiMaurelliUnnithanetal.2021}, such a robot is proposed that is to be lowered into a lunar cave and create a 3D map of the environment.
+The authors choose this format as the moon regolith is known to damage instruments and other components.
+They also present an approach to protect the shell from accumulating dust and dirt.
+However, the spherical format also comes with some disadvantages.
+Lidar sensors often have a minimum scan distance, resulting in a less dense (or empty) point cloud when the scanner looks on the ground, whereas density is higher when looking in other directions.
+The ground itself is likely to be less populated with points, due to weak angles of incidence while mapping it.
+Furthermore, relying on IMU based odometry as a localization technique alone yields inaccurate and noisy pose measurements.
+Therefore, a robust registration procedure is needed for spherical robots, that is able to cope with vast differences is point cloud density and high noise regarding pose measurements.
 
@@
 The proposed registration method minimizes the distances of each point to its corresponding plane as an objective function.
-The method is evaluated on a simulated dataset as well as on experimentally acquired datasets.
+
+The next sections introduce you to the state of the art SLAM algorithms for spherical robots and plane based registration.
+We then present our approach, which includes global plane extraction, point to plane correspondences, and an optimized gradient descent that minimizes point to plane distances.
+Furthermore, we show the results on simulated, as well on real world datasets.

### 2021-07-05 Latex/sections/registration.tex
second reviewers comments addressed

@@
 \section{Registration Algorithm}
+
+This section describes the presented algorithm.
+We first pre-register all individual scans to align them in one global coordinate system.
+Then, planes are extracted from this global scan model.
+The key to improving map quality is to find a model for each individual scan, where each point is correctly identified to belong to one of the globally extracted planes.
+Then, each scan is optimized by finding a 6D pose-transformation that minimizes the distance of each correspondence.
+Note that all of this takes place on a reduced version of the scan, i.e. an octree based reduction where each voxel is only allowed to have a certain amount of points.
+This ensures a homogenous point density, which is favourable for the error function.
+It also decreases runtime, while preserving the planar features of the environment.
 
@@
 After having done the pre-registration, the scans are aligned well enough to make statements about the potential planes in the environment.
-To find the planes in the environment, a Randomized Hough transform with an accumulator ball as described in~\cite{3DRESEARCH2011} is used as this method prefers dominant planes such as the building structure over smaller planar surfaces.
+We extract planes only once from the global scan model, or a portion of that global scan model.
+Sometimes, the first few line scans, or the first half of the global model, is sufficient for plane extration.
+This is because within the initial movements of the robot system, pose error has not accumulated for very long, resulting in a less dense, yet less distorted representation of the environment.
+For mobile robots this yields a quite consistent strategy for global plane extraction, since all the subsequent scans (where pose error keeps accumulating) get matched with the initial planar representation.
+However, especially for large datasets, it is necessary to update the global plane model multiple times while establishing point to plane correspondences.  
+Currently, this is not done, and is one of the primary objectives for future works.
+
+To find the planes in the environment, a Randomized Hough transform (RHT) with an accumulator ball as described in~\cite{3DRESEARCH2011} is used as this method prefers dominant planes such as the building structure over smaller planar surfaces.
+The RHT is combined with a region growing approach, similar plane merging, and a flatness filter that is based on principal component analysis (PCA).
+Once a plane is identified, we calculate the convex hull of that plane, representing the extent in all directions. 
+This way, large planar structures such as walls or the ceiling get identified.
+Note that smaller, non-perpendicular planes such as doors or cupboards potentially get identified, too, if they pass the PCA filter.
+
 In the next subsection, the Hesse-normal form of the plane is required.
 For an ideal plane the orthogonal distance from the origin $\rho_{\mathcal{P}_k}$ is computed via $\vec{n}_{\mathcal{P}_k}\cdot\vec{p}_i$, where $\vec{p}_i = \begin{bmatrix}x_i&y_i&z_i\end{bmatrix}^T$ is an arbitrary point on the plane and $\vec{n}_{\mathcal{P}_k} = \begin{bmatrix}n_{\mathcal{P}_k}^x&n_{\mathcal{P}_k}^y&n_{\mathcal{P}_k}^z\end{bmatrix}^T$ is the normal vector of that plane.
-To find this point, we first define a convex hull that encloses the plane.
-It is defined by the points that lie on the plane with the furthest distance to one another.
-We then choose the center point of the convex hull of the plane as $\vec{p}_i$.
+To find this point, we use the convex hull, as it is defined by the points that lie on the plane with the furthest distance to one another.
+We choose the center point of the convex hull of the plane as $\vec{p}_i$.
 
@@
 
+\todo{@fabi Introduce PPD, think about clustering?}
 There are many ways of solving the segmentation problem of assigning each point to a plane often based on the RANSAC method \cite{Honti2018}, \cite{Gaspers2011} but also deep learning approaches \cite{Engelmann2018}.

### 2021-07-05 Latex/sections/stateOfTheArt.tex
Done reviewer one comments
second reviewers comments addressed

@@
 also propose an online method using a novel combination of a hierarchical graph structure with local multi-resolution maps to overcome problems due to sparse scans~\cite{Droeschel2018}.
+Another intriguing example is the NavVis VLX system \cite{navvis}.
+They perform real time SLAM on a mobile, wearable platform, producing colored point clouds with high accuracy by combining it with a camera system.
+However, this platform has to be moved by a human operator. 
+They also employ artificial markers that can be placed in the environment, which get recognized by the camera system to further improve registration accuracy.
+Obviously, this is not possible in inaccessible, or dangerous environments.
 
@@
 For low-cost LiDARs, this implies slow motion and long integration time.
-Further, none of the approaches exploit the structure inherent to human-made environments, hence missing out on possible advantages.
 
@@
 For plane-to-plane correspondences, this is necessary to gather enough data to measure planes in each scan robustly. 
+In particular this pause is needed when the field-of-view (FOV) is limited, as the detected planes are thin slices of the true planes which are difficult to find correct correspondences for. 
 The resulting scan procedure is stop-scan-and-go.

### 2021-07-05 Latex/sections/registration.tex
projection polygon distance added

@@
     \SetAlgoLined
-    \KwResult{A corrected map}
+    \KwResult{A corrected map and 6D path.}
     1. Pre-register the scans\;
-    2. Extract planes from the registered map\;
+    2. Extract planes from the registered map globally\;
     3. Further improve map by solving the optimization that minimizes the distance of all points to their respective planes\;
@@
 We extract planes only once from the global scan model, or a portion of that global scan model.
-Sometimes, the first few line scans, or the first half of the global model, is sufficient for plane extration.
+The first few line scans, e.g. the first half of the global model, is sufficient for plane extration.
 This is because within the initial movements of the robot system, pose error has not accumulated for very long, resulting in a less dense, yet less distorted representation of the environment.
@@
 To find this point, we use the convex hull, as it is defined by the points that lie on the plane with the furthest distance to one another.
-We choose the center point of the convex hull of the plane as $\vec{p}_i$.
+We choose the center point of the convex hull of the plane as $\vec{a}_{\rho_k}$.
 
@@
 
-\todo{@fabi Introduce PPD, think about clustering?}
+\begin{figure*}
+	\centering
+	\includegraphics[width=\textwidth]{images/project}
+	\caption{Illustration of the Hesse- and polygon projection distance.
+The point p (red) gets projected onto the infinitley extending global plane.
+Both the point projection (green) and the global planes convex hull (blue) get projected into 2D space (grey).
+The Hesse distance, i.e.
+the shortest distance to the infinitley extending plane (red), is shown as well as the minimum distance from the 2D point projection to the polygon (purple).}
+	\label{fig:proj}
+\end{figure*}
+
 There are many ways of solving the segmentation problem of assigning each point to a plane often based on the RANSAC method \cite{Honti2018}, \cite{Gaspers2011} but also deep learning approaches \cite{Engelmann2018}.
@@
 Which is why a simple algorithm is sufficient:
-For each plane, consider adding a correspondence for all points closer to the plane than a threshold $\epsilon$.
-Depending on the threshold value, some points now have multiple correspondences if their distance to multiple planes is less than $\epsilon$ \todo{@Fabi add actual $\epsilon$ values to the experiments}.
-We then omit all points with correspondences to multiple planes from the optimization to avoid introducing errors due to the plane ambiguity.
+For each globaly extracted plane, consider adding a correspondence for all points that are close to that plane.
+To do this, we employ two distance models to represent the distance from a point to a plane.
+
+The Hesse-normal form describes the distance from the $i$-th transformed point $T(\vec{p}_i)$ to the ever extending, infinetly large $k$-th plane in 3D vector space
+\begin{equation}\label{eq:hesse}
+	D_{h}^{i,k} = \vec{n}_{\mathcal{P}_k} \cdot \left[ T(\vec{p}_i) - \vec{a}_{\mathcal{P}_k} \right] ,
+\end{equation}
+with normal $\vec{n}_{\mathcal{P}_k}$ and receiving point $\vec{a}_{\mathcal{P}_k}$ of the $k$-th plane.
+The distance $D_{h}^{i,k}$ of the $i$-th point to the $k$-th plane reflects the length of the line segment, constructed from the $i$-th point to its projection on the $k$-th plane.
+Therefore, the line segment is parallel to the normal vector of the plane by definition.
+Thus, the $i$-th points projection $\widetilde{T}(\vec{p}_i)$ onto the $k$-th plane, which is required later, is easily calculated by shifting the point against normal direction:
+\begin{equation}\label{eq:projection}
+	\widetilde{T}(\vec{p}_i) = T(\vec{p}_i) - D_{h}^{i,k} \cdot \vec{n}_{\mathcal{P}_k}
+\end{equation}
+
+The simplicity of the Hesse distance is at contrast with its inability to represent the extent of the plane.
+It becomes obvious that another distance model, based on the polygonic nature of the plane, is required: the polygon projection distance (PPD).
+The convex hull is able of representing the expansion of the plane in all directions, and thus is utilized as a distance model.
+First, the corresponding point gets projected onto the ever extending, infinetly large plane representation given by the Hesse form.
+Keeping track of the points projection $\widetilde{T}(\vec{p}_k)$ from eq.
+\ref{eq:projection}, i.e.
+the green point from figure \ref{fig:proj}, is essential for calculating a points distance to a polygon in 3D.
+It represents the points "shadow" onto the same plane where the polygon lives in. 
+Then, the 3D polygon, i.e. the points that make up the convex hull, as well as the corresponding 3D point projection, are projected again into a 2D vector space, using the direction of the plane.
+E.g. in figure \ref{fig:proj}, the blue planes normal vector has its mayor component in z-direction (upwards), thus the projection polygon distance is calculated in the xy-plane.
+Using the maximum of the absolute magnitude of the normal vector ensures that the most sensible 2D projection is used for every direction.
+Sunday \cite{pga} presents an optimal algorithm to check wether the point projection lies inside the polygon in 2D. 
+If the point is inside the polygon, its PPD is set to zero.
+If, however, the point is outside the polygon, the shortest distance to the polygon is calculated by looking at the minimum distance of the $i$-th point to each line segment, making up the convex hull of the $k$-th plane in 3D.
+Let the polygon $P_k$ be made up of line segments $s_{m, j}$.
+The line segment $s_{m, j}$ consists of the points $p_m$ and $p_j$.
+The line segment is parameterized by
+\begin{equation}
+\label{eq:line}
+	s_{m, j}(t) = p_m + t \cdot \left( p_j - p_m \right) ,  
+\end{equation}
+where $t \in \left[ 0, 1 \right]$. 
+We setup a distance function, which measures the distance from the query point $p_i$ to an arbitrary point on the line segment.
+\begin{equation}
+	\label{eq:dist}
+	d_{m, j}(t) = \lVert s_{m, j}(t) - p_i \rVert 
+\end{equation} 
+It is now possible to find the shortest distance to the line segment by finding the argument of the function that minimizes this distance.
+Therefore, we set
+\begin{equation}
+	\pardiff{t} d_{m, j}(t)^2 \overset{!}{=} 0 ,
+\end{equation}  
+resulting in 
+\begin{equation}
+	t_0 = \frac{\left( p_j - p_m \right) \cdot \left( p_i - p_m \right)}{\left( p_j - p_m \right)^2} .
+\end{equation}
+Note the possibility of $t_0 \notin \left[ 0, 1 \right]$. 
+In that case, the projection of the point $p_i$ onto the line given by eq. \ref{eq:line} is not between $p_j$ and $p_m$. 
+Instead, its projection onto the line falls outside of the segment.
+By constraining $t_0$ we get 
+\begin{equation}
+\label{eq:minmax}
+	\hat t = \min \left( \max \left( t_0, 0\right) , 1 \right) ,
+\end{equation}
+which is the argument that gives the shortest distance to the line segment, when inserted to eq. \ref{eq:dist}.
+We find the PPD from the point $p_i$ to the polygon $P_k$ by calculating the minimum shortest distance over all line segments that make up the polygon.
+
+Now, we say that a point is "close" to a plane, if the Hesse distance is smaller than $\epsilon_H$ and the PPD is smaller than $\epsilon_P$. 
+Depending on the threshold values, some points are contemplable having correspondences with multiple planes, if both distances fall below their respective threshold.
+\todo{@Fabi add actual $\epsilon$ values to the experiments}.
+We then chose to establish a correspondence to the plane with the smallest PPD.
+If there are still ambiguities, i.e. multiple PPD's with value zero, a correspondence is established with the plane having the smallest PPD and smallest Hesse distance.
 

### 2021-07-06 Latex/sections/registration.tex
Proofread fabis point2plane correspondence section

@@
 	\caption{Illustration of the Hesse- and polygon projection distance.
-The point p (red) gets projected onto the infinitley extending global plane.
+The point p (red) gets projected onto the infinitely extending global plane.
 Both the point projection (green) and the global planes convex hull (blue) get projected into 2D space (grey).
 The Hesse distance, i.e.
-the shortest distance to the infinitley extending plane (red), is shown as well as the minimum distance from the 2D point projection to the polygon (purple).}
+the shortest distance to the infinitely extending plane (red), is shown as well as the minimum distance from the 2D point projection to the polygon (purple).}
 	\label{fig:proj}
@@
 
-There are many ways of solving the segmentation problem of assigning each point to a plane often based on the RANSAC method \cite{Honti2018}, \cite{Gaspers2011} but also deep learning approaches \cite{Engelmann2018}.
+There are many ways of solving the segmentation problem of assigning each point to a plane.
+They are often based on the RANSAC method \cite{Honti2018}, \cite{Gaspers2011} but also deep learning approaches \cite{Engelmann2018} have been successful.
 The quality of the preregistration directly influences the quality of the plane detection and therefore the quality of the final registration result. 
 We assume small enough errors in pre-registration to allow for plane detection.
-Which is why a simple algorithm is sufficient:
-For each globaly extracted plane, consider adding a correspondence for all points that are close to that plane.
+Therefore, a simple algorithm is sufficient:
+For each globally extracted plane, consider adding a correspondence for all points that are close to that plane.
 To do this, we employ two distance models to represent the distance from a point to a plane.
 
-The Hesse-normal form describes the distance from the $i$-th transformed point $T(\vec{p}_i)$ to the ever extending, infinetly large $k$-th plane in 3D vector space
+The Hesse-normal form describes the distance from the $i$-th transformed point $T(\vec{p}_i)$ to infinitely extending, $k$-th plane in 3D vector space
 \begin{equation}\label{eq:hesse}
@@
 \end{equation}
-with normal $\vec{n}_{\mathcal{P}_k}$ and receiving point $\vec{a}_{\mathcal{P}_k}$ of the $k$-th plane.
+with normal $\vec{n}_{\mathcal{P}_k}$ and supporting point $\vec{a}_{\mathcal{P}_k}$ of the $k$-th plane.
 The distance $D_{h}^{i,k}$ of the $i$-th point to the $k$-th plane reflects the length of the line segment, constructed from the $i$-th point to its projection on the $k$-th plane.
-Therefore, the line segment is parallel to the normal vector of the plane by definition.
+Therefore, by definition the line segment is parallel to the normal vector of the plane.
 Thus, the $i$-th points projection $\widetilde{T}(\vec{p}_i)$ onto the $k$-th plane, which is required later, is easily calculated by shifting the point against normal direction:
 \begin{equation}\label{eq:projection}
-	\widetilde{T}(\vec{p}_i) = T(\vec{p}_i) - D_{h}^{i,k} \cdot \vec{n}_{\mathcal{P}_k}
+	\widetilde{T}(\vec{p}_i) = T(\vec{p}_i) - D_{h}^{i,k} \vec{n}_{\mathcal{P}_k}
 \end{equation}
 
-The simplicity of the Hesse distance is at contrast with its inability to represent the extent of the plane.
-It becomes obvious that another distance model, based on the polygonic nature of the plane, is required: the polygon projection distance (PPD).
-The convex hull is able of representing the expansion of the plane in all directions, and thus is utilized as a distance model.
-First, the corresponding point gets projected onto the ever extending, infinetly large plane representation given by the Hesse form.
-Keeping track of the points projection $\widetilde{T}(\vec{p}_k)$ from eq.
-\ref{eq:projection}, i.e.
-the green point from figure \ref{fig:proj}, is essential for calculating a points distance to a polygon in 3D.
-It represents the points "shadow" onto the same plane where the polygon lives in. 
-Then, the 3D polygon, i.e. the points that make up the convex hull, as well as the corresponding 3D point projection, are projected again into a 2D vector space, using the direction of the plane.
-E.g. in figure \ref{fig:proj}, the blue planes normal vector has its mayor component in z-direction (upwards), thus the projection polygon distance is calculated in the xy-plane.
+The simplicity of the Hesse distance is at contrast with its inability to take the extent of the plane into account.
+Therefore, a second distance model is introduced: the polygon projection distance (PPD).
+The convex hull represents the extend of a plane by forming a polygon with the outer most points that are assigned to a given plane. 
+It is able to represent the expansion of the plane in all directions, and thus is utilized as a distance model.
+To find the PPD, first a corresponding point gets projected onto the infinitely extending plane representation given by the Hesse form, i.e.
+$\widetilde{T}(\vec{p}_k)$ is found from eq.~\eqref{eq:projection} (green point in~\ref{fig:proj}).
+Then, the 3D polygon, i.e.
+the points that make up the convex hull, as well as the corresponding 3D point projection, are projected again into a 2D vector space, using the maximal component of the normal vector of the plane.
+E.g. in figure~\ref{fig:proj}, the blue planes normal vector has its mayor component in z-direction (upwards), thus the projection polygon distance is calculated in the xy-plane.
 Using the maximum of the absolute magnitude of the normal vector ensures that the most sensible 2D projection is used for every direction.
-Sunday \cite{pga} presents an optimal algorithm to check wether the point projection lies inside the polygon in 2D. 
+Sunday et al.~\cite{pga} presents an optimal algorithm to check whether the point projection lies inside the polygon in 2D. 
 If the point is inside the polygon, its PPD is set to zero.
@@
 where $t \in \left[ 0, 1 \right]$. 
-We setup a distance function, which measures the distance from the query point $p_i$ to an arbitrary point on the line segment.
+We set up a distance function, which measures the distance from the query point $p_i$ to an arbitrary point on the line segment.
 \begin{equation}
@@
 Note the possibility of $t_0 \notin \left[ 0, 1 \right]$. 
-In that case, the projection of the point $p_i$ onto the line given by eq. \ref{eq:line} is not between $p_j$ and $p_m$. 
+In that case, the projection of the point $p_i$ onto the line given by eq.~\ref{eq:line} is not between $p_j$ and $p_m$. 
 Instead, its projection onto the line falls outside of the segment.
@@
 
-Now, we say that a point is "close" to a plane, if the Hesse distance is smaller than $\epsilon_H$ and the PPD is smaller than $\epsilon_P$. 
-Depending on the threshold values, some points are contemplable having correspondences with multiple planes, if both distances fall below their respective threshold.
+We consider a point as ``close'' to a plane, if the Hesse distance is smaller than some threshold $\epsilon_H$ and the PPD is additionally smaller than some other threshold $\epsilon_P$. 
+Depending on the threshold values, some points might have correspondences with multiple planes, if both distances fall below their respective threshold.
 \todo{@Fabi add actual $\epsilon$ values to the experiments}.
-We then chose to establish a correspondence to the plane with the smallest PPD.
+We then choose to establish a correspondence to the plane with the smallest PPD.
 If there are still ambiguities, i.e. multiple PPD's with value zero, a correspondence is established with the plane having the smallest PPD and smallest Hesse distance.
@@
 
-Assuming we know now the Hesse-normal form of all planes and all assigned points to these planes, we register the points to have a minimal distance to their respective planes.
+Assuming we now know the Hesse-normal form of all planes and all assigned points to these planes, we register the points to have a minimal distance to their respective planes.
 The transformation $T(\vec{p}_i)$ of each point $\vec{p}_i$ with respect to a 6 DoF motion is described in homogeneous coordinates using the roll-pitch-yaw ($\varphi-\vartheta-\psi$) Tait-Brian angles as in~\cite{diebel2006representing}.

### 2021-07-06 Latex/sections/stateOfTheArt.tex
sota modified

@@
 also propose an online method using a novel combination of a hierarchical graph structure with local multi-resolution maps to overcome problems due to sparse scans~\cite{Droeschel2018}.
-Another intriguing example is the NavVis VLX system \cite{navvis}.
+Another intriguing example is the NavVis VLX system~\cite{navvis}.
 They perform real time SLAM on a mobile, wearable platform, producing colored point clouds with high accuracy by combining it with a camera system.
@@
 
+LOAM~\cite{Zhang2014} is the baseline-algorthm that provides a real-time and lowdrift solution based on two parallel registration algorithms using planes and lines.
+Unfortunately, it is not open-source anymore.
+LOAM livox extends the LOAM framework to the the rotating prism scanner with small FoV~\cite{Lin_2020}.
+Zhou, Wang, and Kaess write that it also adopts the parallel computing to achieve real-time global registration.
+Parallel computing needs a powerful CPU, it may be not suitable for an embedded system which has limited computational resources~\cite{Zhou_2021}.
+Thus, they extens their smoothing and mapping to the LiDAR and planes case.
+In their experiments they used a VLP-16 LiDAR to collect indoor datasets.
+
+Further recent planar SLAM approaches include~\cite{Jung_2015,Grant_2018,Geneva_2018, Zhou_2021,   wei2021groundslam}.
+While Wei et al. uses only the ground plane in outdoor experiments.
+Indoors, Jung et al.
+used in 2015 several Hokuyo laser scanners for their kinematic scanning system~\cite{Jung_2015}, while Grant used a single Velodyne HDL-32E sensor with 32 rotating laser/receiver pairs mounted on a backpack and walked through different environments~\cite{Grant_2018}.
+The LIPS system~\cite{Geneva_2018} employes a so-called Closest Point Plane Representation with an Anchor Plane Factor.
+RANSAC is used to find the places.
+Their system couples an eight channel Quanergy M8 LiDAR operating at 10Hz with a Microstrain 3DM-GX3-25 IMU attached to the bottom of the LiDAR operating at 500Hz.
+
+\bigskip
+
 In the following, we propose this combination of point-to-point based pre-registration followed by a point-to-plane based optimization.

### 2021-07-06 Latex/rebuttal.txt
sota modified

@@
 
-We clarified that also the lilmited FOV plays a role in not finding 
+We clarified that also the limited FOV plays a role in not finding 
 planes for robust correspondences. 
 FIXME @Andreas and @Dorit: I (Anton) disagree can you double check?
+
+I (Andreas) seconds the statement, that the limited FOV plays a role
+in not finding planes for robust correspondences.
 
@@
 BIG FIXME: Get more references on plane based slam.
+
+I (Andreas) has added a coupe of references.
 

### 2021-07-07 Latex/sections/conclusion.tex
Clustering added. TODO: results.

@@
 
-However, as section~\ref{sec:experimentalSetup} shows, some outliers remain in the registered point cloud due to the simplistic point-to-plane correspondence model.
-This is to be improved in future studies.
-One idea to do so is to employ a local clustering technique onto the individual scans before registering them globally. 
-Points from the scan with a similar normal get clustered together if their distance is small enough.
-This way, a cluster represents a local plane while at the same time storing all the points that belong to that plane.
-Point-to-plane correspondences then arise from the intersection of local clusters with global planes.
-
 %
@@
 I.e., instead of locking some dimensions entirely from being used for optimization, they are weighted based on the dynamics of the system that encode which noise source is more likely.
+
+%
+The biggest issue with the presented algorithm currently is that the global plane model is established only once as initialization, but never updated afterwards.
+Utilizing the local planar clustering (LPC) of the individual line scans for updating the global plane model step by step is currently the primary objective of future studies.
+One idea is to calculate the convex hulls of the LPC, and combine them with the global planes sequentially.
+Therefore, we expect to increase robustness and autonomy of the registration procedure.
+
 %

### 2021-07-07 Latex/sections/registration.tex
Clustering added. TODO: results.

@@
 
+There are many ways of solving the segmentation problem of assigning each point to a plane.
+They are often based on the RANSAC method \cite{Honti2018}, \cite{Gaspers2011} but also deep learning approaches \cite{Engelmann2018} have been successful.
+The quality of the preregistration directly influences the quality of the plane detection and therefore the quality of the final registration result.
+We assume small enough errors in pre-registration to allow for plane detection.
+We employ two distance models to represent the distance from a point to a plane, and combine them with a local planar clustering (LPC) approach to establish point-to-plane correspondences.
+
+\subsubsection{Distance Models}
+
 \begin{figure*}
@@
 \end{figure*}
-
-There are many ways of solving the segmentation problem of assigning each point to a plane.
-They are often based on the RANSAC method \cite{Honti2018}, \cite{Gaspers2011} but also deep learning approaches \cite{Engelmann2018} have been successful.
-The quality of the preregistration directly influences the quality of the plane detection and therefore the quality of the final registration result. 
-We assume small enough errors in pre-registration to allow for plane detection.
-Therefore, a simple algorithm is sufficient:
-For each globally extracted plane, consider adding a correspondence for all points that are close to that plane.
-To do this, we employ two distance models to represent the distance from a point to a plane.
 
@@
 
-We consider a point as ``close'' to a plane, if the Hesse distance is smaller than some threshold $\epsilon_H$ and the PPD is additionally smaller than some other threshold $\epsilon_P$. 
-Depending on the threshold values, some points might have correspondences with multiple planes, if both distances fall below their respective threshold.
-\todo{@Fabi add actual $\epsilon$ values to the experiments}.
-We then choose to establish a correspondence to the plane with the smallest PPD.
-If there are still ambiguities, i.e. multiple PPD's with value zero, a correspondence is established with the plane having the smallest PPD and smallest Hesse distance.
+\subsubsection{Local Planar Clustering}
+
+The presented algorithm employs LPC on each individual line scan.
+The idea is to find planar features utilizing an optimized approximate k nearest neighbour (AKNN) search for normal calculation as in~\cite{aknnbbs}.
+Thus, normals are calculated for every point in the respective line scan.
+Then, the local normal model is combined with a region growing approach.
+The region growing is implemented by storing the k-nearest neighbors in a queue, which represents the iteration order of the clustering algorithm.
+A region grows if one of these neighbors has a distance shorter then a treshold $d_{growth}$ and additionally has a similar normal.
+The angle between two normalized vectors is
+\begin{equation}
+\label{eq:smallestalpha}
+	\alpha(\vec{n}_1, \vec{n}_2) = \arccos(\vec{n}_1 \cdot \vec{n}_2) \; .
+\end{equation}
+However, the \textit{smallest} angle between two \textit{normal} vectors is 
+\begin{equation}
+\label{eq:alphahat}
+\hat \alpha(\vec{n}_1, \vec{n}_2) = 
+\begin{cases}
+	2\pi - \alpha(\vec{n}_1, \vec{n}_2) & \text{if  } \alpha(\vec{n}_1, \vec{n}_2) > \frac{3}{2} \pi \\
+	\alpha(\vec{n}_1, \vec{n}_2) - \pi & \text{if  } \alpha(\vec{n}_1, \vec{n}_2) > \pi \\
+	\pi - \alpha(\vec{n}_1, \vec{n}_2) & \text{if  } \alpha(\vec{n}_1, \vec{n}_2) > \frac{\pi}{2} 
+\end{cases}
+\; ,
+\end{equation}
+because the opposing normals $-\vec{n}_1$ and $-\vec{n}_2$ always have to be considered, since they correspond to the same plane.
+We say that two normals are similar if the angle $\hat \alpha$ from eq. \ref{eq:alphahat} is smaller than a threshold $\epsilon_{\alpha}$.
+
+Planar areas like walls are therefore connected in one large cluster, while non-feature parts, i.e. non-planar or detached parts of the scan will have their own, smaller cluster.
+You are then left with a rather easy task: filtering the clusters.
+The filtering has two parts.
+First, to combine clusters that are similar, i.e. have a short distance ($d_{growth}$) to each other and similar mean normals.
+Second, to filter out all clusters that contain non-feature points.
+Since all the non-feature points belong to small clusters, the presented algorihtm does this by putting a minimum threshold $N_{c,min}$ on the cluster size.
+Figure \ref{fig:clustering} shows this concept on a real world dataset.
+It was obtained by a floating sphere on a desk in an office environment, as described later in section \ref{sec:experimentalSetup}.
+
+\begin{figure*}
+	\includegraphics[width=\linewidth]{images/clustering}
+	\caption{(Left) Point normals using the AKNN method with $K = 20$.
+(Middle) Result of the region growing, before applying the second filter.
+(Right) Resulting clusters after applying the second filter, using $N_{c,min} = 300$.
+}
+	\label{fig:clustering}
+\end{figure*}
+
+Point-to-Plane correspondences are now established.
+The distance models (Hesse + PPD) are utilized to check wether a cluster overlaps with any of the global planes, i.e.
+any point in the cluster has a Hesse distance smaller than some threshold $\epsilon_H$ and a PPD smaller than some additional threshold $\epsilon_P$, to the global plane.
+For each overlap, we find the smallest angle between the corresponding cluster normal and the global plane normal, according to eq. \ref{eq:alphahat}.
+Finally, a correspondence is established between all points in the cluster and the global plane, if the minimum angle between their normals is smaller than $\epsilon_\alpha$.
 

### 2021-07-07 Latex/sections/results.tex
Cylon results renewed and params added
First LPC results added.

@@
 View from the interior that shows the used planes (top) and a birds-eye view (bottom).
-An animation of the registration process is given at \url{https://youtu.be/8XdIUN_9VpY} .}
+Parameters used for optimization: $\epsilon_H = 50$, $\epsilon_P = 200$, $\epsilon_\alpha = \frac{\pi}{4}$, $K = 20$ for AKNN, $N_{c_min} = 200$, $\alpha_0 = [0.01, 0.01, 0.01, 0, 0, 0]^\tau$, $i = 800$, $k = 1$.
+An animation of the registration process is given at \url{https://youtu.be/Qa2Qi8z0KEk} .}
 	\label{fig:cylon-corrected}

### 2021-07-07 Latex/sections/results.tex
region growing threshold added in results

--- 2021-07-04 Latex/sections/results.tex
@@
 View from the interior that shows the used planes (top) and a birds-eye view (bottom).
-An animation of the registration process is given at \url{https://youtu.be/8XdIUN_9VpY} .}
+Parameters used for optimization: $\epsilon_H = 50$, $\epsilon_P = 200$, $\epsilon_\alpha = \frac{\pi}{4}$, $K = 20$ for AKNN, $d_{growth} = 50$, $N_{c_min} = 200$, $\alpha_0 = [0.01, 0.01, 0.01, 0, 0, 0]^\tau$, $i = 800$, $k = 1$.
+An animation of the registration process is given at \url{https://youtu.be/Qa2Qi8z0KEk} .}
 	\label{fig:cylon-corrected}

### 2021-07-07 Latex/sections/registration.tex
Proofread LPC section

@@
 The presented algorithm employs LPC on each individual line scan.
-The idea is to find planar features utilizing an optimized approximate k nearest neighbour (AKNN) search for normal calculation as in~\cite{aknnbbs}.
+The idea is to find planar features utilizing an optimized approximate k nearest neighbor (AKNN) search for normal calculation as in~\cite{aknnbbs}.
 Thus, normals are calculated for every point in the respective line scan.
@@
 The region growing is implemented by storing the k-nearest neighbors in a queue, which represents the iteration order of the clustering algorithm.
-A region grows if one of these neighbors has a distance shorter then a treshold $d_{growth}$ and additionally has a similar normal.
-The angle between two normalized vectors is
+A region grows if one of these neighbors has a distance shorter then a threshold $d_{growth}$ and additionally has a similar normal.
+We define similarity of two normals via the angle between them, which is:
 \begin{equation}
@@
 Planar areas like walls are therefore connected in one large cluster, while non-feature parts, i.e. non-planar or detached parts of the scan will have their own, smaller cluster.
-You are then left with a rather easy task: filtering the clusters.
-The filtering has two parts.
-First, to combine clusters that are similar, i.e. have a short distance ($d_{growth}$) to each other and similar mean normals.
+Finally the clusters are filtered in two steps.
+First, to combine clusters that are similar, i.e. have a short distance ($d_{growth}$) to each other and similar mean normals (similarity defined as before).
 Second, to filter out all clusters that contain non-feature points.
 Since all the non-feature points belong to small clusters, the presented algorihtm does this by putting a minimum threshold $N_{c,min}$ on the cluster size.
-Figure \ref{fig:clustering} shows this concept on a real world dataset.
-It was obtained by a floating sphere on a desk in an office environment, as described later in section \ref{sec:experimentalSetup}.
+Figure \ref{fig:clustering} shows this concept on the real world dataset described later in section \ref{sec:experimentalSetup}.
 
@@
 Point-to-Plane correspondences are now established.
-The distance models (Hesse + PPD) are utilized to check wether a cluster overlaps with any of the global planes, i.e.
+The distance models (Hesse and PPD) are utilized to check whether a cluster overlaps with any of the global planes, i.e.
 any point in the cluster has a Hesse distance smaller than some threshold $\epsilon_H$ and a PPD smaller than some additional threshold $\epsilon_P$, to the global plane.

### 2021-07-07 Latex/sections/realData.tex
LPC results added, rolling sphere added.

@@
 
-We also tested our algorithm on a real-world dataset that adheres to different motion profiles and laser scanners type with different scanning patterns than the simulated data.
+We also tested our algorithm on two real-world datasets that adhere to different motion profiles and laser scanner types with different scanning patterns.
 
-The dataset is collected by a line scanner, in particular a SICK LMS141 industrial scanner, inside a sphere of a diameter of \SI{20}{\centi\meter} that lies on a floating desk, which is air-pressurized, such that the sphere is hovering.
+The first dataset is collected by a line scanner, in particular a SICK LMS141 industrial scanner, inside a sphere of a diameter of \SI{20}{\centi\meter} that lies on a floating desk, which is air-pressurized, such that the sphere is hovering.
 On this floating desk, the sphere can rotate freely about all axis while being fixed in position. 
@@
 This IMU provides a precision of \SI{76.3}{\micro g} in linear acceleration, \SI{0.02}{\degree\per\second} in angular velocity about the x and y axis, and \SI{0.013}{\degree\per\second} in angular velocity about the z axis \cite{PhidgetSpatial}.
-
 Figure~\ref{fig:float-sat-sphere} shows the experimental setup. The left column of figure~\ref{fig:cylon-corrected} shows the pre-registered resulting point cloud. 
@@
 	\label{fig:float-sat-sphere}
+\end{figure}
+
+The second datasets is collected by a 3D scanner, namely the LIVOX MID 100, which is the same sensor that the simulation uses.
+We put the laserscanner into an acrylic glass spherical shell together with three IMUs and roll it by hand with a slow, constant motion in a home indoor environment.
+Thus, this time almost all six degrees of freedom are used for optimization.
+We exlude a translation in the global up- or downwards direction, though, since we also locked that dimension when collecting the data.
+Figure~\ref{fig:prototype} shows the prototpye.
+The left column of figure~\ref{fig:jasperhome} shows the pre-registered point cloud, applying only the pure transformation found by the recorded pose.
+Although the robot has been moved carefully slow, you see that pose error accumulates in such a way that the same wall is visible multiple times.
+
+\begin{figure}
+  \includegraphics[width=0.49\linewidth]{./images/photoReal1.jpeg}
+  \includegraphics[width=0.50\linewidth]{./images/photoReal3.jpeg}
+  \caption{Prototype used for data acquisition with the rolling sphere.
+The main payload is the Livox Mid-100 laser-scanner.
+For pose-estimation, three IMUs of the manufacturer Phidget are placed inside and a Raspberry Pi 4 for the calculations.
+On the top are two batteries, and on the bottom one voltage stabilizer and breakout box of the laser-scanner.
+}
+  \label{fig:prototype}
 \end{figure}

### 2021-07-07 Latex/sections/registration.tex
LPC results added, rolling sphere added.

@@
 Where
-\begin{align}
-    &\begin{alignedat}{3}
-        \nabla E_\varphi = &n_{\mathcal{P}_k}^y&&(x_i[-S_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] \\&&&+y_i[-S_\varphi C_\psi -C_\varphi S_\vartheta S\psi] \\&&&- z_i C_\varphi C_\vartheta)\\
-        &+ n_{\mathcal{P}_k}^z&&(x_i[C_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] \\&&&+y_i[C_\varphi C_\psi -S_\varphi S_\vartheta S\psi] \\&&&- z_i S_\varphi C_\vartheta)
-    \end{alignedat} \\
-    &\begin{alignedat}{3}
+\begin{equation}
+\begin{split}
+\nabla E_\varphi = n_{\mathcal{P}_k}^y(x_i[-S_\varphi S_\psi+C_\varphi C_\psi S_\vartheta]+y_i[-S_\varphi C_\psi -C_\varphi S_\vartheta S\psi] - z_i C_\varphi C_\vartheta)\\
+        + n_{\mathcal{P}_k}^z(x_i[C_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] +y_i[C_\varphi C_\psi -S_\varphi S_\vartheta S\psi] - z_i S_\varphi C_\vartheta)
+\end{split} \;\; ,
+\end{equation}
+\begin{equation}
+\begin{split}
  \nabla E_\vartheta = & n_{\mathcal{P}_k}^x(-x_iS_\vartheta C_\psi + y_i S_\vartheta S_\psi + z_iC_\vartheta) \\
         &+ n_{\mathcal{P}_k}^y(x_iC_\psi S_\varphi C_\vartheta - y_iS_\varphi C_\vartheta S_\psi + z_i S_\vartheta S_\varphi)  \\
-        &+ n_{\mathcal{P}_i}^z(-x_iC_\varphi C_\psi C_\vartheta + y_iC_\varphi C_\vartheta S_\psi - z_i C_\varphi S_\vartheta)
-    \end{alignedat}\\
-    &\begin{alignedat}{3}
-       \nabla E_\psi =& n_{\mathcal{P}_k}^x&&\left(-x_iC_\vartheta S_\psi - y_iC_\vartheta C_\psi\right) \\
-       &+ n_{\mathcal{P}_k}^y&&(x_i[C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi] \\
-       &&&+ y_i[-C_\varphi S_\psi - S_\varphi S_\theta C_\psi]) \\
-       &+ n_{\mathcal{P}_k}^z&&(x_i[S_\varphi C_\psi + C_\varphi S_\psi S_\theta] \\
-       &&&+ y_i[-S_\psi S_ \varphi + C_\varphi S_\vartheta C_\psi]) 
-    \end{alignedat}
-\end{align}
+        &+ n_{\mathcal{P}_i}^z(-x_iC_\varphi C_\psi C_\vartheta + y_iC_\varphi C_\vartheta S_\psi - z_i C_\varphi 
+\end{split}
+\;\; ,
+\end{equation} 
+\begin{equation}
+\begin{split}
+	\nabla E_\psi = &n_{\mathcal{P}_k}^x\left(-x_iC_\vartheta S_\psi - y_iC_\vartheta C_\psi\right) \\
+       &+ n_{\mathcal{P}_k}^y(x_i[C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi] 
+       + y_i[-C_\varphi S_\psi - S_\varphi S_\theta C_\psi]) \\
+       &+ n_{\mathcal{P}_k}^z(x_i[S_\varphi C_\psi + C_\varphi S_\psi S_\theta] 
+       + y_i[-S_\psi S_ \varphi + C_\varphi S_\vartheta C_\psi]) 
+\end{split}
+\end{equation}
 and $\vec{\Pi}=\begin{bmatrix}\varphi & \vartheta & \psi & t_x & t_y & t_z\end{bmatrix}^T$.

### 2021-07-07 Latex/sections/results.tex
LPC results added, rolling sphere added.

--- 2021-07-04 Latex/sections/results.tex
@@
 The pre-registration is obtained by determining the orientation of the sphere via Madgwick filtered~\cite{madgwick2010efficient} IMU measurements.
-To increase the number of possible point-to-plane correspondences we combine ten temporally successive scans into one meta-scan, which is then globally registered. 
-For this we must assume constant pose over this time duration. 
-However, given the high measurement frequency of the Lidar this introduces only a small error. 
+To increase the number of possible point-to-plane correspondences we combine twenty temporally successive scans into one meta-scan, which is then globally registered. 
+We always use the middlest line scan as a reference coordinate system.
 This does not only speed up convergence due to the proportional effect on the error function but also decreases the risk of transforming the scan in a senseless way.
@@
 	\caption{The point cloud acquired by the floating sphere before (left) and after (right) applying the plane based registration.
-View from the interior that shows the used planes (top) and a birds-eye view (bottom).
-An animation of the registration process is given at \url{https://youtu.be/8XdIUN_9VpY} .}
+View from the interior (top) and a birds-eye view (bottom).
+Parameters used for optimization: $\epsilon_H = 50$, $\epsilon_P = 200$, $\epsilon_\alpha = \frac{\pi}{4}$, $K = 20$ for AKNN, $d_{growth} = 50$, $N_{c_min} = 200$, $\alpha_0 = [0.01, 0.01, 0.01, 0, 0, 0]^\tau$, $i = 8$, $k = 100$.
+An animation of the registration process is given at \url{https://youtu.be/Qa2Qi8z0KEk} .}
 	\label{fig:cylon-corrected}
@@
 
-However, some imprecisions remain. 
-In particular, at the initially very noisy top left corner of the top view, a few points cannot be matched to the plane. 
-The scans that cause these imprecisions are problematic as they measure a number of plane-like structures in the interior of the room. 
-When relabeling those scans, the interior points are associated with a plane, and hence the corresponding distance is minimized.
+\subsection{Rolling Sphere Results}
+
+Figure~\ref{fig:jasperhome} shows the results obtained before and after employing the presented algorithm on the dataset, acquired by the prototype in figure~\ref{fig:prototype}.
+We again combine twenty temporally successive scans into one meta-scan, using the middlest scan as a reference coordinate system, which is then globally registered.
+
+\begin{figure*}
+        \centering
+        \includegraphics[width=0.495\textwidth]{./images/jasper_uncorr_corner}\hfill
+        \includegraphics[width=0.495\textwidth]{./images/jasper_corr_corner}\\
+        \includegraphics[width=0.495\textwidth]{./images/jasper_uncorr_top}\hfill
+        \includegraphics[width=0.495\textwidth]{./images/jasper_corr_top}
+        \caption{The point cloud acquired by the rolling sphere before (left) and after (right) applying the plane based registration.
+View from the interior (top) and a birds-eye view (bottom).
+Parameters used for optimization: $\epsilon_H = 200$, $\epsilon_P = 300$, $\epsilon_\alpha = \frac{\pi}{4}$, $K = 200$ for AKNN, $d_{growth} = 50$, $N_{c_min} = 800$, $\alpha_0 = [0.001, 0.001, 0.001, 1, 0, 1]^\tau$, $i = 1500$, $k = 1$.
+An animation of the registration process is given at \url{https://youtu.be/hYR3xLhIdsE}.}
+        \label{fig:jasperhome}
+\end{figure*}
+After the registration, the walls of the environment are no longer ambiguous, since all the sub-sequent scans got matched with the initial plane representation.
+However, the improved path looks distorted.
+This is because for some scans, especially empty ones where the sensor points to the ground, no planar features got matched with the global plane model, thus their pose is not optimized.
+Improving the path quality is an objective for future works.
 

### 2021-07-07 Latex/rebuttal.txt
Proofread & Added some answers to rebuttall.txt

@@
 TODO: Anton? Have you done it already? 
+Yes - Anton
 
@@
 > resolution?
-> 
+
+The distributions we sample from are specified in the caption of 
+figure 3.
+
 > - In the context of the real experiment in Section 5, a major
@@
 > expectations.
-> 
+
+We added a further experiment that uses nearly all DoF: A sphere rolling
+on a ground plane. This was not done in the previous manuscript because
+we had issues getting the hardware shipped, assembled and the experiment 
+carried out in time.
+
 > - Wouldn't it also be interesting to compare the estimated pose
@@
 > point cloud.
-> 
+
+FIXME: Do we have space & time for this?
+
 > - Why is there no qualitative validation for the results of the real
@@
 > known planes in the registered point cloud?
-> 
+
+FIXME: Can we get a riegl scan of jaspers home? If not corona restrictions were at fault
+We couldn't access the rooms the dataset was acquired in for a terrestrial 
+reference scan.
+
 > - In the Conclusion, the parallelism and perpendicularity within the
@@
 > account.
-> 
+
+We removed the initial mention of parrallelism and perpendicularity of the planes
+as it is of no further interest in the rest of the manuscript.
+
 > Further open questions/comments are:

### 2021-07-07 Latex/sections/conclusion.tex
LPC results added, rolling sphere added.
Proofread & Added some answers to rebuttall.txt

@@
 In this paper, we proposed an approach to mobile mapping using a spherical robot. 
-Using the inherent rotation of the sphere used for locomotion to also rotate a laser scanner also for measurements of the entire environment.
+Given that a spherical robot inherently rotates for locomotion, we use this movement to also rotate a laser scanner, that consequently measures the robots entire environment.
 We propose to post-process the acquire data using a novel registration method for man made environments that exploits the structure of those environments. 
-Given that in human-made environments straight planes are abundantly available, we employ point-to-plane correspondences to improve a pre-registered 3D point cloud. 
-We have evaluated the procedure on a simulated dataset and on an experimentally acquired dataset with different motion profiles. 
-In this evaluation, we have shown that the procedure improves both datasets and yields maps that better resemble human-made, structures. 
+In human-made environments straight planes are abundantly available, hence we employ point-to-plane correspondences to improve a pre-registered 3D point cloud. 
+We have evaluated the procedure on a simulated dataset and on two experimentally acquired datasets with different laser scanners and motion profiles. 
+In this evaluation, we have shown that the procedure improves all datasets and yields maps that better resemble human-made, structures. 
 In particular, the qualitative structure of the environments is reconstructed well. 
@@
 %
-The biggest issue with the presented algorithm currently is that the global plane model is established only once as initialization, but never updated afterwards.
+The biggest issue with the presented algorithm currently is that the global plane model is established only once at initialization, but never updated afterwards.
 Utilizing the local planar clustering (LPC) of the individual line scans for updating the global plane model step by step is currently the primary objective of future studies.
@@
 %
-Currently, we are assembling the sphere of Figure~\ref{fig:robotRender} to test it in real environments.
-Unfortunately, this sphere will be too large to fit our floating setup, thus it will be tested in an office like environment.
-Furthermore, we will test our rolling and scanning spheres on a rail system to achieve repeatable trajectories.
+Furthermore, more experimental evaluation is required, hence we will test our rolling and scanning spheres on a rail system to achieve repeatable trajectories.

### 2021-07-07 Latex/sections/realData.tex
Proofread & Added some answers to rebuttall.txt

@@
 
-The second datasets is collected by a 3D scanner, namely the LIVOX MID 100, which is the same sensor that the simulation uses.
-We put the laserscanner into an acrylic glass spherical shell together with three IMUs and roll it by hand with a slow, constant motion in a home indoor environment.
-Thus, this time almost all six degrees of freedom are used for optimization.
-We exlude a translation in the global up- or downwards direction, though, since we also locked that dimension when collecting the data.
-Figure~\ref{fig:prototype} shows the prototpye.
-The left column of figure~\ref{fig:jasperhome} shows the pre-registered point cloud, applying only the pure transformation found by the recorded pose.
-Although the robot has been moved carefully slow, you see that pose error accumulates in such a way that the same wall is visible multiple times.
+The second datasets is collected by a 3D scanner, namely the LIVOX MID 100, which is the same sensor the simulation uses.
+We put the laser scanner into an acrylic glass spherical shell together with three IMUs  (same specifications as in the first dataset) and manually roll it with a slow, constant motion in a home indoor environment.
+Thus, this time nearly all six degrees of freedom are used for optimization.
+We exclude a translation in the global up- or downwards direction as the sphere rolls on the ground surface, without possibility for up- or downwards motion-
+Figure~\ref{fig:prototype} shows the prototype.
+The left column of figure~\ref{fig:jasperhome} shows the pre-registered point cloud, applying only the transformation defined by the recorded poses.
+Although the robot has been moved carefully, you see that pose error accumulates in such a way that the same wall is sensed multiple times.
 
@@
   \caption{Prototype used for data acquisition with the rolling sphere.
-The main payload is the Livox Mid-100 laser-scanner.
+The main payload is the Livox Mid-100 laser scanner.
 For pose-estimation, three IMUs of the manufacturer Phidget are placed inside and a Raspberry Pi 4 for the calculations.
-On the top are two batteries, and on the bottom one voltage stabilizer and breakout box of the laser-scanner.
+On the top are two batteries, and on the bottom one voltage stabilizer and the breakout box of the laser scanner.
 }

### 2021-07-07 Latex/sections/registration.tex
Proofread & Added some answers to rebuttall.txt

@@
 \begin{split}
-\nabla E_\varphi = n_{\mathcal{P}_k}^y(x_i[-S_\varphi S_\psi+C_\varphi C_\psi S_\vartheta]+y_i[-S_\varphi C_\psi -C_\varphi S_\vartheta S\psi] - z_i C_\varphi C_\vartheta)\\
-        + n_{\mathcal{P}_k}^z(x_i[C_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] +y_i[C_\varphi C_\psi -S_\varphi S_\vartheta S\psi] - z_i S_\varphi C_\vartheta)
+\nabla E_\varphi = &n_{\mathcal{P}_k}^y(x_i[-S_\varphi S_\psi+C_\varphi C_\psi S_\vartheta]\\
+			       & +y_i[-S_\varphi C_\psi -C_\varphi S_\vartheta S\psi] - z_i C_\varphi C_\vartheta)\\
+                   &+ n_{\mathcal{P}_k}^z(x_i[C_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] \\
+                   &+y_i[C_\varphi C_\psi -S_\varphi S_\vartheta S\psi] - z_i S_\varphi C_\vartheta)
 \end{split} \;\; ,
@@
 	\nabla E_\psi = &n_{\mathcal{P}_k}^x\left(-x_iC_\vartheta S_\psi - y_iC_\vartheta C_\psi\right) \\
-       &+ n_{\mathcal{P}_k}^y(x_i[C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi] 
-       + y_i[-C_\varphi S_\psi - S_\varphi S_\theta C_\psi]) \\
-       &+ n_{\mathcal{P}_k}^z(x_i[S_\varphi C_\psi + C_\varphi S_\psi S_\theta] 
-       + y_i[-S_\psi S_ \varphi + C_\varphi S_\vartheta C_\psi]) 
+       &+ n_{\mathcal{P}_k}^y(x_i[C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi]\\
+       &+ y_i[-C_\varphi S_\psi - S_\varphi S_\theta C_\psi]) \\
+       &+ n_{\mathcal{P}_k}^z(x_i[S_\varphi C_\psi + C_\varphi S_\psi S_\theta] \\
+       &+ y_i[-S_\psi S_ \varphi + C_\varphi S_\vartheta C_\psi]) 
 \end{split}
@@
 Although 6D optimization generally works, reducing the optimization space is particularly useful if the source of error in the system is known and a model exists.
-That way, e.g., as we expect a spherical robot to move on a plane the position along the axis perpendicular to the plane is constant and should not be used for optimization.\todo{@Fabi clarify in experiments which locks were active.}
+That way, e.g., as we expect a spherical robot to move on a plane the position along the axis perpendicular to the plane is constant and should not be used for optimization.
 

### 2021-07-07 Latex/sections/results.tex
Proofread & Added some answers to rebuttall.txt

@@
 To increase the number of possible point-to-plane correspondences we combine twenty temporally successive scans into one meta-scan, which is then globally registered. 
-We always use the middlest line scan as a reference coordinate system.
-This does not only speed up convergence due to the proportional effect on the error function but also decreases the risk of transforming the scan in a senseless way.
+We always use the scan at the median index as a reference coordinate system.
+This does not only speed up convergence due to the proportional effect on the error function but also decreases the risk of transforming a single line scan incorrectly.
 Transformations like these happen in particular for a small collection of points as outliers have more influence.
@@
 Figure~\ref{fig:jasperhome} shows the results obtained before and after employing the presented algorithm on the dataset, acquired by the prototype in figure~\ref{fig:prototype}.
-We again combine twenty temporally successive scans into one meta-scan, using the middlest scan as a reference coordinate system, which is then globally registered.
+We again combine twenty temporally successive scans into one meta-scan, using the scan at the median index as a reference coordinate system, which is then globally registered.
 
@@
 \end{figure*}
-After the registration, the walls of the environment are no longer ambiguous, since all the sub-sequent scans got matched with the initial plane representation.
-However, the improved path looks distorted.
+After the registration, the walls of the environment are no longer ambiguous, since all the subsequent scans are matched with the initial plane representation.
+When considering the resulting path, it looks distorted.
 This is because for some scans, especially empty ones where the sensor points to the ground, no planar features got matched with the global plane model, thus their pose is not optimized.

### 2021-07-07 Latex/sections/simulation.tex
Proofread & Added some answers to rebuttall.txt

@@
 The Livox-Mid100 laser scanner consists of three Mid40 scanners arranged to provide an overall field-of-view of \SI{98.4}{\degree} (horizontal) by \SI{38.4}{\degree} (vertical).
-It provides \SI[per-mode=symbol]{300,000}{\pts\per\second} with a range precision of \SI{2}{\centi\meter} up to \SI{20}{\meter} range~\cite{LivoxMid40-100}.
+It provides a rate of \SI[per-mode=symbol]{300,000}{\pts\per\second} with a range precision of \SI{2}{\centi\meter} up to \SI{20}{\meter} range~\cite{LivoxMid40-100}.
 A particular advantage of the Livox laser scanners is their unique scanning pattern.

### 2021-07-12 Latex/rebuttal.txt
Cleaned up rebuttal

@@
 
-FIXME: Do we have space & time for this?
+Unfortunately we couldn't add this due to time constraints.
 
@@
 
-FIXME: Can we get a riegl scan of jaspers home? If not corona restrictions were at fault
 We couldn't access the rooms the dataset was acquired in for a terrestrial 
-reference scan.
+reference scan in the time given.
 
@@
 
-FIXME: chose better example, added todo for Fabi
+We chose a better example to clarify the use case.
 
@@
 
-We haded a citation.
+We added a citation.
 

### 2021-07-12 Latex/sections/results.tex
sequential iteration added + results

@@
 We again combine twenty temporally successive scans into one meta-scan, using the scan at the median index as a reference coordinate system, which is then globally registered.
+We used the first seven of these meta-scans, corresponding to the first full rotation of the spher, to extract global planes.
+All subsequent meta-scans are then matched against the initial global plane model.
+We use both optimizations from section \ref{ssec:furtheropts}, introducing a global up/down lock for gradient descent and applying sequential iteration to eliminate accumulating pose error.
 
@@
         \includegraphics[width=0.495\textwidth]{./images/jasper_corr_corner}\\
+	\includegraphics[width=0.495\textwidth]{./images/jasper_uncorr_color}\hfill
+	\includegraphics[width=0.495\textwidth]{./images/jasper_corr_color}\\
         \includegraphics[width=0.495\textwidth]{./images/jasper_uncorr_top}\hfill

### 2021-07-12 Latex/sections/simulation.tex
images moved

@@
 
-Figure~\ref{fig:simulatedDatasets} shows a noise-free trajectory through a simulated environment and a trajectory with noisy sensors while figure~\ref{fig:robotRender} shows a rendering of the simulated robot as well as the simulation sequence.
+Figure~\ref{fig:simulatedDatasets} shows a noise-free trajectory through a simulated environment and a trajectory with noisy sensors while figure~\ref{fig:robotRender} shows a rendering of the simulation sequence.
 In the noise-free case, the trajectory slightly bends in the direction of disturbance torque sideways while also varying in velocity in the intended direction of travel.
@@
 	\includegraphics[width=0.495\textwidth]{images/sim_frame_08.png}\\
-	\includegraphics[width=0.495\textwidth]{images/ProtoRoll1.png}\hfill
-	\includegraphics[width=0.495\textwidth]{images/ProtoRoll2.png}
 	\caption{Top: Initial sequence of the simulation of the ideal dataset showing the simulated motion of the robot.
-	Below: Rendering of the simulated robot, a Livox laser scanner insider a spherical shell.}
 	\label{fig:robotRender}

### 2021-07-12 Latex/rebuttal.txt
Further cleaned up rebuttal

@@
 We clarified that the shell can enclose all sensor without requiring a 
-number of points of connection as individual enlosings for all sensors
+number of points of connection as individual enclosing for all sensors
 would. Further we added the requirement, that the shell is very 
@@
 planes for robust correspondences.
-FIXME @Andreas and @Dorit: I (Anton) disagree can you double check?
-
-I (Andreas) seconds the statement, that the limited FOV plays a role
-in not finding planes for robust correspondences.
 
@@
 
-FIXME: I dont understand. How should we make any statement about the accuracy potential?
-It depends on the motion profile and hardware being used, thus is not easy to generalize.
-Or am I getting this wrong?
+The accuracy potential strongly depends on the motion profile and 
+hardware being used, thus is not easy to generalize. Therefore it is 
+still omitted.
 
@@
 
-We added four more terrestial, human-made environments as dangerous examples where a compact, spherical robot would really unfold its potential.
+We added four more terrestrial, human-made environments as dangerous examples where a compact, spherical robot would really unfold its potential.
 
@@
 Removed the false statement.
-BIG FIXME: Get more references on plane based slam.
-
-I (Andreas) has added a coupe of references.
+We have added a couple of references.
 
@@
 
-TODO: Anton? Have you done it already? 
-Yes - Anton
+We clarified the equations.
 
@@
 
-We removed the initial mention of parrallelism and perpendicularity of the planes
+We removed the initial mention of parallelism and perpendicularity of the planes
 as it is of no further interest in the rest of the manuscript.
@@
 
-We have clarified indeces k=planes i=points.
+We have clarified indices k=planes i=points.
 
@@
 
-We have corrected indeces and introduced the elements.
+We have corrected indices and introduced the elements.
 
@@
 
-An explaination of alpha was added.
+An explanation of alpha was added.
 
@@
 
-FIXME: -> I don't understand this comment? TODO
+This table was given to quantify the improvement of the map via the algorithm.
+In particular they are put forward as evidence of the improvement.
 
@@
 
-We have specified which scans and that we assuem constant pose.
+We have specified which scans and that we assume constant pose.
 

### 2021-07-12 Latex/rebuttal.txt
Small fixes. TODO ANDREAS/DORIT

--- 2021-07-12 Latex/rebuttal.txt
@@
 
-FIXME: I dont understand. How should we make any statement about the accuracy potential?
-It depends on the motion profile and hardware being used, thus is not easy to generalize.
-Or am I getting this wrong?
+Unfortunatley we could not extract a 'ground trouth' scan for the real world datasets, due to covid pandemic issues.
+Both rooms were inaccessible, thus we evaluate the accuracy of the real world dataset on a later point in time.
 
@@
 > and Systems, DOI: https://doi.org/10.1109/IROS.2006.282527
-
-Removed the false statement.
-BIG FIXME: Get more references on plane based slam.
 
@@
 
-FIXME: Added todo for Fabi
+We indicated that we observed the noise level around the expected planes by the naked eye, as an educated guess. You find that part in the 'Local Planar Clustering' section.
 

### 2021-07-12 Latex/sections/registration.tex
sequential iteration added + results
Small fixes. TODO ANDREAS/DORIT

@@
 The most simple method is to use the data from an intertial measurement unit for a coarse alignment.
-Alternative solutions combine multiple linescans into a metascan and perform registration methods known from terrestrial laser scanning \todo{add source here} or use a small number of iterations of continuous-time SLAM approaches such as the one presented in~\cite{REMSEN2013}.
+Alternative solutions combine multiple linescans into a metascan and perform registration methods known from terrestrial laser scanning or use a small number of iterations of continuous-time SLAM approaches such as the one presented in~\cite{REMSEN2013}.
 This yields maps that can be used for plane registration and hence a point-to-plane correspondences based registration.
@@
 any point in the cluster has a Hesse distance smaller than some threshold $\epsilon_H$ and a PPD smaller than some additional threshold $\epsilon_P$, to the global plane.
+We observe the noise level around the expected planes by the naked eye. 
+Then we set the thresholds accordingly. 
 For each overlap, we find the smallest angle between the corresponding cluster normal and the global plane normal, according to eq. \ref{eq:alphahat}.
@@
 \subsection{Further Optimizations of the Algorithm}
+\label{ssec:furtheropts}
 
@@
 
-Secondly, we employ a continuous iteration, where the scans are processed sequentially. 
+Secondly, we employ a sequential iteration, where the scans are processed one after another. 
 This is especially useful for mobile systems, where pose errors accumulate due to increasing tracking uncertainties.

### 2021-07-12 Latex/sections/registration.tex
Minor changes in math formatting

@@
 \end{align}
-From this we define the function $D(\varphi,\vartheta,\psi,t_x,t_y,t_z, \vec{p}_i)$ that computes the distance of a point $\vec{p}_i$ to its corresponding plane $\mathcal{P}_k$.
+From this we define the function $D(\varphi,\vartheta,\psi,t_x,t_y,t_z, \vec{p}_{i})$ that computes the distance of a point $\vec{p}_{i}$ to its corresponding plane $\mathcal{P}_k$.
 Omitting the arguments of the function for simplicity:
-\begin{align}
-\begin{aligned}
+\begin{equation}
+\begin{split}
     D &= T(\vec{p}_i) \cdot \vec{n}_{\mathcal{P}_k} \\
@@
        &- \rho_{\mathcal{P}_k}
-\end{aligned}
-\end{align}
+\end{split}
+\end{equation}
 

### 2021-07-12 Latex/rebuttal.txt
Added the missing references for pre-registration Small typos corrected A few additions in the rebuttal

--- 2021-07-12 Latex/rebuttal.txt
@@
 >
+
+The authors thank the reviewers for their time to review the paper and to
+provide valuable input. The following lists how we improved the manuscript based
+on their thoughtful comments. Minor remarks have been integrated into the
+document without further comments.
+
 >   
@@
 
+I (Dorit) agree, that the FOV is a more important factor.
+
 > Lines 244-246: The Livox ranging precision is given twice in this
 > sentence.
+
+Corrected.
 > 
@@
 
-FIXME: I dont understand. How should we make any statement about the accuracy potential?
-It depends on the motion profile and hardware being used, thus is not easy to generalize.
-Or am I getting this wrong?
+The accuracy potential depends heavily on the motion profile, the environment
+and the hardware being used, thus is not easy to generalize. However, the
+simulation results show that the algorithm has the potential to improve the map
+quality based on point-distances by approximately a factor of ten. We added this
+to the conclusions.
 
@@
 
-Removed the false statement.
-BIG FIXME: Get more references on plane based slam.
-
-I (Andreas) has added a coupe of references.
+The authors thank the reviewer for spotting this false statement. The state
+of the art has been extended with approaches that consider the structural
+properties of human-made environments.
 
@@
 
-We added an outline for section 3.
+The introduction in Section 3 has been revised and includes now an outline.
 
@@
 
-Globally. I hope we made that clear in the introduction of section 3 now.
+The registration is performed globally. The explanation has been adjusted to clarify this.
 
@@
 
-FIXME! Added todo for Dorit/Andreas as they wrote this
+Two references were added as examples for the first half of the sentence.
 

### 2021-07-12 Latex/sections/registration.tex
Added the missing references for pre-registration Small typos corrected A few additions in the rebuttal

--- 2021-07-12 Latex/sections/registration.tex
@@
 The most simple method is to use the data from an intertial measurement unit for a coarse alignment.
-Alternative solutions combine multiple linescans into a metascan and perform registration methods known from terrestrial laser scanning or use a small number of iterations of continuous-time SLAM approaches such as the one presented in~\cite{REMSEN2013}.
+Alternative solutions combine multiple linescans into a metascan and perform registration methods known from terrestrial laser scanning and distribute the error over the trajectory~\cite{Stoyanov2009,Bosse2009} or use a small number of iterations of continuous-time SLAM approaches such as the one presented in~\cite{REMSEN2013}.
 This yields maps that can be used for plane registration and hence a point-to-plane correspondences based registration.
@@
 any point in the cluster has a Hesse distance smaller than some threshold $\epsilon_H$ and a PPD smaller than some additional threshold $\epsilon_P$, to the global plane.
-We observe the noise level around the expected planes by the naked eye. 
-Then we set the thresholds accordingly. 
 For each overlap, we find the smallest angle between the corresponding cluster normal and the global plane normal, according to eq. \ref{eq:alphahat}.

### 2021-07-12 Latex/rebuttal.txt

@@
 
+Unfortunatley we could not extract a 'ground trouth' scan for the real world datasets, due to covid pandemic issues.
+Both rooms were inaccessible, thus we evaluate the accuracy of the real world dataset on a later point in time.
 The accuracy potential depends heavily on the motion profile, the environment

### 2021-07-13 Latex/sections/registration.tex
equation -> align

--- 2021-07-12 Latex/sections/registration.tex
@@
 The Hesse-normal form describes the distance from the $i$-th transformed point $T(\vec{p}_i)$ to infinitely extending, $k$-th plane in 3D vector space
-\begin{equation}\label{eq:hesse}
+\begin{align}\label{eq:hesse}
 	D_{h}^{i,k} = \vec{n}_{\mathcal{P}_k} \cdot \left[ T(\vec{p}_i) - \vec{a}_{\mathcal{P}_k} \right] ,
-\end{equation}
+\end{align}
 with normal $\vec{n}_{\mathcal{P}_k}$ and supporting point $\vec{a}_{\mathcal{P}_k}$ of the $k$-th plane.
@@
 Thus, the $i$-th points projection $\widetilde{T}(\vec{p}_i)$ onto the $k$-th plane, which is required later, is easily calculated by shifting the point against normal direction:
-\begin{equation}\label{eq:projection}
+\begin{align}\label{eq:projection}
 	\widetilde{T}(\vec{p}_i) = T(\vec{p}_i) - D_{h}^{i,k} \vec{n}_{\mathcal{P}_k}
-\end{equation}
+\end{align}
 
@@
 The line segment is parameterized by
-\begin{equation}
+\begin{align}
 \label{eq:line}
 	s_{m, j}(t) = p_m + t \cdot \left( p_j - p_m \right) ,  
-\end{equation}
+\end{align}
 where $t \in \left[ 0, 1 \right]$. 
 We set up a distance function, which measures the distance from the query point $p_i$ to an arbitrary point on the line segment.
-\begin{equation}
+\begin{align}
 	\label{eq:dist}
 	d_{m, j}(t) = \lVert s_{m, j}(t) - p_i \rVert 
-\end{equation} 
+\end{align} 
 It is now possible to find the shortest distance to the line segment by finding the argument of the function that minimizes this distance.
 Therefore, we set
-\begin{equation}
+\begin{align}
 	\pardiff{t} d_{m, j}(t)^2 \overset{!}{=} 0 ,
-\end{equation}  
+\end{align}  
 resulting in 
-\begin{equation}
+\begin{align}
 	t_0 = \frac{\left( p_j - p_m \right) \cdot \left( p_i - p_m \right)}{\left( p_j - p_m \right)^2} .
-\end{equation}
+\end{align}
 Note the possibility of $t_0 \notin \left[ 0, 1 \right]$. 
@@
 By constraining $t_0$ we get 
-\begin{equation}
+\begin{align}
 \label{eq:minmax}
 	\hat t = \min \left( \max \left( t_0, 0\right) , 1 \right) ,
-\end{equation}
+\end{align}
 which is the argument that gives the shortest distance to the line segment, when inserted to eq. \ref{eq:dist}.
@@
 We define similarity of two normals via the angle between them, which is:
-\begin{equation}
+\begin{align}
 \label{eq:smallestalpha}
 	\alpha(\vec{n}_1, \vec{n}_2) = \arccos(\vec{n}_1 \cdot \vec{n}_2) \; .
-\end{equation}
+\end{align}
 However, the \textit{smallest} angle between two \textit{normal} vectors is 
-\begin{equation}
+\begin{align}
 \label{eq:alphahat}
@@
 \; ,
-\end{equation}
+\end{align}
 because the opposing normals $-\vec{n}_1$ and $-\vec{n}_2$ always have to be considered, since they correspond to the same plane.
@@
 Omitting the arguments of the function for simplicity:
-\begin{equation}
+\begin{align}
 \begin{split}
@@
 \end{split}
-\end{equation}
+\end{align}
 
@@
 Where
-\begin{equation}
+\begin{align}
 \begin{split}
@@
 \end{split} \;\; ,
-\end{equation}
-\begin{equation}
+\end{align}
+\begin{align}
 \begin{split}
@@
 \;\; ,
-\end{equation} 
-\begin{equation}
+\end{align} 
+\begin{align}
 \begin{split}
@@
 \end{split}
-\end{equation}
+\end{align}
 and $\vec{\Pi}=\begin{bmatrix}\varphi & \vartheta & \psi & t_x & t_y & t_z\end{bmatrix}^T$.

### 2021-07-13 Latex/sections/registration.tex
Split -> aligned

--- 2021-07-13 Latex/sections/registration.tex
@@
 \begin{align}
-\begin{split}
+\begin{aligned}
     D &= T(\vec{p}_i) \cdot \vec{n}_{\mathcal{P}_k} \\
@@
        &- \rho_{\mathcal{P}_k}
-\end{split}
+\end{aligned}
 \end{align}
@@
 \begin{align}
-\begin{split}
+&\begin{aligned}
 \nabla E_\varphi = &n_{\mathcal{P}_k}^y(x_i[-S_\varphi S_\psi+C_\varphi C_\psi S_\vartheta]\\
@@
                    &+y_i[C_\varphi C_\psi -S_\varphi S_\vartheta S\psi] - z_i S_\varphi C_\vartheta)
-\end{split} \;\; ,
+\end{aligned} \;\; ,
 \end{align}
 \begin{align}
-\begin{split}
+&\begin{aligned}
 \nabla E_\vartheta = & n_{\mathcal{P}_k}^x(-x_iS_\vartheta C_\psi + y_i S_\vartheta S_\psi + z_iC_\vartheta) \\
         &+ n_{\mathcal{P}_k}^y(x_iC_\psi S_\varphi C_\vartheta - y_iS_\varphi C_\vartheta S_\psi + z_i S_\vartheta S_\varphi)  \\
-        &+ n_{\mathcal{P}_i}^z(-x_iC_\varphi C_\psi C_\vartheta + y_iC_\varphi C_\vartheta S_\psi - z_i C_\varphi 
-\end{split}
-\;\; ,
-\end{align} 
-\begin{align}
-\begin{split}
+        &+ n_{\mathcal{P}_i}^z(-x_iC_\varphi C_\psi C_\vartheta + y_iC_\varphi C_\vartheta S_\psi - z_i C_\varphi)
+\end{aligned}\;\; ,\\
+&\begin{aligned}
 	\nabla E_\psi = &n_{\mathcal{P}_k}^x\left(-x_iC_\vartheta S_\psi - y_iC_\vartheta C_\psi\right) \\
        &+ n_{\mathcal{P}_k}^y(x_i[C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi]\\
-       &+ y_i[-C_\varphi S_\psi - S_\varphi S_\theta C_\psi]) \\
+       &\quad+ y_i[-C_\varphi S_\psi - S_\varphi S_\theta C_\psi]) \\
        &+ n_{\mathcal{P}_k}^z(x_i[S_\varphi C_\psi + C_\varphi S_\psi S_\theta] \\
-       &+ y_i[-S_\psi S_ \varphi + C_\varphi S_\vartheta C_\psi]) 
-\end{split}
+       &\quad+ y_i[-S_\psi S_ \varphi + C_\varphi S_\vartheta C_\psi]) 
+\end{aligned}
 \end{align}

### 2021-07-13 Latex/sections/stateOfTheArt.tex

@@
 The de-facto standard for many SLAM algorithms is the Iterative-Closest-Point (ICP) algorithm~\cite{Besl1992} that employs point-to-point correspondences using closest points, as the name suggests. 
-To overcome the requirements on point-density imposed by the point-to-point correspondences instead other correspondences are used. 
+To overcome the requirements on point-density imposed by the point-to-point corres\-pon\-den\-ces instead other correspondences are used. 
 In human-made environments, planes are abundantly available and hence provide an attractive base for correspondences.
@@
 
-LOAM~\cite{Zhang2014} is the baseline-algorthm that provides a real-time and lowdrift solution based on two parallel registration algorithms using planes and lines.
+LOAM~\cite{Zhang2014} is the baseline-algorithm that provides a real-time and lowdrift solution based on two parallel registration algorithms using planes and lines.
 Unfortunately, it is not open-source anymore.
@@
 Parallel computing needs a powerful CPU, it may be not suitable for an embedded system which has limited computational resources~\cite{Zhou_2021}.
-Thus, they extens their smoothing and mapping to the LiDAR and planes case.
+Thus, they extend their smoothing and mapping to the LiDAR and planes case.
 In their experiments they used a VLP-16 LiDAR to collect indoor datasets.

### 2021-07-13 Latex/Rebuttal/rebuttal.tex
Rebuttal now a tex/pdf file

\documentclass{article}
\usepackage[margin=0.5in]{geometry}
\newenvironment{itquote}
  {\begin{quote}\itshape}
  {\end{quote}\ignorespacesafterend}

\title{Rebuttal - ``Towards Spherical Robots for Mobile Mapping in Human Made Environments''}
\date{\today}

\begin{document}

\maketitle

\begin{itquote}
Manuscript Number: OPHOTO-D-21-00020 
Towards Spherical Robots For Mobile Mapping In Human Made
Environments

Dear Mr. Bredenbeck,

Thank you for submitting your manuscript to ISPRS Open Journal of
Photogrammetry and Remote Sensing.

I have completed my evaluation of your manuscript. The reviewers
recommend reconsideration of your manuscript following revision. I
invite you to resubmit your manuscript after addressing the comments
below. Please resubmit your revised manuscript by Jul 13, 2021. 
When revising your manuscript, please consider all issues mentioned
in the reviewers' comments carefully: please outline in a cover
letter every change made in response to their comments and provide
suitable rebuttals for any comments not addressed. Please note that
your revised submission may need to be re-reviewed.

To submit your revised manuscript, please log in as an author at
https://www.editorialmanager.com/ophoto/, and navigate to the
"Submissions Needing Revision" folder under the Author Main Menu.

ISPRS Open Journal of Photogrammetry and Remote Sensing values your
contribution and I look forward to receiving your revised
manuscript.

Kind regards,

George Vosselman 
Editor-in-Chief 
ISPRS Open Journal of Photogrammetry and Remote Sensing

Editor and Reviewer Comments:

Editor:
As you will see below both reviewers appreciate your work and don't
have concerns regarding the methodology or results. Reviewers, in
particular Reviewer 2, however, do have a large number of
recommendations to improve the readability of your manuscript. Given
the number of recommendations, I ask for a major revision. The
recommendations should, however, be easy to implement.
\end{itquote}

The authors thank the reviewers for their time to review the paper and to
provide valuable input. The following lists how we improved the manuscript based
on their thoughtful comments. Minor remarks have been integrated into the
document without further comments.

\begin{itquote}
Reviewer 1: The manuscript presents the concept of a spherical
mobile mapping system, and a data registration method utilizing the
geometric constraints available in human-made environments,
e.g. planar surfaces.

The manuscript has novel content and is well written, and can be
accepted after minor revision. My comments and suggestions to be
considered by the authors are as follows:

Lines 27-31: Spherical shell can protect the sensors inside, but
this does not prevent the external disturbances such as dust or
other impurity staining the sphere itself, affecting the performance
of the system. So in principle this does not change the situation
from using protective covers on "traditional" mobile mapping
systems.
\end{itquote}

We clarified that the shell can enclose all sensor without requiring a 
number of points of connection as individual enclosing for all sensors
would. Further we added the requirement, that the shell is very 
durable.

\begin{itquote} 
Lines 111-116: Modern kinematic scanning systems collect point cloud
data with high speed, so also kinematic point clouds are often dense
enough for robust plane determination. And on the other hand, sparse
point clouds are not optimal for point-to-point matching due to the
long distances between adjacent points.
\end{itquote}

We clarified that also the limited FOV plays a role in not finding 
planes for robust correspondences.

\begin{itquote}
Lines 244-246: The Livox ranging precision is given twice in this
sentence.
\end{itquote}

Corrected.

\begin{itquote}
Figure 4: The x-scale is given in figure caption, but I would
suggest adding it to the figure also.
\end{itquote}

We added a label to the x-scale

\begin{itquote}
Conclusion: Whereas the title, Abstract and Introduction concentrate
mostly on the spherical mobile mapping, Conclusion deals almost
entirely with the topic of registration, thus feeling somewhat
separated from the rest of the manuscript. Reading the Conclusion,
one could not expect that the title of the manuscript is "Towards
Spherical Robots For Mobile Mapping In Human Made Environments". 
\end{itquote}

We added more context about spherical robots and put the registration 
method into context.

\begin{itquote} 
Reviewer 2: The authors have written a very interesting paper on the
topic of mobile mapping with a spherical robot in human-made
environments. In a rather slim paper and pleasantly concise form,
they present their approach and show first results based on
simulated and real measurement data. The reading flow is mostly
pleasant and no irrelevant information is given. Conversely, the
authors leave out important details in several places, which would
be very helpful for the reader's understanding. In the following,
there are several points for improvement and critical questions that
should be taken into account by the authors. Due to the large number
and the importance of central points, I recommend a major revision.

- A short outline at the end of the introduction is missing and
would give the reader a better overview of the further content and
should therefore be added.
\end{itquote}

We added a short outline, introducing you to the next sections.

\begin{itquote} 
- In the entire article, no reference is made to the targeted or
required accuracy potential with regard to the quality of the pose
determination and the final point cloud, which in my view requires
at least a rough classification.
\end{itquote}

The accuracy potential depends heavily on the motion profile, the environment
and the hardware being used, thus is not easy to generalize. However, the
simulation results show that the algorithm has the potential to improve the map
quality based on point-distances by approximately a factor of ten. We added this
to the conclusions.

\begin{itquote}
- For the short state of the art of existing mobile systems, which
is mainly based on the authors' own systems, in my opinion, at least
the NavVis VLX system could be added, which in comparison is newer
and as a human-operated system for mapping is one of the benchmarks.
\end{itquote}

We added the NavVis VLX as a reference in the state of the art section.

\begin{itquote}
- The authors mention unknown and dangerous environments as a
possible field of application, in which the advantages of the
spherical system should come into play. However, only one space
application is mentioned as a specific example, which may be valid
in itself, but does not fit the topic of the paper very
well. Application scenarios should therefore be supplemented by
terrestrial areas and in particular by human-made environments,
which better motivate the future use and justification of such
systems for the reader.
\end{itquote}

We added four more terrestrial, human-made environments as dangerous examples where a compact, spherical robot would really unfold its potential.

\begin{itquote} 
- The advantages of a spherical mobile platforms are discussed, but
nothing is said about possible disadvantages or
limitations. Moreover, I wonder whether the "unique advantages"
mentioned in the abstract are really so remarkable. For example,
what about the recording constellation, which in my view is
relatively unfavourable on the ground and leads to restricted fields
of view and weak angles of incidence when mapping the
ground. Another advantage stated by the authors is that the shell
protects the sensors, which is reasonable. However, this protection
is only necessary because you are moving directly on the ground. In
addition, can dust and dirt accumulate on the outer shell and
influence the acquisition?
\end{itquote}

We added disadvantages (including your example) in the introduction. They show the need for a reliable, robust registration procedure.
There is an approach to prevent dust and dirt to accumulate on the outer shell. We added a note / reference to that.

\begin{itquote}
- Section 2.1 discusses existing laser scanner-based SLAM
approaches, concluding with the statement (page 2, column 1, line
30-33) that all of these approaches do not include the structural
properties of human-made environments in the calculations. This
statement is not true. There are a variety of SLAM approaches
(e.g. Jung et al. 2015; Nguyen et al.  2006), but also other
applications with recursive state estimation, which make use of this
additional information in the context of georeferencing. This should
be investigated and reworked in more detail.
Jung, J., Yoon, S., Ju, S., and Heo, J. (2015). Development of Kinematic
3D Laser Scanning System for Indoor Mapping and As-Built BIM Using
Constrained SLAM. In: Sensors (Basel, Switzerland), 15(10), DOI:
https://doi.org/10.3390/s151026430
Nguyen, V., Harati, A., Martinelli, A., Siegwart, R., and Tomatis, N.
(2006). Orthogonal SLAM: A Step Toward Lightweight Indoor Autonomous
Navigation. In: IEEE/RSJ International Conference on Intelligent Robots
and Systems, DOI: https://doi.org/10.1109/IROS.2006.282527
\end{itquote}

The authors thank the reviewer for spotting this false statement. The state
of the art has been extended with approaches that consider the structural
properties of human-made environments.

\begin{itquote}
- The introduction in Section 3 should be revised as it is not very
intuitive for the reader and something about the objectives and
motivation for this section is missing. However, this is also
directly related to the general lack of outline beforehand.
\end{itquote}

The introduction in Section 3 has been revised and includes now an outline.

\begin{itquote}
- In section 3, the authors introduce their own algorithm for
registration, but they do not give any further information about the
temporal context in which this takes place. This is probably done in
the context of each epoch, although it is also not clear how an
epoch is defined. Is it a single full rotation of the laser scanner
or that of the sphere? And does the registration always take place
only locally in relation from the current epoch to the last epoch or
with a window approach or even globally?
\end{itquote}

The registration is performed globally. The explanation has been adjusted to clarify this.

\begin{itquote}
- Section 3.2 deals with the extraction of planes, but the following
questions remain unclear: Are the planes always searched for only in
the context of the observational data of the current epoch or with
regard to the already registered global point clouds? How many
planes are needed at all and what demands are made on their spatial
distribution? The abstract mentioned parallel and perpendicular
planes, but this is not discussed further here. Are only these used
and if so, why? Furthermore, what about planes that are not
necessarily parallel or perpendicular (e.g. doors, cupboards)? How
is the flatness of a plane checked? How are outliers dealt with,
especially with regard to existing measurement and motion noise?
\end{itquote}

We added more detail in the plane extraction subsection.

\begin{itquote}
- Throughout Section 3, numerous variables are not introduced and
are not explained in more detail, so that a more precise
interpretation of the individual equations is unfortunately very
difficult.
\end{itquote}

We clarified the equations.

\begin{itquote} 
- What orders of magnitude were finally used for the range noise
model in Section 4.2 and why was this not also done for the angular
resolution?
\end{itquote}

The distributions we sample from are specified in the caption of 
figure 3.

\begin{itquote}
- In the context of the real experiment in Section 5, a major
simplification is made. Even if this simplification is certainly
suitable for the beginning, this special case comes quite surprising
here...especially in comparison to the simulation and the previously
described topic. Have other measurements not yet been carried out or
what is the background? At least this simplification should be
mentioned beforehand, otherwise the reader will have false
expectations.
\end{itquote}

We added a further experiment that uses nearly all DoF: A sphere rolling
on a ground plane. This was not done in the previous manuscript because
we had issues getting the hardware shipped, assembled and the experiment 
carried out in time.

\begin{itquote}
- Wouldn't it also be interesting to compare the estimated pose
parameters with the nominal values from the simulation within the
framework of Section 6.1? So far, the focus has only been on the
point cloud.
\end{itquote}

Unfortunately we couldn't add this due to time constraints.

\begin{itquote}
- Why is there no qualitative validation for the results of the real
data in Section 6.2? A reference scan based on a terrestrial laser
scanner would be a possibility or otherwise based on the flatness of
known planes in the registered point cloud?
\end{itquote}

We couldn't access the rooms the dataset was acquired in for a terrestrial 
reference scan in the time given.

\begin{itquote}
- In the Conclusion, the parallelism and perpendicularity within the
environment is mentioned again, although this did not play a role in
the main part of the paper. Therefore, the question arises why do
the planes have to be parallel or perpendicular at all and why is
this mentioned at the beginning? This is not clear, since no
additional constraints on this additional information have been
introduced so far. There are also already other methods available
that take this idea of geometric constraints for georeferencing into
account.
\end{itquote}

We removed the initial mention of parallelism and perpendicularity of the planes
as it is of no further interest in the rest of the manuscript.

\begin{itquote}
Further open questions/comments are:
Page 1, Line 15: Double use of "protect", which should be replaced
by another word.
\end{itquote}

We removed first instance.

\begin{itquote}
Page 1, Line 21: The link for the "second option" to the use of an
IMU is not necessarily obvious.
\end{itquote}

We added IMU as sensor for coarse estimation.

\begin{itquote}
Page 1, Column 2, Line 52-56: Complicated sentence which should be
reformulated.
\end{itquote}

We have split into two sentences

\begin{itquote}
Page 2, Column 2, Line 15: The term "long stop" is not to be
interpreted clearly in this context. How is "long" defined in this
sense in terms of time or space?
\end{itquote}

The specified stop is required to find enough points for plane
detection.

\begin{itquote}
Page 3, Column 1, Line 6-8: A specific reference should be given for
the first half of the sentence, as reference 19 probably only refers
to the second part.
\end{itquote}

Two references were added as examples for the first half of the sentence.

\begin{itquote}
Page 3, Column 1, Line 23-26: Not all variables were introduced
accurately
\end{itquote}

We have introduced missing variables and clarified rho.

\begin{itquote}
Page 3, Column 1, Line 33-34: Selection of references for this?
\end{itquote}

We have added plane segmentation sources.

\begin{itquote}
Page 3, Column 1, Line 34-35: But what demands are made on the
assignment? This is unclear to the reader, also in connection with
the general quality requirements of the whole approach.
\end{itquote}

We have added a specification for "good enough for plane detection".

\begin{itquote}
Page 3, Column 1, Line 38: How can the threshold value be specified,
since it is probably not identical for all applications and
environments? Later, there is also no indication of the choice of
the value in the own simulated and real data sets.
\end{itquote}

We indicated that we observed the noise level around the expected planes by the naked eye, as an educated guess. You find that part in the 'Local Planar Clustering' section.

\begin{itquote}
Page 3, Column 1, Line 50: What does the index k describe and what
is the difference to the index i from Section 3.2?
\end{itquote}

We have clarified indices k=planes i=points.

\begin{itquote}
Page 3, Equation (1): Again, some variables are not specified. For
example, the source of tx ty tz is unclear.
\end{itquote}

We have introduced tx ty tz properly.

\begin{itquote}
Page 3, Equation (3): Equation unclear regarding the use of the 2 in
the exponent as well as the index.
\end{itquote}

We have specified that we're using the square of the L2 norm.

\begin{itquote}
Page 3, Equations (4)-(5): Here I don't understand how to get to the
elements of the normal vector through the transformation. Moreover,
these individual elements were not introduced in advance.
\end{itquote}

We have corrected indices and introduced the elements.

\begin{itquote}
Page 3, Equation (9): The variable alpha is not described and
explained
\end{itquote}

An explanation of alpha was added.

\begin{itquote}
Page 4, Equation (10): Description of the individual variables is
missing and, moreover, rho has already been used as the distance
parameter of the Hesse-normal form
\end{itquote}

we have replaced rho with zeta and introduced the variables.

\begin{itquote}
Page 4, Equation (12): Is epsilon here the same variable as the
threshold in Section 3.3? Probably not.
\end{itquote}

We have renamed epsilon and clarified the difference to sec 3.3.

\begin{itquote}
Page 4, Column 2, Line 15: "some dimensions": Which ones and why
exactly are they chosen? Are they the same ones all the time?
\end{itquote}

We chose a better example to clarify the use case.

\begin{itquote}
Page 4, Column 2, Line 20-25: Reference is missing
\end{itquote}
We added a citation.

\begin{itquote}
Page 4, Column 2, Line 55: Doubled information regarding the
distance measurement precision 
\end{itquote}

We removed duplicated information.

\begin{itquote}
Page 5, Column 1, Equation (18): Delta t is not introduced and there
is no indication of how large the time steps are in the simulation.
\end{itquote}

We have introduced delta t and give a default value.

\begin{itquote}
Page 5, Column 2, Line 14: Should it be "out" or "our"?
\end{itquote}

We have corrected this.

\begin{itquote}
Page 5, Column 2, Line 27-28: What are the specifications of the
used IMU with regard to its uncertainty budget?
\end{itquote}

We have added precision of IMU and added the source.

\begin{itquote}
Page 6, Figure 1: How are these values to be interpreted and from
where are they derived? This is unclear.
\end{itquote}

We have specified this. It is chosen to still allow plane detection.

\begin{itquote}
Page 8, Column 1, Table 1: What is missing is a classification of
the values in relation to the expectations / demands...which are,
however, still missing in advance
\end{itquote}

This table was given to quantify the improvement of the map via the algorithm.
In particular they are put forward as evidence of the improvement.

\begin{itquote}
Page 8, Column 1, Line 37-43: At this point, an interpretation of
the significant second peak in the right-hand diagram in Figure 4
should be made. This has not been discussed so far.
\end{itquote}

We redid the experiment with a larger amount of points (all that have
a point-to-point distance of less than 3000 cm). The second significant
peak was a artifact of the previous thresholding and hence is now not 
significant anymore.

\begin{itquote}
Page 8, Column 1, Line 51: "ten scans" means ten individual scan
lines or 10 epochs or is this identical? What kind of temporal
offset is there in between and wouldn't a motion model have to be
taken into account for this or is an identical pose assumed for the
ten scans?
\end{itquote}

We have specified which scans and that we assume constant pose.

\begin{itquote}
Page 8, Column 1, Line 55-56: What could be the cause of this?
\end{itquote}

We have specified the outlier dependence.

\begin{itquote}
Page 8, Column 2, Line 44: This sentence always applies and should
not remain here.
\end{itquote}

We have removed this sentence.

\begin{itquote}
Page 9, Figure 4: The two point clouds in the middle of figure 4 are
difficult to interpret in terms of their orientation. Could it be an
option to add a coordinate system?
\end{itquote}
It was specified in the figure label that lateral images have the same
orientation.

\end{document}

### 2021-07-13 tdp/DataSetSim/dataset_sim.cpp
Removed parsing bug

@@ -2,7 +2,7 @@
 #include <iomanip>
 #include <fstream>
 #include <sstream>
-#include <omp.h>
+//#include <omp.h>
 
 #include "dataset_sim.hpp"
 #include "constants.hpp"
@@ -172,6 +172,7 @@ int main(int argc, char const *argv[])
       //Close Files
       scan_file.close();
       pose_file.close();
+      delete[] file_name;
 
       // Console Logging
       if(VERBOSITY_LEVEL == 1)

### 2021-07-13 tdp/DataSetSim/parsing.cpp

@@ -214,12 +214,10 @@ bool parse_options(int argc, char const *argv[],
     // entry
     while(pose_mean.size() < 2){
       pose_mean.push_back(0.0);
-
     }
 
     while(pose_std.size() < 2){
-      pose_mean.push_back(0.0);
-
+      pose_std.push_back(0.0);
     }
 
     if(pose_mean.size() != 2)

### 2021-07-13 tdp/DataSetSim/utils.cpp

@@ -99,7 +99,7 @@ double eucl_dist(vector<double> first, vector<double> second)
 }
 
 /* Helper to normalize a vector */
-void normalize(vector<double> vec)
+void normalize(vector<double> & vec)
 {
   double mag = 0;
   for(double val: vec){ mag += val*val;}

### 2021-07-13 tdp/DataSetSim/utils.hpp

@@ -16,7 +16,7 @@ void world2robot_frame(vector<double> pose, bool rotate_only, vector<double> &wo
 /* Helper function for computing the distance between two
  * points in space using the euclidean distance */
 double eucl_dist(vector<double> first, vector<double> second);
-void normalize(vector<double> vec);
+void normalize(vector<double> & vec);
 
 /* Helper for Radians and degree conversion */
 double deg_to_rad(double deg);

### 2021-07-13 tdp/datasets/processing/plot_path_diff.py
Added plotting file for trajectory difference

@@ -1,0 +1,93 @@
+import sys
+import numpy as np
+import matplotlib.pyplot as plt
+from dataclasses import dataclass
+
+
+@dataclass
+class Pose:
+    x: float
+    y: float
+    z: float
+    yaw: float
+    pitch: float
+    roll: float
+
+
+def norm(angle: float) -> float:
+    normed_angle = angle
+    if normed_angle < -180.0:
+        normed_angle += 360.0
+    elif normed_angle > 180.0:
+        normed_angle -= 360.0
+
+    return normed_angle
+
+
+def load_poses(path: str, num_poses: int) -> list:
+    poses = []
+    for i in range(num_poses):
+        pose = np.matrix.flatten(np.loadtxt(path + f"scan{i:03}.pose"))
+        poses.append(Pose(pose[0], pose[1], pose[2],
+                          pose[3], pose[4], pose[5]))
+
+    return poses
+
+
+def compute_diff(true_traj: list, other_traj: list) -> tuple:
+    position_diff = []
+    orientation_diff = []
+    for i in range(len(true_traj)):
+        pose_t = true_traj[i]
+        pose_o = other_traj[i]
+        position_diff.append(
+            np.sqrt(
+                (pose_t.x - pose_o.x)**2
+                + (pose_t.y - pose_o.y)**2
+                + (pose_t.z - pose_o.z)**2))
+
+        yaw_diff = norm(pose_t.yaw - pose_o.yaw)
+        pitch_diff = norm(pose_t.pitch - pose_o.pitch)
+        roll_diff = norm(pose_t.roll - pose_o.roll)
+        orientation_diff.append(
+            np.sqrt((yaw_diff)**2 + (pitch_diff)**2 + (roll_diff)**2))
+
+    return position_diff, orientation_diff
+
+
+def main():
+    if len(sys.argv) != 4:
+        print("Not both paths given. Usage: ")
+        print("python plot_path_diff.py <path_to_truth>" +
+              " <path_to_other> <num_of_pose_files>")
+        exit(0)
+
+    path_truth = str(sys.argv[1])
+    path_other = str(sys.argv[2])
+    num_poses = int(sys.argv[3])
+
+    true_traj = load_poses(path_truth, num_poses)
+    other_traj = load_poses(path_other, num_poses)
+
+    position_diff, orientation_diff = compute_diff(true_traj, other_traj)
+
+    plt.figure(1)
[…]

### 2021-07-13 tdp/datasets/processing/plot_path_diff.py
Added diff sum output

@@ -71,6 +71,10 @@ def main():
 
     position_diff, orientation_diff = compute_diff(true_traj, other_traj)
 
+    print(f"Sum of Position Difference: {np.sum(position_diff):.2f} cm")
+    print(f"Sum of Orientation Difference: {np.sum(orientation_diff):.2f} °")
+
+
     plt.figure(1)
     plt.plot(position_diff)
     plt.grid()

### 2021-07-13 Latex/Rebuttal/rebuttal.tex
Changed some wording in rebuttal

@@
 We couldn't access the rooms the dataset was acquired in for a terrestrial 
-reference scan in the time given.
+reference scan in the time given, therefore the quantitative analysis was done with the simulation and the real data was analyzed qualitatively.
 
@@
 \end{itquote}
+
 We added a citation.
@@
 \end{itquote}
+
 It was specified in the figure label that lateral images have the same

### 2021-07-13 Latex/Rebuttal/rebuttal.tex
Fixed Noise model and Touch up of rebuttal

@@
 provide valuable input. The following lists how we improved the manuscript based
-on their thoughtful comments. Minor remarks have been integrated into the
-document without further comments.
+on their thoughtful comments.
 
@@
 
-Unfortunately we couldn't add this due to time constraints.
+Whenever the scanner is looking into the ground the number of points is too small
+and thus neglected in the optimization step. This leads to jumps in the
+trajectory, which make the comparison results less meaningful than expected. Due to time constraints this issue could not be resolved.
 

### 2021-07-13 Latex/sections/simulation.tex
yt link added to pdf
Fixed Noise model and Touch up of rebuttal

@@
 	\begin{bmatrix}\varphi\\\vartheta\\\psi\\x\\y\\z\end{bmatrix}_{k} 
-	+ \begin{bmatrix}1\\0\\0\\R\\0\\0\end{bmatrix}\omega \Delta t  + \Delta t \underbrace{\sum_{i = 0}^k \begin{bmatrix}n_\varphi[i] \Delta t\\0\\n_\psi[i]\Delta t\\0\\0\\0\end{bmatrix}}_{\substack{\text{Difference in}\\\text{angular velocity}}}
+	+ \begin{bmatrix}1\\0\\0\\0\\0\\R\end{bmatrix}\omega \Delta t  + \Delta t \underbrace{\sum_{i = 0}^k \begin{bmatrix}n_\varphi[i] \Delta t\\0\\n_\psi[i]\Delta t\\n_\varphi[i] \Delta t R\\0\\n_\psi[i]\Delta tR\end{bmatrix}}_{\substack{\text{Difference in}\\\text{angular velocity}}}
 \end{align}
@@
 	\includegraphics[width=0.495\textwidth]{images/sim_frame_08.png}\\
-	\caption{Top: Initial sequence of the simulation of the ideal dataset showing the simulated motion of the robot.
+	\caption{Top: Initial sequence of the simulation of the ideal dataset showing the simulated motion of the robot.}
 	\label{fig:robotRender}

### 2021-08-20 Latex/sections/introduction.tex
added comma, look of pushing works
Incorporated both reviewers comments

@@
 That way, a sensor fixed inside the spherical structure will cover the entire environment, given the required locomotion without the need for additional actuators for the sensors.
-This requires a solution for the spherical simultaneous localization and mapping (SLAM) problem given the six degrees of freedom of the robot.
+This requires a solution for the spherical simultaneous localization and mapping (SLAM) problem, given the six degrees of freedom of the robot.
 Secondly, a spherical shell that encloses all sensors protects these from possible hazardous environments.  
@@
 This is particularly useful for unknown or dangerous environments.
-E.g., old buildings that are in danger of collapsing, narrow underground funnels, construction sites, or mining shafts. 
+E.g., old buildings that are in danger of collapsing, narrow underground tunnels, construction sites, or mining shafts. 
 The spherical format is, in fact, also suited for space applications.

### 2021-08-20 Latex/sections/realData.tex
images moved
Incorporated both reviewers comments

@@
 We exclude a translation in the global up- or downwards direction as the sphere rolls on the ground surface, without possibility for up- or downwards motion-
-Figure~\ref{fig:prototype} shows the prototype.
+figure~\ref{fig:prototype} shows the prototype.
 The left column of figure~\ref{fig:jasperhome} shows the pre-registered point cloud, applying only the transformation defined by the recorded poses.
@@
 \begin{figure}
-  \includegraphics[width=0.49\linewidth]{./images/photoReal1.jpeg}
-  \includegraphics[width=0.50\linewidth]{./images/photoReal3.jpeg}
-  \caption{Prototype used for data acquisition with the rolling sphere.
+  \includegraphics[width=0.495\linewidth]{./images/photoReal1.jpeg}\hfill
+  \includegraphics[width=0.495\linewidth]{./images/photoReal3.jpeg}\\
+  \includegraphics[width=0.495\linewidth]{./images/ProtoRoll1.png}\hfill
+  \includegraphics[width=0.495\linewidth]{./images/ProtoRoll2.png}
+  \caption{Top: Prototype used for data acquisition with the rolling sphere.
 The main payload is the Livox Mid-100 laser scanner.
@@
 On the top are two batteries, and on the bottom one voltage stabilizer and the breakout box of the laser scanner.
+Below: Rendering of the simulated robot, a Livox laser scanner insider a spherical shell.
 }

### 2021-08-20 Latex/sections/registration.tex
Incorporated both reviewers comments

@@
 The Hesse distance, i.e.
-the shortest distance to the infinitely extending plane (red), is shown as well as the minimum distance from the 2D point projection to the polygon (purple).}
+the shortest distance to the infinitely extending plane (red), is shown as well as the minimum distance from the 3D point projection to the polygon (purple).}
 	\label{fig:proj}
@@
 To find the PPD, first a corresponding point gets projected onto the infinitely extending plane representation given by the Hesse form, i.e.
-$\widetilde{T}(\vec{p}_k)$ is found from eq.~\eqref{eq:projection} (green point in~\ref{fig:proj}).
+$\widetilde{T}(\vec{p}_k)$ is found from eq.~\eqref{eq:projection} (green point in figure~\ref{fig:proj}).
 Then, the 3D polygon, i.e.
@@
 Note the possibility of $t_0 \notin \left[ 0, 1 \right]$. 
-In that case, the projection of the point $p_i$ onto the line given by eq.~\ref{eq:line} is not between $p_j$ and $p_m$. 
+In that case, the projection of the point $p_i$ onto the line given by eq.~\eqref{eq:line} is not between $p_j$ and $p_m$. 
 Instead, its projection onto the line falls outside of the segment.
@@
 \end{align}
-which is the argument that gives the shortest distance to the line segment, when inserted to eq. \ref{eq:dist}.
+which is the argument that gives the shortest distance to the line segment, when inserted to eq. \eqref{eq:dist}.
 We find the PPD from the point $p_i$ to the polygon $P_k$ by calculating the minimum shortest distance over all line segments that make up the polygon.
@@
 because the opposing normals $-\vec{n}_1$ and $-\vec{n}_2$ always have to be considered, since they correspond to the same plane.
-We say that two normals are similar if the angle $\hat \alpha$ from eq. \ref{eq:alphahat} is smaller than a threshold $\epsilon_{\alpha}$.
+We say that two normals are similar if the angle $\hat \alpha$ from eq. \eqref{eq:alphahat} is smaller than a threshold $\epsilon_{\alpha}$.
 
@@
 	\includegraphics[width=\linewidth]{images/clustering}
-	\caption{(Left) Point normals using the AKNN method with $K = 20$.
+	\caption{(Left) Point normals using the AKNN method with $K = 20$ (number of nearest neighbors).
 	(Middle) Result of the region growing, before applying the second filter.
 	(Right) Resulting clusters after applying the second filter, using $N_{c,min} = 300$. 
-}
+	In each image, one color represents a segment.}
 	\label{fig:clustering}
@@
 Then we set the thresholds accordingly. 
-For each overlap, we find the smallest angle between the corresponding cluster normal and the global plane normal, according to eq. \ref{eq:alphahat}.
+For each overlap, we find the smallest angle between the corresponding cluster normal and the global plane normal, according to eq. \eqref{eq:alphahat}.
 Finally, a correspondence is established between all points in the cluster and the global plane, if the minimum angle between their normals is smaller than $\epsilon_\alpha$.

### 2021-08-20 Latex/sections/results.tex
yt link added to pdf
Incorporated both reviewers comments

@@
 
-The plane-based registration is applied to the simulated dataset with noisy pose and range measurements (cf. Figure~\ref{fig:simulatedDatasets}) without further processing.
+The plane-based registration is applied to the simulated dataset with noisy pose and range measurements (cf. figure~\ref{fig:simulatedDatasets}) without further processing.
 Assuming this represents a coarsely pre-registered 3D point cloud, the distances to the ground truth were evaluated before and after the plane-based registration. 
@@
 We again combine twenty temporally successive scans into one meta-scan, using the scan at the median index as a reference coordinate system, which is then globally registered.
-We used the first seven of these meta-scans, corresponding to the first full rotation of the spher, to extract global planes.
+We used the first seven of these meta-scans, corresponding to the first full rotation of the sphere, to extract global planes.
 All subsequent meta-scans are then matched against the initial global plane model.
@@
 Parameters used for optimization: $\epsilon_H = 200$, $\epsilon_P = 300$, $\epsilon_\alpha = \frac{\pi}{4}$, $K = 200$ for AKNN, $d_{growth} = 50$, $N_{c_min} = 800$, $\alpha_0 = [0.001, 0.001, 0.001, 1, 0, 1]^\tau$, $i = 1500$, $k = 1$.
-An animation of the registration process is given at \url{https://youtu.be/hYR3xLhIdsE}.}
+An animation of the registration process is given at \url{https://youtu.be/2J8y7MN10JM}.}
         \label{fig:jasperhome}

### 2021-08-20 Latex/sections/stateOfTheArt.tex
Incorporated both reviewers comments

@@
 Adding a heuristic for closed-loop detection and a global relaxation Borrmann et al. yield highly precise maps of the scanned environment~\cite{RAS2007}.
+Thereby they reduce the position and orientation error of the scanning poses by orders of magnitude in the range of centimeters and hundreds of a degree.
 Zhang et al. propose a real-time solution to the SLAM problem in~\cite{Zhang2014}.
@@
 However, this platform has to be moved by a human operator. 
+In particular, they are able to achieve one sigma of measurements within three millimeters difference the measurements of a ground truth terrestrial laser scan~\cite{navvisAcc}. 
 They also employ artificial markers that can be placed in the environment, which get recognized by the camera system to further improve registration accuracy.
@@
 
-LOAM~\cite{Zhang2014} is the baseline-algorithm that provides a real-time and lowdrift solution based on two parallel registration algorithms using planes and lines.
+Lidar odometry and mapping (LOAM)~\cite{Zhang2014} is the baseline-algorithm that provides a real-time and lowdrift solution based on two parallel registration algorithms using planes and lines.
 Unfortunately, it is not open-source anymore.
-LOAM livox extends the LOAM framework to the the rotating prism scanner with small FoV~\cite{Lin_2020}.
+LOAM livox extends the LOAM framework to the rotating prism scanner with small FoV~\cite{Lin_2020}.
 Zhou, Wang, and Kaess write that it also adopts the parallel computing to achieve real-time global registration.
@@
 The LIPS system~\cite{Geneva_2018} employes a so-called Closest Point Plane Representation with an Anchor Plane Factor.
-RANSAC is used to find the places.
+Random sample consensus (RANSAC) is used to find the planes.
 Their system couples an eight channel Quanergy M8 LiDAR operating at 10Hz with a Microstrain 3DM-GX3-25 IMU attached to the bottom of the LiDAR operating at 500Hz.

### 2021-08-23 mfi21/ieeeconf/root.tex
first feedback implemented

@@
 Therefore it is not possible to implement a physical behavior of the system to the filter, as easily as it is for other robotic applications.
+There exists the Error State Extended Kalman Filter, which works only with IMUs \cite{ahmadi2007orientation,}.
+Still, this requires substantial computational power.
+Furthermore, for the later introduced position estimation for spherical robots, some calculations are used for the orientation as well as position.
+After all, if computational power is not limited, Kalman Filters provide principally better results.
 \subsection{Madgwick Filter}
@@
 The problem is the optimization for movement. The Madgwick filter has a jittering behavior at the standstill of the IMU.
-This does not meet the desired behavior as described in the prelimitations section (\ref{AssumptionsAndPrelimitations}).
+This does not meet the desired behavior as described in the prelimitations section~(\ref{AssumptionsAndPrelimitations}).
 The behavior is shown in comparison to the complementary filter in Figure~\ref{madgVsComp1}.
 This behavior is most prominent during standstill as the Madgwick filter was designed for movement.
+As we do not use a magentometer, we can not use the gyro bias estimation of the Madgwick filter.
 \subsection{Complementary Filter}
@@
 where $\alpha$ is the complementary gain, it determines how much weight the accelerometer data gets and how fast the value drifts towards the measurement of the accelerometer.
+Therefore the gyroscope functions as high-pass and the accelerometer as low-pass part of the basic compelemntary filter \cite{min2015complementary}.
 A commonly used value for $\alpha$ is 0.02, meaning every iteration, 2 percent of the new values come from the accelerometer.
+This correpsonds, with the used \unit{125}[hz] calculation, to frequency-border of the low and high-pass  of \unit{2.5}[hz] \cite{hajdu2017complementary}.
 An $\alpha$ of 1 means the resulting data is completely the one of the accelerometers, and the gyroscope has no influence. \\
@@
 Therefore we calculate factors from the gyroscope measurements and scale the autogain with the maximum $\beta$ (cf. equation \ref{eq:factors} and \ref{eq:weight_autogain}).
-The $g_{k}$ represents the gyro measurement in the corresponding an axis $k$ in rad/s.
+The $g_{k}$ represents the gyro measurement in the corresponding axis $k$, in rad/s.
 Later on, these factors will again be used for position estimation, namely to determine if the sphere rotates or not.
@@
 \includegraphics[width=\linewidth]{./graphics/LTest.eps}
-\caption{Ground position by integration of rotation and the fusion algorithm. An L-shape with 1 meter by 1 meter was performed with provoked slip(rotation without translation)}
+\caption{Ground position (Z-axis, perpendicular to the ground) by integration of rotation and the fusion algorithm.
+An L-shape with 1 meter by 1 meter was performed with provoked slip(rotation without translation)}
 \label{LTest}
@@
 \includegraphics[width=\linewidth]{./graphics/LTestSlide.eps}
-\caption{X-Y-Position by integration of rotation and the fusion algorithm. An L-shape with 1 meter by 1 meter was performed with provoked slide(translateion without rotation)}
+\caption{Ground position (Z-axis, perpendicular to the ground) by integration of rotation and the fusion algorithm.
+An L-shape with 1 meter by 1 meter was performed with provoked slide(translateion without rotation)}
 \label{LTestSlide}

### 2021-08-30 Latex/sections/results.tex
added links to videos

@@
 The color mapping is equivalent in both.
-An animation of the matching process is given at \url{https://youtu.be/7igQdEeCsYk}.}
+An animation of the matching process is given at \url{https://youtu.be/hFx2uGkUdXw}.}
  	\label{fig:simulatedEvaluation}
@@
 Parameters used for optimization: $\epsilon_H = 50$, $\epsilon_P = 200$, $\epsilon_\alpha = \frac{\pi}{4}$, $K = 20$ for AKNN, $d_{growth} = 50$, $N_{c_min} = 200$, $\alpha_0 = [0.01, 0.01, 0.01, 0, 0, 0]^\tau$, $i = 8$, $k = 100$.
-An animation of the registration process is given at \url{https://youtu.be/Qa2Qi8z0KEk} .}
+An animation of the registration process is given at \url{https://youtu.be//ov5Xjgyl9wA} .}
 	\label{fig:cylon-corrected}
@@
 Parameters used for optimization: $\epsilon_H = 200$, $\epsilon_P = 300$, $\epsilon_\alpha = \frac{\pi}{4}$, $K = 200$ for AKNN, $d_{growth} = 50$, $N_{c_min} = 800$, $\alpha_0 = [0.001, 0.001, 0.001, 1, 0, 1]^\tau$, $i = 1500$, $k = 1$.
-An animation of the registration process is given at \url{https://youtu.be/2J8y7MN10JM}.}
+An animation of the registration process is given at \url{https://youtu.be/Mkil0vLk8f8}.}
         \label{fig:jasperhome}

### 2021-09-01 mfi21/ieeeconf/root.tex
implemented reviwers feedback

@@
 This makes them suitable for interpolating the more precise and absolute position reference points of these slower algorithms.
-The limited computational power of space-qualified CPUs further imposes the need for good orientation estimation without complex algorithms like Kalman filters.
+The limited computational power of space-qualified CPUs further imposes the need for good orientation estimation without computational-intensive algorithms like Kalman filters.
 Lastly, the geometry of a spherical robot opens the possibility for the not often used position estimation by IMUs.
@@
 \label{AssumptionsAndPrelimitations}
-The presented approach requires multiple assumptions to be beneficial. Therefore, we classify the calculated pose concerning its usage and define the usage as follows:
+The presented approach works for any spherical robot which moves based on rotation and possesses IMUs.
+Nevertheless, it requires multiple assumptions to be beneficial in comparison to existing approaches.
+Some of these arise from the fundamental behavior of spherical robots and some more specific from the DAEDALUS project.
+Therefore we state the following assumptions and requirements for the approach to become beneficial:
 
-\paragraph*{Real Time Operations}The operation of the robot does not allow post-calculation of the pose. This implies the usage of the onboard computational power.
-\paragraph*{Limitation of Computational Power}
+\paragraph*{1. Real Time Operations}The operation of the robot does not allow post-calculation of the pose. This implies the usage of the onboard computational power.
+\paragraph*{2. Limitation of Computational Power}
 The need for efficient code without massive bottleneck operations, as the inversion of a matrix with the extended Kalman Filter. The code must minimize the usage of non-basic operations.
-\paragraph*{Low Costs of Hardware}
+\paragraph*{3. Low Costs of Hardware}
 We assume biased and noisy data in a non-negligible amount from the sensors.
-\paragraph*{Sphere Shape}
+\paragraph*{4. Sphere Shape}
 The code relies heavily on the assumption of a sphere as a shape with a constant robot radius. We also assume that rotation of the sphere on the ground leads to translation.
-\paragraph*{Slip and Sliding}
+\paragraph*{5. Slip and Sliding}
 We assume the sphere has both slip, i.e., the sphere rotates, but there is no translation, and sliding, i.e., the sphere has translation without rotation in a limited, non-permanent way.
 The limitation is the assumption that sliding and slipping reduce the amount of translation for a given rotation and vice versa, however, not for complete absence.
-\paragraph*{Stability of Pose}
+\paragraph*{6. Stability of Pose}
 For simultaneous localization and mapping (SLAM) purposes using LIDARs, a stable but maybe minimal false pose is preferred over a more exact but noisy and jumping position.
-\paragraph*{Further Processing of Pose}
+\paragraph*{7. Further Processing of Pose}
 The approach should avoid jumping values or abrupt changes of data if not clearly indicated by the sensors.
@@
 This change shall not be represented by an unnatural acceleration in the data.
-\paragraph*{Uncertainty of Iterative Position Integration}
+\paragraph*{8. Uncertainty of Iterative Position Integration}
 IMUs deliver no absolute reference for a position as they do for orientation in the form of the noisy but measurable gravity vector.
 So without external or secondary sensors relying on absolute waypoints, the position estimation will always underlie the integration of errors.
-\paragraph*{Space Suitability}
+\paragraph*{9. Space Suitability}
 As the code is considered to be applied to space missions, the magnetometer data is not considered in the algorithm, despite its positive impact on orientation estimation on earth.
-\paragraph*{Short Term Mobility}
-As the code still relies on integration parts, the sphere is not considered to have a straight motion for a long period but rather makes multiple short paths.
-This will also be necessary, as in further extension, the fusion of a pose estimation by the laser scanner will be merged.
-\paragraph*{Non-reliablity of locomotion}
-The sphere from the DAEDALUS example has two completely different locomotion methods; therefore, we will not take locomotion itself into account of pose estimation.
-Also, the same locomotion approach leads to heavily different behavior with the same input to the controller \cite{rossi2021daedalus}.
-\paragraph*{Classification of Data}
-The generated data is a compromise between simple filtration of data and heavy calculation post-record.
+\paragraph*{10. Short Term Mobility}
+As the code still relies on integration parts, the sphere is not considered to roll continiously for a long period but rather makes multiple short paths.
+This will also be necessary, as in further extension, the fusion of a pose estimation by the laser scanner will be merged, requiring standstill periods for scanning.
+\paragraph*{11. Non-reliablity of locomotion commands}
+For DAEDALUS, the same input of the locomotion controller leads to heavily different behaviors \cite{rossi2021daedalus}.
+Also, the DAEDALUS robot has two utterly different locomotion methods; therefore, we will not take locomotion itself, nor the controller input or output into account of pose estimation.
+
+We conclude that the generated data is a compromise between on-board simple filtration of data, while still being usable for heavy calculation post-record, e.g. map generation.
 It is easy to calculate and adapt, but due to the number of pre-assumptions, as it makes assumptions about the dynamics of the system, it still delivers reliable data which takes physical behavior into account.
 It is real-time capable and does not rely on delay steps.
-\paragraph*{Field of Use}
 The use is restricted to spherical or cylindrical robots. 
@@
 Therefore it is not possible to implement a physical behavior of the system to the filter, as easily as it is for other robotic applications.
-There exists the Error State Extended Kalman Filter, which works only with IMUs \cite{ahmadi2007orientation,}.
+There exists the Error State Extended Kalman Filter, which works only with IMUs \cite{ahmadi2007orientation}.
 Still, this requires substantial computational power.
@@
 It is efficient and easy to calculate but does not consider the physics of the motion.
+It uses a gradient descent algorithm.
 It does nearly provide the sought-after classification for the orientation.
@@
 The basic concept is to have the main part of a value dominated by a gyroscope but, over time, a slow drift towards the measurement of the accelerometer.
-E.g., for pitch, this would result in 
+E.g., for pitch, this results in 
 \begin{equation}
@@
 \includegraphics[width=\linewidth]{./graphics/MadgwickVsComplementaryNoMovement.eps}
-\caption{Quaternion values of Madgwick filter and complementary filter with default gains when IMU is at a standstill. The Madgwick filter has clearly visible jitter, the complementary filter not}
+\caption{Quaternion values of Madgwick filter and complementary filter with default gains when IMU is at a standstill. The Madgwick filter has clearly visible jitter, the complementary filter not.}
 \label{madgVsComp1}
@@
 \subsection{Position Estimation}
-Position estimation purely by IMU is not recommended as the double integration of the accelerometer measurements leads to exponential error summation.
-Therefore there are very few approaches targeting this way of position calculating \cite{kok2017IMU}.
+Position estimation purely by IMU is not recommended as the double integration of the accelerometer measurements leads to exponential error summation \cite{thong2004numerical}.
+Let $a$ be the real acceleration and $\epsilon_a$ the error of the measurment, then
+\begin{equation}
+\iint (a+\epsilon_a) = \iint a+  \iint \epsilon_a = \frac{1}{2}a\cdot t^2+  \frac{1}{2} \epsilon_a \cdot t^2 \,,
+\end{equation}
+shows that the position depends on $\epsilon_a$ exponentially.
+
+Therefore there are very few approaches targeting this way of position calculation \cite{kok2017IMU}.
 The very impressive work in \cite{valencia2017simpleIMU} also uses IMUs for 6D pose estimation.
@@
 \includegraphics[width=\linewidth]{./graphics/imuJasperChartSimliefiedBIG.png}
-\caption{Generalized diagram of the estimation}
+\caption{Diagram of the estimation.}
 \label{generalized}
@@
 The Madgwick filter itself is untouched.
-The standard complimentary filter is adapted to work in quaternion representation.
- As the integration only works in the Euler angles domain, the jumps of values at the border of the domain like 0 and 360 degrees do not match the mostly continuous change of quaternions.
-As the representation is still valid (0 degree and 360-degree roll being the same position), this is not a problem when using at one point any of the filters exclusively.
+The standard complimentary filter is adapted to work in quaternion representation,
+ as the direct integration of the gyroscope give angles in the Euler domain.
+Quaternions were designed under the premise of being free from singularities, i.e., for a continuous change in orientation, there exists a continuous change in quaternion representation \cite{becker2017dealing}.
+This is not the case for the Euler angle, as a transition of a 359-degree roll to a 0-degree roll has no continuous value representation if limited to 360 degrees.
  But combining both leads to contradicting updates from each filter with respect to the other.
@@
 Therefore we calculate factors from the gyroscope measurements and scale the autogain with the maximum $\beta$ (cf. equation \ref{eq:factors} and \ref{eq:weight_autogain}).
-The $g_{k}$ represents the gyro measurement in the corresponding axis $k$, in rad/s.
-Later on, these factors will again be used for position estimation, namely to determine if the sphere rotates or not.
+$g_{k}$ represents the gyro measurement in the corresponding axis $k$, in rad/s. Later on, these factors will again be used for position estimation, namely to determine if the sphere rotates or not.
 These factors are heuristically defined as:
@@
 \end{equation}
-These Values are now coupled to the rotation of both axes other than their own. Meaning there can be a translation in $x$ if there is rotation around $y$ or $z$.
+These values are now coupled to the rotation of both axes other than their own.
 Therefore, the factors from the orientation step, representing the likelihood of rotation, are used as limiting factors.
@@
 When there is no rotation, it will set the velocity to 0.
-During fast movement, this would allow exponential error integration. 
+During fast movement, this allows exponential error integration. 
 This exponential error is taken care of later on.
@@
 \end{bmatrix}
-\cdot r 
-\end{equation}.
+\cdot r \,.
+\end{equation}
 As the gyro measures rad/s the value is divided by $2\pi$.
@@
  The higher the likelihood of sliding, the higher the velocity should be compared to the velocity based on rotation.
- This does not count for slipping, as this would imply the adaption of limitation of the velocity by rotation by the velocity measured by the accelerometer.
+ This does not count for slipping, as this implies the adaption of limitation of the velocity by rotation by the velocity measured by the accelerometer.
  But there is no need for a limitation as there is no double integration with the rotation.\\
@@
 There should always be a mixture of both if there is slipping and sliding.
-$\beta =0.5$ would result in just the mean of both approaches.
+$\beta =0.5$  results in just the mean of both approaches.
 
@@
 \includegraphics[width=\linewidth]{./graphics/AutoVsMad.eps}
-\caption{Quaternion-values of the presented fusion of complementary and Madgwick filter and only the Madgwick filter}
+\caption{Quaternion-values of the presented fusion of complementary and Madgwick filter and only the Madgwick filter.}
 \label{quaternion1}
@@
 As described in Section \ref{sec:SOA}, the Madgwick filter has a rather strong jitter during slow movements and standstill. 
+This comes due to the (too) large steps of gradient descent iterations.
+To compensate for this, one needs to lower the gain, leading to more imprecise results during movement, as large iteration steps are beneficial there.
+Alternatively, one needs to increase calculation frequency, which is in our scenarios limited by the hardware capabilities of the IMU.
+
 To evaluate specifically the change between slow motion and standstill, a further experiment is performed.
@@
 \includegraphics[width=\linewidth]{./graphics/distance.eps}
-\caption{Distance of the proposed and the Madgwick algorithm to the optical tracking system, using the distance definiton of \cite{kuffner2004effectiveDistance}}
+\caption{Distance of the proposed and the Madgwick algorithm to the optical tracking system, using the distance definiton of \cite{kuffner2004effectiveDistance}.}
 \label{distance}
@@
 \includegraphics[width=\linewidth]{./graphics/diffdistance.eps}
-\caption{Derivative of the distance of the proposed and the Madgwick algorithm to the optical tracking system, using the distance definiton of \cite{kuffner2004effectiveDistance}}
+\caption{Derivative of the distance of the proposed and the Madgwick algorithm to the optical tracking system, using the distance definiton of \cite{kuffner2004effectiveDistance}.}
 \label{diffdistance}
@@
 \caption{Ground position (Z-axis, perpendicular to the ground) by integration of rotation and the fusion algorithm.
-An L-shape with 1 meter by 1 meter was performed with provoked slip(rotation without translation)}
+An L-shape with 1 meter by 1 meter was performed with provoked slip(rotation without translation).
+The boxes are manually set points as a reference.
+The distance in meter between these points is written on the lines connecting them.
+The rolling started in the upper right corner and ended in the bottom left corner.
+}
 \label{LTest}
@@
 
-The same experiment was performed with added slide, i.e., more translation than just the rotation would produce.
+The same experiment was performed with added slide, i.e., more translation than just the rotation produces.
 Here the results shown in Figure \ref{LTestSlide} were not as reliably good as with the slip experiment.
@@
 \caption{Ground position (Z-axis, perpendicular to the ground) by integration of rotation and the fusion algorithm.
-An L-shape with 1 meter by 1 meter was performed with provoked slide(translateion without rotation)}
+An L-shape with 1 meter by 1 meter was performed with provoked slide(translation without rotation).The boxes are manually set points as a reference.
+The distance in meter between these points is written on the lines connecting them.
+The rolling started in the lower right corner and ended in the upper left corner.}
 \label{LTestSlide}
@@
 Both images were shot from the exact same point of view.
-}
+Both plots use the same LIDAR data and apply one of the simultaneously generated pose estimations.}
 \label{fig:mapping}
@@
 One major design driver for the presented algorithm is the suitability for spherical SLAM purposes.
-Therefore, a pose estimate is prefered to be slightly wrong, if it is more continuous, i.e. less noisy and with less jumping values, than the exact pose would be.
+Therefore, a pose estimate is prefered to be slightly wrong, if it is more continuous, i.e. less noisy and with less jumping values, than the exact pose has.
 

### 2021-09-01 mfi21/ieeeconf/root.tex
small changes, suggested by reviewer

@@
 \section{Assumptions and Prelimitations}
-\label{AssumptionsAndPrelimitations}
+\label{sec:AssumptionsAndPrelimitations}
 The presented approach works for any spherical robot which moves based on rotation and possesses IMUs.
@@
 Further, the Kalman Filter relies on input to a system to calculate an estimation of the pose for the next iterative step.
-But, as we described in the preliminary section, we can not take the input to locomotion control as a reference to pose estimation as the behavior is not linear or even predictable without exact knowledge of the surface.
+But, as we described in the preliminary Section \ref{sec:AssumptionsAndPrelimitations}, we can not take the input to locomotion control as a reference to pose estimation as the behavior is not linear or even predictable without exact knowledge of the surface.
 Therefore it is not possible to implement a physical behavior of the system to the filter, as easily as it is for other robotic applications.
@@
 The problem is the optimization for movement. The Madgwick filter has a jittering behavior at the standstill of the IMU.
-This does not meet the desired behavior as described in the prelimitations section~(\ref{AssumptionsAndPrelimitations}).
+This does not meet the desired behavior as described in the prelimitations Section~(\ref{AssumptionsAndPrelimitations}).
 The behavior is shown in comparison to the complementary filter in Figure~\ref{madgVsComp1}.

### 2021-09-02 mfi21/ieeeconf/root.tex
small changes

@@
 \usepackage{hyperref}
+\usepackage{siunitx}
 %\usepackage{enumitem}
@@
 Furthermore, the inherent rotation the robot uses for locomotion can be exploited to measure in all directions without having the sensor itself actuated.
-To exploit this rotation in combination with sensor data to create a consistent environment map, a reasonable estimation of the robot pose is required. 
-In such cases the pose estimation are done by inertial measurements from sensors such as accelerometers and gyroscopes, as interpolating instances for calculation-intensive slow algorithms such as optical localization algorithms or as initial estimate for subsequent SLAM.
-The paper presents a pose estimation procedure based on inertial measurements that exploits the known dynamics of a spherical robot. 
+A reasonable estimation of the robot pose is required to exploit this rotation in combination with sensor data to create a consistent environment map. 
+This raises the need for interpolating instances for calculation-intensive slow algorithms such as optical localization algorithms or as an initial estimate for subsequent SLAM.
+In such cases,  inertial measurements from sensors such as accelerometers and gyroscopes generate a pose estimation for these interpolation steps.
+
+The paper presents a pose estimation procedure based on inertial measurements, that exploits the known dynamics of a spherical robot. 
 It emphasizes a low jitter to maintain constant world measurements during standstill and avoids exponentially growing error in position estimates. 
@@
 As an example, the 2021 CDF study LunarCaves by ESA about the feasibility of a spherical robot, called DAEDALUS, for exploration of lunar-lava tubes \cite{rossi2021daedalus} shows the need for orientation and position estimation of spherical robots with the limitations on resources common for space-applications.
-Although a purely Inertial Measurement Unit (IMU)-based estimation is not recommended for various reasons, the lack of an absolute reference among others, IMU-based pose estimation are used as part of an overall multi-sensor-fusion-based estimation.
+Although a purely Inertial Measurement Unit (IMU)-based estimation is not recommended for various reasons, the lack of an absolute reference among others, IMU-based pose estimation is used as part of an overall multi-sensor-fusion-based estimation.
 IMUs are sensors with higher refresh rates than Simultaneous Localization and Mapping (SLAM) based algorithms, such as optical or Light Detection and Ranging (LIDAR) SLAM.
@@
 For pose-estimation, three IMUs of the manufacturer Phidget are placed on the middle plate and a Raspberry Pi 4 for the calculations.
-On the top are two batteries, and on the bottom one voltage stabilizer and breakout box of the laser-scanner.
+On the top are two batteries and on the bottom one voltage stabilizer and breakout box of the laser-scanner.
 }
@@
 
-\paragraph*{1. Real Time Operations}The operation of the robot does not allow post-calculation of the pose. This implies the usage of the onboard computational power.
+\paragraph*{1. Real Time Operations}The operation of the robot does not allow post-calculation of the pose. This implies the usage of onboard computational power.
 \paragraph*{2. Limitation of Computational Power}
@@
 \paragraph*{10. Short Term Mobility}
-As the code still relies on integration parts, the sphere is not considered to roll continiously for a long period but rather makes multiple short paths.
-This will also be necessary, as in further extension, the fusion of a pose estimation by the laser scanner will be merged, requiring standstill periods for scanning.
-\paragraph*{11. Non-reliablity of locomotion commands}
+As the code still relies on integration parts, the sphere is not considered to roll continuously for a long period but rather makes multiple short paths.
+This will also be necessary, as, in further extension, the fusion of a pose estimation by the laser scanner will be merged, requiring standstill periods for scanning.
+\paragraph*{11. Non-reliability of locomotion commands}
 For DAEDALUS, the same input of the locomotion controller leads to heavily different behaviors \cite{rossi2021daedalus}.
@@
 We conclude that the generated data is a compromise between on-board simple filtration of data, while still being usable for heavy calculation post-record, e.g. map generation.
-It is easy to calculate and adapt, but due to the number of pre-assumptions, as it makes assumptions about the dynamics of the system, it still delivers reliable data which takes physical behavior into account.
-It is real-time capable and does not rely on delay steps.
+It is easy to calculate and adapt the specific needs.
+Still, as it makes assumptions about the dynamics of the system, it delivers reliable data which takes physical behavior into account. 
+It is real-time capable and does not rely on delayed steps.
 The use is restricted to spherical or cylindrical robots. 
@@
 The problem is the optimization for movement. The Madgwick filter has a jittering behavior at the standstill of the IMU.
-This does not meet the desired behavior as described in the prelimitations Section~(\ref{AssumptionsAndPrelimitations}).
+This does not meet the desired behavior as described in the prelimitations Section~(\ref{sec:AssumptionsAndPrelimitations}).
 The behavior is shown in comparison to the complementary filter in Figure~\ref{madgVsComp1}.
 This behavior is most prominent during standstill as the Madgwick filter was designed for movement.
-As we do not use a magentometer, we can not use the gyro bias estimation of the Madgwick filter.
+As we do not use a magnetometer, we can not use the gyro bias estimation of the Madgwick filter.
 \subsection{Complementary Filter}
@@
 where $\alpha$ is the complementary gain, it determines how much weight the accelerometer data gets and how fast the value drifts towards the measurement of the accelerometer.
-Therefore the gyroscope functions as high-pass and the accelerometer as low-pass part of the basic compelemntary filter \cite{min2015complementary}.
+Therefore the gyroscope functions as high-pass and the accelerometer as low-pass part of the basic complementary filter \cite{min2015complementary}.
 A commonly used value for $\alpha$ is 0.02, meaning every iteration, 2 percent of the new values come from the accelerometer.
@@
 \includegraphics[width=\linewidth]{./graphics/MadgwickVsComplementaryNoMovement.eps}
-\caption{Quaternion values of Madgwick filter and complementary filter with default gains when IMU is at a standstill. The Madgwick filter has clearly visible jitter, the complementary filter not.}
+\caption{Quaternion values of Madgwick filter and complementary filter with default gains when IMU is at a standstill.
+The Madgwick filter has clearly visible jitter, the complementary filter has not.}
 \label{madgVsComp1}
@@
 Position estimation purely by IMU is not recommended as the double integration of the accelerometer measurements leads to exponential error summation \cite{thong2004numerical}.
-Let $a$ be the real acceleration and $\epsilon_a$ the error of the measurment, then
-\begin{equation}
-\iint (a+\epsilon_a) = \iint a+  \iint \epsilon_a = \frac{1}{2}a\cdot t^2+  \frac{1}{2} \epsilon_a \cdot t^2 \,,
-\end{equation}
-shows that the position depends on $\epsilon_a$ exponentially.
-
 Therefore there are very few approaches targeting this way of position calculation \cite{kok2017IMU}.
@@
 Every rotation will be predicted as translation in a given plane, which is most likely parallel to the ground. 
-For instance, if the sphere rolls up an obstacle, the length of the path is estimated to the ground-plane.\\
+For instance, if the sphere rolls up an obstacle, the length of the path is estimated to be the ground-plane.\\
 Our algorithm uses both approaches of position estimation and combines them, and limits all components according to the other components. 
 This implies a connection between translation and rotation. 
-Overall the extreme situation (full translation, no rotation, or the other way round) will automatically lead to wrong predictions, as there is no way to tell which of the two approaches has at that moment more reliable data and therefore, the combination of both rather then only one is chosen.
+Overall the extreme situation (full translation, no rotation, or the other way round) will automatically lead to wrong predictions, as there is no way to tell which of the two approaches has at that moment more reliable data.
+Therefore, the combination of both rather than only one is chosen.
 
@@
  These values have been determined empirically.
- Future work will investigate on improving those values by developing a suitable heuristic for the thresholds.
+ Future work will investigate improving those values by developing a suitable heuristic for the thresholds.
  The transition between the full complementary and the full Madgwick estimation, as is required by the prelimitations,  can be done by a function suitable for the specific scenario.
@@
 \end{align}
-Thus, the Madgwick gain has a maximum of $\Theta$ and $\alpha$ a maximum of $\theta\cdot \Theta$. This ratio $\theta$ can be adapted to the specific needs.
-We used $\theta = 0.1$ because of the the two often used standard values $\gamma = 0.2$ and $\alpha= 0.02$, which have the ration $0.1$.
+Thus, the Madgwick gain has a maximum of $\Theta$ and $\alpha$ a maximum of $\theta\cdot \Theta$. This ratio $\theta$ can be adapted to specific needs.
+We used $\theta = 0.1$ because of the two often used standard values $\gamma = 0.2$ and $\alpha= 0.02$, which have the ration $0.1$.
 Hence for adapting the position estimation to a given robot, the function for calculating the factors needs to be adapted in such a way that it is a good indication for movement or standstill.
@@
 Therefore, the factors from the orientation step, representing the likelihood of rotation, are used as limiting factors.
-If one of the three factors is 1, this means there is translation in this direction.
+If one of the three factors is 1, this means there is a translation in this direction.
 When there is no rotation, it will set the velocity to 0.
@@
 Its effect on the position is considered when determining the orientation in the world coordinate system.
-$\omega_{z\text{world}}$ needs not to be computed as it has no influence on the position.
+$\omega_{z\text{world}}$ does not need to be computed as it has no influence on the position.
 
@@
 
- Next the $v_{x\text{AccWorld}} $and $v_{y\text{AccWorld}} $ are limited depending on $v_{xz\text{GyroWorld}} $ and $ v_{yz\text{GyroWorld}}$, in our example this was 120 percent of the gyro calculated velocity.
+ Next, the $v_{x\text{AccWorld}} $and $v_{y\text{AccWorld}} $ are limited depending on $v_{xz\text{GyroWorld}} $ and $ v_{yz\text{GyroWorld}}$, in our example, this was 120 percent of the gyro calculated velocity.
  This ensures no exponential error integration when in motion.
@@
 A first experiment is performed to show the overall performance of the Autogain computation for the orientation.
-The IMU was manually rotated in arbitrary directions, and the orientation plotted using the Madgwick filter and the Autogain filter.
+The IMU was manually rotated in arbitrary directions, and the orientation was plotted using the Madgwick filter and the Autogain filter.
 As expected, the overall behavior of the presented algorithm and the Madgwick filter is very similar in the dynamic case, as shown in Figure \ref{quaternion1}.
@@
 The optical system matches the overall orientation of both IMU-based algorithms quite well and lacks the perturbations of the pure Madgwick filter.
-The transformation of the orientation by the optical tracking system and the IMU-based algorithms yield two unknown transformations, the relative orientation of the coordinate system of IMU to optical coordinate system and how the axes are interpreted on the tracked structure itself.
+The transformation of the orientation by the optical tracking system and the IMU-based algorithms yield two unknown transformations, the relative orientation of the coordinate system of IMU to the optical coordinate system and how the axes are interpreted on the tracked structure itself.
 Therefore, the algorithm proposed in \cite{dornaika1998calibration} was used to solve these two unknown transformations on basis of the resulting data, to match the orientations of optical and IMU-based data.
@@
 \includegraphics[width=\linewidth]{./graphics/distance.eps}
-\caption{Distance of the proposed and the Madgwick algorithm to the optical tracking system, using the distance definiton of \cite{kuffner2004effectiveDistance}.}
+\caption{Distance of the proposed and the Madgwick algorithm to the optical tracking system, using the distance definition of \cite{kuffner2004effectiveDistance}.}
 \label{distance}
@@
 \includegraphics[width=\linewidth]{./graphics/diffdistance.eps}
-\caption{Derivative of the distance of the proposed and the Madgwick algorithm to the optical tracking system, using the distance definiton of \cite{kuffner2004effectiveDistance}.}
+\caption{Derivative of the distance of the proposed and the Madgwick algorithm to the optical tracking system, using the distance definition of \cite{kuffner2004effectiveDistance}.}
 \label{diffdistance}
@@
 \caption{Ground position (Z-axis, perpendicular to the ground) by integration of rotation and the fusion algorithm.
-An L-shape with 1 meter by 1 meter was performed with provoked slide(translation without rotation).The boxes are manually set points as a reference.
+An L-shape with 1 meter by 1 meter was performed with provoked slide(translation without rotation).
+The boxes are manually set points as a reference.
 The distance in meter between these points is written on the lines connecting them.
@@
 One major design driver for the presented algorithm is the suitability for spherical SLAM purposes.
-Therefore, a pose estimate is prefered to be slightly wrong, if it is more continuous, i.e. less noisy and with less jumping values, than the exact pose has.
+Therefore, a pose estimate is preferred to be slightly wrong, if it is more continuous, i.e. less noisy and with less jumping values, than the exact pose has.
 
 To test the impact of the filter on 3D mapping, we used the prototype in Figure \ref{fig:prototype}.
-After 2 s of initial standstill, the sphere rolled aprox. 4 m within 5 s in an indoor environment from one corner of the room to the opposing one.
+After 2 s of initial standstill, the sphere rolled approx. 4 m within 5 s in an indoor environment from one corner of the room to the opposing one.
 We extracted the laserscan data, as well as the filtered pose using the presented algorithm.
@@
 The presented filter improves the resulting map compared to the Madgwick/integrator approach.
-In particular, the walls and the ceiling are more recognizable, which is favourable for many SLAM algorithms.
+In particular, the walls and the ceiling are more recognizable, which is favorable for many SLAM algorithms.
 Another improvement is that points were no longer sensed below the ground, i.e. the plane on which the sphere was moving on.
@@
 
-The position estimation, as well as the combination of the integration by rotation of the gyroscope and the linear acceleration of the accelerometer, shows the capability to overcome the problem of the missing absolute reference of IMUs partly, which mainly improves the measurements in the ground plane with slip behavior present.
+The presented position estimation shows the capability to overcome the problem of the missing absolute reference of IMUs partly.
+It mainly improves the measurements in the ground plane with slip behavior present.
 Still, there is a need for parameter tuning. 

### 2021-09-07 mfi21/ieeeconf/root.tex
shortend to 8 side
above -> left:

@@
 \usepackage{hyperref}
-\usepackage{siunitx}
+\usepackage{units}
 %\usepackage{enumitem}
@@
 The behavior is shown in comparison to the complementary filter in Figure~\ref{madgVsComp1}.
-This behavior is most prominent during standstill as the Madgwick filter was designed for movement.
+This jittering behavior is most prominent during standstill as the Madgwick filter was designed for movement.
 As we do not use a magnetometer, we can not use the gyro bias estimation of the Madgwick filter.
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{./graphics/MadgwickVsComplementaryNoMovement.eps}
+%\includegraphics[width=\linewidth]{./graphics/MadgwickVsComplementaryNoMovement.eps}
+\makebox[\linewidth][c]{\includegraphics[width=1.15\linewidth]{./graphics/CompelemntaryVsMadgwick6seconds.eps}}
 \caption{Quaternion values of Madgwick filter and complementary filter with default gains when IMU is at a standstill.
@@
 \begin{figure}
-\begin{center}
-\includegraphics[width=\linewidth]{./graphics/jasperhome1.png} \\
-\vspace{0.2cm}
-\includegraphics[width=\linewidth]{./graphics/jasperhome1madw.png}
-\caption{Above: resulting 3D map using the presented algorithm for pose filtering.
-Below: resulting 3D map using Madgwick + double integrator for pose filtering.
+%\begin{center}
+\includegraphics[width=0.52\linewidth]{./graphics/jasperhome1.png} 
+%\vspace{0.2cm}
+\includegraphics[width=0.45\linewidth]{./graphics/jasperhome1madw.png}
+\caption{Left: resulting 3D map using the presented algorithm for pose filtering.
+Right: resulting 3D map using Madgwick + double integrator for pose filtering.
 Both images were shot from the exact same point of view.
@@
 \label{fig:mapping}
-\end{center}
+%\end{center}
 \end{figure}

### 2021-09-09 mfi21/ieeeconf/root.tex

@@
 \usepackage{amsmath}
-\usepackage{hyperref}
+\usepackage[hidelinks]{hyperref}
 \usepackage{units}
@@
 \thanks{We acknowledge funding from the ESA Contract No.
-4000130925/20/NL/GLC for the ``DAEDALUS -- Descent And Exploration in Deep Autonomy of Lava Underground Structures” Open Space Innovation Platform (OSIP) lunar caves-system study and the Elite Network Bavaria (ENB) for providing funds for the academic program “Satellite Technology''.}
+4000130925/20/NL/GLC for the ``DAEDALUS -- Descent And Exploration in Deep Autonomy of Lava Underground Structures'' Open Space Innovation Platform (OSIP) lunar caves-system study and the Elite Network Bavaria (ENB) for providing funds for the academic program ``Satellite Technology''.}
 }
@@
 A reasonable estimation of the robot pose is required to exploit this rotation in combination with sensor data to create a consistent environment map. 
-This raises the need for interpolating instances for calculation-intensive slow algorithms such as optical localization algorithms or as an initial estimate for subsequent SLAM.
-In such cases,  inertial measurements from sensors such as accelerometers and gyroscopes generate a pose estimation for these interpolation steps.
+This raises the need for interpolating instances for calculation-intensive slow algorithms such as optical localization algorithms or as an initial estimate for subsequent simultaneous localization and mapping (SLAM).
+In such cases,  inertial measurements from sensors such as accelerometers and gyroscopes generate a pose estimate for these interpolation steps.
 
@@
 It emphasizes a low jitter to maintain constant world measurements during standstill and avoids exponentially growing error in position estimates. 
-Evaluating the position and orientation estimates with given ground truth frames shows that we reduce the jitter in orientation and handle slip and partly slide behavior as other commonly used filters such as the Madgwick filter.
+Evaluating the position and orientation estimates with given ground truth frames shows that we reduce the jitter in orientation and handle slip and partly slide behavior better than other commonly used filters such as the Madgwick filter.
 
@@
 This paper introduces an IMU-based orientation and position algorithm for spherical robots with limited computational power and the need for real-time estimations.
-Also, the algorithm does not rely on one specific locomotion approach, but it assumes a rotation of the IMU together with at least the shell of the robot, if not the whole robot. 
-Such as the prototype developed at the computer science chair VII of the University of Wuerzburg,  depicted in Figure \ref{fig:prototype}, which is inspired by previous work such as Daedalus \cite{rossi2021daedalus}.
-The evaluations made in this paper were done with the setup of the IMUs of this prototype.
-The Livox Mid-100 was used as laser scanner.
+The algorithm does not rely on one specific locomotion approach, but it assumes a rotation of the IMU together with at least the shell of the robot, if not the whole robot. 
+The experiments in this paper use the setup of the IMUs of the prototype depicted in Figure \ref{fig:prototype}, which is inspired by previous work such as Daedalus \cite{rossi2021daedalus}.
 Due to the lack of a locomotion system, the sphere needs to be rolled externally. 
-The evaluation will also show and discuss the resulting point clouds of this prototype with the presented pose estimation.
+A Livox Mid-100 laser scanner was used for evaluation purposes.
+In addition to analying the filtered output the evaluation will show and discuss the resulting point clouds of this prototype with the presented pose estimation.
 \begin{figure}[b]
@@
 \label{sec:AssumptionsAndPrelimitations}
-The presented approach works for any spherical robot which moves based on rotation and possesses IMUs.
+The presented approach works for any spherical robot with IMUs that moves based on rotation.
 Nevertheless, it requires multiple assumptions to be beneficial in comparison to existing approaches.
 Some of these arise from the fundamental behavior of spherical robots and some more specific from the DAEDALUS project.
-Therefore we state the following assumptions and requirements for the approach to become beneficial:
+The approach becomes beneficial given the following assumptions and requirements:
 
-\paragraph*{1. Real Time Operations}The operation of the robot does not allow post-calculation of the pose. This implies the usage of onboard computational power.
+\paragraph*{1. Real Time Operations}The operation of the robot requires real time calculation of the pose. This implies the usage of onboard computational power.
 \paragraph*{2. Limitation of Computational Power}
-The need for efficient code without massive bottleneck operations, as the inversion of a matrix with the extended Kalman Filter. The code must minimize the usage of non-basic operations.
+Due to the need for efficient code without massive bottleneck operations, such as matrix inversion in the extended Kalman Filter, the code must minimize the usage of non-basic operations.
 \paragraph*{3. Low Costs of Hardware}
@@
 \paragraph*{4. Sphere Shape}
-The code relies heavily on the assumption of a sphere as a shape with a constant robot radius. We also assume that rotation of the sphere on the ground leads to translation.
+The code relies heavily on the assumption of a sphercal shape with a constant radius. Accordingly, rotation on the ground leads to translation.
 \paragraph*{5. Slip and Sliding}
-We assume the sphere has both slip, i.e., the sphere rotates, but there is no translation, and sliding, i.e., the sphere has translation without rotation in a limited, non-permanent way.
-The limitation is the assumption that sliding and slipping reduce the amount of translation for a given rotation and vice versa, however, not for complete absence.
+We assume the sphere is affected by both slip, i.e., the sphere rotates with no resulting translation, and sliding, i.e., the sphere may translate without rotating in a limited, non-permanent way.
+The limitation refers to the assumption that sliding and slipping reduce the amount of translation for a given rotation and vice versa, however, not for complete absence.
 \paragraph*{6. Stability of Pose}
-For simultaneous localization and mapping (SLAM) purposes using LIDARs, a stable but maybe minimal false pose is preferred over a more exact but noisy and jumping position.
+For SLAM purposes using LIDARs, a stable but maybe minimal false pose is preferred over a more exact but noisy and jumping position.
 \paragraph*{7. Further Processing of Pose}
@@
 \paragraph*{8. Uncertainty of Iterative Position Integration}
-IMUs deliver no absolute reference for a position as they do for orientation in the form of the noisy but measurable gravity vector.
+IMUs deliver no absolute reference for a position as they do for orientation in the form of a noisy but measurable gravity vector.
 So without external or secondary sensors relying on absolute waypoints, the position estimation will always underlie the integration of errors.
 \paragraph*{9. Space Suitability}
-As the code is considered to be applied to space missions, the magnetometer data is not considered in the algorithm, despite its positive impact on orientation estimation on earth.
+As the algorithm is considered to be applied for space missions, the magnetometer data is not considered in the algorithm, despite its positive impact on orientation estimation on Earth.
 \paragraph*{10. Short Term Mobility}
-As the code still relies on integration parts, the sphere is not considered to roll continuously for a long period but rather makes multiple short paths.
-This will also be necessary, as, in further extension, the fusion of a pose estimation by the laser scanner will be merged, requiring standstill periods for scanning.
+The approach is targeted at a spherical robot for data collection with a laser scanner, requiring standstill periods for scanning.
+In further extension, the fusion of a pose estimation by the laser scanner will be merged.
+Thus, the sphere is considered to move in multiple short paths and not roll continuously for a long perod.
+Relying on integration parts, the IMU-based algorithm benefits from this assumption.
 \paragraph*{11. Non-reliability of locomotion commands}
-For DAEDALUS, the same input of the locomotion controller leads to heavily different behaviors \cite{rossi2021daedalus}.
-Also, the DAEDALUS robot has two utterly different locomotion methods; therefore, we will not take locomotion itself, nor the controller input or output into account of pose estimation.
+For DAEDALUS, the same input of the locomotion controller leads to heavily different behaviors depending on the pose of the robot and the surface structure \cite{rossi2021daedalus}.
+Also, the DAEDALUS robot has two utterly different locomotion methods; therefore, we will not take locomotion itself, nor the controller input or output into account for pose estimation.
 
-We conclude that the generated data is a compromise between on-board simple filtration of data, while still being usable for heavy calculation post-record, e.g. map generation.
-It is easy to calculate and adapt the specific needs.
-Still, as it makes assumptions about the dynamics of the system, it delivers reliable data which takes physical behavior into account. 
-It is real-time capable and does not rely on delayed steps.
-The use is restricted to spherical or cylindrical robots. 
-It is optimized for gathering laser data where noise position while standing means a lot of calculation in the map creation, and therefore, even if slightly false, a stable position is preferred.
+We conclude that the generated data has to be a compromise between simple on-board filtration of data, while still being usable for heavy calculation post-record, e.g. map generation.
+It has to be easy to calculate and adapt to the specific needs.
+Making assumptions about the dynamics of the system, it must nevertheless deliver reliable estimates which take physical behavior into account.
+It shall be real-time capable and does not rely on delayed steps.
+The resulting algorithm is restricted to spherical or cylindrical robots. 
+It is optimized for gathering laser data where noisy pose estimates while standing deteriorate the map creation process, and therefore, even if slightly false, a stable position is preferred.
 
@@
 \subsection{Kalman Filter}
-The Kalman Filter or the more suitable Extended Kalman Filter (EKF) uses state prediction and merges with sensor data for orientation estimation \cite{Kalman1960}.
+The Kalman Filter or the more suitable Extended Kalman Filter (EKF) use state prediction and correction based on sensor data for orientation estimation \cite{Kalman1960}.
 For the defined limitations, the Kalman Filter is not suitable.
-First, the EKF requires a considerable amount of computational power, mainly because of its matrix inversion. For this, the needed calculations grow exponentially for input data.
+First, the EKF requires a considerable amount of computational power, mainly due to matrix inversion. For this, the needed calculations grow exponentially with input data.
 Further, the Kalman Filter relies on input to a system to calculate an estimation of the pose for the next iterative step.
-But, as we described in the preliminary Section \ref{sec:AssumptionsAndPrelimitations}, we can not take the input to locomotion control as a reference to pose estimation as the behavior is not linear or even predictable without exact knowledge of the surface.
+But, as described in Section \ref{sec:AssumptionsAndPrelimitations}, we cannot take the input to locomotion control as a reference for pose estimation as the behavior is not linear or even predictable without exact knowledge of the surface.
 Therefore it is not possible to implement a physical behavior of the system to the filter, as easily as it is for other robotic applications.
-There exists the Error State Extended Kalman Filter, which works only with IMUs \cite{ahmadi2007orientation}.
-Still, this requires substantial computational power.
-Furthermore, for the later introduced position estimation for spherical robots, some calculations are used for the orientation as well as position.
+The Error State Extended Kalman Filter, which works solely with IMU data \cite{ahmadi2007orientation}, requires substantial computational power.
+For the later introduced position estimation for spherical robots, some calculations are used for the orientation as well as for the position.
 After all, if computational power is not limited, Kalman Filters provide principally better results.
@@
 It uses a gradient descent algorithm.
-It does nearly provide the sought-after classification for the orientation.
-The problem is the optimization for movement. The Madgwick filter has a jittering behavior at the standstill of the IMU.
+It nearly provides the sought-after classification for the orientation.
+The behavior is shown in comparison to the complementary filter in Figure~\ref{madgVsComp1}.
+Being optimized for movement the Madgwick filter has a jittering behavior at the standstill of the IMU.
 This does not meet the desired behavior as described in the prelimitations Section~(\ref{sec:AssumptionsAndPrelimitations}).
-The behavior is shown in comparison to the complementary filter in Figure~\ref{madgVsComp1}.
-This jittering behavior is most prominent during standstill as the Madgwick filter was designed for movement.
-As we do not use a magnetometer, we can not use the gyro bias estimation of the Madgwick filter.
+As we do not use a magnetometer, we cannot use the gyro bias estimation of the Madgwick filter.
 \subsection{Complementary Filter}
-A complementary filter is a fundamental approach of combining precise but not exact with nonprecise but exact data.
+A complementary filter is a fundamental approach of combining precise but not exact with non-precise but exact data.
 Therefore it is widely used for combining noisy accelerometer data with gravity as a reference with the precise but drifting measurement of gyroscopes.
@@
 \begin{equation}
-\theta_n = (1-\alpha) \cdot (\theta_{n-1}+ \Delta\theta_{gyro})+ \alpha \cdot \theta_{acc}
+\theta_n = (1-\alpha) \cdot (\theta_{n-1}+ \Delta\theta_{gyro})+ \alpha \cdot \theta_{acc},
 \end{equation}
-where $\alpha$ is the complementary gain, it determines how much weight the accelerometer data gets and how fast the value drifts towards the measurement of the accelerometer.
+where $\alpha$ is the complementary gain. 
+It determines how much weight the accelerometer data gets and how fast the value drifts towards the measurement of the accelerometer.
 Therefore the gyroscope functions as high-pass and the accelerometer as low-pass part of the basic complementary filter \cite{min2015complementary}.
 A commonly used value for $\alpha$ is 0.02, meaning every iteration, 2 percent of the new values come from the accelerometer.
-This correpsonds, with the used \unit{125}[hz] calculation, to frequency-border of the low and high-pass  of \unit{2.5}[hz] \cite{hajdu2017complementary}.
-An $\alpha$ of 1 means the resulting data is completely the one of the accelerometers, and the gyroscope has no influence. \\
+This corresponds, with the used \unit{125}[Hz] calculation, to frequency-border of the low and high-pass  of \unit{2.5}[Hz] \cite{hajdu2017complementary}.
+With $\alpha=1$ the result is the accelerometer data and the gyroscope has no influence. \\
 The complementary filter is best for orientation calculation without position change as the accelerometer value relies on calculating the gravity value.
-With an acceleration in one direction, this leads to a miscalculation of the gravitational vector and, therefore, a false orientation estimation.
-Simple orientation estimations of quadrocopters often just change the $\alpha$ gain to 0 while moving to avoid the miscalculation of the gravity vector \cite{mahony2005complementary}.
+Acceleration in one direction leads to a miscalculation of the gravitational vector and therefore false orientation estimates.
+Simple orientation estimations of quadrocopters often change the $\alpha$ gain to 0 while moving to avoid the miscalculation of the gravity vector \cite{mahony2005complementary}.
 
@@
 \section{Filter-concept}
-The pose orientation for the spherical robot is a concept with interconnected components.
-Both orientation and position estimates are computed based on the gyroscope and accelerometer measurements, taking into account the limitations of the system described in Section \ref{AssumptionsAndPrelimitations}.
+The pose estimation for spherical robots is a concept with interconnected components.
+Both orientation and position estimates are computed based on gyroscope and accelerometer measurements, taking into account the limitations of the system described in Section \ref{sec:AssumptionsAndPrelimitations}.
 The generalized overview of orientation and position estimation is shown in Figure \ref{generalized}.
+
+\begin{figure}%[hb]
+\includegraphics[width=\linewidth]{./graphics/imuJasperChartSimliefiedBIG.png}
+\caption{Diagram of the estimation.}
+\label{generalized}
+\end{figure}
 
@@
 The general idea is to distinguish between motion and standstill.
-In motion, the Madgwick filter is used, and without motion, the complementary filter.
+In motion the Madgwick filter is used and without motion the complementary filter.
 This approach of distinguishing the evaluation algorithm depending on the state of motion of a robot has been done before, like in \cite{Hertzberg2012}.
@@
 As both filters are applied with weights, this is done by an allocation of a given gain towards one or the other filter depending on the likelihood the sphere is moving or standing still. 
-Therefore, this requires some adapting mechanism and an estimation of how likely the sphere is to rotate and hence translate.
+Therefore, this requires an adapting mechanism and an estimation of how likely the sphere is to rotate and hence translate.
 Both filters share the same meaning of the gain, describing the influence of the accelerometer on the orientation.
@@
 \subsection{Position}
-As there is no absolute reference of position neither with calculation by gyroscope nor by the accelerometer, the filter can only try to improve the calculation of the velocity leading to the change of position.
-The straightforward approach is to double integrate the accelerometer, which leads to the earlier described unwanted behavior of exponential error integration and makes the position estimation completely unusable.
-The second straightforward approach multiplies the rotation speed with the known radius of the sphere and takes this as velocity, integrating once.
-The more precise and less noisy data of the gyroscope leads to a usable estimation of position. 
-There is no such extensive error integration, and more importantly: no exponentially growing error with time.
-However, this approach neglects slip and sliding, as it is one of our limitations and vertical movement.
-If the sphere slips, this approach will predict the full translation.
-Also, the other way around, if there is translation but just slight rotation because of sliding, it will predict too little translation. 
+As there is no absolute reference of position neither with calculation by gyroscope nor by the accelerometer, the filter can only try to improve the calculation of the velocity leading to the change in position.
+The straightforward approach is to double integrate the accelerometer, which leads to the earlier described unwanted behavior of exponential error integration and makes the position estimates completely unusable.
+
+An alternative approach multiplies the rotation speed with the known radius of the sphere and takes this as velocity, integrating once.
+The more precise and less noisy data of the gyroscope leads to usable position estimates. 
+There is no such extensive error integration, and more importantly: no exponentially growing error with time. However, if the sphere slips, this approach will predict the full translation.
+Also, if there is translation but just slight rotation because of sliding, it will predict too little translation. 
 Furthermore, integrating the rotation does not let us estimate vertical movement.
+Thus, this approach neglects potential slip and sliding (limitation 5) as well as vertical movement. 
 Every rotation will be predicted as translation in a given plane, which is most likely parallel to the ground. 
-For instance, if the sphere rolls up an obstacle, the length of the path is estimated to be the ground-plane.\\
-Our algorithm uses both approaches of position estimation and combines them, and limits all components according to the other components. 
+For instance, if the sphere rolls up an obstacle, the length of the path is estimated to be the ground-plane.
+
+Our algorithm combines both approaches of position estimation and limits all components according to the other components. 
 This implies a connection between translation and rotation. 
@@
 Therefore, the combination of both rather than only one is chosen.
-
-\begin{figure}[hb]
-\includegraphics[width=\linewidth]{./graphics/imuJasperChartSimliefiedBIG.png}
-\caption{Diagram of the estimation.}
-\label{generalized}
-\end{figure}
 
@@
 The orientation is a symbiosis of the Madgwick filter and complementary filter.
-For no- or slow-moving, the complementary filter is chosen; for fast-moving, the Madgwick filter.
+The complementary filter is chosen for no or slow motion, the Madgwick filter for fast motion.
 The Madgwick filter itself is untouched.
-The standard complimentary filter is adapted to work in quaternion representation,
- as the direct integration of the gyroscope give angles in the Euler domain.
+The standard complimentary filter works with Euler angles (roll, pitch, yaw (RPY)) resulting from the direct integration of the gyroscope while the Madgwick filter outputs quaternions.
 Quaternions were designed under the premise of being free from singularities, i.e., for a continuous change in orientation, there exists a continuous change in quaternion representation \cite{becker2017dealing}.
-This is not the case for the Euler angle, as a transition of a 359-degree roll to a 0-degree roll has no continuous value representation if limited to 360 degrees.
- But combining both leads to contradicting updates from each filter with respect to the other.
- Therefore the normal way of the complementary filter of the new orientation is 
+This is not the case for Euler angles, as a transition of a 359-degree roll to a 0-degree roll has no continuous value representation if limited to 360 degrees.
+Updating the orientation for the complementary filter implies a transition of each angle independently.
+Combining both representations leads to contradicting updates from each filter with respect to the other.
+Thus, the complementary filter is adapted to work in quaternion representation.
 %\begin{equation}
@@
 %\end{equation}
-not used, as it implies a transition of each value independently.
 Hence, the RPY-orientation is transformed to a quaternion, and then a quaternion slerp is performed \cite{dam1998quaternions}.
- The quaternion slerp works over the representation of two quaternions on a sphere. The transition from one quaternion to the other is now not done directly but always over the surface of the sphere.
+ The quaternion slerp utilizes the representation of two quaternions on a sphere. The transition from one quaternion to the other is not done directly but always over the surface of the sphere.
  This leads to the same handling of quaternions as the Madgwick-filter without endangering robustness due to using two different quaternions with very different values representing nearly the same position.
@@
  As the common values widely used for both values differ by one order of magnitude, the gain shifted to the complementary filter is divided by ten and the other way around.
- The values of the gyroscope axes determine in which manner to shift the overall gain to both Filters. 
- The defining values are the threshold until which value the algorithm should use a purely complementary filter and at which threshold a full Madgwick filter should determine the orientation. 
+ The values of the gyroscope axes determine in which manner to shift the overall gain to both filters. 
+ The defining values are the two thresholds for using solely the complementary filter on the one hand or a full Madgwick filter on the other hand. 
  These values have been determined empirically.
@@
  This avoids large accelerations of the orientation due to the rapid change from one algorithm to another.
- In our experiments, when starting from stillstand, the incipient Madgwick filter has a more rapid impact on the orientation values than the incipient complementary filter when stopping rotation. 
+ In our experiments, when starting from standstill, the incipient Madgwick filter has a more rapid impact on the orientation values than the incipient complementary filter when stopping rotation. 
 Therefore the transition between $\alpha$-gain of the complementary filter and the Madgwick-gain was chosen to have quadratic behavior. 
@@
 \label{eq:factors}
-\right.
+\right. ,
 \end{equation}
@@
 \begin{align}
-\beta &= \max(f_x,f_y,f_z)\label{eq:weight_autogain}\\
+\beta &= \max(f_x,f_y,f_z).\label{eq:weight_autogain}\\
 \intertext{Then the Madgwick gain $\gamma$ for a given autogain $\Theta$ is calculated by} 
-\gamma &= \Theta \cdot \beta \\
+\gamma &= \Theta \cdot \beta. \\
 \intertext{The complementary filter gain $\alpha$ is calculated by}
@@
 Thus, the Madgwick gain has a maximum of $\Theta$ and $\alpha$ a maximum of $\theta\cdot \Theta$. This ratio $\theta$ can be adapted to specific needs.
-We used $\theta = 0.1$ because of the two often used standard values $\gamma = 0.2$ and $\alpha= 0.02$, which have the ration $0.1$.
+We used $\theta = 0.1$ as the ratio between the two often used standard values $\gamma = 0.2$ and $\alpha= 0.02$.
 Hence for adapting the position estimation to a given robot, the function for calculating the factors needs to be adapted in such a way that it is a good indication for movement or standstill.
@@
 \end{bmatrix}
-= R \cdot \begin{bmatrix}
+= \M R \cdot \begin{bmatrix}
 0 \\0  \\ - 1
@@
 If there are non-zero components, there is an acceleration in that direction.
-Given the accelerometer measurement $\V a$ this leads to the
+Given the accelerometer measurement $\V a$ this leads to the velocity vector
 \begin{equation}
@@
 g'_z
-\end{bmatrix} \right)\, dt
+\end{bmatrix} \right)\, dt.
 \end{equation}
 These values are now coupled to the rotation of both axes other than their own.
-Therefore, the factors from the orientation step, representing the likelihood of rotation, are used as limiting factors.
-If one of the three factors is 1, this means there is a translation in this direction.
-When there is no rotation, it will set the velocity to 0.
-During fast movement, this allows exponential error integration. 
-This exponential error is taken care of later on.
-
+Therefore, the factors from the orientation step, representing the likelihood of rotation, are used as limiting factors:
+%TODO So würde man das programmieren, aber mathematisch ist das nicht korrekt.
 \begin{equation}
@@
 \max(f_y, f_z) \\ \max(f_x, f_z) \\ \max(f_x, f_y) 
-\end{bmatrix}
+\end{bmatrix}.
 \end{equation}
+If one of the three factors is 1, this means there is a translation in this direction.
+When there is no rotation, it will set the velocity to 0.
+During fast motion, this allows exponential error integration. 
+This exponential error is taken care of later on.
+
 The rotation into the world frame is computed as:
@@
 \intertext{The next step requires the velocity calculated by the rotation.
-Therefore the rotations (direct values from the gyro) are rotated into the world frame, i.e.,}
+Therefore the rotations (direct values from the gyro) are rotated into the world frame:}
 \begin{bmatrix}
@@
 \end{bmatrix}
-\cdot r \,.
+\cdot r \,,
 \end{equation}
+where $r$ is the radius of the sphere.
 As the gyro measures rad/s the value is divided by $2\pi$.
@@
 
- Next, the $v_{x\text{AccWorld}} $and $v_{y\text{AccWorld}} $ are limited depending on $v_{xz\text{GyroWorld}} $ and $ v_{yz\text{GyroWorld}}$, in our example, this was 120 percent of the gyro calculated velocity.
+ Next, the $v_{x\text{AccWorld}} $ and $v_{y\text{AccWorld}} $ are limited depending on $v_{xz\text{GyroWorld}} $ and $ v_{yz\text{GyroWorld}}$.
+In our experiments, this was 120 percent of the gyro calculated velocity.
  This ensures no exponential error integration when in motion.
@@
  To avoid large errors and unlimited double error integration, $v_{z\text{AccWorld}} $ is limited to the mean of $v_{xz\text{GyroWorld}} $ and $ v_{yz\text{GyroWorld}}$.
-With the Pythagorean theorem the velocity in $x$ and $y$ is calculated accepting the velocity in $z$-direction, which can not be derived by the velocity from gyro:
+With the Pythagorean theorem the velocity in $x$ and $y$ is calculated accepting the velocity in $z$-direction, which cannot be derived by the velocity from gyro:
  \begin{equation}
@@
 \sqrt{v_{xz\text{GyroWorld}}^2-v_{z\text{AccWorld}}^2}  \\ \sqrt{v_{yz\text{GyroWorld}}^2-v_{z\text{AccWorld}}^2 } 
-\end{bmatrix}
-\end{equation}.
+\end{bmatrix}.
+\end{equation}
 The last step to get the position is the integration of the velocity:
@@
  v_{z\text{AccWorld}} 
- \end{bmatrix} dt
-\end{equation}.
-Depending on the practical circumstances and quality of measurement, the factor $\beta$ is chosen, that the position relies more on the accelerometer calculated velocity or the gyro.
+ \end{bmatrix} dt.
+\end{equation}
+Depending on the practical circumstances and quality of measurement, the factor $\beta$ determines whether the position relies more on the accelerometer calculated velocity or the gyro.
 There should always be a mixture of both if there is slipping and sliding.
-$\beta =0.5$  results in just the mean of both approaches.
+$\beta =0.5$  results in the mean of both approaches.
 
@@
 Given that we use the fast inverse square root algorithm \cite{FastInvSqrt} for normalization, the trigonometric functions are considered the remaining bottleneck.
-Overall there are $4\times\arctan$, $4\times\sin$, $4\times\cos$, $1\times\arcsin$, $1\times\arccos$ and $5$ divisions. 
-The IMU measurements are collected at 125 Hz and are processed without problems on a raspberry pi.
+Overall there are 14 trigonometric function calls ($4\times \arctan$, $4\times\sin$, $4\times\cos$, $1\times \arcsin$, $1\times\arccos$) and $5$ divisions. 
+The IMU measurements are collected at \unit[125]{Hz} and are processed without problems on a raspberry pi.
 Hence no overwhelming computational load is to be expected.
@@
 \section{Evaluation}
-The proposed algorithm is evaluated by two sets of experiments performed with three Phidgets 3/3/3 IMUs mounted on a round acrylic glass plate.
+The proposed algorithm is evaluated by two sets of experiments performed with three Phidgets 3/3/3 IMUs mounted on a circular acrylic glass plate with a diameter of \unit[29]{cm}.
 For position experiments, the plate is put inside an acrylic glass sphere with the same radius as the glass plate.
@@
 \end{figure}
-Therefore the more expressive part is the change between motion and standstill, as in motion, both filters use the same algorithm but differ during standstill and slow motion.
+The more expressive part is the change between motion and standstill, as in motion, both filters use the same algorithm but differ during standstill and slow motion.
 As described in Section \ref{sec:SOA}, the Madgwick filter has a rather strong jitter during slow movements and standstill. 
-This comes due to the (too) large steps of gradient descent iterations.
+This is due to the (too) large steps of gradient descent iterations.
 To compensate for this, one needs to lower the gain, leading to more imprecise results during movement, as large iteration steps are beneficial there.
@@
 To evaluate specifically the change between slow motion and standstill, a further experiment is performed.
-Figure \ref{distance} show the distance to a precise external reference over a short excerpt of the experiment with three phases of slow motion and rotation at second 7, 9 and 11. 
+Figure \ref{distance} shows the distance to a precise external reference over a short excerpt of the experiment with three phases of slow motion and rotation at second 7, 9 and 11. 
 For external reference, we used a precise optical tracking system. 
@@
 \caption{Ground position (Z-axis, perpendicular to the ground) by integration of rotation and the fusion algorithm.
-An L-shape with 1 meter by 1 meter was performed with provoked slip(rotation without translation).
-The boxes are manually set points as a reference.
-The distance in meter between these points is written on the lines connecting them.
+An L-shape with \unit[1]{m} by \unit[1]{m} was performed with provoked slip (rotation without translation).
+For references the boxes are manually set points with the distance in meter between these points written on the lines connecting them.
 The rolling started in the upper right corner and ended in the bottom left corner.
@@
 This can be seen in Figure \ref{LTestSlide} where the one side of the L was interpreted exactly as long as the pure integration, i.e., approx.
-50 cm, and therefore the slide was not recognized, and the other part was correctly interpreted longer than just the integration and with 1.02 m as long as the actual trajectory.
+\unit[50]{cm}, and therefore the slide was not recognized, and the other part was correctly interpreted longer than just the integration and with \unit[1.02]{m} as long as the actual trajectory.
 This is to be blamed on the execution of the experiment, as it was not perfectly ensured only to start sliding with ongoing rotation, which is hard to achieve when manually guiding a sphere.
@@
 \caption{Ground position (Z-axis, perpendicular to the ground) by integration of rotation and the fusion algorithm.
-An L-shape with 1 meter by 1 meter was performed with provoked slide(translation without rotation).
-The boxes are manually set points as a reference.
-The distance in meter between these points is written on the lines connecting them.
+An L-shape with \unit[1]{m} by \unit[1]{m} was performed with provoked slide(translation without rotation).
+For reference the boxes are manually set points with the distance in meter between these points written on the lines connecting them.
 The rolling started in the lower right corner and ended in the upper left corner.}
@@
 To test the impact of the filter on 3D mapping, we used the prototype in Figure \ref{fig:prototype}.
-After 2 s of initial standstill, the sphere rolled approx. 4 m within 5 s in an indoor environment from one corner of the room to the opposing one.
+After \unit[2]{s} of initial standstill, the sphere rolled approx. \unit[4]{m} within \unit[5]{s} in an indoor environment from one corner of the room to the opposing one.
 We extracted the laserscan data, as well as the filtered pose using the presented algorithm.
 Without further pre- or post-processing, the pose data is combined with the laserscan data to create a 3D map.
-Figure \ref{fig:mapping} compares the result with another 3D map, using the in the same run recorded pose-estimation by the Madgwick filter and simple rotation integration.
-Therefore the same laserscan-data is mapped with different pose estimations.
+Figure \ref{fig:mapping} compares the result with another 3D map, using the pose estimation by the Madgwick filter and simple rotation integration on the same data.
+Therefore the same laserscan data is mapped with different pose estimates.
 The presented filter improves the resulting map compared to the Madgwick/integrator approach.
@@
 It is more detailed than the rest of the map because of the non-repeating flower-shaped scanning pattern, where point density increases with time.
-Note that the same detailed part is also present in the image below.
+Note that the same detailed part is also present in the right image.
 However, it is out of view (at the bottom right corner) as the Madgwick filter misses the initial pose estimation by a lot.
@@
 Needless to say, a lot of work remains to be done.
-The fusion of complementary and Madgwick orientation filter solves the shortcomings of the Madgwick when at a standstill, particularly the jitter.
+The fusion of complementary and Madgwick orientation filter solves the shortcomings of the Madgwick filter when at a standstill, particularly the jitter.
 A first qualitative evaluation shows the functionality of the filter.
-In further experiments, a quantitative Orientation comparison with a synchronized optical system needs to be performed.
+In further experiments, a quantitative orientation comparison with a synchronized optical system needs to be performed.
 Also, as further work, all puzzle pieces, this filter, a locomotion system, a scanning system need to be combined to an overall working prototype.
@@
 Still, there is a need for parameter tuning. 
-Sliding was not detected all the time reliably.
+Sliding was not all the time detected reliably.
 Here a mechanism for a delayed rotation needs to be investigated, that acceleration shortly before a rotation is integrated and taken into account once rotation starts.
